### 2019-02-19 draft/paper.tex
created template draft

\documentclass[useAMS,usenatbib]{mn2e}
\bibliographystyle{mn2e}
\pdfoutput=1
\pdfminorversion=5

%\usepackage{widetext}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{amssymb,amsmath,bm}
\usepackage{color}
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks,linkcolor=red,citecolor=blue,urlcolor=blue ]{hyperref}
\usepackage{multirow}
\usepackage[utf8]{inputenc}
\usepackage{balance}
\usepackage{enumitem}
\usepackage{lipsum}
\newcommand{\nv}{\hat{\bf n}}
\newcommand{\kalo}{Karhunen-Lo\`{e}ve~}
\newcommand{\cfh}{CFHTLenS~}
\newcommand{\jcap}{JCAP}
\newcommand{\mnras}{MNRAS}
\newcommand{\aap}{A\&A}
\newcommand{\aaps}{A\&AS}
\newcommand{\apjs}{ApJS}
\newcommand{\apj}{ApJ}
\newcommand{\apjl}{ApJL}
\newcommand{\prd}{Phys.~Rev.~D}
\newcommand{\prl}{Phys.~Rev.~Lett.}
\newcommand{\aj}{Astron. Journal}
\newcommand{\pasp}{Publications of the ASP}
\newcommand{\nar}{New Astronomy Review}
\newcommand{\procspie}{Proceedings of the SPIE}
\newcommand{\physrep}{Physics Reports}

\newcommand{\todo}[1]{{\bf TODO: #1}}
%\newcommand{\todo}[1]{$\,$}

\title[Covariance]{Covariance}
\author[Carlos Garc\`{i}a Garc\`{i}a]{Carlos Garc\`{i}a Garc\`{i}a$^1$\thanks{carlosgarcia@iff.csic.es}\\
$^{1}$Oxford Astrophysics, Department of Physics, Keble Road, Oxford, OX1 3RH, UK
}

\begin{document}
  \date{\today}
  \pagerange{1--18} \pubyear{2019}
  \maketitle

\begin{abstract}
\end{abstract}

\begin{keywords}
  cosmology: large-scale structure of the Universe -- methods: data analysis
\end{keywords}

\section{Introduction}

\section{Analytical Covariance}
\begin{itemize}
\item Naive Covariance
\item Approximation 1
\item Approximation 2
\end{itemize}

\section{Results}
For both flat and curved sky we should discuss:
\begin{itemize}
\item Simulations (sims, methods)
\item Approximation 1 - spin 0
\item Approximation 1 - spin 2
\item Approximation 2
\item Naive
\end{itemize}

For both flat and curved plots:
\begin{itemize}
\item Cov Naive
\item Cov approx 1
\item Cov approx 2
\item Cov approx 2 + deprojection
\end{itemize}

Plots:
\begin{itemize}
\item Plot $\chi^2$
\item KS($\chi^2$)
\item Eigenvalues
\end{itemize}

\section{Discussion}\label{sec:discussion}

\section*{Acknowledgements}

\setlength{\bibhang}{2.0em}
\setlength\labelwidth{0.0em}
\bibliography{paper}

\end{document}

### 2019-03-04 draft/paper.tex
First paragraphs on results section

@@
 
+\newcommand{\pcl}[3]{\hat C_{#1}^{#2 #3}}
+\newcommand{\avg}[1]{\langle #1 \rangle}
+\newcommand{\fsky}{f_{\mbox{sky}}}
+\newcommand{\clth}{C_l^{th}}
+\newcommand{\clf}{C_l^{fore}}
+
 \title[Covariance]{Covariance}
@@
 \item Naive Covariance
+  \begin{equation}
+    C^{abcd} = \frac{\pcl{l}ad \pcl{l}bc + \pcl{l}ac \pcl{l}bd}{f_{sky} (2l +
+      1) \Delta l} \delta_{ll'}
+  \end{equation}
 \item Approximation 1
@@
 
+In this section we will show the performance of each approximation and up to
+which point they hold. We will compare the analytical covariance matrices with
+those from simulated power spectra.
+
+The simulations were generated as follow. First, we starts from some realistic
+large scale structure power spectra ($\clth$) and mask, which determines the
+number of modes. To this theoretical power spectra, in some cases, we
+added some foregrounds to study their impact on the approximated covariances.
+The mock power spectra was computed as follows:
+\begin{eqnarray}
+  \clf = A (l + 1)^\beta\,,
+\end{eqnarray}
+where $A$ was fixed so that $\clf/\clth = 0.1$ at $l=400$ and $\beta$ was
+randomly chosen from $U[-3, -1)$ for large scale effects and fixed to $\beta =
+0$ for small scale ones.
+
+The resulting power spectra was then used to generate a map of the sky and
+two random fields (one of spin-0 and other of spin-2), in turn. Their
+correlations give  are used to obtain the simulated power spectra. In order to
+find the real power spectra, it is necessary to subtract the contribution of
+the mask, which mixes the closest modes:
+\begin{equation}
+  C^{obs}_l = \sum_{l'} M_{ll'} C_{l'}\,
+\end{equation}
+where $M$ depends on the mask. In general, $M$ is not invertible and, in order
+to be able to recover the true power spectra, $C_l$, one needs to bin the
+$l$-space. The bin width is given by the characteristics of the window
+function and must be, approximately, of the size of the range of the mixed
+modes. That way, each $\tilde C_l$ is almost independent of the others
+$\tilde C_{l'}$, and $M$ is invertible.
+
+The code to generate simulated power spectra and analyze the posterior
+results is publicly available in
+\url{https://github.com/damonge/PCLCovariance} and uses
+NaMaster~\cite{2018arXiv180909603A}, in order to compute the necessary fields
+and their correlations.
+
 \section{Discussion}\label{sec:discussion}
@@
 \section*{Acknowledgements}
+We would like to thank Eva-Maria Mueller for useful discussion. CGG is
+supported by AYA2015-67854-P from the Ministry of Industry, Science and
+Innovation of Spain and the FEDER funds and by the Spanish grant, partially
+funded by the ESF, BES-2016-077038.
 
@@
 \setlength\labelwidth{0.0em}
-\bibliography{paper}
+\bibliography{paper.bib}
 

### 2019-03-07 draft/paper.tex
script for theory power spectra

--- 2019-02-19 draft/paper.tex
@@
 \title[Covariance]{Covariance}
-\author[Carlos Garc\`{i}a Garc\`{i}a]{Carlos Garc\`{i}a Garc\`{i}a$^1$\thanks{carlosgarcia@iff.csic.es}\\
-$^{1}$Oxford Astrophysics, Department of Physics, Keble Road, Oxford, OX1 3RH, UK
+\author[C.
+Garc\'{i}a Garc\'{i}a et al.]{Carlos Garc\'{i}a Garc\'{i}a$^1$\thanks{carlosgarcia@iff.csic.es}, David Alonso$^2$\thanks{david.alonso@physics.ox.ac.uk}, Emilio Bellini$^2$\thanks{emilio.bellini@physics.ox.ac.uk}\\
+$^{1}$IFF\\
+$^{2}$Oxford Astrophysics, Department of Physics, Keble Road, Oxford, OX1 3RH, UK
 }
@@
 \section{Analytical Covariance}
+This section presents the analytical calculation of the pseudo-$C_\ell$ covariance.
+The treatment and approximations used are similar to those presented in \todo{cite} for spin-0 quantities, in \todo{cite} for particular spin-0 and spin-2 correlations and in \todo{cite} for the 3D density field.
+We reproduce the formalism here in order to provide a complete set of formulas applicable to any cross-correlation in both curved and flat skies.
+
 \begin{itemize}

### 2019-05-07 lsst_cov.py
started lsst_cov

@@ -1,0 +1,290 @@
+import numpy as np
+import pyccl as ccl
+import os
+
+
+def nofz(z,z0,sz,ndens):
+    return np.exp(-0.5*((z-z0)/sz)**2)*ndens/np.sqrt(2*np.pi*sz**2)
+
+# pz single bin
+z_1  = np.array([np.linspace(0,3,512)])
+pz_1 = np.array([nofz(z_1[0],0.955,0.13,7.55)])
+
+# pz new bins
+z  = np.array([np.linspace(0,3,512),np.linspace(0,3,512),np.linspace(0,3,512)])
+pz = np.array([nofz(z[0],0.955,0.13,7.55),nofz(z[1],1.355,0.13,7.55),nofz(z[1],1.755,0.13,7.55)])
+z_2  = np.array([np.linspace(0,3,512),np.linspace(0,3,512)])
+pz_2 = np.array([nofz(z[0],0.955,0.13,7.55),nofz(z[1],1.355,0.13,7.55)])
+
+pars = {
+    'h'       : 0.67,
+    'Omega_c' : 0.27,
+    'Omega_b' : 0.045,
+    'A_s'     : 2.1e-9,
+    'n_s'     : 0.96,
+    'w_0'     : -1.,
+    'w_a'     : -0.
+}
+
+
+
+ell_bp = np.load('./data/run_sph_ells.npz')['lsims'].astype(int)
+cov_th = np.load('./data/run_sph_covThTTTEEE.npz')['arr_0']
+cov_sim = np.load('./data/run_sph_covTTTEEE_clsims_0001-20000.npz')['arr_0']
+print(cov_th.shape)
+print(cov_sim.shape)
+
+
+
+def get_cosmo_ccl(pars):
+    cosmo = ccl.Cosmology(
+        h        = pars['h'],
+        Omega_c  = pars['Omega_c'],
+        Omega_b  = pars['Omega_b'],
+        A_s      = pars['A_s'],
+        n_s      = pars['n_s'],
+        w0       = pars['w_0'],
+        wa       = pars['w_a']
+        )
+    return cosmo
+
+
+def get_tracers_ccl(cosmo, z, pz):
+    n_bins = pz.shape[0]
+    bz = np.ones(pz.shape)
+    # Tracers
+    tracers = []
[…]

### 2019-05-10 lsst_cov.py
implemented covariance for lsst

@@ -11,10 +11,34 @@ z_1  = np.array([np.linspace(0,3,512)])
 pz_1 = np.array([nofz(z_1[0],0.955,0.13,7.55)])
 
 # pz new bins
-z  = np.array([np.linspace(0,3,512),np.linspace(0,3,512),np.linspace(0,3,512)])
-pz = np.array([nofz(z[0],0.955,0.13,7.55),nofz(z[1],1.355,0.13,7.55),nofz(z[1],1.755,0.13,7.55)])
 z_2  = np.array([np.linspace(0,3,512),np.linspace(0,3,512)])
-pz_2 = np.array([nofz(z[0],0.955,0.13,7.55),nofz(z[1],1.355,0.13,7.55)])
+pz_2 = np.array([nofz(z_2[0],0.955,0.13,7.55),nofz(z_2[1],1.355,0.13,7.55)])
+z_3  = np.array([np.linspace(0,3,512),np.linspace(0,3,512),np.linspace(0,3,512)])
+pz_3 = np.array([nofz(z_3[0],0.955,0.13,7.55),nofz(z_3[1],1.355,0.13,7.55),nofz(z_3[2],1.755,0.13,7.55)])
+z_n  = np.array([
+    np.linspace(0,3,512),
+    np.linspace(0,3,512),
+    np.linspace(0,3,512),
+    np.linspace(0,3,512),
+    np.linspace(0,3,512),
+    np.linspace(0,3,512),
+    np.linspace(0,3,512),
+    np.linspace(0,3,512),
+    np.linspace(0,3,512),
+    np.linspace(0,3,512)
+    ])
+pz_n = np.array([
+    nofz(z_n[0],0.955,0.13,7.55),
+    nofz(z_n[1],1.155,0.13,7.55),
+    nofz(z_n[2],1.255,0.13,7.55),
+    nofz(z_n[3],1.355,0.13,7.55),
+    nofz(z_n[4],1.455,0.13,7.55),
+    nofz(z_n[5],1.655,0.13,7.55),
+    nofz(z_n[6],1.755,0.13,7.55),
+    nofz(z_n[7],1.855,0.13,7.55),
+    nofz(z_n[8],1.955,0.13,7.55),
+    nofz(z_n[9],2.055,0.13,7.55)
+    ])
 
 pars = {
     'h'       : 0.67,
@@ -23,17 +47,18 @@ pars = {
     'A_s'     : 2.1e-9,
     'n_s'     : 0.96,
[…]

### 2019-05-15 lsst_cov.py
Few changes in lsst cov

@@ -1,6 +1,7 @@
 import numpy as np
 import pyccl as ccl
 import os
+import sys
 
 
 def nofz(z,z0,sz,ndens):
@@ -56,9 +57,6 @@ cov_th_tmp = np.load('./data/run_sph_covThTTTEEE_short.npz')['arr_0']
 cov_sim = np.load('./data/run_sph_covTTTEE_short_clsims_0001-20000.npz')['arr_0']
 cov_th = np.zeros((1023,1023))
 
-idx = np.triu_indices(1023)
-cov_th[idx] = cov_th_tmp[idx]
-cov_th.T[idx] = cov_th_tmp[idx]
 
 
 def get_cosmo_ccl(pars):
@@ -168,27 +166,6 @@ def get_cov_N(cov_lambda):
     return cov
 
 
-def get_cov_F(cov_c, cov_n, n_bte, n_ells):
-    cov_c_l = unflatten_covmat(cov_c, n_bte, n_ells)
-    cov_c_l = np.diagonal(cov_c_l,axis1=2,axis2=5)
-    cov_n_l = np.diagonal(cov_n,axis1=2,axis2=5)
-    cov_f = cov_c_l/cov_n_l
-    return cov_f
-
-
-def get_cov_R(cov_c, n_bte, n_ells):
-    cov_c_l = unflatten_covmat(cov_c, n_bte, n_ells)
-    cov_c_l = flatten_ell_covmat(cov_c_l, n_bte, n_ells)
-    cov_c_l_d = np.diagonal(cov_c_l,axis1=1,axis2=3)
-    cov_c_l_2 = np.multiply.outer(cov_c_l_d, cov_c_l_d)
-    cov_c_l_2 = np.diagonal(cov_c_l_2,axis1=0,axis2=3)
-    cov_c_l_2 = np.diagonal(cov_c_l_2,axis1=0,axis2=2)
-    cov_c_l_2 = np.moveaxis(cov_c_l_2,[0,1],[-3,-1])
-    cov_r = cov_c_l/np.sqrt(cov_c_l_2)
-    cov_r = unflatten_ell_covmat(cov_r, n_bte, n_ells)
-    return cov_r
-
-
 def get_idx_cov(pos, idx):
     if pos>=len(idx[0]):
         raise IOError('pos is larger than the number of elements. Maximum: {}'.format(len(idx[0])-1))
[…]

### 2019-05-15 notebooks/compute_covSims_same_mask.py
added notebooks for TTTEEE cov

@@ -1,0 +1,38 @@
+#!/usr/bin/python
+import numpy as np
+import os
+
+# pylint: disable=C0103
+
+prefix = 'run_sph_2b_same_mask'
+run_path = os.path.join('./simulations_outputs/', prefix, prefix)
+outdir = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
+output_path = os.path.join(outdir, prefix)
+
+
+##################
+ells = np.loadtxt(run_path + '_ells.txt')
+
+lmax = (ells < 2*512).sum()
+
+cl_ar = np.load(run_path + '_cl_0001-20000.npz')['arr_0']
+print(cl_ar.shape)
+
+Cls = [0, 1, 2] * 2  # 0 for T, 1 for E and 2 for B
+Cls_Bs_ar = []
+
+c = 0
+for i, Cli in enumerate(Cls):
+    for Clj in Cls[i:]:
+        Cls_Bs_ar.append([Cli, Clj])
+
+Cls_Bs_ar = np.array(Cls_Bs_ar)
+
+cl_ar_noBs = cl_ar[np.all(Cls_Bs_ar != 2, axis=1)]
+
+cl_for_C = np.concatenate(cl_ar_noBs[:, :, :lmax].swapaxes(1, 2))
+
+C = np.cov(cl_for_C)
+
+fname = run_path + '_covSims_TTTEEE_short' + ' _0001-20000.npz' # sims_suffix
+np.savez_compressed(fname, C)

### 2019-05-15 notebooks/join_covsTh_2bins.py

@@ -1,0 +1,104 @@
+#!/usr/bin/python
+import numpy as np
+import os
+
+# pylint: disable=C0103
+
+##############################################################################
+##############################################################################
+##############################################################################
+
+prefix = 'run_sph_2b_same_mask'
+run_path = os.path.join('./simulations_outputs/', prefix, prefix)
+output_dir = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
+out_run_path = os.path.join(output_dir, prefix)
+
+##############################################################################
+##############################################################################
+##############################################################################
+
+ells = np.loadtxt(run_path + '_ells.txt')
+
+lmax = (ells < 2*512).sum()
+
+# We want the covariance matrix of:
+# T1T1 T1E1 E1E1 T1T2 T1E2 E1T2 E1E2 T2T2 T2E2 E2E2
+
+nbins = 2
+nmaps = 4
+maps_bins = [0, 0, 1, 1]
+maps_spins = [0,  2] * nbins
+
+cl_indices = []
+cl_spins = []
+cl_bins = []
+for i in range(nmaps):
+    si = maps_spins[i]
+    for j in range(i, nmaps):
+        sj = maps_spins[j]
+        cl_indices.append([i, j])
+        cl_spins.append([si, sj])
+        cl_bins.append([maps_bins[i], maps_bins[j]])
+
+cov_indices = []
+cov_spins = []
+cov_bins = []
+for i, clij in enumerate(cl_indices):
+    for j, clkl in enumerate(cl_indices[i:]):
+        cov_indices.append(cl_indices[i] + cl_indices[i + j])
+        cov_spins.append(cl_spins[i] + cl_spins[i + j])
+        cov_bins.append(cl_bins[i] + cl_bins[i + j])
+
+cov_indices = np.array(cov_indices)
+cov_spins = np.array(cov_spins)
+cov_bins = np.array(cov_bins)
+
+##############################################################################
+##############################################################################
+##############################################################################
[…]

### 2019-05-15 notebooks/compute_covSims_same_mask.py
fix name

@@ -34,5 +34,5 @@ cl_for_C = np.concatenate(cl_ar_noBs[:, :, :lmax].swapaxes(1, 2))
 
 C = np.cov(cl_for_C)
 
-fname = run_path + '_covSims_TTTEEE_short' + ' _0001-20000.npz' # sims_suffix
+fname = output_path + '_covSims_TTTEEE_short_0001-20000.npz' # sims_suffix
 np.savez_compressed(fname, C)

### 2019-05-15 notebooks/run_sph_2b_same_mask_do_all_check_Full_cov.py
added script to do all checks for Full covs

@@ -1,0 +1,46 @@
+#!/usr/bin/python
+import common as co
+import numpy as np
+import os
+
+# pylint: disable=C0103
+
+prefix = 'run_sph_2b_same_mask'
+run_path = os.path.join('./simulations_outputs/', prefix, prefix)
+outdir = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
+output_path = os.path.join(outdir, prefix)
+figures_dir = os.path.join('./simulations_outputs/', prefix, 'figures')
+figures_path = os.path.join(figures_dir, prefix)
+
+
+##################
+ells = np.loadtxt(run_path + '_ells.txt')
+
+lmax = (ells < 2*512).sum()
+
+CovSims_path = output_path + '_covSims_TTTEEE_short_0001-20000.npz'
+CovTh_path = output_path + '_covTh_TTTEEE_short_2bins_same_mask.npz'
+
+CovSims_Full = np.load(CovSims_path)['arr_0']
+CovTh_Full = np.load(CovTh_path)['arr_0']
+
+cl_ar = np.load(run_path + '_cl_0001-20000.npz')['arr_0']
+Cls = [0, 1, 2] * 2  # 0 for T, 1 for E and 2 for B
+Cls_Bs_ar = []
+
+c = 0
+for i, Cli in enumerate(Cls):
+    for Clj in Cls[i:]:
+        Cls_Bs_ar.append([Cli, Clj])
+
+Cls_Bs_ar = np.array(Cls_Bs_ar)
+cl_ar_noBs = cl_ar[np.all(Cls_Bs_ar != 2, axis=1)]
+
+cl_for_C = np.concatenate(cl_ar_noBs[:, :, :lmax].swapaxes(1, 2))
+lbins_Full = np.concatenate([ells[:lmax]]*10)
+
+##################
+foutput = figures_path + '_Efstathiou'
+chi2_Full, corr_Full = co.do_all_checks(lbins_Full, cl_for_C, CovSims_Full,
+                                        CovTh_Full, modes="All TTTEEE",
+                                        row_cov=False, foutput=foutput + '_TTTEEE_Full')

### 2019-05-15 notebooks/run_sph_2b_same_mask_do_all_check_Full_cov.py
swap axis for cls

@@ -36,8 +36,8 @@ for i, Cli in enumerate(Cls):
 Cls_Bs_ar = np.array(Cls_Bs_ar)
 cl_ar_noBs = cl_ar[np.all(Cls_Bs_ar != 2, axis=1)]
 
-cl_for_C = np.concatenate(cl_ar_noBs[:, :, :lmax].swapaxes(1, 2))
-lbins_Full = np.concatenate([ells[:lmax]]*10)
+cl_for_C = np.concatenate(cl_ar_noBs[:, :, :lmax].swapaxes(1, 2)).T
+lbins_Full = np.concatenate([ells[:lmax]] * cl_ar_noBs.shape[0])
 
 ##################
 foutput = figures_path + '_Efstathiou'

### 2019-05-15 notebooks/run_sph_2b_same_mask_do_all_check_Full_cov.py
remove swap

@@ -36,7 +36,7 @@ for i, Cli in enumerate(Cls):
 Cls_Bs_ar = np.array(Cls_Bs_ar)
 cl_ar_noBs = cl_ar[np.all(Cls_Bs_ar != 2, axis=1)]
 
-cl_for_C = np.concatenate(cl_ar_noBs[:, :, :lmax].swapaxes(1, 2)).T
+cl_for_C = np.concatenate(cl_ar_noBs[:, :, :lmax])
 lbins_Full = np.concatenate([ells[:lmax]] * cl_ar_noBs.shape[0])
 
 ##################

### 2019-05-15 notebooks/run_sph_2b_same_mask_do_all_check_Full_cov.py
Revert "remove swap"
This reverts commit 41de8d492172bc12d53dffc9c6468ac5414be296.

@@ -36,7 +36,7 @@ for i, Cli in enumerate(Cls):
 Cls_Bs_ar = np.array(Cls_Bs_ar)
 cl_ar_noBs = cl_ar[np.all(Cls_Bs_ar != 2, axis=1)]
 
-cl_for_C = np.concatenate(cl_ar_noBs[:, :, :lmax])
+cl_for_C = np.concatenate(cl_ar_noBs[:, :, :lmax].swapaxes(1, 2)).T
 lbins_Full = np.concatenate([ells[:lmax]] * cl_ar_noBs.shape[0])
 
 ##################

### 2019-05-15 notebooks/join_covsTh_2bins.py
fixed order of cov dim. Still sth wrong. Perhaps triu_indices?

@@ -78,15 +78,15 @@ for i, indices in enumerate(cov_indices):
     nb2 = get_nelems_spin(s_b2)
 
     fname = out_run_path + '_cov_c{}{}{}{}_{}{}{}{}.npz'.format(*cov_spins[i], *cov_bins[i])
-    c = np.load(fname)['arr_0'].reshape((na1 * na2, l_bpw, nb1 * nb2, l_bpw))
+    c = np.load(fname)['arr_0'].reshape((l_bpw, na1 * na2, l_bpw, nb1 * nb2))
 
     #  TT -> 0; TE -> 0;  EE -> 0
-    cov_arr[i] = c[0, :, 0, :]
+    cov_arr[i] = c[:, 0, :, 0]
 
-cov_mat = np.empty((len(cl_bins), l_bpw, len(cl_bins), l_bpw))
+cov_mat = np.empty((l_bpw, len(cl_bins), l_bpw, len(cl_bins)))
 i, j = np.triu_indices(len(cl_bins))
-cov_mat[i, :,  j] = cov_arr
-cov_mat[j, :,  i] = cov_arr
+cov_mat[:, i, :, j] = cov_arr
+cov_mat[:, j, :, i] = cov_arr
 cov_mat = cov_mat.reshape((len(cl_bins) * l_bpw, len(cl_bins) * l_bpw))
 
 fname = out_run_path + '_covTh_TTTEEE_2bins_same_mask.npz'
@@ -94,10 +94,10 @@ np.savez_compressed(fname, cov_mat)
 
 ####################### Short version #############################
 
-cov_mat = np.empty((len(cl_bins), lmax, len(cl_bins), lmax))
+cov_mat = np.empty((lmax, len(cl_bins), lmax, len(cl_bins)))
 i, j = np.triu_indices(len(cl_bins))
-cov_mat[i, :,  j] = cov_arr[:, :lmax, :lmax]
-cov_mat[j, :,  i] = cov_arr[:, :lmax, :lmax]
+cov_mat[:, i, :, j] = cov_arr[:, :lmax, :lmax]
+cov_mat[:, j, :, i] = cov_arr[:, :lmax, :lmax]
 cov_mat = cov_mat.reshape((len(cl_bins) * lmax, len(cl_bins) * lmax))
 
 fname = out_run_path + '_covTh_TTTEEE_short_2bins_same_mask.npz'

### 2019-05-16 notebooks/join_covsTh_2bins.py
correct order is (n, l, n, l)

@@ -83,10 +83,10 @@ for i, indices in enumerate(cov_indices):
     #  TT -> 0; TE -> 0;  EE -> 0
     cov_arr[i] = c[:, 0, :, 0]
 
-cov_mat = np.empty((l_bpw, len(cl_bins), l_bpw, len(cl_bins)))
+cov_mat = np.empty((len(cl_bins), l_bpw, len(cl_bins), l_bpw))
 i, j = np.triu_indices(len(cl_bins))
-cov_mat[:, i, :, j] = cov_arr
-cov_mat[:, j, :, i] = cov_arr
+cov_mat[i, :, j, :] = cov_arr
+cov_mat[j, :, i, :] = cov_arr.swapaxes(1, 2)
 cov_mat = cov_mat.reshape((len(cl_bins) * l_bpw, len(cl_bins) * l_bpw))
 
 fname = out_run_path + '_covTh_TTTEEE_2bins_same_mask.npz'
@@ -94,11 +94,20 @@ np.savez_compressed(fname, cov_mat)
 
 ####################### Short version #############################
 
-cov_mat = np.empty((lmax, len(cl_bins), lmax, len(cl_bins)))
+cov_mat = np.empty((len(cl_bins), lmax, len(cl_bins), lmax))
 i, j = np.triu_indices(len(cl_bins))
-cov_mat[:, i, :, j] = cov_arr[:, :lmax, :lmax]
-cov_mat[:, j, :, i] = cov_arr[:, :lmax, :lmax]
+cov_mat[i, :, j, :] = cov_arr[:, :lmax, :lmax]
+cov_mat[j, :, i, :] = cov_arr[:, :lmax, :lmax].swapaxes(1, 2)
 cov_mat = cov_mat.reshape((len(cl_bins) * lmax, len(cl_bins) * lmax))
 
 fname = out_run_path + '_covTh_TTTEEE_short_2bins_same_mask.npz'
 np.savez_compressed(fname, cov_mat)
+
+# cov_mat = np.empty((len(cl_bins), len(cl_bins), lmax, lmax))
+# i, j = np.triu_indices(len(cl_bins))
+# cov_mat[i, j] = cov_arr[:, :lmax, :lmax]
+# cov_mat[j, i] = cov_arr[:, :lmax, :lmax].swapaxes(1, 2)
+# cov_mat = np.block([[mat for mat in cov_mati] for cov_mati in cov_mat])
+#
[…]

### 2019-05-17 draft/paper.tex
added introductory paragraph for results + update acknowledgments

--- 2019-03-07 draft/paper.tex
@@
 
+\newcommand{\pcl}[3]{\hat C_{#1}^{#2 #3}}
+\newcommand{\avg}[1]{\langle #1 \rangle}
+\newcommand{\fsky}{f_{\mbox{sky}}}
+\newcommand{\clth}{C_l^{th}}
+\newcommand{\clf}{C_l^{fore}}
+\newcommand{\cl}{C_l}
+
 \title[Covariance]{Covariance}
 \author[C.
-Garc\'{i}a Garc\'{i}a et al.]{Carlos Garc\'{i}a Garc\'{i}a$^1$\thanks{carlosgarcia@iff.csic.es}, David Alonso$^2$\thanks{david.alonso@physics.ox.ac.uk}, Emilio Bellini$^2$\thanks{emilio.bellini@physics.ox.ac.uk}\\
+Garc\'{i}a-Garc\'{i}a et al.]{Carlos Garc\'{i}a-Garc\'{i}a$^1$\thanks{carlosgarcia@iff.csic.es}, David Alonso$^2$\thanks{david.alonso@physics.ox.ac.uk}, Emilio Bellini$^2$\thanks{emilio.bellini@physics.ox.ac.uk}\\
 $^{1}$IFF\\
@@
 
+In this section we will show that our approximation is good and reliable; i.e.
+that can be used to obtain the cosmological parameters within the same error
+as using the covariance matrix from simulations.
+
+In order to generate the simulated $\cl$ we computed the correlations between
+Gaussian fields obtained from a fiducial power spectra. The covariance
+matrices were obtained from 20000 simulations, for the band powers determined
+by the fraction of sky observed and the maximum band power we were able to
+recover the fiducial power spectra for; i.e. $\sim 2 \times 512$. This
+corresponds to $l_{\mbox{bpw}} = 1023$. For the lecture of the sky maps and
+masks we used \textit{HEALPix}\footnote{\url{https://healpix.jpl.nasa.gov/}}
+and
+\textit{NaMaster}~\cite{2018arXiv180909603A}\footnote{\url{https://github.com/LSSTDESC/NaMaster}}
+for the computation of the pseudo-$\cl$ and the covariances. Our code is
+publicly available~\footnote{\url{https://github.com/damonge/PCLCovariance}}.
+
+% The simulations were generated as follow. First, we starts from some realistic
+% large scale structure power spectra ($\clth$) and mask, which determines the
+% number of modes. To this theoretical power spectra, in some cases, we
+% added some foregrounds to study their impact on the approximated covariances.
+% The mock power spectra was computed as follows:
+% \begin{eqnarray}
+%   \clf = A (l + 1)^\beta\,,
+% \end{eqnarray}
+% where $A$ was fixed so that $\clf/\clth = 0.1$ at $l=400$ and $\beta$ was
+% randomly chosen from $U[-3, -1)$ for large scale effects and fixed to $\beta =
+% 0$ for small scale ones.
+
+% The resulting power spectra was then used to generate a map of the sky and
+% two random fields (one of spin-0 and other of spin-2), in turn. Their
+% correlations give  are used to obtain the simulated power spectra. In order to
+% find the real power spectra, it is necessary to subtract the contribution of
+% the mask, which mixes the closest modes:
+% \begin{equation}
+%   C^{obs}_l = \sum_{l'} M_{ll'} C_{l'}\,
+% \end{equation}
+% where $M$ depends on the mask. In general, $M$ is not invertible and, in order
+% to be able to recover the true power spectra, $C_l$, one needs to bin the
+% $l$-space. The bin width is given by the characteristics of the window
+% function and must be, approximately, of the size of the range of the mixed
+% modes. That way, each $\tilde C_l$ is almost independent of the others
+% $\tilde C_{l'}$, and $M$ is invertible.
+
 \section{Discussion}\label{sec:discussion}
@@
 \section*{Acknowledgements}
+We would like to thank Eva-Maria Mueller for useful discussion. CGG is
+supported the Spanish grant BES-2016-077038, partially funded by the ESF and by
+AYA2015-67854-P from the Ministry of Industry, Science and Innovation of Spain
+and the FEDER funds. He was partially supported by a Balzan Fellowship while
+in Oxford. He would like to thank New College and the Department of Physics at
+Oxford for their hospitality. DA acknowledges support from STFC through an
+Ernest Rutherford Fellowship, grant reference ST/P004474/1.
 

### 2019-05-17 draft/paper.tex
more suggestive title + IFF address

@@
 
-\title[Covariance]{Covariance}
+\title[Reliable analytic approximation of pseudo-$\cl$ covariances]{Reliable analytic approximation of pseudo-$\cl$ covariances}
 \author[C.
-Garc\'{i}a-Garc\'{i}a et al.]{Carlos Garc\'{i}a-Garc\'{i}a$^1$\thanks{carlosgarcia@iff.csic.es}, David Alonso$^2$\thanks{david.alonso@physics.ox.ac.uk}, Emilio Bellini$^2$\thanks{emilio.bellini@physics.ox.ac.uk}\\
-$^{1}$IFF\\
+Garc\'{i}a-Garc\'{i}a et al.]{Carlos Garc\'{i}a-Garc\'{i}a$^{1,\,2}$\thanks{carlosgarcia@iff.csic.es}, David Alonso$^2$\thanks{david.alonso@physics.ox.ac.uk}, Emilio Bellini$^2$\thanks{emilio.bellini@physics.ox.ac.uk}\\
+$^{1}$ Instituto de Física Fundamental, Consejo Superior de Investigaciones
+Científicas, c/. Serrano 123, E–28006, Madrid, Spain\\
 $^{2}$Oxford Astrophysics, Department of Physics, Keble Road, Oxford, OX1 3RH, UK

### 2019-05-17 draft/paper.tex
added paragraph about foregrounds

@@
 and
-\textit{NaMaster}~\cite{2018arXiv180909603A}\footnote{\url{https://github.com/LSSTDESC/NaMaster}}
+\textit{NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}}~\cite{2018arXiv180909603A}
 for the computation of the pseudo-$\cl$ and the covariances. Our code is
 publicly available~\footnote{\url{https://github.com/damonge/PCLCovariance}}.
+
+The same results are obtained even in the presence of foregrounds. In
+particular, we tested the case with 100 foregrounds both at low scales and
+large scales, adding a mock power spectra computed as $\clf = A (l +
+1)^{\beta}$. Here, $A$ was fixed so that the ratio of the foreground power
+spectra and the theoretical one ($\clth$) was $\clf/\clth = 0.1$ at $l=400$
+and $\beta$ was randomly chosen from $U[-3, -1)$ for large scale effects and
+fixed to $\beta = 0$ for small scale ones.
 

### 2019-05-18 draft/paper.tex
added paragraph about chi2 dists for TTTT and EEEE

@@
 
-% The simulations were generated as follow. First, we starts from some realistic
-% large scale structure power spectra ($\clth$) and mask, which determines the
-% number of modes. To this theoretical power spectra, in some cases, we
-% added some foregrounds to study their impact on the approximated covariances.
-% The mock power spectra was computed as follows:
-% \begin{eqnarray}
-%   \clf = A (l + 1)^\beta\,,
-% \end{eqnarray}
-% where $A$ was fixed so that $\clf/\clth = 0.1$ at $l=400$ and $\beta$ was
-% randomly chosen from $U[-3, -1)$ for large scale effects and fixed to $\beta =
-% 0$ for small scale ones.
+Apart from checking that we recover the same cosmological predictions from the
+covariance matrices from both the analytical approximation and the
+simulations, we studied the differences between them.
+
+In this sense, one can see in Fig.~\ref{fig:TTTT_EEEE_chi2} that, we fully
+recover the $\chi^2$ distribution (defining $\chi^2 = \cl - \langle \cl
+\rangle C^{-1} \cl - \langle \cl \rangle$) for the TTTT case; whereas for the
+worst case (i.e. EEEE) we are able to recover its shape, although sifted
+toward lower values. This might be consequence of the fact that the inverse of
+the covariance matrix is biased. It is interesting to note that even
+the approximation of the mixing matrix as that of the case with spin 0, gives
+good results.
+
+\begin{figure*} %[htb]
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_TTTT_chi2.png}~
+  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_EEEE_chi2.png}
+  \caption{Comparison of the distributions found using different
+    covariances for the TTTT (left) and EEEE (right) cases.}
+  \label{fig:TTTT_EEEE_chi2}
+\end{figure*}
 

### 2019-05-18 draft/paper.tex
added paragraph about corr diff

@@
 
+The goodness of the approximation is also seen in the correlation matrices.
+The subtraction of the correlation matrix from the simulations and that from
+the best analytical approximation (which we have just seen gives almost same
+results as the spin-0 one) shows no structure but for the case of the larger
+scales of the EEEE modes (see Fig.~\ref{fig:TTTT_EEEE_corr}).
+
+\begin{figure*} %[htb]
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/run_sph_Efstathiou_TTTT_correlation_difference.png}~
+  \includegraphics[width=\columnwidth]{./figures/run_sph_Efstathiou_EEEE_correlation_difference.png}
+  \caption{Differences between the simulations and analytical approximation of
+    the correlation matrices for the TTTT (left) and EEEE (right) cases.}
+  \label{fig:TTTT_EEEE_chi2}
+\end{figure*}
+
 % The resulting power spectra was then used to generate a map of the sky and

### 2019-05-18 draft/paper.tex
added all plots for sph case

@@
 the approximation of the mixing matrix as that of the case with spin 0, gives
-good results.
+good results. The eigenvalues of the covariance matrix
+(Fig.~\ref{fig:TTTT_EEEE_eigv}) are also recovered within the 1\%, for all
+scales in the TTTT case, and for the smaller scales in the EEEE case. Finally,
+the analytical approximations are shown to describe the scale coupling (see
+Fig.~\ref{fig:TTTT_rows} that comes from the fact we are masking some regions
+of the sky.
 
@@
 \end{figure*}
+
+\begin{figure*} %[htb]
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_TTTT_reldev_eigval.png}~
+  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_EEEE_reldev_eigval.png}
+  \caption{Comparison of the covariance matrix eigenvalues for the TTTT (left)
+    and EEEE (right) cases.} 
+  \label{fig:TTTT_EEEE_eigv}
+\end{figure*}
+
+\begin{figure} %[htb]
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_TTTT_rows_cov_matrix.png}~
+  \caption{Mask induced scale correlations for the TTTT case.}
+  \label{fig:TTTT_rows}
+\end{figure}
 
@@
     the correlation matrices for the TTTT (left) and EEEE (right) cases.}
-  \label{fig:TTTT_EEEE_chi2}
+  \label{fig:TTTT_EEEE_corr}
 \end{figure*}
+
+These approximations do not work well for the B-modes, though. We can see in
+Fig.~\ref{fig:BBBB} that we are not able to recover with accuracy the first
+diagonal elements.
+
+\begin{figure} %[htb]
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_BBBB_check_diagonal_terms_4.png}
+  \caption{The analytical approximation are not able to reproduce accurately
+    the terms involving the B-modes. Here we show the most extreme case, the
+    corresponding to the BBBB modes.}
+  \label{fig:BBBB}
+\end{figure}
 

### 2019-05-18 notebooks/common.py
improved figures

@@ -9,7 +9,7 @@ import os
 DEFAULT_COLOR_CYCLE = plt.rcParams['axes.prop_cycle'].by_key()['color']
 DPI = 500
 FIGSIZE = (4, 3)
-FIGSIZE2 = (4, 6)
+FIGSIZE2 = (4, 4)
 
 # Diagonal check
 def check_Covariance_diagonal_terms(CovSims, CovTh, labelsTh, principal=True, foutput=None):
@@ -39,11 +39,11 @@ def check_Covariance_diagonal_terms(CovSims, CovTh, labelsTh, principal=True, fo
             if principal:
                 break
 
-    ax[0].set_ylabel('abs(Diag terms covariance matrix)')
+    ax[0].set_ylabel('abs(Diag terms cov.)')
     ax[0].set_yscale('log')
     ax[0].legend(loc=0)
 
-    ax[1].set_ylabel('Rel.dev. wrt. sim. cov.')
+    ax[1].set_ylabel('Rdev. wrt. sim. cov.')
     ax[1].set_xlabel('Element')
     ax[1].legend(loc=0)
 

### 2019-05-18 draft/paper.tex
slightly improve explanation row plot
added paragraph about the full 2-bins cov

@@
       1) \Delta l} \delta_{ll'}
+    \label{eq:naive}
   \end{equation}
@@
 Fig.~\ref{fig:TTTT_rows} that comes from the fact we are masking some regions
-of the sky.
+of the sky. The larger relative deviation for the better approximations are
+due to the fact that there is a division by $0$, while for the naive
+approximation, which is defined diagonal (see Eq.~\ref{eq:naive}), the
+relative deviation is exactly $-1$ for all values with $l \neq l'$.
 
@@
 
+Finally, the analytical covariance also works for 2-bin correlations. The
+$\chi^2$ distribution for the covariance matrix of the modes TT, TE and EE
+among two bins is shown in Fig.~\ref{fig:TTTEEE_chi2}. The eigenvalues
+are recovered well except for the lower ones that diverge up to $20\%$ (see
+Fig.~\ref{fig:TTTEEE_eigv}, this is not translated to any substantial
+difference in the correlation matrices (see Fig.~\ref{fig:TTTEEE_corr}).
+
+\begin{figure} %[htb]
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/run_sph_2b_same_mask_Efstathiou_TTTEEE_Full_chi2.png}
+  \caption{$\chi^2$ distribution check for the full analytical covariance
+    matrix for the modes TT, TE and EE among two bins.}
+  \label{fig:TTTEEE_chi2}
+\end{figure}
+
+\begin{figure} %[htb]
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/run_sph_2b_same_mask_Efstathiou_TTTEEE_Full_reldev_eigval.png}
+  \caption{Relative difference between the analytical and simulations
+    covariance matrix eigenvalues for the modes TT, TE and EE among two bins.}
+  \label{fig:TTTEEE_eigv}
+\end{figure}
+
+\begin{figure} %[htb]
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/run_sph_2b_same_mask_Efstathiou_TTTEEE_Full_correlation_difference.png}
+  \caption{Difference of the correlation matrices of the simulated $\cl$ and
+    the analytical one for the modes TT, TE and EE among two bins..}
+  \label{fig:TTTEEE_corr}
+\end{figure}
+
 % The resulting power spectra was then used to generate a map of the sky and

### 2019-05-20 notebooks/general-plots-paper.py
script to generate some general plots

@@ -1,0 +1,138 @@
+#!/usr/bin/python
+import healpy as hp
+import numpy as np
+import templates as tp
+import matplotlib.pyplot as plt
+import os
+
+# pylint: disable=C0103
+
+DPI = 500
+FSIZE1 = (4, 3)
+
+outdir = './general-plots/'
+
+if not os.path.isdir(outdir):
+    os.mkdir(outdir)
+
+##############################################################################
+######################### Fig with z-bins ####################################
+##############################################################################
+
+fname = os.path.join(outdir, 'pz.png')
+
+nell = 30000
+
+def nofz(z, z0, sz, ndens):
+    return np.exp(-0.5*((z-z0)/sz)**2)*ndens/np.sqrt(2*np.pi*sz**2)
+
+z = np.linspace(0, 2, 512)
+pz1 = nofz(z, 0.955, 0.13, 7.55)
+pz2 = nofz(z, 0.755, 0.13, 7.55)
+ndens1 = np.sum(pz1)*np.mean(z[1:]-z[:-1])*(180*60./np.pi)**2
+ndens2 = np.sum(pz2)*np.mean(z[1:]-z[:-1])*(180*60./np.pi)**2
+
+f, ax = plt.subplots(1, 1, figsize=FSIZE1)
+ax.plot(z, pz1, label='Bin 1')
+ax.plot(z, pz2, label='Bin 2')
+
+ax.set_ylabel('P(z)')
+ax.set_xlabel('z')
+
+ax.legend(loc=0)
+plt.tight_layout()
+plt.savefig(fname, dpi=DPI)
+plt.close()
+
+##############################################################################
+######################### Theoretical sph 2b Cls #############################
+##############################################################################
+
+fname = os.path.join(outdir, 'cls-sph-2b.png')
+
+ell_cl1bin = np.loadtxt('./data/cls_lss.txt')
+
+ell_cl2bin_file = np.load('./data/cls_lss_2bins.npz')
+ell, cl2bin, nls2bin = ell_cl2bin_file['ls'], ell_cl2bin_file['cls'], ell_cl2bin_file['nls']
+
+f, axs = plt.subplots(4, 4, figsize=(8, 6), gridspec_kw={'hspace': 0, 'wspace': 0}, sharex=True, sharey='row')
+
+ci = 0
+for i in range(cl2bin.shape[0]):
+    if i in (2, 5):
+        continue
+    cj = ci
+    for j in range(i, cl2bin.shape[0]):
[…]

### 2019-05-20 notebooks/general-plots-paper.py
added code for comparison between fiducial theoretical cl's 1 and 2 bins case full sky

@@ -50,7 +50,6 @@ plt.close()
 
 fname = os.path.join(outdir, 'cls-sph-2b.png')
 
-ell_cl1bin = np.loadtxt('./data/cls_lss.txt')
 
 ell_cl2bin_file = np.load('./data/cls_lss_2bins.npz')
 ell, cl2bin, nls2bin = ell_cl2bin_file['ls'], ell_cl2bin_file['cls'], ell_cl2bin_file['nls']
@@ -89,6 +88,55 @@ for i in range(axs.shape[0]):
     axs[i, -1].text(1, 0.5, labels[i], rotation=270,
                     transform=axs[i, -1].transAxes)
 
+
+plt.tight_layout()
+plt.savefig(fname, dpi=DPI)
+plt.close()
+
+##############################################################################
+########################## Compare 1bin sph Cls ##############################
+##############################################################################
+
+fname = os.path.join(outdir, 'compare-sph-cls-used-1bin.png')
+
+clTh = np.loadtxt('./data/cls_lss.txt', unpack=True)
+lTh = clTh[0][:3*512]
+clTh_TT = (clTh[1] + clTh[5])[:3*512]
+clTh_EE = (clTh[2] + clTh[6])[:3*512]
+clTh_TE = (clTh[4] + clTh[-1])[:3*512]
+
+f, axs = plt.subplots(1, 3, figsize=(8, 3), gridspec_kw={'hspace': 0, 'wspace': 0}, sharex=True, sharey='row')
+
+cl2bins_1 = (cl2bin + nls2bin)[:3, :3]
+cl2bins_2 = (cl2bin + nls2bin)[3:, 3:]
+i, j = np.triu_indices(3)
+
+#### TT #####
+axs[0].loglog(lTh, clTh_TT, label='1-bin Fid.')
+axs[0].loglog(ell, cl2bins_1[0, 0], label='1st 2-bin Fid.', ls='--')
+axs[0].loglog(ell, cl2bins_2[0, 0], label='2nd 2-bin Fid.', ls='--')
+axs[0].set_ylabel(r'$C_l$')
+
+axs[0].text(0.5, 0.9, r'$C_l^{gg}$', transform=axs[0].transAxes)
+
+#### TE #####
+axs[1].loglog(lTh, clTh_TE, label='1-bin Fid.')
[…]

### 2019-05-20 draft/paper.tex
added more info about simulations

@@
 \newcommand{\cl}{C_l}
+\newcommand{\red}[1]{{\color{red} #1}}
 
@@
 
-In order to generate the simulated $\cl$ we computed the correlations between
-Gaussian fields obtained from a fiducial power spectra. The covariance
-matrices were obtained from 20000 simulations, for the band powers determined
-by the fraction of sky observed and the maximum band power we were able to
-recover the fiducial power spectra for; i.e. $\sim 2 \times 512$. This
-corresponds to $l_{\mbox{bpw}} = 1023$. For the lecture of the sky maps and
-masks we used \textit{HEALPix}\footnote{\url{https://healpix.jpl.nasa.gov/}}
-and
+\subsection{Simulations}
+
+Simulations were drawn from Gaussian fields that followed a fiducial power
+spectrum. The fiducial cosmology (\red{For the 2-bin case}) is given by
+$\Omega_{cdm} = 0.27$, $\Omega_b = 0.045$, $h = 0.67$, $A_s = 2.1 10^{-9}$ and
+$n_s = 0.9$, where each term correspond, respectively, to the amount of cold
+dark matter and baryons today, the reduced Hubble constant, the primordial
+curvature fluctuation amplitude and the spectral index. The redshift bins are
+shown in Fig.~\ref{fig:pz} and the theoretical $\cl$ are in
+Fig.~\ref{fig:cl-2bins}. Since all parts of sky will not be equally observed, a
+realist mask (see Fig.~\ref{fig:mask}) was applied.  The results shown here
+were obtained using the same mas for each redshift bin, as it is expected to
+be that way in real observations. We checked, however, that using different
+masks yielded same results.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/pz.png}
+  \caption{Galaxy distribution redshift bins.}
+  \label{fig:pz}
+\end{figure}
+
+\begin{figure*}
+  \centering
+  \includegraphics[width=\textwidth]{./figures/cls-sph-2b.png}
+  \caption{Fiducial power spectra. We show only the $T$ and $E$ modes.}
+  \label{fig:cl-2bins}
+\end{figure*}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/mask-lss1.png}
+  \caption{Example of sky mask used in our analysis}
+  \label{fig:mask}
+\end{figure}
+
+The covariance matrices were obtained from 20000 simulations, for the band
+powers determined by the fraction of sky observed and the maximum band power
+we were able to recover the fiducial power spectra for; i.e. $\sim 2 \times
+512$. This corresponds to $l_{\mbox{bpw}} = 1023$. To work with the sky
+maps and masks we used
+\textit{HEALPix}\footnote{\url{https://healpix.jpl.nasa.gov/}}, whereas 
 \textit{NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}}~\cite{2018arXiv180909603A}
-for the computation of the pseudo-$\cl$ and the covariances. Our code is
-publicly available~\footnote{\url{https://github.com/damonge/PCLCovariance}}.
+was used for the computation of the pseudo-$\cl$ and the covariances. Our code
+is publicly
+available~\footnote{\url{https://github.com/damonge/PCLCovariance}}.
 
@@
 particular, we tested the case with 100 foregrounds both at low scales and
-large scales, adding a mock power spectra computed as $\clf = A (l +
-1)^{\beta}$. Here, $A$ was fixed so that the ratio of the foreground power
-spectra and the theoretical one ($\clth$) was $\clf/\clth = 0.1$ at $l=400$
-and $\beta$ was randomly chosen from $U[-3, -1)$ for large scale effects and
-fixed to $\beta = 0$ for small scale ones.
+large scales, adding a mock power spectra (e.g. Fig.~\ref{fig:fore}) computed as 
+\begin{equation}
+  \clf = A (l + 1)^{\beta}\,.
+  \label{eq:fore}
+\end{equation}
+Here, $A$ was fixed so that the ratio of the foreground power spectra and the
+theoretical one ($\clth$) was $\clf/\clth = 0.1$ at $l=400$ and $\beta$ was
+randomly chosen from $U[-3, -1)$ for large scale effects and fixed to $\beta =
+0$ for small scale ones.
 
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{./figures/foreground.png}
+  \caption{$\cl$ foregrounds used in order to check the reliability of the
+    analytical approximations in their presence. They did not affect. The
+    limiting cases for the large-scale contaminants are given by $\beta=-1,\,
+    -3$ in Eq.~\ref{eq:fore}}
+  \label{fig:fore}
+\end{figure}
+
+\subsection{$\chi^2$ distribution and covariance matrices}
 Apart from checking that we recover the same cosmological predictions from the
@@
 
+\appendix
+
+\section{Flat sky}
+
 \setlength{\bibhang}{2.0em}

### 2019-05-20 notebooks/general-plots-paper.py
use 1bin cl's for fores + rotate right labels cl 2bins

@@ -85,7 +85,7 @@ for i in range(axs.shape[0]):
     axs[i, i].xaxis.set_tick_params(labelbottom=True)
     axs[i, i].yaxis.set_tick_params(labelleft=True)
     axs[0, i].set_title(labels[i])
-    axs[i, -1].text(1, 0.5, labels[i], rotation=270,
+    axs[i, -1].text(1, 0.5, labels[i], fontsize=12,
                     transform=axs[i, -1].transAxes)
 
 
@@ -159,18 +159,22 @@ plt.close()
 ############################# Foregrounds ####################################
 ##############################################################################
 
-cl_ss = tp.create_cl_templates(ell, cl2bin[0, 0] + nls2bin[0, 0], exp_range=(0, 0), N=1)
-cl_ls1 = tp.create_cl_templates(ell, cl2bin[0, 0] + nls2bin[0, 0], exp_range=(-3, -3), N=1)[0]
-cl_ls2 = tp.create_cl_templates(ell, cl2bin[0, 0] + nls2bin[0, 0], exp_range=(-1, -1), N=1)[0]
+# cl_ss = tp.create_cl_templates(ell, cl2bin[0, 0] + nls2bin[0, 0], exp_range=(0, 0), N=1)
+# cl_ls1 = tp.create_cl_templates(ell, cl2bin[0, 0] + nls2bin[0, 0], exp_range=(-3, -3), N=1)[0]
+# cl_ls2 = tp.create_cl_templates(ell, cl2bin[0, 0] + nls2bin[0, 0], exp_range=(-1, -1), N=1)[0]
+
+cl_ss = tp.create_cl_templates(lTh, clTh_TT,  exp_range=(0, 0), N=1)
+cl_ls1 = tp.create_cl_templates(lTh, clTh_TT, exp_range=(-3, -3), N=1)[0]
+cl_ls2 = tp.create_cl_templates(lTh, clTh_TT, exp_range=(-1, -1), N=1)[0]
 
 fname = os.path.join(outdir, 'foreground.png')
 
 f, ax = plt.subplots(1, 1, figsize=FSIZE1)
-ax.loglog(ell, cl_ls1, c='orange')
-ax.loglog(ell, cl_ls2, c='orange', label='Large scales')
-ax.fill_between(ell, cl_ls1, cl_ls2, facecolor="orange", alpha=0.5)
+ax.loglog(lTh, cl_ls1, c='orange')
[…]

### 2019-05-20 draft/paper.tex
stuff about math

--- 2019-05-18 draft/paper.tex
@@
 \usepackage{bm}
-\usepackage{amssymb,amsmath,bm}
+\usepackage{amssymb,amsmath,bm,dsfont}
 \usepackage{color}
@@
 \newcommand{\avg}[1]{\langle #1 \rangle}
-\newcommand{\fsky}{f_{\mbox{sky}}}
-\newcommand{\clth}{C_l^{th}}
-\newcommand{\clf}{C_l^{fore}}
-\newcommand{\cl}{C_l}
+\newcommand{\fsky}{f_{\mbox{\rm sky}}}
+\newcommand{\clth}{C_\ell^{th}}
+\newcommand{\clf}{C_\ell^{fore}}
+\newcommand{\cl}{C_\ell}
+\newcommand{\summ}[1]{\sum_{\bf #1}\Delta #1^2}
+\newcommand{\Ylm}[3]{{\sf E}^{#1}_{\bf #2}({\bf #3})}
 
-\title[Reliable analytic approximation of pseudo-$\cl$ covariances]{Reliable analytic approximation of pseudo-$\cl$ covariances}
+\title[Disconnected pseudo-$C_\ell$ covariances for projected datasets]{Disconnected pseudo-$C_\ell$ covariances for projected datasets}
 \author[C.
@@
 
-\section{Analytical Covariance}
+\section{Analytical Gaussian covariances}
+
+  \subsection{Preliminaries}
+    We will deal with spin-0 and spin-2 fields defined on a 2-dimensional space.
+We will use a generic notation that can then be applied to both curved and flat skies.
+Given a field ${\bf a}({\bf x})$, with 1 (spin-0) or 2 components (spin-2), and defined on the coordinates ${\bf x}$, we define its generalized Fourier coefficients as
+    \begin{equation}
+      {\bf a}_{\bf k}=\summ{x}\,\Ylm{\dag}{k}{x}\,{\bf a}({\bf x}),
+    \end{equation}
+    where the operator $\summ{x}$ denotes a integral or sum of all coordinates ${\bf x}$, and $\Ylm{\dag}{k}{x}$ are a set of orthogonal functions.
+We will also assume that the $\Ylm{\dag}{k}{x}$ are a \emph{complete} set of basis functions, in which case:
+    \begin{align}
+      &\summ{x}\,\Ylm{\dag}{k}{x}\Ylm{}{l}{x}=\mathds{1}\Delta^x({\bf k},{\bf l}),\\
+      &\summ{k}\,\Ylm{}{k}{x}\Ylm{\dag}{k}{y}=\mathds{1}\Delta^k({\bf x},{\bf y}),
+    \end{align}
+    where $\summ{k}$ denotes an integral over all possible generalised Fourier coefficients ${\bf k}$, and $\Delta^x$ and $\Delta^k$ are generalized delta functions:
+    \begin{align}
+      &\summ{k}\,f({\bf k})\Delta^x({\bf k},{\bf l})\equiv f({\bf l}),\\
+      &\summ{x}\,f({\bf x})\Delta^k({\bf x},{\bf y})\equiv f({\bf y}).
+    \end{align}
+
+    Finally, we will assume that all fields are Gaussian stochastic fields that are additionally statistically isotropic.
+As a consequence of this, different generalised Fourier modes are uncorrelated:
+    \begin{equation}
+      \langle {\bf a}_{\bf k}{\bf b}^\dag_{\bf l}\rangle\equiv {\sf C}^{ab}_k\,F\,\Delta^x({\bf k},{\bf l}),
+    \end{equation}
+    where $F$ is a volume factor (see below) and ${\sf C}^{ab}_k$ is the power spectrum.
+
+    All the functions and operators above can be specialised to fields defined on the sphere or the 2D plane (flat sky) as follows:
+
+  \subsection{The pseudo-$C_\ell$ method}
+
+  \subsection{Covariance matrices}
+
+This section presents the analytical approximations used here to estimate the pseudo-$C_\ell$ Gaussian covariance matrix.
+
 This section presents the analytical calculation of the pseudo-$C_\ell$ covariance.

### 2019-05-20 draft/paper.tex

--- 2019-05-20 draft/paper.tex
@@
 \newcommand{\avg}[1]{\langle #1 \rangle}
-\newcommand{\fsky}{f_{\mbox{sky}}}
-\newcommand{\clth}{C_l^{th}}
-\newcommand{\clf}{C_l^{fore}}
-\newcommand{\cl}{C_l}
+\newcommand{\fsky}{f_{\mbox{\rm sky}}}
+\newcommand{\clth}{C_\ell^{th}}
+\newcommand{\clf}{C_\ell^{fore}}
+\newcommand{\cl}{C_\ell}
+\newcommand{\summ}[1]{\sum_{\bf #1}\Delta #1^2}
+\newcommand{\Ylm}[3]{{\sf E}^{#1}_{\bf #2}({\bf #3})}
 \newcommand{\red}[1]{{\color{red} #1}}
 
-\title[Reliable analytic approximation of pseudo-$\cl$ covariances]{Reliable analytic approximation of pseudo-$\cl$ covariances}
+\title[Disconnected pseudo-$C_\ell$ covariances for projected datasets]{Disconnected pseudo-$C_\ell$ covariances for projected datasets}
 \author[C.

### 2019-05-21 draft/paper.tex
bs

@@
 \usepackage{lipsum}
-\newcommand{\nv}{\hat{\bf n}}
+\newcommand{\nv}{\hat{\mathbf{\theta}}}
 \newcommand{\kalo}{Karhunen-Lo\`{e}ve~}
@@
     \end{align}
+    For a spin-$s$ quantity, $\Ylm{}{l}{x}$ can be written in terms of a differential operator $\eth$ and a set of scalar orthogonal functions $q({\bf l},{\bf x})$ as:
+    \begin{equation}
+      \Ylm{}{l}{x}=-\frac{\beta_{\ell,s}}{2}\left(
+      \begin{array}{cc}
+        \eth^s+\bar{\eth}^s & i(\eth^s-\bar{\eth}^s)\\
+        -i(\eth^s-\bar{\eth}^s) & \eth^s+\bar{\eth}^s
+      \end{array}\right) q({\bf l},{\bf x}).
+    \end{equation}
 
@@
 
+    \begin{table*}
+     \centering
+     \begin{tabular}{|c|c|c|}
+     \hline
+      Mathematical object & Curved sky & Flat sky (continuum $\rightarrow$ discretized) \\
+      \hline
+      \(\displaystyle {\bf l} \) & \(\displaystyle (\ell,m) \) & \(\displaystyle (l_x,l_y) \) \\
+      \(\displaystyle \summ{l} \) & \(\displaystyle \sum_{\ell=0}^{\infty} \sum_{m=-\ell}^{\ell} \) & \(\displaystyle \int \frac{dl^2}{2\pi} \rightarrow \sum_{\bf k}\frac{2\pi}{L_xL_y}\) \\
+      \(\displaystyle \Delta^x({\bf l},{\bf l}')\) & \(\displaystyle \delta^K_{\ell\ell'}\delta^K_{mm'}\) & \(\displaystyle 2\pi\,\delta^D({\bf l}-{\bf l}')\rightarrow \delta^K_{l_xl'_x}\delta^K_{l_yl'_y}\frac{L_xL_y}{(2\pi)^2}2\pi\) \\
+      \(\displaystyle {\bf x}\) & \(\displaystyle \nv\equiv(\theta,\varphi) \) & \(\displaystyle (x,y)\) \\
+      \(\displaystyle \summ{x} \) & \(\displaystyle \int_0^{\phi} d\phi \int_{-1}^{1} d(\cos\theta)\) & \(\displaystyle \int \frac{dx^2}{2\pi}\rightarrow\sum_{\bf x}\frac{L_xL_y}{2\pi N_xN_y}\) \\
+      \(\displaystyle \Delta^k(\nv,\nv')\) & \(\displaystyle \delta^D(\cos\theta-\cos\theta')\delta^D(\varphi-\varphi')\) & \(\displaystyle 2\pi\,\delta^D({\bf x}-{\bf y})\rightarrow\delta^K_{xx'}\delta^K_{yy'}\frac{N_xN_y}{L_xL_y}2\pi\) \\
+      \(\displaystyle q({\bf l},{\bf x})\) & \(\displaystyle Y_{\ell m}(\nv)\) & \(\displaystyle e^{i{\bf l}\cdot{\bf x}}\) \\
+      \(\displaystyle \eth\,_sf\) & \(\displaystyle -(\sin\theta)^s\left(\partial_\theta+i\frac{\partial_\varphi}{\sin\theta}\right)(\sin\theta)^{-s}\,_sf(\nv)\) & \(\displaystyle (\partial_x-i\partial_y)\,_sf\) \\
+      \(\displaystyle \bar{\eth}\,_sf\) & \(\displaystyle -(\sin\theta)^{-s}\left(\partial_\theta-i\frac{\partial_\varphi}{\sin\theta}\right)(\sin\theta)^s\,_sf(\nv)\) & \(\displaystyle (\partial_x+i\partial_y)\,_sf\) \\
+      \(\displaystyle F \) & \(\displaystyle 1 \) & \(\displaystyle (2\pi)^{-1}\) \\
+      \hline
+     \end{tabular}
+     \caption{Lookup table describing the generalized notation used here in for quantities defined on the sphere (second column) and on the flat 2D plane (third column).
+For the flat-sky case, we also provide expressions for a discretized, finite 2D plane with }
+
+    \end{table*}
+
     All the functions and operators above can be specialised to fields defined on the sphere or the 2D plane (flat sky) as follows:
+    Curved sky:
 
@@
 
-\subsection{$\chi^2$ distribution and covariance matrices}
+\subsection{Visual inspection}
+\todo{Rows plot para todas las combinaciones}
+\todo{diff-correlation TTTT BBBB}
+\todo{Eigenvalue plot}
+
+\subsection{Quantitative checks}
+\todo{Plot $\chi^2$ TT-TT, TE-TE, EE-EE}
+
+\subsection{Contaminant deprojection}
+\todo{}
+
+\subsection{Impact on parameter estimation}
+\todo{$\chi^2$ plot para 2 bines}
+\todo{diff-correlaton plot para 2 bines ver Fig. 8  de 1809.09603}
+\todo{Contours}
+
+\subsection{$\chi^2$ distribution and covariance matrices - old section}
 Apart from checking that we recover the same cosmological predictions from the
@@
 \section{Flat sky}
+\todo{Plot mascara}
+\todo{Plot chi2 TT-TT TE-TE, EE-EE}
 

### 2019-05-21 notebooks/general-plots-paper.py
added mask-lss2

@@ -151,7 +151,20 @@ fname = os.path.join(outdir, 'mask-lss1.png')
 fmask = "./data/mask_lss_sph1.fits"
 mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=512)
 
-hp.mollview(mask_lss)
+hp.mollview(mask_lss, title="", cbar=False, coord=['G', 'C'])
+plt.savefig(fname, dpi=DPI)
+plt.close()
+
+##############################################################################
+############################# Mask sph1 ######################################
+##############################################################################
+
+fname = os.path.join(outdir, 'mask-lss2.png')
+
+fmask = "./data/mask_lss_sph2.fits"
+mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=512)
+
+hp.mollview(mask_lss, title="", coord=['G', 'C'])
 plt.savefig(fname, dpi=DPI)
 plt.close()
 

### 2019-05-21 notebooks/general-plots-paper.py
added masks flat

@@ -169,6 +169,32 @@ plt.savefig(fname, dpi=DPI)
 plt.close()
 
 ##############################################################################
+############################# Mask flat1 ######################################
+##############################################################################
+
+fname = os.path.join(outdir, 'mask-lss_flat1.png')
+
+fmask = "./data/mask_lss_flat.fits"
+mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=512)
+
+hp.mollview(mask_lss, title="", coord=['G', 'C'])
+plt.savefig(fname, dpi=DPI)
+plt.close()
+
+##############################################################################
+############################# Mask flat2 ######################################
+##############################################################################
+
+fname = os.path.join(outdir, 'mask-lss_flat2.png')
+
+fmask = "./data/mask_lss_flat2.fits"
+mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=512)
+
+hp.mollview(mask_lss, title="", coord=['G', 'C'])
+plt.savefig(fname, dpi=DPI)
+plt.close()
+
+##############################################################################
 ############################# Foregrounds ####################################
 ##############################################################################
 

### 2019-05-21 notebooks/general-plots-paper.py
fixed all-cls + fixed flat mask

@@ -1,4 +1,5 @@
 #!/usr/bin/python
+import flatmaps as fm
 import healpy as hp
 import numpy as np
 import templates as tp
@@ -56,6 +57,7 @@ ell, cl2bin, nls2bin = ell_cl2bin_file['ls'], ell_cl2bin_file['cls'], ell_cl2bin
 
 f, axs = plt.subplots(4, 4, figsize=(8, 6), gridspec_kw={'hspace': 0, 'wspace': 0}, sharex=True, sharey='row')
 
+labels = [r'$\delta_1$', r'$\gamma_{E,\, 1}$', r'$\delta_2$', r'$\gamma_{E,\, 2}$']
 ci = 0
 for i in range(cl2bin.shape[0]):
     if i in (2, 5):
@@ -65,8 +67,10 @@ for i in range(cl2bin.shape[0]):
         if j in (2, 5):
             continue
         axs[ci, cj].loglog(ell, cl2bin[i, j], label='Th. w/o noise')
-        axs[ci, cj].loglog(ell, cl2bin[i, j] + nls2bin[ci, cj], label='Th.', ls='--')
+        axs[ci, cj].loglog(ell, cl2bin[i, j] + nls2bin[i, j], label='Th.', ls='--')
         axs[ci, cj].loglog(ell, nls2bin[i, j], label='Noise')
+        axs[ci, cj].text(0.8, 0.9, "({}, {})".format(labels[ci], labels[cj]),
+                         transform=axs[ci, cj].transAxes)
         cj += 1
 
     ci += 1
@@ -77,16 +81,15 @@ i, j = np.triu_indices(axs.shape[0])
 for ax in axs[j, i].reshape((-1)):
     ax.set_visible(False)
 
-labels = ['T1', 'E1', 'T2', 'E2']
 for i in range(axs.shape[0]):
     axs[i, i].set_xlabel('l')
     axs[i, i].set_ylabel(r'$C_l$')
     axs[i, i].set_visible(True)
     axs[i, i].xaxis.set_tick_params(labelbottom=True)
     axs[i, i].yaxis.set_tick_params(labelleft=True)
-    axs[0, i].set_title(labels[i])
-    axs[i, -1].text(1, 0.5, labels[i], fontsize=12,
-                    transform=axs[i, -1].transAxes)
+    # axs[0, i].set_title(labels[i])
[…]

### 2019-05-21 draft/paper.tex
section 2.2 done

@@
 
+\newcommand{\PCL}{pseudo-$C_\ell$~}
 \newcommand{\pcl}[3]{\hat C_{#1}^{#2 #3}}
@@
 
-  \subsection{Preliminaries}
+  \subsection{Preliminaries}\label{ssec:theory.prelim}
+    \todo{cites everywhere}
     We will deal with spin-0 and spin-2 fields defined on a 2-dimensional space.
@@
     \begin{equation}
-      \langle {\bf a}_{\bf k}{\bf b}^\dag_{\bf l}\rangle\equiv {\sf C}^{ab}_k\,F\,\Delta^x({\bf k},{\bf l}),
+      \langle {\bf a}_{\bf k}{\bf b}^\dag_{\bf l}\rangle\equiv {\sf C}^{ab}_\ell\,F\,\Delta^x({\bf k},{\bf l}),
     \end{equation}
@@
       \(\displaystyle {\bf l} \) & \(\displaystyle (\ell,m) \) & \(\displaystyle (l_x,l_y) \) \\
-      \(\displaystyle \summ{l} \) & \(\displaystyle \sum_{\ell=0}^{\infty} \sum_{m=-\ell}^{\ell} \) & \(\displaystyle \int \frac{dl^2}{2\pi} \rightarrow \sum_{\bf k}\frac{2\pi}{L_xL_y}\) \\
+      \(\displaystyle \summ{l} \) & \(\displaystyle \sum_{\ell=0}^{\infty} \sum_{m=-\ell}^{\ell} \) & \(\displaystyle \int \frac{dl^2}{2\pi} \rightarrow \sum_{\bf l}\frac{2\pi}{L_xL_y}\) \\
       \(\displaystyle \Delta^x({\bf l},{\bf l}')\) & \(\displaystyle \delta^K_{\ell\ell'}\delta^K_{mm'}\) & \(\displaystyle 2\pi\,\delta^D({\bf l}-{\bf l}')\rightarrow \delta^K_{l_xl'_x}\delta^K_{l_yl'_y}\frac{L_xL_y}{(2\pi)^2}2\pi\) \\
@@
      \end{tabular}
-     \caption{Lookup table describing the generalized notation used here in for quantities defined on the sphere (second column) and on the flat 2D plane (third column).
-For the flat-sky case, we also provide expressions for a discretized, finite 2D plane with }
-
+     \caption{Lookup table describing the generalized notation introduced in Section \ref{ssec:theory.prelim}p for quantities defined on the sphere (second column) and on the flat 2D plane (third column).
+For the flat-sky case, we also provide expressions for a discretized, finite 2D plane with periodic boundary conditions.
+In this case, the map has dimensions $(L_x,L_y)$ subdivided into $(N_x,N_y)$ equi-spaced pixels in $(x,y)$.
+$\delta^D$ and $\delta^K$ are the Dirac and Kronecker delta functions respectively.}\label{tab:notation}
     \end{table*}
 
-    All the functions and operators above can be specialised to fields defined on the sphere or the 2D plane (flat sky) as follows:
-    Curved sky:
+    All the functions and operators above can be specialised to fields defined on the sphere or the 2D plane (flat sky approximation) as described in Table \ref{tab:notation}.
 
   \subsection{The pseudo-$C_\ell$ method}
+    This section provides a very brief introduction to the \PCL power spectrum estimator. Further details can be found in \todo{cite}. 
+    In any practical situation we do not have access to maps of a given field ${\bf a}$ over the full sky, but rather to a weighted or masked version of them
+    \begin{equation}
+      \tilde{\bf a}({\bf x})\equiv w^a({\bf x}){\bf a}({\bf x}),
+    \end{equation}
+    where $w^a$ is commonly called the ``mask''. Due to the convolution theorem, the generalized Fourier coefficients of the masked field will be a convolution of the mask and true field coefficients:
+    \begin{align}\nonumber
+      \tilde{\bf a}_{\bf l}=&\summ{k}\left[\summ{x}\,w^a({\bf x})\,\Ylm{\dag}{l}{x}\,\Ylm{}{k}{x}\right]{\bf a}_{\bf k}\\
+                      \equiv&\summ{l}\,^a{\sf M}_{{\bf l}{\bf k}}\,{\bf a}_{\bf k},
+    \end{align}
+    where we have defined the mode-coupling coefficients $\,^a{\sf M}_{{\bf l}{\bf k}}$ in the second line.
+
+    Correlating the generalized Fourier coefficients of two masked fields therefore yields a mode-coupled version of their true underlying power spectrum:
+    \begin{align}\nonumber
+      \left\langle\tilde{\bf a}_{\bf l}\tilde{\bf b}^\dag_{\bf l}\right\rangle=&\summ{k}\summ{q}\,^a{\sf M}_{{\bf l}{\bf k}}\left\langle{\bf a}_{\bf k}{\bf b}^\dag_{\bf q}\right\rangle\,^b{\sf M}^\dag_{{\bf l}{\bf q}}\\
+                                                                              =&F\,\summ{k}\,^a{\sf M}_{{\bf l}{\bf k}}{\sf C}^{ab}_k\,^b{\sf M}^\dag_{{\bf l}{\bf k}}
+    \end{align}
+
+    The \PCL estimator then proceeds in two steps:
+    \begin{enumerate}
+      \item We first bin different ${\bf l}$ modes into sets of them called bandpowers (typically bands of similar $\ell$ or annuli of flat-sky Fourier modes spanning a narrow range of radii).
+Let us denote a given bandpower by its index $q$.
+The binned pseudo-power spectrum is:
+      \begin{equation}
+        \tilde{\sf C}^{ab}_q=\sum_{{\bf l}\in q} B_q^{\bf l}\,\tilde{\bf a}_{\bf l}\tilde{\bf b}^\dag_{\bf l},
+      \end{equation}
+      where the bandpower weights are normalized such that $\sum_{{\bf l}\in q}B_q^{\bf l}=1$.
+      \item Then, the correlation between bandpowers induced by the mode-coupling coefficients is partially reversed by multiplying $\tilde{\sf C}^{ab}_q$ by the so-called binned ``mode-coupling matrix'' $\mathcal{M}$, giving the final estimator
+      \begin{equation}
+        {\rm vec}\left(\hat{\sf C}^{ab}_q\right) = \sum_{q'} \left(\mathcal{M}^{-1}\right)_{qq'}\,{\rm vec}\left(\tilde{\sf C}^{ab}_{q'}\right).
+      \end{equation}
+
+      The main advantage of the \PCL estimator is that the mode-coupling matrix $\mathcal{M}$ is directly related to the coupling coefficients $\,^a{\sf M}_{{\bf l}{\bf l}'}$ (see \todo{cite}), and can be computed analytically making use of methods that scale like $\ell_{\rm max}^3$.
+    \end{enumerate}
+
+    Before we move on to covariances, it is worth considering the case of unmasked field (i.e $w^a({\bf x})=1$ everywhere).
+In this case $\,^a{\sf M}_{{\bf l}{\bf l}'}=\mathds{1}\Delta^x({\bf k},{\bf l})$, and therefore different modes are uncorrelated (as should have been obvious).
+In a non-ideal case where the mask is still sufficiently well behaved (i.e.
+masks without too much small-scale structure), we can still expect the coupling coefficients $\,^a{\sf M}_{{\bf l}{\bf l}'}$ to be sharply peaked around ${\bf l}={\bf l}'$.
 
   \subsection{Covariance matrices}
-
     This section presents the analytical approximations used here to estimate the pseudo-$C_\ell$ Gaussian covariance matrix.

### 2019-05-21 draft/paper.tex
added mask, fixed cls plot

--- 2019-05-21 draft/paper.tex
@@
 shown in Fig.~\ref{fig:pz} and the theoretical $\cl$ are in
-Fig.~\ref{fig:cl-2bins}. Since all parts of sky will not be equally observed, a
-realist mask (see Fig.~\ref{fig:mask}) was applied.  The results shown here
+Fig.~\ref{fig:cl-2bins}. Since all parts of sky will not be equally observed,
+realist masks (see Fig.~\ref{fig:mask}) were applied.  The results shown here
 were obtained using the same mas for each redshift bin, as it is expected to
@@
   \centering
-  \includegraphics[width=\columnwidth]{./figures/mask-lss1.png}
-  \caption{Example of sky mask used in our analysis}
+  \includegraphics[width=\columnwidth]{./figures/mask-lss1.png}\\
+  \includegraphics[width=\columnwidth]{./figures/mask-lss2.png}
+  \caption{Sky mask used in our analysis}
   \label{fig:mask}

### 2019-05-22 notebooks/general-plots-paper.py
remove l<0 from cls plot, remove cbar + text in mask plot

@@ -57,7 +57,7 @@ ell, cl2bin, nls2bin = ell_cl2bin_file['ls'], ell_cl2bin_file['cls'], ell_cl2bin
 
 f, axs = plt.subplots(4, 4, figsize=(8, 6), gridspec_kw={'hspace': 0, 'wspace': 0}, sharex=True, sharey='row')
 
-labels = [r'$\delta_1$', r'$\gamma_{E,\, 1}$', r'$\delta_2$', r'$\gamma_{E,\, 2}$']
+labels = [r'$\delta_1$', r'$\gamma_{E,1}$', r'$\delta_2$', r'$\gamma_{E,2}$']
 ci = 0
 for i in range(cl2bin.shape[0]):
     if i in (2, 5):
@@ -66,11 +66,12 @@ for i in range(cl2bin.shape[0]):
     for j in range(i, cl2bin.shape[0]):
         if j in (2, 5):
             continue
-        axs[ci, cj].loglog(ell, cl2bin[i, j], label='Th. w/o noise')
-        axs[ci, cj].loglog(ell, cl2bin[i, j] + nls2bin[i, j], label='Th.', ls='--')
-        axs[ci, cj].loglog(ell, nls2bin[i, j], label='Noise')
-        axs[ci, cj].text(0.8, 0.9, "({}, {})".format(labels[ci], labels[cj]),
-                         transform=axs[ci, cj].transAxes)
+        axs[ci, cj].loglog(ell[ell >= 2], cl2bin[i, j, ell >= 2], label='Th. w/o noise')
+        axs[ci, cj].loglog(ell[ell >= 2], cl2bin[i, j, ell >= 2] + nls2bin[i, j, ell >= 2], label='Th.', ls='--')
+        axs[ci, cj].loglog(ell[ell >= 2], nls2bin[i, j, ell >= 2], label='Noise')
+        axs[ci, cj].text(0.78, 0.9, "({}, {})".format(labels[ci], labels[cj]),
+                         transform=axs[ci, cj].transAxes,
+                         fontsize=9, horizontalalignment='center')
         cj += 1
 
     ci += 1
@@ -154,7 +155,7 @@ fname = os.path.join(outdir, 'mask-lss1.png')
 fmask = "./data/mask_lss_sph1.fits"
[…]

### 2019-05-22 notebooks/general-plots-paper.py
png -> pdf + script for cov rows

@@ -20,7 +20,7 @@ if not os.path.isdir(outdir):
 ######################### Fig with z-bins ####################################
 ##############################################################################
 
-fname = os.path.join(outdir, 'pz.png')
+fname = os.path.join(outdir, 'pz.pdf')
 
 nell = 30000
 
@@ -49,7 +49,7 @@ plt.close()
 ######################### Theoretical sph 2b Cls #############################
 ##############################################################################
 
-fname = os.path.join(outdir, 'cls-sph-2b.png')
+fname = os.path.join(outdir, 'cls-sph-2b.pdf')
 
 
 ell_cl2bin_file = np.load('./data/cls_lss_2bins.npz')
@@ -101,7 +101,7 @@ plt.close()
 ########################## Compare 1bin sph Cls ##############################
 ##############################################################################
 
-fname = os.path.join(outdir, 'compare-sph-cls-used-1bin.png')
+fname = os.path.join(outdir, 'compare-sph-cls-used-1bin.pdf')
 
 clTh = np.loadtxt('./data/cls_lss.txt', unpack=True)
 lTh = clTh[0][:3*512]
@@ -150,7 +150,7 @@ plt.close()
 ############################# Mask sph1 ######################################
 ##############################################################################
 
-fname = os.path.join(outdir, 'mask-lss1.png')
+fname = os.path.join(outdir, 'mask-lss1.pdf')
 
 fmask = "./data/mask_lss_sph1.fits"
 mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=512)
@@ -163,7 +163,7 @@ plt.close()
 ############################# Mask sph1 ######################################
 ##############################################################################
 
-fname = os.path.join(outdir, 'mask-lss2.png')
+fname = os.path.join(outdir, 'mask-lss2.pdf')
 
 fmask = "./data/mask_lss_sph2.fits"
 mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=512)
@@ -176,7 +176,7 @@ plt.close()
 ############################# Mask flat1 ######################################
 ##############################################################################
 
-fname = os.path.join(outdir, 'mask-lss_flat1.png')
+fname = os.path.join(outdir, 'mask-lss_flat1.pdf')
[…]

### 2019-05-22 notebooks/row-sph-1bin-cov.py

@@ -1,0 +1,193 @@
+from matplotlib import pyplot as plt
+from scipy.interpolate import interp1d
+import numpy as np
+import os
+
+# pylint: disable=C0103
+
+DPI = 500
+DEFAULT_COLOR_CYCLE = plt.rcParams['axes.prop_cycle'].by_key()['color']
+
+##############################################################################
+############################# All rows sph ###################################
+##############################################################################
+
+prefix = 'run_sph_2b'
+run_path = os.path.join('./simulations_outputs/', prefix, prefix)
+sims_suffix = '_cl_0001-20000.npz'
+
+outdir = './general-plots/'
+
+if not os.path.isdir(outdir):
+    os.mkdir(outdir)
+
+##############################################################################
+
+ell = np.loadtxt(run_path + '_ells.txt')
+ell_cl2bin_file = np.load('./data/cls_lss_2bins.npz')
+ell_th, cl2bin, nls2bin = ell_cl2bin_file['ls'], ell_cl2bin_file['cls'], ell_cl2bin_file['nls']
+
+clth = (cl2bin + nls2bin)[:3, :3]
+
+clth = interp1d(ell_th, clth)(ell)
+i, j = np.triu_indices(3)
+clth_ar = clth[i, j]
+
+##############################################################################
+
+CovSims_path = run_path + '_cov_b1' + sims_suffix
+Csims = np.load(CovSims_path)['arr_0']
+
+CsimOld = Csims.copy()
+nlbins = int(Csims.shape[0] / 6)
+Csims = Csims.reshape((6, nlbins, 6, nlbins))
+
+##############################################################################
+c0000 = np.load(run_path+'_c0000_b1.npz')['arr_0']
+c0002 = np.load(run_path+'_c0002_b1.npz')['arr_0']
+c0022 = np.load(run_path+'_c0022_b1.npz')['arr_0']
+c0202 = np.load(run_path+'_c0202_b1.npz')['arr_0']
[…]

### 2019-05-22 notebooks/row-sph-1bin-cov.py
use upto lmax=2*512 bpws + black color & star for scatter plot naive

@@ -33,14 +33,20 @@ clth = interp1d(ell_th, clth)(ell)
 i, j = np.triu_indices(3)
 clth_ar = clth[i, j]
 
+lmax = (ell < 2*512).sum()
+nlbins_orig = int(ell.shape[0])
+nlbins = lmax
+ell = ell[:lmax]
+
 ##############################################################################
 
 CovSims_path = run_path + '_cov_b1' + sims_suffix
 Csims = np.load(CovSims_path)['arr_0']
 
-CsimOld = Csims.copy()
-nlbins = int(Csims.shape[0] / 6)
-Csims = Csims.reshape((6, nlbins, 6, nlbins))
+Csims = Csims.reshape((6, nlbins_orig, 6, nlbins_orig))
+Csims = Csims[:, :lmax, :, :lmax]
+
+CsimOld = Csims.reshape(6 * nlbins, 6 * nlbins)
 
 ##############################################################################
 c0000 = np.load(run_path+'_c0000_b1.npz')['arr_0']
@@ -77,16 +83,16 @@ Cth_ar = np.array([CovTh_TT, CovTh_TTTE, CovTh_TTTB, CovTh_TTEE, CovTh_TTEB, Cov
 
 Cth = np.empty((6, nlbins, 6, nlbins))
 i, j = np.triu_indices(6)
-Cth[i, :,  j, :] = Cth_ar
-Cth[j, :,  i, :] = Cth_ar
+Cth[i, :,  j, :] = Cth_ar[:, :lmax, :lmax]
+Cth[j, :,  i, :] = Cth_ar[:, :lmax, :lmax]
 
 ##############################################################################
 
 CthN = np.empty((6, nlbins, 6, nlbins))
 CthN_ar = np.load(run_path + '_covNaive_ar_b1.npz')['arr_0']
 i, j = np.triu_indices(6)
-CthN[i, :,  j, :] = CthN_ar
-CthN[j, :,  i, :] = CthN_ar
+CthN[i, :,  j, :] = CthN_ar[:, :lmax, :lmax]
+CthN[j, :,  i, :] = CthN_ar[:, :lmax, :lmax]
 
 ##############################################################################
 
@@ -114,7 +120,7 @@ def ax_plot_row(ax, lbins, CovSims, CovTh, CovThN, normalization, index=20, dx=5
     indexi = index
[…]

### 2019-05-22 notebooks/general-plots-paper.py
added 2bins corr-diff plot code

@@ -231,3 +231,41 @@ plt.tight_layout()
 plt.savefig(fname, dpi=DPI)
 plt.close()
 
+##############################################################################
+############################# 2bins corr diff ################################
+##############################################################################
+
+Cth = np.load('../run_sph_2b_same_mask/full_covariance/run_sph_2b_same_mask_covTh_TTTEEE_short_2bins_same_mask.npz')['arr_0']
+Csims = np.load('../run_sph_2b_same_mask/full_covariance/run_sph_2b_same_mask_covSims_TTTEEE_short_0001-20000.npz')['arr_0']
+
+ell = np.loadtxt('../run_sph_2b_same_mask/run_sph_2b_same_mask_ells.txt')
+
+CorrSims = Csims/np.sqrt(np.diag(Csims)[:, None] * np.diag(Csims)[None, :])
+CorrTh = Cth/np.sqrt(np.diag(Cth)[:, None] * np.diag(Cth)[None, :])
+
+
+labels_cov = []
+for i in range(4):
+    for j in range(i, 4):
+        labels_cov.append(labels[i] + labels[j])
+
+
+f, ax = plt.subplots(1, 1, figsize=(8, 8))
+nlbins = CorrSims.shape[0] / 10
+print(nlbins)
+cb = ax.imshow(CorrTh - CorrSims, vmin=-0.02, vmax=0.02)
+f.colorbar(cb)
+for i in range(1, 10):
+    ax.plot([0, 10*nlbins-1], [i*nlbins, i*nlbins], 'k-', lw=0.5)
+    ax.plot([i*nlbins, i*nlbins], [0, 10*nlbins-1], 'k-', lw=0.5)
+
+ticks = [(i + 0.5) * nlbins for i in range(10)]
+ax.set_xticks(ticks)
+ax.set_xticklabels(labels_cov)
+ax.set_yticks(ticks)
+ax.set_yticklabels(labels_cov)
+
+fname = os.path.join(outdir, 'sph_2bin_diff_corr.pdf')
+plt.savefig(fname, dpi=DPI)
+# plt.show()
+plt.close()

### 2019-05-22 notebooks/common.py
default pdf

@@ -6,6 +6,8 @@ import numpy as np
 import pymaster as nmt
 import os
 
+plt.rcParams["savefig.format"] = 'pdf'
+
 DEFAULT_COLOR_CYCLE = plt.rcParams['axes.prop_cycle'].by_key()['color']
 DPI = 500
 FIGSIZE = (4, 3)
@@ -231,6 +233,7 @@ def plot_rows_cov_matrix(lbins, CovSims, CovTh, normalization, labels, index=20,
         Xi = X
         for _ in range(peaks):
             Xi = lbins[indexi - dx:indexi + dx + 1]
+            Y = CovSims[indexi, indexi - dx:indexi + dx + 1]/normalization
             Yth = covth[indexi, indexi - dx:indexi + dx + 1]/normalization
             ax[0].plot(Xi, Yth, c=c[i+1], label=labels[i] if not _ else '')
             ax[1].plot(Xi, Yth/Y - 1, c=c[i+1])

### 2019-05-23 notebooks/general-plots-paper.py
added chi2 for case with contaminants

@@ -1,4 +1,6 @@
 #!/usr/bin/python
+from scipy import stats
+import common as co
 import flatmaps as fm
 import healpy as hp
 import numpy as np
@@ -232,6 +234,49 @@ plt.savefig(fname, dpi=DPI)
 plt.close()
 
 ##############################################################################
+############################ Chi2 foregrounds ################################
+##############################################################################
+
+CovSims = np.load('../run_sph/run_sph_covTTTEE_short_clsims_0001-20000.npz')['arr_0']
+CovSims_fore = np.load('../run_sph_contaminants_ls100_ss100/run_sph_contaminants_ls100_ss100_covTTTEE_short_clsims_0001-20000.npz')['arr_0']
+CovTh = np.load('../run_sph_contaminants_ls100_ss100/run_sph_contaminants_ls100_ss100_covThTTTEEE_short.npz')['arr_0']
+
+clsims = np.load('../run_sph_contaminants_ls100_ss100/run_sph_contaminants_ls100_ss100_clsims_0001-20000.npz')
+clTT = np.reshape(clsims['cl00'], np.array(np.shape(clsims['cl00']))[[0, 2]])
+clTE = clsims['cl02'][:, 0, :]
+clEE = clsims['cl22'][:, 0, :]
+lbins = clsims['l']
+lmax = (clsims['l'] < 2*512).sum()
+
+clTTTEEE = np.concatenate([clTT[:, :lmax], clTE[:, :lmax], clEE[:, :lmax]], axis=1)
+
+chi2_list = co.get_chi2(clTTTEEE, [np.linalg.inv(CovSims), np.linalg.inv(CovSims_fore), np.linalg.inv(CovTh)])
+
+f, ax = plt.subplots(1, 1, figsize=FSIZE1)
+bins = np.linspace(np.min(chi2_list), np.max(chi2_list), 60)
+_, x, _ = ax.hist(chi2_list[0], bins=bins, histtype='step', density=True,
+                  label='Sim. w/o cont.', ls='-')
+_, x, _ = ax.hist(chi2_list[1], bins=bins, histtype='step', density=True,
+                  label='Sim. w/ cont.', ls='-.')
+_, x, _ = ax.hist(chi2_list[2], bins=bins, histtype='step', density=True,
+                  label='NKA', ls='-')
+
+ax.plot(x[:-1], stats.chi2.pdf(x[:-1], clTTTEEE.shape[1]), ls='--', label=r'$\chi^2$ pdf')
+
[…]

### 2019-05-23 notebooks/general-plots-paper.py
commented-out lines for comparison 1bin th cls + change name of diff corr matrix

@@ -103,50 +103,50 @@ plt.close()
 ########################## Compare 1bin sph Cls ##############################
 ##############################################################################
 
-fname = os.path.join(outdir, 'compare-sph-cls-used-1bin.pdf')
-
-clTh = np.loadtxt('./data/cls_lss.txt', unpack=True)
-lTh = clTh[0][:3*512]
-clTh_TT = (clTh[1] + clTh[5])[:3*512]
-clTh_EE = (clTh[2] + clTh[6])[:3*512]
-clTh_TE = (clTh[4] + clTh[-1])[:3*512]
-
-f, axs = plt.subplots(1, 3, figsize=(8, 3), gridspec_kw={'hspace': 0, 'wspace': 0}, sharex=True, sharey='row')
-
-cl2bins_1 = (cl2bin + nls2bin)[:3, :3]
-cl2bins_2 = (cl2bin + nls2bin)[3:, 3:]
-i, j = np.triu_indices(3)
-
-#### TT #####
-axs[0].loglog(lTh, clTh_TT, label='1-bin Fid.')
-axs[0].loglog(ell, cl2bins_1[0, 0], label='1st 2-bin Fid.', ls='--')
-axs[0].loglog(ell, cl2bins_2[0, 0], label='2nd 2-bin Fid.', ls='--')
-axs[0].set_ylabel(r'$C_l$')
-
-axs[0].text(0.5, 0.9, r'$C_l^{gg}$', transform=axs[0].transAxes)
-
-#### TE #####
-axs[1].loglog(lTh, clTh_TE, label='1-bin Fid.')
-axs[1].loglog(ell, cl2bins_1[0, 1], label='1st 2-bin Fid.', ls='--')
-axs[1].loglog(ell, cl2bins_2[0, 1], label='2nd 2-bin Fid.', ls='--')
-axs[1].text(0.5, 0.9, r'$C_l^{g\gamma}$', transform=axs[1].transAxes)
-#axs[1].set_ylabel(r'$C_l^{g\gamma}$')
-
-#### EE #####
-axs[2].loglog(lTh, clTh_EE, label='1-bin Fid.')
-axs[2].loglog(ell, cl2bins_1[1, 1], label='1st 2-bin Fid.', ls='--')
-axs[2].loglog(ell, cl2bins_2[1, 1], label='2nd 2-bin Fid.', ls='--')
[…]

### 2019-05-23 notebooks/general-plots-paper.py
added chi2 TT TE EE for sph and flat

@@ -309,3 +309,190 @@ fname = os.path.join(outdir, 'run_sph_2b_same_mask_NKA_diff_corr.pdf')
 plt.savefig(fname, dpi=DPI)
 # plt.show()
 plt.close()
+
+##############################################################################
+####################### TTTT, TETE, EEEE chi2 plot ###########################
+##############################################################################
+
+prefix = 'run_sph_2b'
+run_path = os.path.join('./simulations_outputs/', prefix, prefix)
+sims_suffix = '_cl_0001-20000.npz'
+
+##############################################################################
+
+ell = np.loadtxt(run_path + '_ells.txt')
+lmax = (ell < 2*512).sum()
+nlbins = lmax
+
+clsims = np.load(run_path + '_b1' + sims_suffix)['arr_0']
+
+clTT = clsims[0, 0, :, :lmax]
+clEE = clsims[1, 1, :, :lmax]
+clTE = clsims[0, 1, :, :lmax]
+
+##############################################################################
+
+CovSims_path = run_path + '_cov_b1' + sims_suffix
+Csims = np.load(CovSims_path)['arr_0']
+
+nlbins_orig = int(Csims.shape[0] / 6)
+Csims = Csims.reshape((6, nlbins_orig, 6, nlbins_orig))
+Csims = Csims[:, :lmax, :, :lmax]
+
+Csims_TT = Csims[0, :, 0, :]
+Csims_TE = Csims[1, :, 1, :]
+Csims_EE = Csims[3, :, 3, :]
+
+##############################################################################
+c0000 = np.load(run_path+'_c0000_b1.npz')['arr_0']
+c0202 = np.load(run_path+'_c0202_b1.npz')['arr_0']
+c2222 = np.load(run_path+'_c2222_b1.npz')['arr_0']
+
+CovTh_TT = c0000[:lmax, :lmax]
+CovTh_TETE, CovTh_TETB = c0202[:, 0, :, [0, 1] ]
+CovTh_EEEE, CovTh_EEEB, CovTh_EEBE, CovTh_EEBB = c2222[:, 0, :, [0, 1, 2, 3] ]
+
+CovTh_TE = CovTh_TETE[:lmax, :lmax]
+CovTh_EE = CovTh_EEEE[:lmax, :lmax]
+
[…]

### 2019-05-23 notebooks/general-plots-paper.py
encapsulate code in functions

@@ -21,478 +21,508 @@ if not os.path.isdir(outdir):
 ##############################################################################
 ######################### Fig with z-bins ####################################
 ##############################################################################
+def pz():
+    fname = os.path.join(outdir, 'pz.pdf')
 
-fname = os.path.join(outdir, 'pz.pdf')
+    nell = 30000
 
-nell = 30000
+    def nofz(z, z0, sz, ndens):
+        return np.exp(-0.5*((z-z0)/sz)**2)*ndens/np.sqrt(2*np.pi*sz**2)
 
-def nofz(z, z0, sz, ndens):
-    return np.exp(-0.5*((z-z0)/sz)**2)*ndens/np.sqrt(2*np.pi*sz**2)
+    z = np.linspace(0, 2, 512)
+    pz1 = nofz(z, 0.955, 0.13, 7.55)
+    pz2 = nofz(z, 0.755, 0.13, 7.55)
+    ndens1 = np.sum(pz1)*np.mean(z[1:]-z[:-1])*(180*60./np.pi)**2
+    ndens2 = np.sum(pz2)*np.mean(z[1:]-z[:-1])*(180*60./np.pi)**2
 
-z = np.linspace(0, 2, 512)
-pz1 = nofz(z, 0.955, 0.13, 7.55)
-pz2 = nofz(z, 0.755, 0.13, 7.55)
-ndens1 = np.sum(pz1)*np.mean(z[1:]-z[:-1])*(180*60./np.pi)**2
-ndens2 = np.sum(pz2)*np.mean(z[1:]-z[:-1])*(180*60./np.pi)**2
+    f, ax = plt.subplots(1, 1, figsize=FSIZE1)
+    ax.plot(z, pz1, label='Bin 1')
+    ax.plot(z, pz2, label='Bin 2')
 
-f, ax = plt.subplots(1, 1, figsize=FSIZE1)
-ax.plot(z, pz1, label='Bin 1')
-ax.plot(z, pz2, label='Bin 2')
+    ax.set_ylabel('P(z)')
+    ax.set_xlabel('z')
 
-ax.set_ylabel('P(z)')
-ax.set_xlabel('z')
-
-ax.legend(loc=0)
-plt.tight_layout()
-plt.savefig(fname, dpi=DPI)
-plt.close()
+    ax.legend(loc=0)
+    plt.tight_layout()
+    plt.savefig(fname, dpi=DPI)
+    plt.close()
 
 ##############################################################################
 ######################### Theoretical sph 2b Cls #############################
[…]

### 2019-05-23 notebooks/general-plots-paper.py
beautiful l's and improved th. cls plot

@@ -58,7 +58,7 @@ def cls_2b():
     ell_cl2bin_file = np.load('./data/cls_lss_2bins.npz')
     ell, cl2bin, nls2bin = ell_cl2bin_file['ls'], ell_cl2bin_file['cls'], ell_cl2bin_file['nls']
 
-    f, axs = plt.subplots(4, 4, figsize=(8, 6), gridspec_kw={'hspace': 0, 'wspace': 0}, sharex=True, sharey='row')
+    f, axs = plt.subplots(4, 4, figsize=(9, 6), gridspec_kw={'hspace': 0, 'wspace': 0}, sharex=True, sharey='row')
 
     labels = [r'$\delta_1$', r'$\gamma_{E,1}$', r'$\delta_2$', r'$\gamma_{E,2}$']
     ci = 0
@@ -86,11 +86,12 @@ def cls_2b():
         ax.set_visible(False)
 
     for i in range(axs.shape[0]):
-        axs[i, i].set_xlabel('l')
-        axs[i, i].set_ylabel(r'$C_l$')
+        axs[i, i].set_xlabel('$\ell$')
+        axs[i, i].set_ylabel(r'$C_\ell$')
         axs[i, i].set_visible(True)
-        axs[i, i].xaxis.set_tick_params(labelbottom=True)
-        axs[i, i].yaxis.set_tick_params(labelleft=True)
+        axs[i, i].xaxis.set_tick_params(labelbottom=True, labelsize=8)
+        axs[i, i].yaxis.set_tick_params(labelleft=True, labelsize=8,
+                                        labelrotation=45, pad=-3)
         # axs[0, i].set_title(labels[i])
         # axs[i, -1].text(1, 0.5, labels[i], fontsize=12,
         #                 transform=axs[i, -1].transAxes)
@@ -126,28 +127,28 @@ def compare_cls_1bin():
     axs[0].loglog(lTh, clTh_TT, label='1-bin Fid.')
     axs[0].loglog(ell, cl2bins_1[0, 0], label='1st 2-bin Fid.', ls='--')
     axs[0].loglog(ell, cl2bins_2[0, 0], label='2nd 2-bin Fid.', ls='--')
-    axs[0].set_ylabel(r'$C_l$')
+    axs[0].set_ylabel(r'$C_\ell$')
 
[…]

### 2019-05-23 notebooks/general-plots-paper.py
added fiducial cl + default color cycle

@@ -12,6 +12,7 @@ import os
 
 DPI = 500
 FSIZE1 = (4, 3)
+DEFAULT_COLOR_CYCLE = plt.rcParams['axes.prop_cycle'].by_key()['color']
 
 outdir = './general-plots/'
 
@@ -224,16 +225,19 @@ def foregrounds():
     fname = os.path.join(outdir, 'foreground.pdf')
 
     f, ax = plt.subplots(1, 1, figsize=FSIZE1)
-    ax.loglog(ell, cl_ls1, c='orange')
-    ax.loglog(ell, cl_ls2, c='orange', label='Large scales')
-    ax.fill_between(ell, cl_ls1, cl_ls2, facecolor="orange", alpha=0.5)
+    color = DEFAULT_COLOR_CYCLE
 
-    ax.loglog(ell, cl_ss[0], c='b', label='Small scales')
+    ax.loglog(ell, cl2bin[0, 0] + nls2bin[0, 0], label='Fiducial')
+    ax.loglog(ell, cl_ls1, c=color[1])
+    ax.loglog(ell, cl_ls2, c=color[1], label='Large scales cont.')
+    ax.fill_between(ell, cl_ls1, cl_ls2, facecolor=color[1], alpha=0.5)
+
+    ax.loglog(ell, cl_ss[0], c=color[2], label='Small scales cont.')
 
     ax.set_xlabel('$\ell$')
     ax.set_ylabel(r'$C_\ell^{gg}$')
 
-    ax.legend(loc=0)
+    ax.legend(loc=0, fontsize=8)
 
     plt.tight_layout()
     plt.savefig(fname, dpi=DPI)

### 2019-05-23 notebooks/run_sph_2b_same_mask_do_all_check_Full_cov.py
save chi2 & corr from Full TTTEEE 2bins

@@ -44,3 +44,6 @@ foutput = figures_path + '_Efstathiou'
 chi2_Full, corr_Full = co.do_all_checks(lbins_Full, cl_for_C, CovSims_Full,
                                         CovTh_Full, modes="All TTTEEE",
                                         row_cov=False, foutput=foutput + '_TTTEEE_Full')
+
+np.savez_compressed(output_path + '_chi2_sims_th_TTTEEE_short_Full.npz', chi2_Full)
+np.savez_compressed(output_path + '_corr_sims_th_TTTEEE_short_Full.npz', corr_Full)

### 2019-05-23 notebooks/general-plots-paper.py
added function for NKA chi2 TTTEEE Full case

@@ -516,6 +516,39 @@ def chi2_flat_TT_TE_EE():
     plt.close()
 
 ##############################################################################
+############################ Chi2 TTTEEE Full ################################
+##############################################################################
+
+def chi2_NKA_TTTEEE_full():
+    chi2_list = np.load('../run_sph_2b_same_mask/full_covariance/run_sph_2b_same_mask_chi2_sims_th_TTTEEE_short_Full.npz')['arr_0']
+
+    ell = np.loadtxt('../run_sph_2b_same_mask/run_sph_2b_same_mask_ells.txt')
+    lbins = ell
+    lmax = (ell < 2*512).sum()
+
+    f, ax = plt.subplots(1, 1, figsize=FSIZE1)
+    bins = np.linspace(np.min(chi2_list), np.max(chi2_list), 60)
+    _, x, _ = ax.hist(chi2_list[0], bins=bins, histtype='step', density=True,
+                      label='Sim.', ls='-')
+    _, x, _ = ax.hist(chi2_list[1], bins=bins, histtype='step', density=True,
+                      label='NKA', ls='-')
+
+    ax.plot(x[:-1], stats.chi2.pdf(x[:-1], lmax * 10), ls='--', label=r'$\chi^2$ pdf')
+
+    ax.set_xlabel(r'$\chi^2$')
+    ax.set_ylabel('$10^3$ pdf')
+
+    y_vals = ax.get_yticks()
+    ax.set_yticklabels([str(x * 1000) for x in y_vals])
+
+    ax.legend(loc='upper right', fontsize='9') # , frameon=False)
+    plt.tight_layout()
+    fname = os.path.join(outdir, 'run_sph_2b_same_mask_NKA_TTTEEE_Full_chi2.pdf')
+    plt.savefig(fname, dpi=DPI)
+    # plt.show()
+    plt.close()
+
+##############################################################################
 ############################## What to plot ##################################
 ##############################################################################
 
@@ -529,3 +562,5 @@ if __name__ == '__main__':
     cls_2b()
     chi2_sph_TT_TE_EE()
     chi2_flat_TT_TE_EE()
+    chi2_NKA_TTTEEE_full()
+

### 2019-05-25 notebooks/general-plots-paper.py
improved plots

@@ -70,8 +70,8 @@ def cls_2b():
         for j in range(i, cl2bin.shape[0]):
             if j in (2, 5):
                 continue
-            axs[ci, cj].loglog(ell[ell >= 2], cl2bin[i, j, ell >= 2], label='Th. w/o noise')
-            axs[ci, cj].loglog(ell[ell >= 2], cl2bin[i, j, ell >= 2] + nls2bin[i, j, ell >= 2], label='Th.', ls='--')
+            axs[ci, cj].loglog(ell[ell >= 2], cl2bin[i, j, ell >= 2], label='Signal')
+            axs[ci, cj].loglog(ell[ell >= 2], cl2bin[i, j, ell >= 2] + nls2bin[i, j, ell >= 2], label='Signal + Noise', ls='--')
             axs[ci, cj].loglog(ell[ell >= 2], nls2bin[i, j, ell >= 2], label='Noise')
             axs[ci, cj].text(0.78, 0.9, "({}, {})".format(labels[ci], labels[cj]),
                              transform=axs[ci, cj].transAxes,
@@ -164,6 +164,7 @@ def masks():
 
     fmask = "./data/mask_lss_sph1.fits"
     mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=512)
+    mask_lss[mask_lss == 0] = hp.UNSEEN
 
     hp.mollview(mask_lss, title="", cbar=False, coord=['G', 'C'], notext=True)
     plt.savefig(fname, dpi=DPI)
@@ -177,6 +178,7 @@ def masks():
 
     fmask = "./data/mask_lss_sph2.fits"
     mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=512)
+    mask_lss[mask_lss == 0] = hp.UNSEEN
 
     hp.mollview(mask_lss, title="", coord=['G', 'C'], cbar=False, notext=True)
     plt.savefig(fname, dpi=DPI)
@@ -191,10 +193,11 @@ def masks():
     fmask = "./data/mask_lss_flat.fits"
 
     fmi, mask_hsc = fm.read_flat_map(fmask)
+    mask_hsc[mask_hsc == 0] = hp.UNSEEN
 
     fmi.view_map(mask_hsc, addColorbar=False)
 
-    plt.savefig(fname, dpi=DPI)
[…]

### 2019-05-26 notebooks/general-plots-paper.py
added reldev eigv plot + commented out 2nd mask flat sky

@@ -204,16 +204,16 @@ def masks():
 ############################# Mask flat2 ######################################
 ##############################################################################
 
-    fname = os.path.join(outdir, 'mask-lss_flat2.pdf')
+    # fname = os.path.join(outdir, 'mask-lss_flat2.pdf')
 
-    fmask = "./data/mask_lss_flat_2.fits"
+    # fmask = "./data/mask_lss_flat_2.fits"
 
-    fmi, mask_hsc = fm.read_flat_map(fmask)
-    mask_hsc[mask_hsc == 0] = hp.UNSEEN
+    # fmi, mask_hsc = fm.read_flat_map(fmask)
+    # mask_hsc[mask_hsc == 0] = hp.UNSEEN
 
-    fmi.view_map(mask_hsc)
-    plt.savefig(fname, dpi=DPI)
-    plt.close()
+    # fmi.view_map(mask_hsc)
+    # plt.savefig(fname, dpi=DPI)
+    # plt.close()
 
 ##############################################################################
 ############################# Foregrounds ####################################
@@ -581,6 +581,39 @@ def chi2_NKA_TTTEEE_full():
 ############################## What to plot ##################################
 ##############################################################################
 
+def eigv_sph_1bin():
+    CovSims = np.load('../run_sph_2b/run_sph_2b_covTTTEE_b1_short_cl_0001-20000.npz')['arr_0']
+    CovTh = np.load('../run_sph_2b/run_sph_2b_covThTTTEEE_b1_short.npz')['arr_0']
+
+    f, ax = plt.subplots(2, 1, sharex=True, figsize=(4,4))
+
+    Eval_Sims, Evec_Sims = co.diagonalize(CovSims)
+    X = np.arange(len(Eval_Sims))
+    ax[0].plot(X, Eval_Sims, label='Sim.')
+    ax[1].plot(X, 0 * X)
+
+    Eval_Th, Evec_Th = co.diagonalize(CovTh)
+
+    ax[0].plot(X, Eval_Th, label='NKA')
+    ax[1].plot(X, Eval_Th/Eval_Sims - 1)
+
+    ax[0].set_ylabel('Eigenvalue')
+    ax[0].legend(loc=0)
+
+    ax[1].set_xlabel('Element')
[…]

### 2019-05-27 notebooks/general-plots-paper.py
improved cl-cont plot

@@ -226,12 +226,12 @@ def foregrounds():
     cl_ls1 = tp.create_cl_templates(ell, cl2bin[0, 0] + nls2bin[0, 0], exp_range=(-3, -3), N=1)[0]
     cl_ls2 = tp.create_cl_templates(ell, cl2bin[0, 0] + nls2bin[0, 0], exp_range=(-1, -1), N=1)[0]
 
-    fname = os.path.join(outdir, 'foreground.pdf')
+    fname = os.path.join(outdir, 'contaminants_cl.pdf')
 
     f, ax = plt.subplots(1, 1, figsize=FSIZE1)
     color = DEFAULT_COLOR_CYCLE
 
-    ax.loglog(ell, cl2bin[0, 0] + nls2bin[0, 0], label='Fiducial')
+    ax.loglog(ell, cl2bin[0, 0] + nls2bin[0, 0], label='Signal')
     ax.loglog(ell, cl_ls1, c=color[1])
     ax.loglog(ell, cl_ls2, c=color[1], label='Large-scale cont.')
     ax.fill_between(ell, cl_ls1, cl_ls2, facecolor=color[1], alpha=0.5)
@@ -239,7 +239,7 @@ def foregrounds():
     ax.loglog(ell, cl_ss[0], c=color[2], label='Small-scale cont.')
 
     ax.set_xlabel('$\ell$')
-    ax.set_ylabel(r'$C_\ell^{gg}$')
+    ax.set_ylabel(r'$C_\ell^{\delta \delta}$')
 
     ax.legend(loc=0, fontsize=8)
 

### 2019-05-31 notebooks/compute_covTh_sph_same_mask_spin0.py
added codes to compute Full Cov matrix with spin0 app

@@ -1,0 +1,136 @@
+#!/usr/bin/python
+
+import pymaster as nmt
+import numpy as np
+import healpy as hp
+import os
+
+#pylint: disable=C0103
+##############################################################################
+##############################################################################
+# This script uses the covariance workspace created by the other script for NKA
+##############################################################################
+##############################################################################
+
+##############################################################################
+##############################################################################
+##############################################################################
+
+prefix = 'run_sph_2b_same_mask'
+run_path = os.path.join('./simulations_outputs/', prefix, prefix)
+full_cov_path_NKA = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
+
+output_dir = os.path.join('./simulations_outputs/', prefix, 'full_covariance_spin0')
+
+if not os.path.isdir(output_dir):
+    os.makedirs(output_dir)
+out_run_path = os.path.join(output_dir, prefix)
+data_folder = './data/'
+
+nside = 512
+
+##############################################################################
+##############################################################################
+##############################################################################
+
+f = np.load(os.path.join(data_folder, "cls_lss_2bins.npz"))
+lTh = f['ls']
+clTh = (f['cls'] + f['nls'])
+clTh_all = (f['cls'] + f['nls'])[np.triu_indices(f['cls'].shape[0])]
+
+fmasks = ["data/mask_lss_sph1.fits", "data/mask_lss_sph1.fits"]
+
+mask_lss_ar = []
+
+for fmask in fmasks:
+    mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=nside)
+    mask_lss_ar.append(mask_lss)
+
+#Set up binning scheme
+fsky = np.mean(np.product(mask_lss_ar, axis=0))
+d_ell = int(1./fsky)
+b = nmt.NmtBin(nside, nlb=d_ell)
+
+if not os.path.isfile(run_path + '_ells.txt'):
+    np.savetxt(run_path + '_ells.txt', b.get_effective_ells())
+
+##############################################################################
[…]

### 2019-05-31 notebooks/run_sph_2b_same_mask_do_all_check_Full_cov_spin0.py
added script to make plots for spin0 full cov approximation

@@ -1,0 +1,53 @@
+#!/usr/bin/python
+import common as co
+import numpy as np
+import os
+
+# pylint: disable=C0103
+
+prefix = 'run_sph_2b_same_mask'
+run_path = os.path.join('./simulations_outputs/', prefix, prefix)
+
+outdir = os.path.join('./simulations_outputs/', prefix, 'full_covariance_spin0')
+output_path = os.path.join(outdir, prefix)
+
+figures_dir = os.path.join('./simulations_outputs/', prefix, 'figures')
+figures_path = os.path.join(figures_dir, prefix)
+
+outdir_NKA = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
+output_path_NKA = os.path.join(outdir_NKA, prefix)
+
+##################
+ells = np.loadtxt(run_path + '_ells.txt')
+
+lmax = (ells < 2*512).sum()
+
+CovSims_path = output_path_NKA + '_covSims_TTTEEE_short_0001-20000.npz'
+CovTh_path = output_path + '_covTh_TTTEEE_short_2bins_same_mask.npz'
+
+CovSims_Full = np.load(CovSims_path)['arr_0']
+CovTh_Full = np.load(CovTh_path)['arr_0']
+
+cl_ar = np.load(run_path + '_cl_0001-20000.npz')['arr_0']
+Cls = [0, 1, 2] * 2  # 0 for T, 1 for E and 2 for B
+Cls_Bs_ar = []
+
+c = 0
+for i, Cli in enumerate(Cls):
+    for Clj in Cls[i:]:
+        Cls_Bs_ar.append([Cli, Clj])
+
+Cls_Bs_ar = np.array(Cls_Bs_ar)
+cl_ar_noBs = cl_ar[np.all(Cls_Bs_ar != 2, axis=1)]
+
+cl_for_C = np.concatenate(cl_ar_noBs[:, :, :lmax].swapaxes(1, 2)).T
+lbins_Full = np.concatenate([ells[:lmax]] * cl_ar_noBs.shape[0])
+
+##################
+foutput = figures_path + '_Spin0'
+chi2_Full, corr_Full = co.do_all_checks(lbins_Full, cl_for_C, CovSims_Full,
+                                        CovTh_Full, modes="All TTTEEE",
+                                        row_cov=False, foutput=foutput + '_TTTEEE_Full')
+
[…]

### 2019-05-31 notebooks/run_sph_2b_same_mask_do_all_check_Full_cov_spin0.py
fixed filepath name

@@ -23,7 +23,7 @@ ells = np.loadtxt(run_path + '_ells.txt')
 lmax = (ells < 2*512).sum()
 
 CovSims_path = output_path_NKA + '_covSims_TTTEEE_short_0001-20000.npz'
-CovTh_path = output_path + '_covTh_TTTEEE_short_2bins_same_mask.npz'
+CovTh_path = output_path + '_covTh_TTTEEE_short_2bins_same_mask_spin0.npz'
 
 CovSims_Full = np.load(CovSims_path)['arr_0']
 CovTh_Full = np.load(CovTh_path)['arr_0']

### 2019-05-31 notebooks/general-plots-paper.py
added function to compare chi NKA & spin0

@@ -577,6 +577,38 @@ def chi2_NKA_TTTEEE_full():
     # plt.show()
     plt.close()
 
+def chi2_Spin0_NKA_TTTEEE_full():
+    chi2_list = np.load('../run_sph_2b_same_mask/full_covariance/run_sph_2b_same_mask_chi2_sims_th_TTTEEE_short_Full.npz')['arr_0']
+    chi2_spin0 = np.load('../run_sph_2b_same_mask/full_covariance_spin0/run_sph_2b_same_mask_chi2_sims_th_TTTEEE_short_Full_spin0.npz')['arr_0'][1]
+
+    ell = np.loadtxt('../run_sph_2b_same_mask/run_sph_2b_same_mask_ells.txt')
+    lbins = ell
+    lmax = (ell < 2*512).sum()
+
+    f, ax = plt.subplots(1, 1, figsize=FSIZE1)
+    bins = np.linspace(np.min(chi2_list), np.max(chi2_list), 60)
+    _, x, _ = ax.hist(chi2_list[0], bins=bins, histtype='step', density=True,
+                      label='Sim.', ls='-')
+    _, x, _ = ax.hist(chi2_list[1], bins=bins, histtype='step', density=True,
+                      label='NKA', ls='-')
+    _, x, _ = ax.hist(chi2_spin0, bins=bins, histtype='step', density=True,
+                      label='Spin-0', ls='-')
+
+    ax.plot(x[:-1], stats.chi2.pdf(x[:-1], lmax * 10), ls='--', label=r'$\chi^2$ pdf')
+
+    ax.set_xlabel(r'$\chi^2$')
+    ax.set_ylabel('$10^3$ pdf')
+
+    y_vals = ax.get_yticks()
+    ax.set_yticklabels([str(x * 1000) for x in y_vals])
+
+    ax.legend(loc='upper right', fontsize='9') # , frameon=False)
+    plt.tight_layout()
+    fname = os.path.join(outdir, 'run_sph_2b_same_mask_Spin0_NKA_TTTEEE_Full_chi2.pdf')
+    plt.savefig(fname, dpi=DPI)
+    # plt.show()
+    plt.close()
+
 ##############################################################################
 ############################## What to plot ##################################
 ##############################################################################
@@ -626,4 +658,5 @@ if __name__ == '__main__':
     chi2_flat_TT_TE_EE_TB_EB_BB()
[…]

### 2019-06-10 notebooks/row-sph-1bin-cov.py
transpose l's for cov matrix + larger dots

@@ -84,7 +84,7 @@ Cth_ar = np.array([CovTh_TT, CovTh_TTTE, CovTh_TTTB, CovTh_TTEE, CovTh_TTEB, Cov
 Cth = np.empty((6, nlbins, 6, nlbins))
 i, j = np.triu_indices(6)
 Cth[i, :,  j, :] = Cth_ar[:, :lmax, :lmax]
-Cth[j, :,  i, :] = Cth_ar[:, :lmax, :lmax]
+Cth[j, :,  i, :] = Cth_ar[:, :lmax, :lmax].swapaxes(1, 2)
 
 ##############################################################################
 
@@ -92,7 +92,7 @@ CthN = np.empty((6, nlbins, 6, nlbins))
 CthN_ar = np.load(run_path + '_covNaive_ar_b1.npz')['arr_0']
 i, j = np.triu_indices(6)
 CthN[i, :,  j, :] = CthN_ar[:, :lmax, :lmax]
-CthN[j, :,  i, :] = CthN_ar[:, :lmax, :lmax]
+CthN[j, :,  i, :] = CthN_ar[:, :lmax, :lmax].swapaxes(1, 2)
 
 ##############################################################################
 
@@ -120,7 +120,7 @@ def ax_plot_row(ax, lbins, CovSims, CovTh, CovThN, normalization, index=20, dx=5
     indexi = index
     Xi = lbins[indexi : indexi + (peaks * 2 * dx) : 2 * dx]
     Yth = np.diag(CovThN)[indexi : indexi + (peaks * 2 * dx) : 2 * dx]/normalization[indexi : indexi + (peaks * 2 * dx) : 2 * dx]
-    ax.scatter(Xi, Yth, marker="*", s=6, c='k', label='Naive', zorder=3)
+    ax.scatter(Xi, Yth, marker="*", s=9, c='k', label='Naive', zorder=3)
 
     # dl = lbins[1] - lbins[0]
     # Naive = 1. / (0.4 * (2 * lbins  + 1) * dl)

### 2019-06-10 notebooks/general-plots-paper.py
improved plots + removed frames from legends

@@ -1,5 +1,6 @@
 #!/usr/bin/python
 from scipy import stats
+from mpl_toolkits.axes_grid1 import make_axes_locatable
 import common as co
 import flatmaps as fm
 import healpy as hp
@@ -40,10 +41,11 @@ def pz():
     ax.plot(z, pz1, label='Bin 1')
     ax.plot(z, pz2, label='Bin 2')
 
-    ax.set_ylabel('P(z)')
-    ax.set_xlabel('z')
+    ax.set_ylabel(r'$P(z)$ [arcmin$^{-2}$]')
+    ax.set_xlabel(r'$z$')
 
-    ax.legend(loc=0)
+    ax.legend(loc=0, frameon=False)
+    ax.set_ylim(bottom=0)
     plt.tight_layout()
     plt.savefig(fname, dpi=DPI)
     plt.close()
@@ -71,7 +73,7 @@ def cls_2b():
             if j in (2, 5):
                 continue
             axs[ci, cj].loglog(ell[ell >= 2], cl2bin[i, j, ell >= 2], label='Signal')
-            axs[ci, cj].loglog(ell[ell >= 2], cl2bin[i, j, ell >= 2] + nls2bin[i, j, ell >= 2], label='Signal + Noise', ls='--')
+            # axs[ci, cj].loglog(ell[ell >= 2], cl2bin[i, j, ell >= 2] + nls2bin[i, j, ell >= 2], label='Signal + Noise', ls='--')
             axs[ci, cj].loglog(ell[ell >= 2], nls2bin[i, j, ell >= 2], label='Noise')
             axs[ci, cj].text(0.78, 0.9, "({}, {})".format(labels[ci], labels[cj]),
                              transform=axs[ci, cj].transAxes,
@@ -80,7 +82,7 @@ def cls_2b():
 
         ci += 1
 
-    axs[0, 0].legend(loc=0)
+    axs[0, 0].legend(loc=0, frameon=False)
 
     i, j = np.triu_indices(axs.shape[0])
     for ax in axs[j, i].reshape((-1)):
@@ -146,7 +148,7 @@ def compare_cls_1bin():
     axs[2].text(0.5, 0.9, r'$C_\ell^{\gamma\gamma}$', transform=axs[2].transAxes)
     #axs[2].set_ylabel(r'$C_\ell^{\gamma\gamma}$')
 
[…]

### 2019-06-10 notebooks/row-sph-1bin-cov.py

@@ -1,3 +1,4 @@
+from mpl_toolkits.axes_grid1 import make_axes_locatable
 from matplotlib import pyplot as plt
 from scipy.interpolate import interp1d
 import numpy as np
@@ -182,17 +183,21 @@ CorrTh = Cth/np.sqrt(np.diag(Cth)[:, None] * np.diag(Cth)[None, :])
 f, ax = plt.subplots(1, 1, figsize=(4, 4))
 
 cb = ax.imshow(CorrTh - CorrSims, vmin=-0.02, vmax=0.02)
-f.colorbar(cb)
+divider = make_axes_locatable(ax)
+cax = divider.append_axes("right", size="5%", pad=0.05)
+cbar = f.colorbar(cb, cax=cax)
+cbar.ax.tick_params(labelsize=8)
 for i in range(1, 6):
     ax.plot([0, 6*nlbins-1], [i*nlbins, i*nlbins], 'k-', lw=0.5)
     ax.plot([i*nlbins, i*nlbins], [0, 6*nlbins-1], 'k-', lw=0.5)
 
 ticks = [(i + 0.5) * nlbins for i in range(6)]
 ax.set_xticks(ticks)
-ax.set_xticklabels(labels)
+ax.set_xticklabels(labels, fontsize=8)
 ax.set_yticks(ticks)
-ax.set_yticklabels(labels)
+ax.set_yticklabels(labels, fontsize=8)
 
+plt.tight_layout()
 fname = os.path.join(outdir, 'sph_1bin_diff_corr.pdf')
 plt.savefig(fname, dpi=DPI)
 # plt.show()

### 2019-06-11 notebooks/general-plots-paper.py
plot 4x4 corr diff

@@ -319,7 +319,7 @@ def corr_diff_2bins():
 
 
     f, ax = plt.subplots(1, 1, figsize=(8, 8))
-    nlbins = CorrSims.shape[0] / 10
+    nlbins = int(CorrSims.shape[0] / 10)
     cb = ax.imshow(CorrTh - CorrSims, vmin=-0.02, vmax=0.02)
     divider = make_axes_locatable(ax)
     cax = divider.append_axes("right", size="5%", pad=0.05)
@@ -338,6 +338,32 @@ def corr_diff_2bins():
     plt.savefig(fname, dpi=DPI)
     # plt.show()
     plt.close()
+
+    f, ax = plt.subplots(1, 1, figsize=(4, 4))
+    mat = (CorrTh - CorrSims).reshape((10, nlbins, 10, nlbins))
+    ixgrid = np.ix_([2, 3, 5, 6], [2, 3, 5, 6])
+    mat = mat.swapaxes(1, 2)[ixgrid].swapaxes(1, 2).reshape((4 * nlbins, 4 * nlbins))
+    cb = ax.imshow(mat, vmin=-0.02, vmax=0.02)
+    divider = make_axes_locatable(ax)
+    cax = divider.append_axes("right", size="5%", pad=0.05)
+    cbar = f.colorbar(cb, cax=cax)
+    cbar.ax.tick_params(labelsize=8)
+    for i in range(1, 4):
+        ax.plot([0, 4*nlbins-1], [i*nlbins, i*nlbins], 'k-', lw=0.5)
+        ax.plot([i*nlbins, i*nlbins], [0, 4*nlbins-1], 'k-', lw=0.5)
+
+    ticks = [(i + 0.5) * nlbins for i in range(4)]
+    ax.set_xticks(ticks)
+    ax.set_xticklabels(np.array(labels_cov)[[2, 3, 5, 6]], fontsize=8)
+    ax.set_yticks(ticks)
+    ax.set_yticklabels(np.array(labels_cov)[[2, 3, 5, 6]], fontsize=8)
+
+    plt.tight_layout()
+    fname = os.path.join(outdir, 'run_sph_2b_same_mask_NKA_diff_corr_short.pdf')
+    plt.savefig(fname, dpi=DPI)
+    # plt.show()
+    plt.close()
+
 
 ##############################################################################
 ####################### TTTT, TETE, EEEE chi2 plot ###########################

### 2019-06-11 notebooks/general-plots-paper.py
added Spin0 reldev

@@ -650,6 +650,7 @@ def chi2_Spin0_NKA_TTTEEE_full():
 def eigv_sph_1bin():
     CovSims = np.load('../run_sph_2b/run_sph_2b_covTTTEE_b1_short_cl_0001-20000.npz')['arr_0']
     CovTh = np.load('../run_sph_2b/run_sph_2b_covThTTTEEE_b1_short.npz')['arr_0']
+    CovTh0 = np.load('../run_sph_2b/run_sph_2b_covSpin0TTTEEE_b1_short.npz')['arr_0']
 
     f, ax = plt.subplots(2, 1, sharex=True, figsize=(4,4))
 
@@ -659,15 +660,19 @@ def eigv_sph_1bin():
     ax[1].plot(X, 0 * X)
 
     Eval_Th, Evec_Th = co.diagonalize(CovTh)
+    Eval_Th0, Evec_Th0 = co.diagonalize(CovTh0)
 
     ax[0].plot(X, Eval_Th, label='NKA')
     ax[1].plot(X, Eval_Th/Eval_Sims - 1)
+
+    ax[0].plot(X, Eval_Th0, label='Spin-0')
+    ax[1].plot(X, Eval_Th0/Eval_Sims - 1)
 
     ax[0].set_ylabel('Eigenvalue')
     ax[0].legend(loc=0, frameon=False)
 
     ax[1].set_xlabel('Element')
-    ax[1].set_ylabel(r'$\tilde e_{i}^{NKA} / \tilde e_{i}^{Sim.} - 1$')
+    ax[1].set_ylabel(r'$\tilde e_{i}^{Analytical} / \tilde e_{i}^{Sim.} - 1$')
 
     ax[0].set_yscale('log')
     ax[1].set_yscale('linear')

### 2019-06-11 notebooks/general-plots-paper.py
updated plots, save Spin0 TTTEEE cov, add ylabel to chi2 1bin plot

@@ -569,6 +569,7 @@ def chi2_flat_TT_TE_EE_TB_EB_BB():
         i += 1
 
     axs[0, 0].set_ylabel('pdf')
+    axs[1, 0].set_ylabel('pdf')
 
     # axs[0].legend(loc='center right', fontsize='9') # , frameon=False)
     axs[0, 0].legend(loc='center right', fontsize='8', frameon=False)
@@ -644,7 +645,7 @@ def chi2_Spin0_NKA_TTTEEE_full():
     plt.close()
 
 ##############################################################################
-############################## What to plot ##################################
+############################## 1bin cov eigv #################################
 ##############################################################################
 
 def eigv_sph_1bin():
@@ -686,16 +687,16 @@ def eigv_sph_1bin():
     plt.close()
 
 if __name__ == '__main__':
-    pz()
-    masks()
-    foregrounds()
-    chi2_foregrounds()
-    corr_diff_2bins()
-    # compare_cls_1bin()
-    cls_2b()
-    chi2_sph_TT_TE_EE()
+    # pz()
+    # masks()
+    # foregrounds()
+    # chi2_foregrounds()
+    # corr_diff_2bins()
+    # # compare_cls_1bin()
+    # cls_2b()
+    # chi2_sph_TT_TE_EE()
     chi2_flat_TT_TE_EE_TB_EB_BB()
-    chi2_NKA_TTTEEE_full()
-    eigv_sph_1bin()
-    chi2_Spin0_NKA_TTTEEE_full()
+    # chi2_NKA_TTTEEE_full()
+    # eigv_sph_1bin()
+    # chi2_Spin0_NKA_TTTEEE_full()
 

### 2019-06-11 notebooks/general-plots-paper.py
uncommented lines

@@ -687,16 +687,16 @@ def eigv_sph_1bin():
     plt.close()
 
 if __name__ == '__main__':
-    # pz()
-    # masks()
-    # foregrounds()
-    # chi2_foregrounds()
-    # corr_diff_2bins()
-    # # compare_cls_1bin()
-    # cls_2b()
-    # chi2_sph_TT_TE_EE()
+    pz()
+    masks()
+    foregrounds()
+    chi2_foregrounds()
+    corr_diff_2bins()
+    # compare_cls_1bin()
+    cls_2b()
+    chi2_sph_TT_TE_EE()
     chi2_flat_TT_TE_EE_TB_EB_BB()
-    # chi2_NKA_TTTEEE_full()
-    # eigv_sph_1bin()
-    # chi2_Spin0_NKA_TTTEEE_full()
+    chi2_NKA_TTTEEE_full()
+    eigv_sph_1bin()
+    chi2_Spin0_NKA_TTTEEE_full()
 

### 2019-06-12 notebooks/general-plots-paper.py
plot diag cov conts.

@@ -298,6 +298,53 @@ def chi2_foregrounds():
     plt.close()
 
 ##############################################################################
+######################### Foregrounds diag cov ###############################
+##############################################################################
+
+def foregrounds_cov_diag():
+    lbpw = np.loadtxt('../run_sph/run_sph_ells.txt')
+    CovSims = np.load('../run_sph/run_sph_covTTTEE_short_clsims_0001-20000.npz')['arr_0']
+    CovSims_fore = np.load('../run_sph_contaminants_ls100_ss100/run_sph_contaminants_ls100_ss100_covTTTEE_short_clsims_0001-20000.npz')['arr_0']
+    CovTh = np.load('../run_sph_contaminants_ls100_ss100/run_sph_contaminants_ls100_ss100_covThTTTEEE_short.npz')['arr_0']
+
+    nlbins = int(CovSims.shape[0] / 3)
+    X = lbpw[:nlbins]
+
+    f, axs = plt.subplots(2, 3, sharex=True, sharey='row', gridspec_kw={'height_ratios': [2, 1], 'hspace': 0, 'wspace': 0}, figsize=(6, 3))
+    for i in range(3):
+        index1 = i * nlbins
+        index2 = (i + 1) * nlbins
+
+        Y0 = np.diag(CovSims[index1 : index2, index1 : index2])
+        axs[0, i].loglog(X, Y0, label='Sim')
+        axs[1, i].semilogx(X, Y0/Y0 - 1, label='Sim')
+
+        Y = np.diag(CovSims_fore[index1 : index2, index1 : index2])
+        axs[0, i].loglog(X, Y, ls='--', label='Sim w/ cont.')
+        axs[1, i].semilogx(X, Y/Y0 - 1, ls='--', label='Sim w/ cont.')
+
+        Y = np.diag(CovTh[index1 : index2, index1 : index2])
+        axs[0, i].loglog(X, Y, label='NKA')
+        axs[1, i].semilogx(X, Y/Y0 - 1, label='NKA')
+
+    axs[0, 0].set_ylabel('$C_{\ell\ell}$')
+    axs[1, 0].set_ylabel(r'$\frac{C_{\ell\ell}}{C_{\ell\ell}^{Sim.}} - 1$')
+
+    label = [r'$\delta \delta$', r'$\delta \gamma_E$', r'$\gamma_E \gamma_E$']
+    for i in range(3):
+        axs[1, i].set_xlabel('$\ell$')
[…]

### 2019-06-13 notebooks/general-plots-paper.py
removed reldev panel

@@ -310,29 +310,32 @@ def foregrounds_cov_diag():
     nlbins = int(CovSims.shape[0] / 3)
     X = lbpw[:nlbins]
 
-    f, axs = plt.subplots(2, 3, sharex=True, sharey='row', gridspec_kw={'height_ratios': [2, 1], 'hspace': 0, 'wspace': 0}, figsize=(6, 3))
+    # f, axs = plt.subplots(2, 3, sharex=True, sharey='row', gridspec_kw={'height_ratios': [2, 1], 'hspace': 0, 'wspace': 0}, figsize=(6, 3))
+    f, axs = plt.subplots(1, 3, sharex=True, sharey='row', gridspec_kw={'hspace': 0, 'wspace': 0}, figsize=(8, 3))
+    axs = axs.reshape((1, 3))
     for i in range(3):
         index1 = i * nlbins
         index2 = (i + 1) * nlbins
 
         Y0 = np.diag(CovSims[index1 : index2, index1 : index2])
         axs[0, i].loglog(X, Y0, label='Sim')
-        axs[1, i].semilogx(X, Y0/Y0 - 1, label='Sim')
+        # axs[1, i].semilogx(X, Y0/Y0 - 1, label='Sim')
 
         Y = np.diag(CovSims_fore[index1 : index2, index1 : index2])
         axs[0, i].loglog(X, Y, ls='--', label='Sim w/ cont.')
-        axs[1, i].semilogx(X, Y/Y0 - 1, ls='--', label='Sim w/ cont.')
+        # axs[1, i].semilogx(X, Y/Y0 - 1, ls='--', label='Sim w/ cont.')
 
         Y = np.diag(CovTh[index1 : index2, index1 : index2])
         axs[0, i].loglog(X, Y, label='NKA')
-        axs[1, i].semilogx(X, Y/Y0 - 1, label='NKA')
+        # axs[1, i].semilogx(X, Y/Y0 - 1, label='NKA')
 
     axs[0, 0].set_ylabel('$C_{\ell\ell}$')
-    axs[1, 0].set_ylabel(r'$\frac{C_{\ell\ell}}{C_{\ell\ell}^{Sim.}} - 1$')
+    # axs[1, 0].set_ylabel(r'$\frac{C_{\ell\ell}}{C_{\ell\ell}^{Sim.}} - 1$')
 
     label = [r'$\delta \delta$', r'$\delta \gamma_E$', r'$\gamma_E \gamma_E$']
     for i in range(3):
[…]

### 2019-06-14 get_grid/get_grid.py
Uploaded get_grid and plots

@@ -1,0 +1,160 @@
+import os, sys
+import re, pickle
+import numpy as np
+import get_grid_tools as tools
+
+# Call the parser
+args = tools.argument_parser()
+
+
+# Run grid
+if args.mode=='run':
+
+
+    # Folders
+    data_dir   = os.path.abspath(args.data)
+    output_dir = os.path.abspath(args.output)
+
+    # Parameters
+    n_bins       = args.n_bins
+    cov_m    = args.cov_m
+    ell_min  = args.ell_min
+    var_pars = np.array(args.pars)
+
+    if len(args.n_grid)==len(var_pars):
+        n_grid  = args.n_grid
+    elif len(args.n_grid)==1:
+        n_grid = np.full(var_pars.shape,args.n_grid[0])
+    else:
+        raise IOError('n_grid should either contain one element or an array with the same dimension of params!')
+
+    # Initialize varying parameters
+    all_pars = tools.default_pars.copy()
+    is_s8 = 'sigma8' in var_pars
+    is_As = 'ln10_A_s' in var_pars
+    if is_s8 and is_As:
+        raise IOError('sigma8 and ln10_A_s can not be varied at the same time. Choose one of the two!')
+    elif is_s8:
+        all_pars.pop('ln10_A_s',None)
+    else:
+        all_pars.pop('sigma8',None)
+
+    # Create name file output
+    s1 = '_'.join(var_pars)
+    s2 = '_'.join([str(i) for i in n_grid])
+    grid_fn = '{}/chi2_{}_ng_{}_lmin_{}_nb_{}_cov_{}.npz'.format(output_dir,s1,s2,ell_min,n_bins,cov_m)
+
+    # Load data
+    data, data_ref = tools.load_data(data_dir,cov_m,n_bins,ell_min)
+
+    # Get range of variation for each parameter
+    range_m_M = tools.get_range(var_pars,all_pars,data_ref)
+
+    sets = {
+        'n_bins'         : n_bins,
+        'cov_m'          : cov_m,
+        'ell_min'        : ell_min,
+        'var_pars'       : var_pars,
+        'n_grid'         : n_grid,
+        'all_pars'       : all_pars,
+        'grid_fn'        : os.path.basename(grid_fn),
+        'range_m_M'      : range_m_M
+    }
+
+    # Save settings to file
+    sets_out = open(grid_fn.replace('.npz','.pickle'),'wb')
[…]

### 2019-06-14 get_grid/get_grid_tools.py

@@ -1,0 +1,243 @@
+import os, sys
+import argparse
+import numpy as np
+import pyccl as ccl
+
+
+default_pars = {
+    'h'               :  0.67,
+    'Omega_c'         :  0.27,
+    'Omega_b'         :  0.045,
+    'ln10_A_s'        :  3.044522438,
+    'sigma8'          :  0.840421163375,
+    'n_s'             :  0.96,
+    'w_0'             : -1.0,
+    'w_a'             :  0.0
+}
+
+
+def flatten_cls(cls, n_bte, n_ells):
+    flat_cls = np.moveaxis(cls,[-3,-2,-1],[0,1,2])
+    flat_cls = flat_cls[np.triu_indices(n_bte)]
+    flat_cls = flat_cls.reshape(((n_bte+1)*n_bte*n_ells/2,)+cls.shape[:-3])
+    return flat_cls
+
+def unflatten_cls(cls, n_bte, n_ells):
+    tmp_cls = cls.reshape((n_bte+1)*n_bte/2,n_ells)
+    unflat_cls = np.zeros((n_bte,n_bte,n_ells))
+    unflat_cls[np.triu_indices(n_bte)] = tmp_cls
+    unflat_cls = np.moveaxis(unflat_cls,[0,1],[1,0])
+    unflat_cls[np.triu_indices(n_bte)] = tmp_cls
+    return unflat_cls
+
+def flatten_covmat(cov, n_bte, n_ells):
+    flat_cov = flatten_cls(cov, n_bte, n_ells)
+    flat_cov = flatten_cls(flat_cov, n_bte, n_ells)
+    return flat_cov
+
+def unflatten_covmat(cov, n_bte, n_ells):
+    unflat_cov = np.apply_along_axis(unflatten_cls, -1, cov, n_bte, n_ells)
+    unflat_cov = np.apply_along_axis(unflatten_cls, -4, unflat_cov, n_bte, n_ells)
+    return unflat_cov
+
+
+def get_cosmo_ccl(pars):
+    try:
+        cosmo = ccl.Cosmology(
+            h        = pars['h'],
+            Omega_c  = pars['Omega_c'],
+            Omega_b  = pars['Omega_b'],
+            A_s      = (10.**(-10.))*np.exp(pars['ln10_A_s']),
+            n_s      = pars['n_s'],
+            w0       = pars['w_0'],
+            wa       = pars['w_a']
+            )
+    except:
+        cosmo = ccl.Cosmology(
[…]

### 2019-06-14 lsst_cov.py

@@ -4,60 +4,6 @@ import os
 import sys
 
 
-def nofz(z,z0,sz,ndens):
-    return np.exp(-0.5*((z-z0)/sz)**2)*ndens/np.sqrt(2*np.pi*sz**2)
-
-# pz single bin
-z_1  = np.array([np.linspace(0,3,512)])
-pz_1 = np.array([nofz(z_1[0],0.955,0.13,7.55)])
-
-# pz new bins
-z_2  = np.array([np.linspace(0,3,512),np.linspace(0,3,512)])
-pz_2 = np.array([nofz(z_2[0],0.955,0.13,7.55),nofz(z_2[1],1.355,0.13,7.55)])
-z_3  = np.array([np.linspace(0,3,512),np.linspace(0,3,512),np.linspace(0,3,512)])
-pz_3 = np.array([nofz(z_3[0],0.955,0.13,7.55),nofz(z_3[1],1.355,0.13,7.55),nofz(z_3[2],1.755,0.13,7.55)])
-z_n  = np.array([
-    np.linspace(0,3,512),
-    np.linspace(0,3,512),
-    np.linspace(0,3,512),
-    np.linspace(0,3,512),
-    np.linspace(0,3,512),
-    np.linspace(0,3,512),
-    np.linspace(0,3,512),
-    np.linspace(0,3,512),
-    np.linspace(0,3,512),
-    np.linspace(0,3,512)
-    ])
-pz_n = np.array([
-    nofz(z_n[0],0.955,0.13,7.55),
-    nofz(z_n[1],1.155,0.13,7.55),
-    nofz(z_n[2],1.255,0.13,7.55),
-    nofz(z_n[3],1.355,0.13,7.55),
-    nofz(z_n[4],1.455,0.13,7.55),
-    nofz(z_n[5],1.655,0.13,7.55),
-    nofz(z_n[6],1.755,0.13,7.55),
-    nofz(z_n[7],1.855,0.13,7.55),
-    nofz(z_n[8],1.955,0.13,7.55),
-    nofz(z_n[9],2.055,0.13,7.55)
-    ])
-
-pars = {
-    'h'       : 0.67,
-    'Omega_c' : 0.27,
-    'Omega_b' : 0.045,
-    'A_s'     : 2.1e-9,
-    'n_s'     : 0.96,
-    'w_0'     : -1.,
-    'w_a'     : -0.,
-    'fsky'    : 1.
-}
-
-ell_bp = np.load('./data/run_sph_ells.npz')['lsims'].astype(int)
[…]

### 2019-06-15 notebooks/row-sph-1bin-cov.py
Naive -> MC

@@ -121,7 +121,7 @@ def ax_plot_row(ax, lbins, CovSims, CovTh, CovThN, normalization, index=20, dx=5
     indexi = index
     Xi = lbins[indexi : indexi + (peaks * 2 * dx) : 2 * dx]
     Yth = np.diag(CovThN)[indexi : indexi + (peaks * 2 * dx) : 2 * dx]/normalization[indexi : indexi + (peaks * 2 * dx) : 2 * dx]
-    ax.scatter(Xi, Yth, marker="*", s=9, c='k', label='Naive', zorder=3)
+    ax.scatter(Xi, Yth, marker="*", s=9, c='k', label='MC', zorder=3)
 
     # dl = lbins[1] - lbins[0]
     # Naive = 1. / (0.4 * (2 * lbins  + 1) * dl)

### 2019-06-15 notebooks/general-plots-paper.py
fixed Figure 9

@@ -501,7 +501,8 @@ def chi2_sph_TT_TE_EE():
 
         i += 1
 
-    axs[0].set_ylabel('pdf')
+    axs[0].set_ylabel('$p(\chi^2)$')
+    axs[0].set_yticks([])
 
     # axs[0].legend(loc='center right', fontsize='9') # , frameon=False)
     axs[0].legend(loc='center right', fontsize='9', frameon=False)

### 2019-06-15 notebooks/general-plots-paper.py
pdf -> p(chi2) + no yticks for chi2 plots

@@ -285,10 +285,11 @@ def chi2_foregrounds():
     ax.plot(x[:-1], stats.chi2.pdf(x[:-1], clTTTEEE.shape[1]), ls='--', label=r'$\chi^2$ pdf')
 
     ax.set_xlabel(r'$\chi^2$')
-    ax.set_ylabel('$10^3$ pdf')
+    ax.set_ylabel('$p(\chi^2)$')
+    ax.set_yticks([])
 
-    y_vals = ax.get_yticks()
-    ax.set_yticklabels([str(x * 1000) for x in y_vals])
+    # y_vals = ax.get_yticks()
+    # ax.set_yticklabels([str(x * 1000) for x in y_vals])
 
     ax.legend(loc='upper right', fontsize='8', frameon=False)
     plt.tight_layout()
@@ -619,8 +620,10 @@ def chi2_flat_TT_TE_EE_TB_EB_BB():
 
         i += 1
 
-    axs[0, 0].set_ylabel('pdf')
-    axs[1, 0].set_ylabel('pdf')
+    axs[0, 0].set_ylabel('$p(\chi^2)$')
+    axs[0, 0].set_yticks([])
+    axs[1, 0].set_ylabel('$p(\chi^2)$')
+    axs[1, 0].set_yticks([])
 
     # axs[0].legend(loc='center right', fontsize='9') # , frameon=False)
     axs[0, 0].legend(loc='center right', fontsize='8', frameon=False)
@@ -651,10 +654,11 @@ def chi2_NKA_TTTEEE_full():
     ax.plot(x[:-1], stats.chi2.pdf(x[:-1], lmax * 10), ls='--', label=r'$\chi^2$ pdf')
 
     ax.set_xlabel(r'$\chi^2$')
-    ax.set_ylabel('$10^3$ pdf')
+    ax.set_ylabel('$p(\chi^2)$')
+    ax.set_yticks([])
 
-    y_vals = ax.get_yticks()
-    ax.set_yticklabels([str(x * 1000) for x in y_vals])
+    # y_vals = ax.get_yticks()
+    # ax.set_yticklabels([str(x * 1000) for x in y_vals])
 
     ax.legend(loc='upper right', fontsize='9', frameon=False)
     plt.tight_layout()
@@ -683,10 +687,11 @@ def chi2_Spin0_NKA_TTTEEE_full():
[…]

### 2019-06-15 notebooks/general-plots-paper.py
improved flat sky mask

@@ -193,13 +193,16 @@ def masks():
     fname = os.path.join(outdir, 'mask-lss_flat1.pdf')
 
     fmask = "./data/mask_lss_flat.fits"
+    f, ax = plt.subplots(figsize=(5, 2))
 
     fmi, mask_hsc = fm.read_flat_map(fmask)
     mask_hsc[mask_hsc == 0] = hp.UNSEEN
 
-    fmi.view_map(mask_hsc, addColorbar=False)
+    fmi.view_map(mask_hsc, ax=ax, addColorbar=False)
 
-    plt.savefig(fname, dpi=DPI, bbox_iches='tight')
+    plt.tight_layout()
+    plt.savefig(fname, dpi=DPI, bbox_iches='tight',
+                pad_inches=0)
     plt.close()
 
 ##############################################################################

### 2019-06-17 notebooks/general-plots-paper.py
plot corr diff of 2bin diff mask case

@@ -352,10 +352,10 @@ def foregrounds_cov_diag():
 
 
 ##############################################################################
-############################# 2bins corr diff ################################
+###################### 2bins corr diff (same mask) ###########################
 ##############################################################################
 
-def corr_diff_2bins():
+def corr_diff_2bins_same_mask():
     Cth = np.load('../run_sph_2b_same_mask/full_covariance/run_sph_2b_same_mask_covTh_TTTEEE_short_2bins_same_mask.npz')['arr_0']
     Csims = np.load('../run_sph_2b_same_mask/full_covariance/run_sph_2b_same_mask_covSims_TTTEEE_short_0001-20000.npz')['arr_0']
 
@@ -418,6 +418,44 @@ def corr_diff_2bins():
     # plt.show()
     plt.close()
 
+##############################################################################
+####################### 2bins corr diff (diff mask) ##########################
+##############################################################################
+
+def corr_diff_2bins_diff_mask():
+    Cth = np.load('../run_sph_2b/run_sph_2b_covThTTTEEE_short.npz')['arr_0']
+    Csims = np.load('../run_sph_2b/run_sph_2b_covTTTEEE_short.npz')['arr_0']
+
+    ell = np.loadtxt('../run_sph_2b/run_sph_2b_ells.txt')
+
+    CorrSims = Csims/np.sqrt(np.diag(Csims)[:, None] * np.diag(Csims)[None, :])
+    CorrTh = Cth/np.sqrt(np.diag(Cth)[:, None] * np.diag(Cth)[None, :])
+
+
+    labels = [r'$\delta_1 \delta_2$', r'$\delta_1 \gamma_{E,2}$', r'$\gamma_{E,1} \delta_2$', r'$\gamma_{E,1} \gamma_{E,2}$']
+
+    f, ax = plt.subplots(1, 1, figsize=(4, 4))
+    nlbins = int(CorrSims.shape[0] / 4)
+    cb = ax.imshow(CorrTh - CorrSims, vmin=-0.02, vmax=0.02)
+    divider = make_axes_locatable(ax)
+    cax = divider.append_axes("right", size="5%", pad=0.05)
+    cbar = f.colorbar(cb, cax=cax)
+    cbar.ax.tick_params(labelsize=8)
+    for i in range(1, 4):
[…]

### 2019-06-17 notebooks/general-plots-paper.py
label Cov as Cov

@@ -333,7 +333,7 @@ def foregrounds_cov_diag():
         axs[0, i].loglog(X, Y, label='NKA')
         # axs[1, i].semilogx(X, Y/Y0 - 1, label='NKA')
 
-    axs[0, 0].set_ylabel('$C_{\ell\ell}$')
+    axs[0, 0].set_ylabel('Cov$_{\ell\ell}$')
     # axs[1, 0].set_ylabel(r'$\frac{C_{\ell\ell}}{C_{\ell\ell}^{Sim.}} - 1$')
 
     label = [r'$\delta \delta$', r'$\delta \gamma_E$', r'$\gamma_E \gamma_E$']

### 2019-06-17 notebooks/general-plots-paper.py
added mean shift information for 1b case (Fig.9)

@@ -527,6 +527,8 @@ def chi2_sph_TT_TE_EE():
              r"$(\gamma_E \gamma_E, \gamma_E \gamma_E)$"]
     i = 0
     for ax, chi2_list in zip(axs, [chi2_TT_ar, chi2_TE_ar, chi2_EE_ar]):
+        # means = np.mean(chi2_list, axis=1)
+        # print('Mean shift: ' + str(means/means[0] -1))
         bins = np.linspace(np.min(chi2_list), np.max(chi2_list), 60)
         _, x, _ = ax.hist(chi2_list[0], bins=bins, histtype='step', density=True,
                           label='Sims.')
@@ -550,7 +552,7 @@ def chi2_sph_TT_TE_EE():
     axs[0].legend(loc='center right', fontsize='9', frameon=False)
     plt.tight_layout()
     fname = os.path.join(outdir, 'run_sph_2b_1stbin_chi2_TT_TE_EE.pdf')
-    plt.savefig(fname, dpi=DPI)
+    # plt.savefig(fname, dpi=DPI)
     # plt.show()
     plt.close()
 

### 2019-06-17 draft/paper.tex

--- 2019-05-21 draft/paper.tex
@@
-\documentclass[useAMS,usenatbib]{mn2e}
-\bibliographystyle{mn2e}
+\documentclass[a4paper,11pt]{article}
+
 \pdfoutput=1
-\pdfminorversion=5
-
-%\usepackage{widetext}
+\usepackage{jcappub}
 \usepackage{graphicx}
-\usepackage{dcolumn}
 \usepackage{bm}
@@
 \usepackage[dvipsnames]{xcolor}
-\usepackage[colorlinks,linkcolor=red,citecolor=blue,urlcolor=blue ]{hyperref}
-\usepackage{multirow}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{lipsum}
-\newcommand{\nv}{\hat{\mathbf{\theta}}}
-\newcommand{\kalo}{Karhunen-Lo\`{e}ve~}
-\newcommand{\cfh}{CFHTLenS~}
+\usepackage[caption=false]{subfig}
+
 \newcommand{\jcap}{JCAP}
@@
 \newcommand{\physrep}{Physics Reports}
+\newcommand{\pasj}{Publications of the ASJ}
+
+\newcommand{\nv}{\hat{\boldsymbol{\theta}}}
+\newcommand{\summ}[1]{\sum_{\bf #1}\Delta #1^2}
+\newcommand{\Ylm}[3]{{\sf E}^{#1}_{\bf #2}({\bf #3})}
+\newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}
 
 \newcommand{\todo}[1]{{\bf TODO: #1}}
-%\newcommand{\todo}[1]{$\,$}
-
-\newcommand{\PCL}{pseudo-$C_\ell$~}
-\newcommand{\pcl}[3]{\hat C_{#1}^{#2 #3}}
-\newcommand{\avg}[1]{\langle #1 \rangle}
-\newcommand{\fsky}{f_{\mbox{\rm sky}}}
-\newcommand{\clth}{C_\ell^{th}}
-\newcommand{\clf}{C_\ell^{fore}}
-\newcommand{\cl}{C_\ell}
-\newcommand{\summ}[1]{\sum_{\bf #1}\Delta #1^2}
-\newcommand{\Ylm}[3]{{\sf E}^{#1}_{\bf #2}({\bf #3})}
 \newcommand{\red}[1]{{\color{red} #1}}
 
-\title[Disconnected pseudo-$C_\ell$ covariances for projected datasets]{Disconnected pseudo-$C_\ell$ covariances for projected datasets}
-\author[C.
-Garc\'{i}a-Garc\'{i}a et al.]{Carlos Garc\'{i}a-Garc\'{i}a$^{1,\,2}$\thanks{carlosgarcia@iff.csic.es}, David Alonso$^2$\thanks{david.alonso@physics.ox.ac.uk}, Emilio Bellini$^2$\thanks{emilio.bellini@physics.ox.ac.uk}\\
-$^{1}$ Instituto de Física Fundamental, Consejo Superior de Investigaciones
-Científicas, c/. Serrano 123, E–28006, Madrid, Spain\\
-$^{2}$Oxford Astrophysics, Department of Physics, Keble Road, Oxford, OX1 3RH, UK
-}
+\title{\boldmath Disconnected pseudo-$C_\ell$ covariances for projected large-scale structure data}
+
+\author[a]{Carlos Garc\'{i}a-Garc\'{i}a}
+\author[b]{David Alonso}
+\author[b]{Emilio Bellini}
+
+\affiliation[a]{Instituto de Física Fundamental, Consejo Superior de Investigaciones Científicas, c/. Serrano 123, E–28006, Madrid, Spain}
+\affiliation[b]{Oxford Astrophysics, Department of Physics, Keble Road, Oxford, OX1 3RH, UK}
+
+\emailAdd{carlosgarcia@iff.csic.es}
+\emailAdd{david.alonso@physics.ox.ac.uk}
+\emailAdd{emilio.bellini@physics.ox.ac.uk}
+
+\abstract{The disconnected part of the power spectrum covariance matrix (also known as the ``Gaussian'' covariance) is the dominant contribution on large scales for galaxy clustering and weak lensing datasets.
+The presence of a complicated sky mask causes non-trivial correlations between different Fourier/harmonic modes, which must be accurately characterised in order to obtain reliable cosmological constraints.
+Unfortunately, an exact calculation of these correlations involve $O(\ell_{\rm max}^6)$ operations that become computationally impractical very quickly.
+We present an implementation of approximate methods to estimate the Gaussian covariance matrix of spin-0 and spin-2 flat- and curved-sky fields, expanding on existing algorithms.
+These methods achieve an $O(\ell_{\rm max}^3)$ scaling, which makes the computation of the covariance matrix as fast as the computation of the power spectrum itself.
+We quantify the accuracy of these method on large-scale structure and weak lensing data, characterised by comparatively complex survey geometries, making use of a large number of Gaussian but otherwise realistic simulations.
+We show that, using the covariance matrix computing using these approximations we are able to recover the true posterior distribution cosmological parameters to high accuracy.
+We also quantify the shortcomings of these methods, which become unreliable on the very largest scales, as well as for covariance matrix elements involving lensing $B$ modes.
+The algorithms presented here are implemented in the public code {\tt NaMaster} \url{https://github.com/LSSTDESC/NaMaster}.}
 
 \begin{document}
-  \date{\today}
-  \pagerange{1--18} \pubyear{2019}
   \maketitle
+  \flushbottom
 
-\begin{abstract}
-\end{abstract}
+  \section{Introduction}\label{sec:intro}
+    The two-point correlation of different fields projected on the celestial sphere is one of the most common observables used in the analysis of large datasets in astrophysics, from studies of the Cosmic Microwave Background \citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,1998PhRvD..57.2117B,2001PhRvD..64f3001T,2002ApJ...567....2H,2003PhRvD..67b3001W} to large-scale structure and weak lensing surveys \cite{2000MNRAS.317L..23H,2001ApJ...555..547H,2001MNRAS.325.1603E,2011arXiv1112.5723H,2016MNRAS.456.1508K,2018MNRAS.476.1050B,2018arXiv181208182X,2019MNRAS.tmp.1446C,2019PASJ...71...43H}.
+Using these two-point functions achieves a high level of data compression (with respect to the size of the raw datasets -- time-ordered data, images or catalogs), and they can be directly used to constrain cosmological and astrophysical parameters assuming that one can model their likelihood.
+This is usually done by assuming that the two-point functions are Gaussianly distributed, which is often good approximation due to the central limit theorem \cite{2008PhRvD..77j3013H,2018MNRAS.477.4879S}.
+In this case, the only obstacle that remains is being able to estimate the covariance matrix of a set of two-point correlators.
+Since the form of this covariance directly affects the posterior parameter uncertainties, a precise determination of it is of paramount importance.
+In large-scale structure experiments this has often been resolved by making use of one's own data through resampling techniques \cite{tukey1958,1984MNRAS.210P..19B,1993ApJ...406L..47H,2002ApJ...571..172Z,2016arXiv160600233E}, or by generating a large number of mock realizations \cite{2013MNRAS.428.1036M,2018MNRAS.479...94A,2019MNRAS.485.2806B}.
+With the advent of the larger current and future surveys, the increasing size of the data vector and of the volume to be simulated has made this solution impractical, and fully analytical and hybrid estimators are now being used.
 
-\begin{keywords}
-  cosmology: large-scale structure of the Universe -- methods: data analysis
-\end{keywords}
+    The problem of producing accurate analytical estimates of the covariance matrix for large-scale structure data has seen significant progress in the last few years \cite{2001ApJ...554...56C,2009ApJ...701..945S,2009MNRAS.395.2065T,2013PhRvD..87l3504T,2014MNRAS.441.2456T,2016PhRvD..94f3533P}.
+As described in \cite{2018JCAP...10..053B}, the covariance matrix recieves three main contributions:
+    \begin{itemize}
+      \item {\bf Gaussian covariance:} this is the contribution to the covariance from the disconnected part of the trispectrum of the different fields involved (also called the ``disconnected'' covariance).
+In simpler terms, this is the covariance matrix one would obtain if all fields involved were Gaussianly distributed.
+      \item {\bf Connected non-Gaussian covariance:} this is the contribution
+        from the connected trispectrum (which would vanish if all fields were
+        Gaussianly distributed).
+      \item {\bf Super-sample covariance:} this is the additional coupling between different scales induced by density fluctuations on scales larger than the volume mapped.
+    \end{itemize}
 
-\section{Introduction}
+    On most scales relevant for cosmological studies, the Gaussian contribution dominates the error budget, although the connected and super-sample terms cannot be neglected \cite{2018JCAP...10..053B}.
+The Gaussian contribution is trivial to compute for fields observed over the full sky:
+    \begin{equation}\label{eq:cov_naive}
+      {\rm Cov}\left(C^{ab}_\ell,C^{cd}_\ell\right)=\delta^K_{\ell\ell'}\frac{C^{ac}_\ell C^{bd}_\ell+C^{ad}_\ell C^{bc}_\ell}{2\ell+1},
+    \end{equation}
+    where $C^{xy}_\ell$ is the angular power spectrum between two maps $x$ and $y$ on multipole $\ell$.
+Unfortunately, the presence of a sky mask in general induces non-trivial couplings between different $\ell$s, which must be accurately estimated in order to produce unbiased evaluations of the parameter likelihood \cite{2018MNRAS.479.4998T}.
 
-\section{Analytical Gaussian covariances}
+    In this paper we will present and generalise methods developed in the context of CMB experiments to account for the impact of survey geometry on the Gaussian part of the power spectrum covariance matrix \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}, and will study in detail the performance of these methods for large-scale structure and weak lensing datasets.
+We have also implemented these approximations in the public code {\tt NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}} \cite{2019MNRAS.484.4127A}, making the computation of accurate Gaussian covariance matrices significantly simpler.
 
+    The paper is structured as follows: Section \ref{sec:theory} presents the methods and approximations used to calculate accurate covariances.
+In Section \ref{sec:results} we test the methods against Gaussian simulations and study their performance as well as their impact on the final cosmological parameter estimation.
+We then summarize our results and conclude in Section \ref{sec:discussion}.
+Appendix \ref{app:flat} presents the performance of these methods for flat-sky approximation, and we provide technical details of the software implementation in Appendix \ref{app:namaster}.
+
+  \section{Analytical Gaussian covariances}\label{sec:theory}
     \subsection{Preliminaries}\label{ssec:theory.prelim}
-    \todo{cites everywhere}
       We will deal with spin-0 and spin-2 fields defined on a 2-dimensional space.
@@
       \end{equation}
-    where the operator $\summ{x}$ denotes a integral or sum of all coordinates ${\bf x}$, and $\Ylm{\dag}{k}{x}$ are a set of orthogonal functions.
+      where the operator $\summ{x}$ denotes an integral or sum over all values of the coordinates ${\bf x}$, and $\Ylm{\dag}{k}{x}$ are a set of orthogonal functions.
 We will also assume that the $\Ylm{\dag}{k}{x}$ are a \emph{complete} set of basis functions, in which case:
@@
       \end{align}
-    where $\summ{k}$ denotes an integral over all possible generalised Fourier coefficients ${\bf k}$, and $\Delta^x$ and $\Delta^k$ are generalized delta functions:
+      where $\summ{k}$ denotes an integral over all possible generalised Fourier coefficients ${\bf k}$, and $\Delta^x$ and $\Delta^k$ are generalized delta functions, defined through their action on functions of ${\bf x}$ or ${\bf k}$:
       \begin{align}
@@
       \end{align}
-    For a spin-$s$ quantity, $\Ylm{}{l}{x}$ can be written in terms of a differential operator $\eth$ and a set of scalar orthogonal functions $q({\bf l},{\bf x})$ as:
+      For a spin-$s$ quantity, $\Ylm{}{l}{x}$ can be written in terms of two spin-raising and spin-lowering operators, $\eth$ and $\bar{\eth}$, and a set of scalar orthogonal functions $q({\bf l},{\bf x})$ as:
       \begin{equation}
@@
           -i(\eth^s-\bar{\eth}^s) & \eth^s+\bar{\eth}^s
-      \end{array}\right) q({\bf l},{\bf x}).
+        \end{array}\right) q({\bf l},{\bf x}),
       \end{equation}
+      where $\beta_{\ell,s}$ is a normalization factor defined in Table \ref{tab:notation}.
 
@@
 As a consequence of this, different generalised Fourier modes are uncorrelated:
+      \begin{equation}\label{eq:iso}
+        \langle {\bf a}_{\bf k}{\bf b}^\dag_{\bf l}\rangle\equiv {\sf C}^{ab}_\ell\,K\,\Delta^x({\bf k},{\bf l}),
+      \end{equation}
+      where $K$ is a volume factor (see below) and ${\sf C}^{ab}_k$ is the power spectrum. Defined this way, the power spectrum is a matrix, with elements
       \begin{equation}
-      \langle {\bf a}_{\bf k}{\bf b}^\dag_{\bf l}\rangle\equiv {\sf C}^{ab}_\ell\,F\,\Delta^x({\bf k},{\bf l}),
+        \left\langle a^\alpha_{\bf k}\left(b^\beta_{\bf l}\right)^*\right\rangle\equiv \left({\sf C}^{ab}_\ell\right)_{\alpha\beta}\,K\,\Delta^x({\bf k},{\bf l}),
       \end{equation}
-    where $F$ is a volume factor (see below) and ${\sf C}^{ab}_k$ is the power spectrum.
+      where $a^\alpha$ is the $\alpha$-th element of field ${\bf a}$.
+It will often be useful in what follows to think of ${\sf C}^{ab}_\ell$ as a 1-dimensional vector that we will denote by ${\rm vec}({\sf C}^{ab}_\ell)$.
+To do so, we simply map the two indices $(\alpha,\beta)$ into a single number $A$, such that ${\rm vec}({\sf C}^{ab}_\ell)_A = ({\sf C}^{ab}_\ell)_{\alpha\beta}$.
 
@@
           \hline
-      Mathematical object & Curved sky & Flat sky (continuum $\rightarrow$ discretized) \\
+          Symbol & Curved sky & Flat sky (continuum $\rightarrow$ discretized) \\
           \hline
@@
           \(\displaystyle \summ{x} \) & \(\displaystyle \int_0^{\phi} d\phi \int_{-1}^{1} d(\cos\theta)\) & \(\displaystyle \int \frac{dx^2}{2\pi}\rightarrow\sum_{\bf x}\frac{L_xL_y}{2\pi N_xN_y}\) \\
-      \(\displaystyle \Delta^k(\nv,\nv')\) & \(\displaystyle \delta^D(\cos\theta-\cos\theta')\delta^D(\varphi-\varphi')\) & \(\displaystyle 2\pi\,\delta^D({\bf x}-{\bf y})\rightarrow\delta^K_{xx'}\delta^K_{yy'}\frac{N_xN_y}{L_xL_y}2\pi\) \\
+          \(\displaystyle \Delta^k({\bf x},{\bf y})\) & \(\displaystyle \delta^D(\cos\theta-\cos\theta')\delta^D(\varphi-\varphi')\) & \(\displaystyle 2\pi\,\delta^D({\bf x}-{\bf y})\rightarrow\delta^K_{xx'}\delta^K_{yy'}\frac{N_xN_y}{L_xL_y}2\pi\) \\
           \(\displaystyle q({\bf l},{\bf x})\) & \(\displaystyle Y_{\ell m}(\nv)\) & \(\displaystyle e^{i{\bf l}\cdot{\bf x}}\) \\
@@
           \(\displaystyle \bar{\eth}\,_sf\) & \(\displaystyle -(\sin\theta)^{-s}\left(\partial_\theta-i\frac{\partial_\varphi}{\sin\theta}\right)(\sin\theta)^s\,_sf(\nv)\) & \(\displaystyle (\partial_x+i\partial_y)\,_sf\) \\
-      \(\displaystyle F \) & \(\displaystyle 1 \) & \(\displaystyle (2\pi)^{-1}\) \\
+          \(\displaystyle \beta_{\ell,s}\) & \(\displaystyle \sqrt{\frac{(\ell-2)!}{(\ell+2)!}}\) & \(\displaystyle \ell^{-s}\) \\
+          \(\displaystyle K \) & \(\displaystyle 1 \) & \(\displaystyle (2\pi)^{-1}\) \\
           \hline
         \end{tabular}
-     \caption{Lookup table describing the generalized notation introduced in Section \ref{ssec:theory.prelim}p for quantities defined on the sphere (second column) and on the flat 2D plane (third column).
+        \caption{Lookup table describing the generalized notation introduced in Section \ref{ssec:theory.prelim} for quantities defined on the sphere (second column) and on the flat 2D plane (third column).
 For the flat-sky case, we also provide expressions for a discretized, finite 2D plane with periodic boundary conditions.
@@
 
-  \subsection{The pseudo-$C_\ell$ method}
-    This section provides a very brief introduction to the \PCL power spectrum estimator. Further details can be found in \todo{cite}. 
+    \subsection{The pseudo-$C_\ell$ method}\label{ssec:theory.pcl}
+      This section provides a very brief introduction to the pseudo-$C_\ell$ power spectrum estimator.
+Further details can be found in e.g.
+\cite{2002ApJ...567....2H,2005MNRAS.360.1262B,2019MNRAS.484.4127A}.
       In any practical situation we do not have access to maps of a given field ${\bf a}$ over the full sky, but rather to a weighted or masked version of them
       \begin{equation}
-      \tilde{\bf a}({\bf x})\equiv w^a({\bf x}){\bf a}({\bf x}),
+        \tilde{\bf a}({\bf x})\equiv w_a({\bf x}){\bf a}({\bf x}),
+        \label{eq:mask}
       \end{equation}
-    where $w^a$ is commonly called the ``mask''. Due to the convolution theorem, the generalized Fourier coefficients of the masked field will be a convolution of the mask and true field coefficients:
+      where $w_a$ is commonly called the ``mask''.
+Due to the convolution theorem, the generalized Fourier coefficients of the masked field will be a convolution of the mask and true field coefficients:
       \begin{align}\nonumber
         \tilde{\bf a}_{\bf l}=&\summ{k}\left[\summ{x}\,w^a({\bf x})\,\Ylm{\dag}{l}{x}\,\Ylm{}{k}{x}\right]{\bf a}_{\bf k}\\
-                      \equiv&\summ{l}\,^a{\sf M}_{{\bf l}{\bf k}}\,{\bf a}_{\bf k},
+                        \equiv&\summ{k}\,^a{\sf M}_{{\bf l}{\bf k}}\,{\bf a}_{\bf k},
       \end{align}
@@
         \left\langle\tilde{\bf a}_{\bf l}\tilde{\bf b}^\dag_{\bf l}\right\rangle=&\summ{k}\summ{q}\,^a{\sf M}_{{\bf l}{\bf k}}\left\langle{\bf a}_{\bf k}{\bf b}^\dag_{\bf q}\right\rangle\,^b{\sf M}^\dag_{{\bf l}{\bf q}}\\
-                                                                              =&F\,\summ{k}\,^a{\sf M}_{{\bf l}{\bf k}}{\sf C}^{ab}_k\,^b{\sf M}^\dag_{{\bf l}{\bf k}}
+                                                                              =&K\,\summ{k}\,^a{\sf M}_{{\bf l}{\bf k}}{\sf C}^{ab}_k\,^b{\sf M}^\dag_{{\bf l}{\bf k}} \label{eq:M}
       \end{align}
 
-    The \PCL estimator then proceeds in two steps:
+      The pseudo-$C_\ell$ estimator then proceeds in two steps:
       \begin{enumerate}
-      \item We first bin different ${\bf l}$ modes into sets of them called bandpowers (typically bands of similar $\ell$ or annuli of flat-sky Fourier modes spanning a narrow range of radii).
+        \item We first bin different ${\bf l}$ modes into sets of them called bandpowers (typically bands of similar $\ell$ or annuli of flat-sky Fourier modes spanning a range of radii).
 Let us denote a given bandpower by its index $q$.
 The binned pseudo-power spectrum is:
-      \begin{equation}
+        \begin{equation}\label{eq:bandpowers}
           \tilde{\sf C}^{ab}_q=\sum_{{\bf l}\in q} B_q^{\bf l}\,\tilde{\bf a}_{\bf l}\tilde{\bf b}^\dag_{\bf l},
         \end{equation}
-      where the bandpower weights are normalized such that $\sum_{{\bf l}\in q}B_q^{\bf l}=1$.
+        where the bandpower weights are normalized such that $\sum_{{\bf l}\in q}B_q^{\bf l}=(K\Delta^x({\bf l},{\bf l}))^{-1}$.
         \item Then, the correlation between bandpowers induced by the mode-coupling coefficients is partially reversed by multiplying $\tilde{\sf C}^{ab}_q$ by the so-called binned ``mode-coupling matrix'' $\mathcal{M}$, giving the final estimator
@@
 
-      The main advantage of the \PCL estimator is that the mode-coupling matrix $\mathcal{M}$ is directly related to the coupling coefficients $\,^a{\sf M}_{{\bf l}{\bf l}'}$ (see \todo{cite}), and can be computed analytically making use of methods that scale like $\ell_{\rm max}^3$.
+        The main advantage of the pseudo-$C_\ell$ estimator is that the mode-coupling matrix $\mathcal{M}$ is directly related to the coupling coefficients $\,^a{\sf M}_{{\bf l}{\bf l}'}$, and can be computed analytically making use of methods that scale like $\ell_{\rm max}^3$ (see e.g.
+\cite{2002ApJ...567....2H}).
       \end{enumerate}
+
+      For completeness, the mode-coupling matrices for flat-sky and curved-sky fields are given by \cite{2019MNRAS.484.4127A}:
+      \begin{itemize}
+        \item {\bf Curved sky.} After averaging over the harmonic number $m$, the mode-coupling matrices are:
+        \begin{equation}\label{eq:mcms}
+          \left\langle\frac{1}{2\ell+1}\sum_{m=-\ell}^\ell {\rm vec}\left[\tilde{\bf a}_{\bf l}\tilde{\bf b}^\dag_{\bf l}\right]\right\rangle=\sum_{\ell'}{\sf M}^{s_as_b}_{\ell\ell'} {\rm vec}\left[{\sf C}^{ab}_{\ell'}\right],
+        \end{equation}
+        with
+        \begin{align}
+          &{\sf M}^{00}_{\ell\ell'}=(2\ell'+1)\,\Xi^{00}_{\ell\ell'}(w_a,w_b),\hspace{12pt}
+          {\sf M}^{02}_{\ell\ell'}=(2\ell'+1)\,\Xi^{0+}_{\ell\ell'}(w_a,w_b)\,\mathds{1},\\
+          &{\sf M}^{22}_{\ell\ell'}=(2\ell'+1)
+          \left(
+          \begin{array}{cccc}
+            \Xi^{++}_{\ell\ell'} & 0 & 0 & \Xi^{--}_{\ell\ell'} \\
+            0 & \Xi^{++}_{\ell\ell'} & -\Xi^{--}_{\ell\ell'} & 0 \\
+            0 & -\Xi^{--}_{\ell\ell'} & \Xi^{++}_{\ell\ell'} & 0 \\
+            \Xi^{--}_{\ell\ell'} & 0 & 0 & \Xi^{++}_{\ell\ell'}
+          \end{array}\right),
+        \end{align}
+        where
+        \begin{align}\label{eq:coeff_mcm}
+          &\Xi^{00}_{\ell\ell'}(w,v)  \equiv\sum_{\ell''}\frac{P^{wv}_{\ell''}}{4\pi}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}^2\\
+          &\Xi^{0+}_{\ell\ell'}(w,v)  \equiv\sum_{\ell''}\frac{P^{wv}_{\ell''}}{4\pi}\wtj{\ell}{\ell'}{\ell''}{0}{0}{0}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}\\
+          &\Xi^{\pm\pm}_{\ell\ell'}(w,v) \equiv\sum_{\ell''}\frac{P^{wv}_{\ell''}}{4\pi}\wtj{\ell}{\ell'}{\ell''}{2}{-2}{0}^2\frac{1\pm(-1)^{\ell+\ell'+\ell''}}{2}.
+        \end{align}
+        Here the 2-by-3 matrix-like quantities are the Wigner 3-$j$ symbols, and
+        \begin{equation}
+          P^{vw}_\ell\equiv\sum_{m=-\ell}^\ell v_{\ell m}w^*_{\ell m}.
+        \end{equation}
+        \item {\bf Flat sky.} In this case the averaging over the Fourier-space azimuth happens while binning into bandpowers, and therefore the unbinned mode-coupling matrix is defined before binning.
+Assuming flat bandpowers, such that $B_q^{\bf l}=(2\pi)^2/(L_xL_yN_q)$, where $N_q$ is the number of Fourier-space modes in the $q$-th bandpower:
+        \begin{equation}
+          \left\langle{\rm vec}\left[\tilde{\sf C}^{ab}_q\right]\right\rangle=\sum_{{\bf l}\in q}\frac{1}{N_q} \sum_{\bf k}{\sf M}^{s_as_b}_{{\bf l}{\bf k}}\,{\rm vec}\left[{\sf C}^{ab}_k\right]
+        \end{equation}
+        with
+        \begin{align}
+          {\sf M}^{00}_{{\bf l}\,{\bf k}}&\equiv \bar{\Xi}^{00}_{{\bf l}{\bf k}},\\
+          {\sf M}^{02}_{{\bf l}\,{\bf k}}&\equiv 
+          \left(
+          \begin{array}{cc}
+            \bar{\Xi}^{0+}_{{\bf l}{\bf k}} & -\bar{\Xi}^{0-}_{{\bf l}{\bf k}} \\
+            \bar{\Xi}^{0-}_{{\bf l}{\bf k}} &  \bar{\Xi}^{0+}_{{\bf l}{\bf k}}
+          \end{array}\right)\\\label{eq:m22_flat}
+          {\sf M}^{22}_{{\bf l}\,{\bf k}}&\equiv 
+          \left(
+          \begin{array}{cccc}
+            \bar{\Xi}^{++}_{{\bf l}{\bf k}} & -\bar{\Xi}^{+-}_{{\bf l}{\bf k}} & -\bar{\Xi}^{+-}_{{\bf l}{\bf k}} &  \bar{\Xi}^{--}_{{\bf l}{\bf k}} \\
+            \bar{\Xi}^{+-}_{{\bf l}{\bf k}} &  \bar{\Xi}^{++}_{{\bf l}{\bf k}} & -\bar{\Xi}^{--}_{{\bf l}{\bf k}} & -\bar{\Xi}^{+-}_{{\bf l}{\bf k}} \\
+            \bar{\Xi}^{+-}_{{\bf l}{\bf k}} & -\bar{\Xi}^{--}_{{\bf l}{\bf k}} &  \bar{\Xi}^{++}_{{\bf l}{\bf k}} & -\bar{\Xi}^{+-}_{{\bf l}{\bf k}} \\
+            \bar{\Xi}^{--}_{{\bf l}{\bf k}} &  \bar{\Xi}^{+-}_{{\bf l}{\bf k}} &  \bar{\Xi}^{+-}_{{\bf l}{\bf k}} & \bar{\Xi}^{++}_{{\bf l}{\bf k}}
+          \end{array}\right),
+        \end{align}
+        where
+        \begin{align}\label{eq:coeff_mcm_flat}
+          &\bar{\Xi}^{00}_{{\bf l}{\bf k}}=\left(\frac{2\pi}{L_xL_y}\right)^2 (w_a)_{{\bf l}-{\bf k}}(w_b)^*_{{\bf l}-{\bf k}},\\
+          &\bar{\Xi}^{0+}_{{\bf l}{\bf k}}=\left(\frac{2\pi}{L_xL_y}\right)^2 (w_a)_{{\bf l}-{\bf k}}(w_b)^*_{{\bf l}-{\bf k}}\cos2\Delta\varphi,\\
+          &\bar{\Xi}^{0-}_{{\bf l}{\bf k}}=\left(\frac{2\pi}{L_xL_y}\right)^2 (w_a)_{{\bf l}-{\bf k}}(w_b)^*_{{\bf l}-{\bf k}}\sin2\Delta\varphi,\\
+          &\bar{\Xi}^{++}_{{\bf l}{\bf k}}=\left(\frac{2\pi}{L_xL_y}\right)^2 (w_a)_{{\bf l}-{\bf k}}(w_b)^*_{{\bf l}-{\bf k}}\cos^22\Delta\varphi,\\
+          &\bar{\Xi}^{+-}_{{\bf l}{\bf k}}=\left(\frac{2\pi}{L_xL_y}\right)^2 (w_a)_{{\bf l}-{\bf k}}(w_b)^*_{{\bf l}-{\bf k}}\cos2\Delta\varphi\,\sin2\Delta\varphi,\\
+          &\bar{\Xi}^{--}_{{\bf l}{\bf k}}=\left(\frac{2\pi}{L_xL_y}\right)^2 (w_a)_{{\bf l}-{\bf k}}(w_b)^*_{{\bf l}-{\bf k}}\sin^22\Delta\varphi,
+        \end{align}
+        and $\Delta\varphi$ is the relative angle between ${\bf l}$ and ${\bf k}$.
+      \end{itemize}
 
@@
 
-  \subsection{Covariance matrices}
-    This section presents the analytical approximations used here to estimate the pseudo-$C_\ell$ Gaussian covariance matrix.
+    \subsection{Covariance matrices}\label{ssec:theory.pclcov}
+      So far we have not assumed anything about the statistics of the fields, other than the fact that they are isotropic (Eq.
+\ref{eq:iso}).
+This section presents a method to estimate the disconnected part of the power spectrum covariance for the pseudo-$C_\ell$ estimator.
 
-This section presents the analytical calculation of the pseudo-$C_\ell$ covariance.
-The treatment and approximations used are similar to those presented in \todo{cite} for spin-0 quantities, in \todo{cite} for particular spin-0 and spin-2 correlations and in \todo{cite} for the 3D density field.
-We reproduce the formalism here in order to provide a complete set of formulas applicable to any cross-correlation in both curved and flat skies.
+      Let $A$ and $F$ be the vector indices corresponding to the pairs of field indices $(\alpha,\beta)$ and $(\phi,\gamma)$ respectively, and let us start by considering the covariance
+      \begin{equation}
+        \Sigma^{AF}_{{\bf l}{\bf l}'}\equiv\left\langle\tilde{a}^\alpha_{\bf l}\tilde{b}^{\beta *}_{\bf l}\tilde{f}^\phi_{{\bf l}'}\tilde{g}^{\gamma *}_{{\bf l}'}\right\rangle-\left\langle\tilde{a}^\alpha_{\bf l}\tilde{b}^{\beta *}_{\bf l}\right\rangle\left\langle\tilde{f}^\phi_{{\bf l}'}\tilde{g}^{\gamma *}_{{\bf l}'}\right\rangle
+      \end{equation}
+      The covariance of the binned bandpowers $\tilde{\sf C}$ can then be computed as
+      \begin{equation}\label{eq:cov_binned}
+        {\rm Cov}\left({\rm vec}\left(\tilde{\sf C}^{ab}_q\right)_A,{\rm vec}\left(\tilde{\sf C}^{fg}_{q'}\right)_F\right)=\sum_{{\bf l}\in q} B^{\bf l}_q\sum_{{\bf l}'\in q'} B^{{\bf l}'}_{q'}\,\Sigma^{AF}_{{\bf l}{\bf l}'}.
+      \end{equation}
+      which can then be used to estimate the covariance of the mode-decoupled bandpowers multiplying it by the inverse mode-coupling matrix twice. I.e., schematically:
+      \begin{equation}
+        {\rm Cov}\left(\hat{\sf C}\right)=\mathcal{M}^{-1}\cdot{\rm Cov}\left(\tilde{C}\right)\cdot\left(\mathcal{M}^{-1}\right)^{T},
+      \end{equation}
+      where we have suppressed all indices for simplicity. The problem of estimating the pseudo-$C_\ell$ covariance therefore reduces to estimating $\Sigma^{AC}_{{\bf l}{\bf l}'}$.
 
-\begin{itemize}
-\item Naive Covariance
+      We now make use of Wick's theorem, which states that, for Gaussian fields $\langle a\,b\,f\,g\rangle=\langle a\,b\rangle\langle f\,g\rangle+\langle a\,f\rangle\langle b\,g\rangle+\langle a\,g\rangle\langle b\,f\rangle$.
+In this case, the expression for $\Sigma^{AF}_{{\bf l}{\bf l'}}$ reads:
+      \begin{align}\nonumber
+        \Sigma^{AF}_{{\bf l}{\bf l}'}=&\left[K\summ{k}\,^aM^{\alpha\alpha'}_{{\bf l}{\bf k}}\,^gM^{\gamma\gamma'*}_{{\bf l}'{\bf k}}C^{ag,(\alpha'\gamma')}_k\right]\left[K\summ{q}\,^bM^{\beta\beta'*}_{{\bf l}{\bf q}}\,^fM^{\phi\phi'}_{{\bf l}'{\bf q}}C^{bf,(\beta'\phi')}_q\right]^*\\\label{eq:covar_general}
+        &+\left((g,\gamma)\leftrightarrow(f,\phi)\right),
+      \end{align}
+      where we implicitly sum over repeated indices (e.g.
+$\alpha'$), and the second term is equivalent to the first one after swapping the roles of fields ${\bf f}$ and ${\bf g}$.
+Without any further approximations, for each pair $({\bf l},{\bf l}')$, we would need to perform two 2-dimensional integrals, and therefore the calculation would scale like $\ell_{\rm max}^6$, quickly becoming unfeasible.
+
+      Under the assumption that the coupling coefficients $M_{{\bf l}{\bf k}}$ are sharply peaked around ${\bf l}={\bf k}$, we can simplify the expression above approximating the power spectra as constants within the support of the coupling coefficients.
+Explicitly, we approximate
+      \begin{equation}\nonumber
+        C^{ag,(\alpha'\gamma')}_kC^{bf,(\beta',\phi')}_q\simeq C^{ag,(\alpha'\gamma')}_{(\ell}C^{bf,(\beta',\phi')}_{\ell')}\equiv\frac{1}{2}\left(C^{ag,(\alpha'\gamma')}_\ell C^{bf,(\beta',\phi')}_{\ell'}+C^{ag,(\alpha'\gamma')}_{\ell'} C^{bf,(\beta',\phi')}_\ell\right).
+      \end{equation}
+      In this case, the expression for $\Sigma^{AF}_{{\bf l}{\bf l}'}$ simplifies to
+      \begin{align}\label{eq:covar_efstathiou}
+        \Sigma^{AF}_{{\bf l}{\bf l}'}=K^2C^{ag,(\alpha'\gamma')}_{(\ell}C^{bf,(\beta',\phi')}_{\ell')}\,^{ag}W^{\alpha\gamma,\alpha'\gamma'}_{{\bf l}{\bf l}'}\left(\,^{bf}W^{\beta\phi,\beta'\phi'}_{{\bf l}{\bf l}'}\right)^*+\left((g,\gamma)\leftrightarrow(f,\phi)\right),
+      \end{align}
+      where we have defined the covariance coupling coefficients
+      \begin{equation}\label{eq:covar_coupling}
+        \,^{ab}W^{\alpha\beta,\alpha'\beta'}_{{\bf l}{\bf l}'}=\summ{k}\,^aM^{\alpha\alpha'}_{{\bf l}{\bf k}}\left(\,^bM^{\beta\beta'}_{{\bf l}'{\bf k}}\right)^*.
+      \end{equation}
+
+      In order to compute these coefficients, let us start by defining the quantities
       \begin{equation}
-    C^{abcd} = \frac{\pcl{l}ad \pcl{l}bc + \pcl{l}ac \pcl{l}bd}{f_{sky} (2l +
-      1) \Delta l} \delta_{ll'}
-    \label{eq:naive}
+        \,^{ab}I^{\pm s_a,\pm s_b}_{{\bf l}{\bf l'}}\equiv\summ{k}\summ{x}\summ{y}w_a({\bf x})w_b({\bf y})\,q^{\pm s_a}_{{\bf l}{\bf k}}({\bf x})\,\left(q^{\pm s_b}_{{\bf l}'{\bf k}}({\bf y})\right)^*,
       \end{equation}
-\item Approximation 1
-\item Approximation 2
-\end{itemize}
+      where $s_a$ is the spin of field ${\bf a}$, and
+      \begin{equation}
+        q^{\pm s}_{{\bf l}{\bf k}}({\bf x})\equiv \frac{\beta_{\ell,s}\beta_{k,s}}{2}\left(\left[\eth^s q({\bf l},{\bf x})\right]^* \eth^sq({\bf k},{\bf x})\pm\left[\bar{\eth}^sq({\bf l},{\bf x})\right]^*\bar{\eth}^sq({\bf k},{\bf x})\right).
+      \end{equation}
 
-\section{Results}
-For both flat and curved sky we should discuss:
-\begin{itemize}
-\item Simulations (sims, methods)
-\item Approximation 1 - spin 0
-\item Approximation 1 - spin 2
-\item Approximation 2
-\item Naive
-\end{itemize}
+      Now, in what follows, we will be concerned with the auto- and cross-correlations of spin-0 and spin-2 fields.
+Thus, to simplify the notation, we will enumerate the different types of coupling coefficients that exist for a spin-0 field with a single component that we will call $\delta$, in analogy to the projected galaxy overensity, and for a spin-2 field $\mathbf{\gamma}$ with $E$ and $B$ components $\gamma_E$ and $\gamma_B$, in analogy to the cosmic shear field.
+With this setup, all the possible non-zero $W^{\alpha\beta,\alpha'\beta'}_{{\bf l}{\bf l}'}$ can be expressed in terms of the $I^{\pm s_a,\pm s_b}_{{\bf l}{\bf l}'}$ as follows:
+      \begin{align}\label{eq:coeff}
+        &W^{\delta\delta,\delta\delta}_{{\bf l},{\bf l}'}=I^{0,0}_{{\bf l}{\bf l}'}; \\
+        &W^{\delta\gamma_E,\delta\gamma_E}_{{\bf l},{\bf l}'}=W^{\delta\gamma_B,\delta\gamma_B}_{{\bf l},{\bf l}'}=I^{0,+2}_{{\bf l},{\bf l}'};\hspace{12pt}
+        W^{\delta\gamma_E,\delta\gamma_B}_{{\bf l},{\bf l}'}=-W^{\delta\gamma_B,\delta\gamma_E}_{{\bf l},{\bf l}'}=-i\,I^{0,-2}_{{\bf l},{\bf l}'};\\
+        &W^{\gamma_E\gamma_E,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_E\gamma_B,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_E,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_B,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=I^{+2,+2}_{{\bf l}{\bf l}'}\\
+        &W^{\gamma_E\gamma_E,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_B,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_E,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_E\gamma_B,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=I^{-2,-2}_{{\bf l}{\bf l}'}\\
+        &W^{\gamma_E\gamma_E,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_E\gamma_B,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_E,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_B,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=-i\,I^{+2,-2}_{{\bf l}{\bf l}'}\\
+        &W^{\gamma_E\gamma_E,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_E\gamma_B,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_E,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_B,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=i\,I^{-2,+2}_{{\bf l}{\bf l}'}.
+      \end{align}
+      Thus, in principle we only need to compute 7 different types of terms ($I^{0,0}$, $I^{0,\pm2}$, $I^{\pm2,\pm2}$ and $I^{\pm2,\mp2})$).
+In order to simplify these expressions further we follow \cite{2017A&A...602A..41C} and neglect all gradients of the masks, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
+      \begin{align}\nonumber
+        \summ{x}w({\bf x})\left(\bar{\eth}^2q({\bf l},{\bf x})\right)^*\,\bar{\eth}^2q({\bf k},{\bf x})
+        &=\summ{x}w({\bf x})\eth^2 q^*({\bf l},{\bf x})\,\bar{\eth}^2q({\bf k},{\bf x})\\\nonumber
+        &=\summ{x}q^*({\bf l},{\bf x})\,\eth^2\left(\bar{\eth}^2q({\bf k},{\bf x})w({\bf x})\right)\\\nonumber
+        &\simeq\summ{x}q^*({\bf l},{\bf x})\,\left(\eth^2\bar{\eth}^2q({\bf k},{\bf x})\right)w({\bf x})\\
+        &=\frac{1}{\beta^2_{k,2}}\summ{x}q^*({\bf l},{\bf x})\,q({\bf k},{\bf x})w({\bf x})\\\nonumber
+        &=\summ{x}q^*({\bf l},{\bf x})\,\left(\bar{\eth}^2\eth^2q({\bf k},{\bf x})\right)w({\bf x})\\\nonumber
+        &=\summ{x}\bar{\eth}^2\left(w({\bf x})\,q^*({\bf l},{\bf x})\right)\,\eth^2q({\bf k},{\bf x})\\
+        &\simeq\summ{x}w({\bf x})\,\left(\eth^2\,q({\bf l},{\bf x})\right)^*\,\eth^2q({\bf k},{\bf x}),
+      \end{align}
+      where we have made repeated use of integration by parts.
 
-For both flat and curved plots:
-\begin{itemize}
-\item Cov Naive
-\item Cov approx 1
-\item Cov approx 2
-\item Cov approx 2 + deprojection
-\end{itemize}
+      Using these identities together with the completeness relation of the basis functions, it is possible to simplify the expressions for the $I^{\pm s_a,\pm s_b}$ \todo{cite}:
+      \begin{align}
+        &\,^{ab}I^{0,0}_{{\bf l}{\bf l}'}=\,^{ab}I^{0,+2}_{{\bf l}{\bf l}'}=\,^{ab}J^0_{{\bf l}{\bf l}'}; \hspace{12pt}
+        \,^{ab}I^{+2,+2}_{{\bf l}{\bf l}'}=\,^{ab}J^+_{{\bf l}{\bf l}'};\\
+        &\,^{ab}I^{+2,-2}_{{\bf l}{\bf l}'}=\,^{ab}I^{-2,+2}_{{\bf l}{\bf l}'}=\,^{ab}J^-_{{\bf l}{\bf l}'}; \hspace{12pt}
+        \,^{ab}I^{0,-2}_{{\bf l}{\bf l}'}=\,^{ab}I^{-2,-2}_{{\bf l}{\bf l}'}=0,
+      \end{align}
+      where  we have defined
+      \begin{align}\label{eq:covar_coupling_J}
+        &\,^{ab}J^0_{{\bf l}{\bf l}'}=\summ{x}\left(w_a w_b\right)({\bf x})\,q({\bf l},{\bf x})\,\left(q({\bf l}',{\bf x})\right)^*,\\
+        &\,^{ab}J^\pm_{{\bf l}{\bf l}'}=\summ{x}\left(w_a w_b\right)({\bf x})\,q^{\pm2}_{{\bf l}{\bf l}'}({\bf x}).
+      \end{align}
+      Thus, the only surviving non-zero coupling coefficients are
+      \begin{align}
+        &W^{\delta\delta,\delta\delta}_{{\bf l},{\bf l}'}=W^{\delta\gamma_E,\delta\gamma_E}_{{\bf l},{\bf l}'}=W^{\delta\gamma_B,\delta\gamma_B}_{{\bf l},{\bf l}'}=J^0_{{\bf l}{\bf l}'};\\
+        &W^{\gamma_E\gamma_E,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_E\gamma_B,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_E,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_B,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=J^+_{{\bf l}{\bf l}'}\\\nonumber
+        &W^{\gamma_E\gamma_E,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_E\gamma_B,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_E,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=-W^{\gamma_B\gamma_B,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=\\
+        &\hspace{20pt}=-W^{\gamma_E\gamma_E,\gamma_B\gamma_E}_{{\bf l},{\bf l}'}=-W^{\gamma_E\gamma_B,\gamma_B\gamma_B}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_E,\gamma_E\gamma_E}_{{\bf l},{\bf l}'}=W^{\gamma_B\gamma_B,\gamma_E\gamma_B}_{{\bf l},{\bf l}'}=-i\,J^-_{{\bf l}{\bf l}'}.
+      \end{align}
 
-Plots:
-\begin{itemize}
-\item Plot $\chi^2$
-\item KS($\chi^2$)
-\item Eigenvalues
-\end{itemize}
+      We have reduced the problem of computing the covariance in Eq.
+\ref{eq:covar_general} to the problem of computing the coupling coefficients \ref{eq:covar_coupling} entering Eq.
+\ref{eq:covar_efstathiou}, and we have now shown that there are only 3 independent coefficients, given by Eq.
+\ref{eq:covar_coupling_J}.
+In order to simplify the calculation further, it is now useful to inspect these results for the specific case of fields defined on the sphere.
 
-In this section we will show that our approximation is good and reliable; i.e.
-that can be used to obtain the cosmological parameters within the same error
-as using the covariance matrix from simulations.
+      \subsubsection{Covariances for curved skies}
+        As described in Section \ref{ssec:theory.pcl}, in the curved-sky estimator it is common to first average over the harmonic number $m$ as part of the bandpower binning operation.
+Let us therefore define
+        \begin{equation}
+          \tilde{C}^{a_\alpha b_\beta}_\ell=\frac{1}{2\ell+1}\sum_{m=-\ell}^\ell (\tilde{a}_\alpha)_{\ell m} (\tilde{b}_\beta)_{\ell m}^*.
+        \end{equation}
+        The covariance of these objects under the approximation in Eq. \ref{eq:covar_efstathiou}, can be computed as:
+        \begin{equation}\label{eq:cov_sph}
+          {\rm Cov}\left(\tilde{C}^{a_\alpha b_\beta}_\ell,\tilde{C}^{f_\phi g_\gamma}_{\ell'}\right) = C^{a_{\alpha'} f_{\phi'}}_{(\ell}C^{b_{\beta'} g_{\gamma'}}_{\ell')}\,^{af}_{bg}\mathcal{W}^{\alpha\phi,\alpha'\phi'}_{\beta\gamma,\beta'\gamma'}(\ell,\ell')+\left((g,\gamma)\leftrightarrow(f,\phi)\right),
+        \end{equation}
+        where we have defined the symbols
+        \begin{equation}
+          ^{af}_{bg}\mathcal{W}^{\alpha\phi,\alpha'\phi'}_{\beta\gamma,\beta'\gamma'}(\ell,\ell')=\equiv\sum_{m=-\ell}^\ell \sum_{m'=-\ell'}^{\ell'}\frac{\,^{af}W^{\alpha\phi,\alpha'\phi'}_{{\bf l}{\bf l}'}\left(\,^{bg}W^{\beta\gamma,\beta'\gamma'}_{{\bf l}{\bf l}'}\right)^*}{(2\ell+1)(2\ell'+1)}.
+        \end{equation}
+        These quantities involve terms of the form:
+        \begin{equation}
+          \,^{af}_{bg}\mathcal{J}^{XY}_{\ell\ell'}\equiv\sum_{m=-\ell}^\ell \sum_{m'=-\ell'}^{\ell'}\frac{\,^{af}J^X_{{\bf l}{\bf l}'}\left(\,^{bg}J^Y_{{\bf l}{\bf l}'}\right)^*}{(2\ell+1)(2\ell'+1)}
+        \end{equation}
+        Making use of the completeness relation for the Wigner 3-j symbols, it is possible to show, that these can be written as:
+        \begin{align}
+          &\,^{af}_{bg}\mathcal{J}^{00}_{\ell\ell'}=\Xi^{00}_{\ell\ell'}(w_aw_f,w_bw_g),\hspace{12pt}
+           \,^{af}_{bg}\mathcal{J}^{0+}_{\ell\ell'}=\Xi^{0+}_{\ell\ell'}(w_aw_f,w_bw_g),\\
+          &\,^{af}_{bg}\mathcal{J}^{++}_{\ell\ell'}=\Xi^{++}_{\ell\ell'}(w_aw_f,w_bw_g),\hspace{12pt}
+           \,^{af}_{bg}\mathcal{J}^{--}_{\ell\ell'}=\Xi^{--}_{\ell\ell'}(w_aw_f,w_bw_g),\\
+          &\,^{af}_{bg}\mathcal{J}^{+-}_{\ell\ell'}=
+           \,^{af}_{bg}\mathcal{J}^{-+}_{\ell\ell'}=
+           \,^{af}_{bg}\mathcal{J}^{0-}_{\ell\ell'}=0,
+        \end{align}
+        where the $\Xi^{XY}_{\ell\ell'}$ are defined in Eq. \ref{eq:coeff_mcm}.
+        Thus, computationally speaking, the problem of computing covariance matrices reduces to that of computing the same coupling coefficients needed for the computation of the pseudo-$C_\ell$ power spectra themselves, except now they involve product of two masks, rather than the masks alone.
+Using the same notation as in Eq.
+\ref{eq:coeff}, the only non-zero $\mathcal{W}$ coefficients are:
+        \begin{align}
+          &\mathcal{W}^{\delta  \delta  ,\delta  \delta  }_{\delta  \delta  ,\delta  \delta  }=
+           \mathcal{W}^{\delta  \delta  ,\delta  \delta  }_{\delta  \gamma_X,\delta  \gamma_X}=
+           \mathcal{W}^{\delta  \gamma_X,\delta  \gamma_X}_{\delta  \gamma_Y,\delta  \gamma_Y}=\mathcal{J}^{00}\\
+          &\mathcal{W}^{\delta  \delta  ,\delta  \delta  }_{\gamma_X\gamma_Y,\gamma_X\gamma_Y}=
+           \mathcal{W}^{\delta  \gamma_Z,\delta  \gamma_Z}_{\gamma_X\gamma_Y,\gamma_X\gamma_Y}=\mathcal{J}^{0+}\\
+          &\mathcal{W}^{\gamma_X\gamma_Y,\gamma_X\gamma_Y}_{\gamma_W\gamma_Z,\gamma_W\gamma_Z}=\mathcal{J}^{++}\\\nonumber
+          &\mathcal{W}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_Y\gamma_E,\gamma_Y\gamma_B}=
+           \mathcal{W}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=
+           \mathcal{W}^{\gamma_B\gamma_X,\gamma_E\gamma_X}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=\\
+          &\hspace{12pt}
+           \mathcal{W}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_Y\gamma_B,\gamma_Y\gamma_E}=
+           \mathcal{W}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=
+           \mathcal{W}^{\gamma_E\gamma_X,\gamma_B\gamma_X}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=-\mathcal{J}^{--}\\\nonumber
+          &\mathcal{W}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_Y\gamma_E,\gamma_Y\gamma_B}=
+           \mathcal{W}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=
+           \mathcal{W}^{\gamma_E\gamma_X,\gamma_B\gamma_X}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=\\
+          &\hspace{12pt}
+           \mathcal{W}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_Y\gamma_B,\gamma_Y\gamma_E}=
+           \mathcal{W}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=
+           \mathcal{W}^{\gamma_B\gamma_X,\gamma_E\gamma_X}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=\mathcal{J}^{--},
+        \end{align}
+        where $(X,Y,Z,W)$ stand for either $E$ or $B$, and where we have suppressed all redundant indices (including $\ell\ell'$).
+Any pseudo-$C_\ell$ covariance element can then be found by replacing these results in Eq.
+\ref{eq:cov_sph}.
+Some explicit examples for common terms can be found in Appendix A.3.3 of \cite{2017A&A...602A..41C}.
 
-\subsection{Simulations}
+      \subsubsection{Covariances for flat skies}
+        Similar results hold in the case of flat skies.
+As mentioned in Section \ref{ssec:theory.pcl}, in this case averaging over the Fourier-space azimuth happens while binning into bandpowers.
+Under the assumption that the underlying power spectra are roughly constant within each bandpower, in this case the covariance matrix of the bandpowers defined in Eq.
+\ref{eq:bandpowers} takes the form:
+        \begin{equation}\label{eq:cov_flat}
+          {\rm Cov}\left(\tilde{C}^{a_\alpha b_\beta}_q,\tilde{C}^{f_\phi g_\gamma}_{q'}\right) = C^{a_{\alpha'} f_{\phi'}}_{(q}C^{b_{\beta'} g_{\gamma'}}_{q')}\sum_{{\bf l}\in q} \frac{1}{N_q} \sum_{{\bf l}' \in q'} \frac{1}{N_{q'}}\,\,^{af}_{bg}\bar{\mathcal{W}}^{\alpha\phi,\alpha'\phi'}_{\beta\gamma,\beta'\gamma'}({\bf l},{\bf l}')+\left((g,\gamma)\leftrightarrow(f,\phi)\right),
+        \end{equation}
+        where the coefficients $\bar{\mathcal{W}}$ are related to the mode-coupling coefficients $\bar{\Xi}$ defined in \ref{eq:coeff_mcm_flat} in the same way that the curved-sky coefficients $\mathcal{W}$ were related to the $\Xi$:
+        \begin{align}
+          &\bar{\mathcal{W}}^{\delta  \delta  ,\delta  \delta  }_{\delta  \delta  ,\delta  \delta  }=
+           \bar{\mathcal{W}}^{\delta  \delta  ,\delta  \delta  }_{\delta  \gamma_X,\delta  \gamma_X}=
+           \bar{\mathcal{W}}^{\delta  \gamma_X,\delta  \gamma_X}_{\delta  \gamma_Y,\delta  \gamma_Y}=\bar{\mathcal{J}}^{00}\\
+          &\bar{\mathcal{W}}^{\delta  \delta  ,\delta  \delta  }_{\gamma_X\gamma_Y,\gamma_X\gamma_Y}=
+           \bar{\mathcal{W}}^{\delta  \gamma_Z,\delta  \gamma_Z}_{\gamma_X\gamma_Y,\gamma_X\gamma_Y}=\bar{\mathcal{J}}^{0+}\\
+          &\bar{\mathcal{W}}^{\gamma_X\gamma_Y,\gamma_X\gamma_Y}_{\gamma_W\gamma_Z,\gamma_W\gamma_Z}=\bar{\mathcal{J}}^{++}\\\nonumber
+          &\bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_Y\gamma_E,\gamma_Y\gamma_B}=
+           \bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=
+           \bar{\mathcal{W}}^{\gamma_B\gamma_X,\gamma_E\gamma_X}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=\\
+          &\hspace{12pt}
+           \bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_Y\gamma_B,\gamma_Y\gamma_E}=
+           \bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=
+           \bar{\mathcal{W}}^{\gamma_E\gamma_X,\gamma_B\gamma_X}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=-\bar{\mathcal{J}}^{--}\\\nonumber
+          &\bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_Y\gamma_E,\gamma_Y\gamma_B}=
+           \bar{\mathcal{W}}^{\gamma_X\gamma_B,\gamma_X\gamma_E}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=
+           \bar{\mathcal{W}}^{\gamma_E\gamma_X,\gamma_B\gamma_X}_{\gamma_B\gamma_Y,\gamma_E\gamma_Y}=\\
+          &\hspace{12pt}
+           \bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_Y\gamma_B,\gamma_Y\gamma_E}=
+           \bar{\mathcal{W}}^{\gamma_X\gamma_E,\gamma_X\gamma_B}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=
+           \bar{\mathcal{W}}^{\gamma_B\gamma_X,\gamma_E\gamma_X}_{\gamma_E\gamma_Y,\gamma_B\gamma_Y}=\bar{\mathcal{J}}^{--},
+        \end{align}
+        and, as before
+        \begin{align}
+          &\,^{af}_{bg}\bar{\mathcal{J}}^{00}_{{\bf l}{\bf l}'}=\bar{\Xi}^{00}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\hspace{12pt}
+           \,^{af}_{bg}\bar{\mathcal{J}}^{0+}_{{\bf l}{\bf l}'}=\bar{\Xi}^{0+}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\\
+          &\,^{af}_{bg}\bar{\mathcal{J}}^{++}_{{\bf l}{\bf l}'}=\bar{\Xi}^{++}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\hspace{12pt}
+           \,^{af}_{bg}\bar{\mathcal{J}}^{--}_{{\bf l}{\bf l}'}=\bar{\Xi}^{--}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\\
+          &\,^{af}_{bg}\bar{\mathcal{J}}^{+-}_{{\bf l}{\bf l}'}=
+           \,^{af}_{bg}\bar{\mathcal{J}}^{-+}_{{\bf l}{\bf l}'}=
+           \,^{af}_{bg}\bar{\mathcal{J}}^{0-}_{{\bf l}{\bf l}'}=0,
+        \end{align}
 
-Simulations were drawn from Gaussian fields that followed a fiducial power
-spectrum. The fiducial cosmology (\red{For the 2-bin case}) is given by
-$\Omega_{cdm} = 0.27$, $\Omega_b = 0.045$, $h = 0.67$, $A_s = 2.1 10^{-9}$ and
-$n_s = 0.9$, where each term correspond, respectively, to the amount of cold
-dark matter and baryons today, the reduced Hubble constant, the primordial
-curvature fluctuation amplitude and the spectral index. The redshift bins are
-shown in Fig.~\ref{fig:pz} and the theoretical $\cl$ are in
-Fig.~\ref{fig:cl-2bins}. Since all parts of sky will not be equally observed, a
-realist mask (see Fig.~\ref{fig:mask}) was applied.  The results shown here
-were obtained using the same mas for each redshift bin, as it is expected to
-be that way in real observations. We checked, however, that using different
-masks yielded same results.
+    \subsection{Approximate covariances}\label{ssec:theory.approx}
+      When presenting our results in Section \ref{sec:results}, we will compare the true covariance matrix, estimated from a large number of Gaussian simulations, with the analytical covariance estimated under different approximations.
+In descending order of complexity, these are:
+      \begin{enumerate}
+        \item The narrow-kernel approximation (labelled {\bf NKA} here), described in the previous sections.
+This approximation assumes that the support of the harmonic-space masks (represented by the mode-coupling coefficients in e.g.
+Eq.
+\ref{eq:mcms}) is small compared to the variation of the true power spectrum with $\ell$.
+Additionally, it neglects all derivatives of the sky mask when accounting for the spin nature of the fields involved.
+        \item The {\bf spin-0} approximation corresponds to a simplified version of the NKA in which the spin nature of all fields involved is completely ignored, and all fields, including the $E$- and $B$-mode components of a spin-2 field, are treated as spin-0 quantities.
+        \item The mode-counting approximation (labelled {\bf MC} here), commonly known as the {\sl Knox formula} \cite{1995PhRvD..52.4307K}, which applies the result found for full-sky observations (Eq.
+\ref{eq:cov_naive}) to masked fields corrected by an overall factor that accounts to the loss of modes due to the sky mask.
+In this case, the covariance is simply given by:
+        \begin{equation}\label{eq:cov_naive_bpw}
+          {\rm Cov}\left(C^{ab}_q,C^{cd}_{q'}\right)=\delta^K_{qq'}\frac{C^{ad}_qC^{bc}_q+C^{ac}_qC^{bd}_q}{(2\ell_q+1)\,f_{\rm sky}\,N_q},
+        \end{equation}
+        where $\ell_q$ is the mean multipole in the $q$-th bandpower, $N_q$ is the number of multipoles assigned to it, and $f_{\rm sky}$ is the available sky fraction.
+      \end{enumerate}
 
+  \section{Results}\label{sec:results}
+    \subsection{Simulations}\label{ssec:results.sims}
       \begin{figure}
         \centering
-  \includegraphics[width=\columnwidth]{./figures/pz.png}
-  \caption{Galaxy distribution redshift bins.}
+        \includegraphics[width=0.6\textwidth]{./figures/pz.pdf}
+        \caption{}
         \label{fig:pz}
@@
         \centering
-  \includegraphics[width=\textwidth]{./figures/cls-sph-2b.png}
+        \includegraphics[width=\textwidth]{./figures/cls-sph-2b.pdf}
         \caption{Fiducial power spectra. We show only the $T$ and $E$ modes.}
@@
         \centering
-  \includegraphics[width=\columnwidth]{./figures/mask-lss1.png}
-  \caption{Example of sky mask used in our analysis}
-  \label{fig:mask}
+        \includegraphics[width=0.6\textwidth]{./figures/mask-lss1.pdf}
+        \includegraphics[width=0.6\textwidth]{./figures/mask-lss2.pdf}
+        \caption{Sky masks used in our analysis for the two different redshift bins.
+The same mask is assumed for the galaxy overdensity $\delta$ and cosmic shear $\mathbf{\gamma}$ for simplicity.} \label{fig:mask}
       \end{figure}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.7\textwidth]{./figures/contaminants_cl.pdf}
+        \caption{Input (signal $+$ noise) power spectrum for the galaxy overdensity in the first redshift bin (blue line).
+The orange band shows the range of power spectra used for the 100 large-scale contaminant templates used to study the impact of contaminant deprojection on the covariance matrix.
+The green line shows the power spectrum used to generate small-scale contaminants.
+The amplitude of the contaminat power spectra was fixed by imposing a $10\%$ contamination level in the power spectrum at $\ell=400$.
+The same approach was used for the cosmic shear maps, with independent contaminants defined in $\gamma_E$ and $\gamma_B$.} \label{fig:clcont}
+      \end{figure}
+      In order to quantitatively study the performance of the analytical approximations for the power spectrum covariance matrix introduced in Section \ref{sec:theory}, we generate a large number of Gaussian simulations including a number of realistic observational effects.
 
-The covariance matrices were obtained from 20000 simulations, for the band
-powers determined by the fraction of sky observed and the maximum band power
-we were able to recover the fiducial power spectra for; i.e. $\sim 2 \times
-512$. This corresponds to $l_{\mbox{bpw}} = 1023$. To work with the sky
-maps and masks we used
-\textit{HEALPix}\footnote{\url{https://healpix.jpl.nasa.gov/}}, whereas 
-\textit{NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}}~\cite{2018arXiv180909603A}
-was used for the computation of the pseudo-$\cl$ and the covariances. Our code
-is publicly
-available~\footnote{\url{https://github.com/damonge/PCLCovariance}}.
+      Each simulation is a set of maps corresponding to a number of spin-0 and spin-2 fields that are drawn as Gaussian random fields following a set of input power spectra that include all relevant cross-correlations between different fields.
+We generate simulations for two types of fields modelled after the two main large-scale structure observables of photometric redshift surveys:
+      \begin{itemize}
+        \item Spin-0 fields, corresponding to maps of the overdensity of galaxies within a given redshift bin projected on the sphere.
+In keeping with the notation introduced in Section \ref{ssec:theory.pclcov}, we will label these fields as $\delta^a$, where the index $a$ denotes the redshift bin.
+        \item Spin-2 fields, corresponding to maps of the cosmic shear measured from the projected shapes of galaxies in a given redshift bin.
+We will label these fields $\mathbf{\gamma}^a=(\gamma_E^a,\gamma_B^a)$ where, again, the index $a$ denotes the redshift bin.
+      \end{itemize}
+      The cross-correlation between two of these fields can be written as:
+      \begin{equation}
+        C^{ab}_\ell = S^{ab}_\ell+N^{ab}_\ell,
+      \end{equation}
+      where $S$ and $N$ are the power spectra of the cosmological signal and noise. We model the signal part for $\delta$ and $\gamma_E$ as:
+      \begin{equation}
+        S^{ab}_\ell = \int d\chi \frac{W^a_\ell(\chi)\,W^b_\ell(\chi)}{\chi^2}\,P\left(k=\frac{\ell+1/2}{\chi},z\right),
+      \end{equation}
+      where $\chi$ is the comoving radial distance, $z\equiv z(\chi)$ is the corresponding redshift in the lightcone, $P(k,z)$ is the matter power spectrum and the window functions are given by \cite{2004PhRvD..70d3009H}\footnote{See \cite{2019ApJS..242....2C} and references therein for details about these theory predictions}:
+      \begin{align}
+        \nonumber
+        &W^\delta_\ell(\chi)=b(z)\,H(z)\,p_z(z),\\
+        &W^\gamma_\ell(\chi)=f_\ell\,\frac{3H_0^2\Omega_M}{2}(1+z)\chi\,\int dz'\,p_z(z')\,\frac{\chi(z')-\chi}{\chi(z')},
+      \end{align}
+      where $H(z)$ is the expansion rate in units where the speed of light is $c=1$, $H_0\equiv H(z=0)$, $b(z)$ is the linear galaxy bias, $p_z$ is the normalized redshift distribution of galaxies within the redshift bin, and
+      \begin{equation}
+        f_\ell\equiv\frac{\sqrt{(\ell+2)(\ell+1)\ell(\ell-1)}}{(\ell+1/2)^2}.
+      \end{equation}
+      We assume zero signal for the shear $B$-modes.
 
-The same results are obtained even in the presence of foregrounds. In
-particular, we tested the case with 100 foregrounds both at low scales and
-large scales, adding a mock power spectra (e.g. Fig.~\ref{fig:fore}) computed as 
+      The noise power spectrum is diagonal (i.e. zero between different fields and redshift bins), and is given by:
       \begin{equation}
-  \clf = A (l + 1)^{\beta}\,.
-  \label{eq:fore}
+        N_\ell^{\delta\delta} = \frac{1}{n_\Omega},\hspace{12pt}
+        N_\ell^{\gamma_E\gamma_E} = N_\ell^{\gamma_B\gamma_B} = \frac{\sigma_\gamma^2}{n_\Omega},
       \end{equation}
-Here, $A$ was fixed so that the ratio of the foreground power spectra and the
-theoretical one ($\clth$) was $\clf/\clth = 0.1$ at $l=400$ and $\beta$ was
-randomly chosen from $U[-3, -1)$ for large scale effects and fixed to $\beta =
-0$ for small scale ones.
+      where $n_\Omega$ is the mean number density of galaxies in units of sterad$^{-2}$ (see below), and $\sigma_\gamma=0.28$ is the intrinsic shape scatter per ellipticity component.
+
+      We consider the case of auto- and cross-correlations between all fields in two redshift bins, with redshift distributions modelled as Gaussians with width $\sigma_z=0.13$ centered around redshifts $0.75$ and $0.95$, with number densities $n_\Omega=7.5\,{\rm arcmin}^{-2}$.
+The corresponding redshift distributions $p_z(z)$ are shown in Figure \ref{fig:pz}.
+For simplicity we use a constant bias $b(z)=1$.
+For these specifications, we generate signal power spectra using the Core Cosmology Library ({\tt CCL} \cite{2019ApJS..242....2C}) for cosmological parameters $(\Omega_c,\Omega_b,h,A_s,n_s)=(0.27,0.045,0.67,2.1\times10^{-9},0.96)$.
+The resulting non-zero signal and noise power spectra are shown in Fig.
+\ref{fig:cl-2bins}.
+
+      After generating a set of Gaussian maps, we mask them making use of a realistic sky mask.
+This mask has three main components: a cut in declination based on the expected sky coverage of LSST \cite{2014SPIE.9150E..15D}, a more conservative Galactic cut using the dust reddening data from \cite{1998ApJ...500..525S} and a set of 100 randomly positioned holes with a radius of 1 degree.
+To explore the case of cross-correlations between fields with different masks, we generated two masks for the two redshift bins above, consisting of two different sets of random holes.
+These masks are shown in Figure \ref{fig:mask}.
+
+      We have also explored the impact of the presence of sky contaminants on the estimate of the covariance matrix.
+As described in \cite{2019MNRAS.484.4127A}, the presence of a small contamination from observational systematics in the data can be accounted for through a technique known as {\sl mode deprojection} \todo{Cites}.
+In this method, the contamination is modeled as a linear contribution at the map level from contaminants with a known template.
+Mode deprojection then consists on projecting the data onto the subspace of modes that are perpendicular to those templates, effectively removing all modes from the map that ``look like'' any of the contaminants.
+This removal of modes has been shown to provide unbiased estimates of the power spectrum, however it could potentially affect the power spectrum uncertainties due to the loss of statistical power.
+To study this effect, we have also generated contaminant maps of two types:
+      \begin{itemize}
+        \item {\bf Large-scale contaminants:} Gaussian random maps with a red spectrum of the form $C_\ell\propto (\ell+1)^\beta$, where $\beta$ is a random number chosen within the range $\beta\in(-1,-3)$.
+        \item {\bf Small-scale contaminants:} Gaussian random maps with a flat spectrum $C_\ell={\rm const.}$
+      \end{itemize}
+      In both cases we fixed the amplitude of the contaminant power spectrum
+      such that it would yield a $10\%$ contamination in the data power
+      spectrum at $\ell=400$. In both cases, for spin-2 fields, we assumed the
+      same power spectrum for $E$ and $B$ modes, with no cross-correlation
+      between them. The resulting contaminant power spectra are shown in
+      Figure \ref{fig:clcont}.
+When exploring the effects of mode deprojection, we generate 100 contaminant maps of both types and add them to each simulated realization.
+We then deproject the full set of 100 contaminant templates from the simulated maps and compute the corresponding unbiased power spectra.
+Note that our fiducial results do not include the effects of mode deprojection.
+These are discussed in detail separately.
+
+      For our fiducial results, we generated a set of 20,000 random simulations.
+This number was chosen in order to recover the covariance matrix for all possible auto- and cross-correlations with sufficient accuracy.
+All simulations were generated as {\tt HEALPix} maps with resolution $N_{\rm side}=512$.
+We then used {\tt NaMaster} to compute all possible power spectra for each simulation using narrow bandpowers of width $\Delta\ell=3$ from $\ell=2$ to $\ell=1023$.
+Finally, we used the power spectra from the simulations to estimate the sample covariance matrix
+      \begin{equation}
+        {\rm Cov}({\bf C}) =\frac{1}{N_{\rm sim}-1}\sum_{i=1}^{N_{\rm sim}}\left({\bf C}_i-\bar{\bf C}\right)\cdot\left({\bf C}_i-\bar{\bf C}\right)^T,
+      \end{equation}
+      where ${\bf C}_i$ is the vector of all possible power spectra for the $i$-th realization, and $\bar{\bf C}$ is the mean of this vector over all realizations.
+The comparison of this sample covariance with the analytical approximations described above is presented in the next sections.
+
+    \subsection{Qualitative comparison}\label{ssec:results.visual}
+      \begin{figure}
+        \centering
+        \includegraphics[width=\textwidth]{./figures/all_rows_sph_1bin.pdf}
+        \caption{Mode coupling of different $C_\ell$'s. The covariances have
+          been normalized by the MC approximation numerator (Eq.~\ref{eq:cov_naive_bpw}).}
+        \label{fig:rows_1bin}
+      \end{figure}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.7\textwidth]{./figures/sph_1bin_diff_corr.pdf}
+        \caption{Difference between the correlation matrices associated to the sample covariance matrix and its estimate with the NKA method including all possible correlations between $\delta$ and $\mathbf{\gamma}$ in a single redshift bin.} \label{fig:diff_corr_1bin}
+      \end{figure}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.9\textwidth]{./figures/run_sph_2b_NKA_diff_corr_short.pdf}
+        \caption{Same as Fig. \ref{fig:diff_corr_1bin} for all cross-correlations between $\delta$ and $\gamma_E$ measured in two different redshift bins.}
+        \label{fig:corr_diff_2bins}
+      \end{figure}
+      As a first step, we visually compare the main properties of the sample covariance matrix estimated from the simulations and the NKA, Spin-0 and MC approximations described in the previous section.
+Figure \ref{fig:rows_1bin} shows four rows of the covariance matrix of different auto- and cross-correlations.
+The rows correspond to bandpowers centered on multipoles $\ell_q\simeq60,\,90,\,120$ and $150$.
+The upper panels show results for the non-zero power spectra ($\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$), with the solid blue, dashed orange and dotted green lines showing results for the sample covariance matrix, the analytical covariance using the NKA approximation and their difference respectively.
+For comparison, the black stars show the diagonal covariance matrix elements predicted by the MC approximation (Eq.
+\ref{eq:cov_naive_bpw}).
+We find an excellent agreement between the simulated and analytical covariances, with very small deviations in the amplitude of the diagonal and first few off-diagonal elements.
+
+      The bottom panel in Fig.
+\ref{fig:rows_1bin} shows the same rows of the covariance matrix for power spectra involving $B$-modes (and therefore with zero signal expectation value).
+In this case we find significant differences, at the level of $30-50\%$, on the covariance matrix elements, with the analyticial prediction underestimating the error bars overall.
+This is expected and can be understood as follows: the presence of a sky mask mixes $E$ and $B$ modes.
+Although this mixing can be accounted for at the level of the power spectrum through the pseudo-$C_\ell$ estimator, the leaked modes contribute to the variance.
+This is particularly significant for power spectra involving $B$-modes, since the $E$-mode amplitude is significantly larger, especially at $\ell\lesssim200$, as can be seen in Fig.
+\ref{fig:cl-2bins}.
+Thus, if the effects of $E$-$B$ mixing caused by the sky mask are not accurately accounted for in the estimation of the covariance matrix for power spectra involving $B$-mode maps, we can expect a misestimation of the contribution to the covariance from the leaked $E$ modes that would underpredict the uncertainties.
+This is not a problem for power spectra involving only $E$ modes, since the only $B$ modes that leak into them are those associated with noise, and they have the same amplitude as the noise $E$ modes move into the $B$-mode map.
+
+      This is further illustrated by Fig.
+\ref{fig:diff_corr_1bin}.
+The figure shows, for the case of a single redshift bin, the difference between the correlation matrices associated with the sample covariance matrix and the NKA estimate\footnote{The correlation matrix is defined as $r_{ij} = {\rm Cov}_{ij}/\sqrt{{\rm Cov}_{ii}{\rm Cov}_{jj}}$}.
+While the differences between both matrices are small for all elements involving $\delta$ and $\gamma_E$, all terms involving $B$-modes show a significant disagreement, particularly the  $\gamma_E\gamma_E$-$\gamma_B\gamma_B$, $\gamma_E\gamma_B$-$\gamma_E\gamma_B$ and $\gamma_B\gamma_B$-$\gamma_B\gamma_B$ boxes.
+
+      We thus conclude that while the NKA estimator is able to recover the covariance matrix for the non-zero power spectrum elements (i.e.
+those involving $\delta$ and $\gamma_E$) with high accuracy, a more sophisticated approach would be needed in order to obtain a precise estimate of the uncertainties for components involving $B$-modes.
+This is not a major concern, since $B$-mode power spectra are predominantly used as null tests, while cosmological parameter constraints are driven by the analysis of $\delta$ and $\gamma_E$.
+For completeness, Fig.
+\ref{fig:corr_diff_2bins} shows the difference between the correlation matrices for the sample covariance and the NKA estimator for all non-zero cross-correlations between different bins in the case of two redshift bins with different small-scale masks, where we find a similarly good agreement.
+
+    \subsection{Quantitative comparison}\label{ssec:results.quant}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_NKA_TTTEEE_reldev_eigval_1stbin.pdf}
+        \caption{Covariance matrix eigenvalues and the relative difference with respect to the eigenvalues from the simulations and NKA covariances.} \label{fig:eigv_1bin}
+      \end{figure}
+      \begin{figure}
+        \centering
+        \includegraphics[width=\textwidth]{./figures/run_sph_2b_1stbin_chi2_TT_TE_EE.pdf}
+        \caption{$\chi^2$ distributions for the cases without 
+          $\gamma_B$-modes. We compare the distributions obtained with the
+          covariance matrix computed directly from the simulated power spectra
+          and the two different analytical methods: NKA and Spin-0. In addition,
+          the theoretical $\chi^2$ distribution has been included. We observe
+          the simulations' distribution perfectly follows the former and
+          negligible differences between the approximations. It must be noted
+          their shift with respect to the true distribution when spin-2 fields
+          are considered. It will be shown in Section~\ref{ssec:results.cosmo}
+          it will have no effect on the cosmological predictions.}
+        \label{fig:chi2_1bin} \end{figure}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_same_mask_Spin0_NKA_TTTEEE_Full_chi2.pdf}
+        \caption{Same as Figure \ref{fig:chi2_1bin} for the combination of all correlations between $\delta$ and $\gamma_E$ for two redshift bins.}
+        \label{fig:chi2_2bins}
+      \end{figure}
+      \begin{figure}
+        \centering
+        \includegraphics[width=\textwidth]{./figures/contaminants_diag_cov.pdf}
+        \caption{Diagonal of the covariance matrix estimated from simulations
+          without contaminants, with 200 contaminants deprojected and the NKA
+          method. The differences between the simulated cases are much
+          smaller than those associated with the NKA estimator (shown in
+          green). The NKA method is therefore sufficient to approximate the
+          covariance even in the presence of contaminant deprojection.}
+        \label{fig:diag_conts}
+      \end{figure}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.6\textwidth]{./figures/contaminants_chi2.pdf}
+        \caption{$\chi^2$ distributions for simulations with no systematic contaminants (blue), and for simulations containing contaminants after deprojection of 200 contaminant templates (orange).
+The covariance matrices associated to both sets of simulations are essentially the same, and the differences between them are much smaller than those associated with the NKA estimator (shown in green).
+The NKA method is therefore sufficient to approximate the covariance even in the presence of contaminant deprojection.}
+        \label{fig:chi2_conts}
+      \end{figure}
+      In order to quantify the validity of our analytical approximations, we need to compare the NKA covariance with the sample covariance estimated from the simulations.
+However, comparing two matrices is not as straightforward as comparing their matrix elements one by one.
+The covariance between far-away bandpowers is expected to be very close to zero, and therefore a direct comparison of those elements would easily yield large relative differences simply due to the statistical noise in the sample covariance matrix.
+We will therefore quantify the differences between the different covariances making use of scalar quantities formed from them.
+The impact of the analytical approximations on the final parameter constraints will then be described in detail in Section \ref{ssec:results.cosmo}.
+
+      As a first test to quantify the differences between covariance matrices, we compute the relative difference between their eigenvalues.
+This is shown in Fig.
+\ref{fig:eigv_1bin} for a data vector combining all auto- and cross-correlations between $\delta$ and $\gamma_E$ for a single redshift bin.
+The eigenvalues of both matrices are roughly similar, with relative differences of about $5\%$.
+The figure also shows the eigenvalues of the covariance matrix estimated using the Spin-0 approximation, which achieves a similar level of precision (even marginally higher in some cases).
+
+      Another scalar quantity that can be used to compare different covariances is the $\chi^2$.
+For a given random data vector ${\bf d}$ with mean ${\bf m}$ and covariance matrix ${\rm Cov}$, this is given by
+      \begin{equation}
+        \chi^2=\left({\bf d}-{\bf m}\right)^T\cdot{\rm Cov}^{-1}\cdot\left({\bf d}-{\bf m}\right).
+      \end{equation}
+      We compute this quantity for a data vector ${\bf d}$ composed of
+      different auto- and cross-correlations for each of the 20,000 Gaussian
+      simulations, with ${\bf m}$ given by the mean over all simulations and
+      different choices of covariance matrix. Figure \ref{fig:chi2_1bin} shows
+      the distribution of $\chi^2$ values for the three non-zero power spectra
+      in the case of a single bin: $\delta$-$\delta$, $\delta$-$\gamma_E$ and
+      $\gamma_E$-$\gamma_E$. The histograms show the distribution for the
+      sample covariance matrix (blue), and the analytical NKA and Spin-0
+      estimators (orange and green respectively). We additionally plot the
+      theoretical $\chi^2$ distribution under the assumption that the
+      underlying data vector is Gaussianly distributed (red dashed lines). In
+      the simplest case of purely spin-0 quantities (leftmost panel), we find
+      an excellent agreement between the different distributions. In the cases
+      involving the spin-2 fields, we see noticeable differences between the
+      distributions found with the sample covariance and the approximate ones.
+      These differences are small, corresponding to less of a $2$ and $4\%$
+      shift in the mean $\chi^2$ for the $\delta$ - $\gamma_E$ and
+      $\gamma_E$ - $\gamma_E$ cases, respectively, and a negligible variation in
+      the width of the distributions. We therefore expect these differents to
+      have a negligible effect on the posterior parameter distributions, as we
+      will show in Section \ref{ssec:results.cosmo}. The fact that these
+      differences appear only for power spectra involving spin-2 quantities
+      indicate that the NKA and spin-0 methods are imperfect at describing the
+      additional mode coupling caused in the presence of a mask for
+      higher-spin fields. This is rather obvious in the case of the spin-0
+      approximation but, interestingly, we find that the NKA and Spin-0
+      predictions yield results that are almost indistinguishable from each
+      other. We therefore conclude that the additional approximation made in
+      the NKA method for spin-2 fields -- neglecting the spatial derivatives
+      of the mask -- is effectively equivalent to ignoring the spin nature of
+      the fields involved. For completeness, Figure \ref{fig:chi2_2bins} shows
+      the distribution of $\chi^2$ values for a data vector composed of all
+      possible auto- and cross-correlations of $\delta$ and $\gamma_E$ for the
+      case of two redshift bins, where similar conclusions hold.
+
+      We have also explored the impact of contaminant deprojection on the
+      different covariance matrix estimates. The loss of modes due to
+      deprojection can potentially increase the variance of the power spectrum
+      estimates, affecting the accuracy with which an analytical estimator
+      would be able to recover the covariance. Figure \ref{fig:diag_conts}
+      shows the diagonal of the covariance matrix for the $\delta$-$\delta$,
+      $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ power spectra. Each panel
+      displays the diagonal for the sample covariance matrix estimated from
+      simulations without contaminants or contaminant deprojection (blue
+      line), the sample covariance from simulations with contaminants and
+      contaminant deprojection (dashed orange line) and for the NKA covariance
+      (green line). We see that the power spectrum uncertainties
+      are almost indistinguishable with or without deprojection, and that
+      those relative differences are much smaller than the differences between
+      the sample covariance and the NKA estimator. We therefore conclude that,
+      except in the case where a very large set of contaminant maps are
+      deprojected (comparable with the number of unmasked pixels in the map),
+      the analytical approximation to the covariance matrix should be as
+      accurate as in the absence of contaminants (i.e. accurate enough).
+      This is quantitatively shown in Fig.~\ref{fig:chi2_conts} through
+      comparison of the $\chi^2$ distributions drawn with covariance matrices
+      estimated with simulations without contaminants, simulations with
+      contaminants deprojected and the NKA method.
+
+    \subsection{Impact on parameter estimation}\label{ssec:results.cosmo}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.6\textwidth]{./figures/contours_2bin.pdf}
+        \caption{1$\sigma$ and 2$\sigma$ contours for $\Omega_M$ and $\sigma_8$ found from two different estimates of the power spectrum likelihood for a dataset containing all possible cross-power spectra between $\delta$ and $\gamma_E$ for two redshift bins.
+The solid black and dashed red contours correspond to Gaussian likelihoods estimated using the sample covariance matrix and the covariance estimated using the NKA method respectively.
+We find a remarkably good agreement between both likelihoods, highlighting the negligible impact of the approximations involved in the NKA method on the final parameter estimates.
+\todo{finish this plot.}} \label{fig:contours}
+      \end{figure}
+      Ultimately, the most important test to judge the accuracy of the analytical covariance matrix estimators implemented here is to study their impact on the posterior distribution of cosmological parameters derived from power spectrum measurements.
+Assuming flat priors, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
+      \begin{equation}\label{eq:parpos}
+        -2\log p(\vec{\theta})=\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)^T\cdot{\rm Cov}^{-1}\cdot\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)+{\rm const.},
+      \end{equation}
+      where $\hat{\bf C}$ is a vector of power spectrum measurements, ${\rm Cov}$ is their covariance matrix, and ${\bf C}(\vec{\theta})$ is their theoretical prediction for parameters $\vec{\theta}$.
+
+      We explore $\log p(\vec{\theta})$ for of two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins.
+In this simple two-dimensional scenario we simply sample the distribution in a regular grid of 200 by 200 points for each parameter.
+We construct a data vector $\hat{\bf C}$ from the theoretical prediction for the experimental setup described in Section \ref{ssec:results.sims}, and produce theoretical predictions for it at each grid point using {\tt CCL}.
+Note that, when evaluating the posterior for the sample covariance matrix, one needs to correct for the finite number of simulations used to construct the covariance.
+In most situations this can be done simply by rescaling the inverse covariance matrix by a factor  given by \cite{2007A&A...464..399H}
+      \begin{equation}
+        {\rm Cov}^{-1}\,\longrightarrow \frac{N_s-2-N_{\rm data}}{N_s-1}\,{\rm Cov}^{-1},
+      \end{equation}
+      where $N_s=20,000$ is the number of samples used to estimate the covariance, and $N_{\rm data}=3510$ is the number of data points.
+
+      The 68\% and 95\% confidence level contours associated with the posterior distributions for the sample covariance and the NKA estimator are shown in Fig.
+\ref{fig:contours}.
+We find that both distributions agree with each other remarkably well, and that the $1\sigma$ errors for each parameter agree for both covariances up to \todo{$x\%$}.
+Note that, since we have not included any statistical noise in the data vector $\hat{\bf C}$, the relative bias between both distributions is zero by construction (since Eq.
+\ref{eq:parpos} is bounded from below by zero).
+When adding Gaussian statistical noise compatible with the sample covariance matrix, we observe small differences (smaller than 0.3$\sigma$) in the best fit parameters found with both covariance matrices.
+These differences, however, are not systematic, and we have verified that averaging over several noise realizations does not yield biased best-fit parameters.
+
+      We therefore conclude that the analytical approximations for the power spectrum covariance matrix explored here are able to reproduce the true posterior distribution of cosmological parameters to very high accuracy.
+
+  \section{Discussion}\label{sec:discussion}
+    Estimating accurate covariance matrices for projected two-point correlators is an ubiquitous problem in modern cosmology \todo{cites}, but it is particularly relevant for large-scale structure datasets.
+This problem is further complicated in this case, in comparison with e.g.
+CMB experiments, by two factors: the fact that the fields involved are non-Gaussian at some level and the higher complexity of the sky masks used in optical datasets.
+The main impact of survey geometry is the statistical coupling it induces between different Fourier/harmonic modes, which must be accurately characterized in order to obtain reliable estimates of the posterior distribution for cosmological parameters.
+In this paper we have focused on the impact of survey geometry on the Gaussian (i.e.
+disconnected) part of the covariance matrix.
+
+    We have described and generalized existing analytical approaches to estimate the covariance matrix for pseudo-$C_\ell$ power spectrum estimators \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}, and implemented them in the public code {\tt NaMaster} \cite{2019MNRAS.484.4127A}, making it straightforward to fully account for the effects of survey geometry on the data uncertainties.
+With these approximation, computationally speaking, the problem of estimating a covariance matrix is as complex as that of computing the power spectrum itself, and scales with the number of pixels in the map as $N_{\rm pix}^{3/2}$.
+
+    The main finding of this paper is the excellent performance of the analytical methods described in Section \ref{ssec:theory.pclcov}.
+We have shown that, in spite of the higher complexity of large-scale structure masks, the NKA estimator (see Section \ref{ssec:theory.approx}) is able to recover the covariance matrix for all power spectra with non-zero signal expectation value (i.e.
+those involving the galaxy overdensity $\delta$ or the $E$-mode shear $\gamma_E$), as well as the posterior distribution for cosmological parameters, to a high degree of accuracy.
+
+    More in detail, we have also found that the impact of contaminant deprojection on the covariance matrix, through the corresponding loss of modes, is negligible unless a very large number of contaminant templates are removed.
+This simplifies the procedure to estimate covariance matrices for galaxy clustering data, which are particularly sensitive to a large number of astrophysical and observational systematics.
+Additionally, we have found that, although the NKA estimator is accurate enough, it is not able to perfectly capture the additional effects of mode coupling that are present for spin-2 fields, and that a simpler approach treating the shear $E$ modes as a spin-0 object is able to reach similar levels of accuracy.
+Due to the imperfect treatment of the $E/B$ mixing caused by the sky mask in the NKA estimator, we also find that the predicted covariance matrix for any power spectra involving $B$ modes differs significantly from the true sample covariance, and therefore this approach cannot be used to reliably estimate the uncertainties of $B$-mode power spectra.
+
+    In spite of these inaccuracies, we find that the approximations described in this paper are able to provide estimates of the power spectrum covariance matrix that are sufficiently accurate for current and future tomographic large-scale structure cosmological datasets.
+The main advantage of this approach is the computational cost, which is comparable to that of estimating the power spectrum in the first place, and which scales with pixel resolution in a similar way.
+This is therefore significantly less time-consuming than generating large numbers of mock datasets (even simple Gaussian or log-normal realizations), and more reliable than the traditional jackknife resampling techniques.
+The method, in all its generality, is currently implemented in the public code {\tt NaMaster}.
+Future extensions to this work will focus on improving the estimator for power spectra involving $B$ modes, and incorporating the impact of $E/B$-mode purification \todo{cites}.
+
+  \acknowledgments
+  We thank Thibaut Louis, Eva-Maria Mueller, Francisco Javier S\'anchez and An\v ze Slosar for useful comments and discussion.
+CGG is supported the Spanish grant BES-2016-077038, partially funded by the ESF and by AYA2015-67854-P from the Ministry of Industry, Science and Innovation of Spain and the FEDER funds.
+He was partially supported by a Balzan Fellowship while in Oxford.
+He would like to thank New College and the Department of Physics at Oxford for their hospitality.
+DA acknowledges support from STFC through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
+EB acknowledges support from the Beecroft Trust.
+
+  \appendix
+  \section{Flat sky}\label{app:flat}
+    \begin{figure}
+      \centering
+      \includegraphics[width=0.6\textwidth]{./figures/mask-lss_flat1.pdf}
+      \caption{\todo{write}} \label{fig:mask_flat}
+    \end{figure}
 
@@
       \centering
-  \includegraphics[width=\columnwidth]{./figures/foreground.png}
-  \caption{$\cl$ foregrounds used in order to check the reliability of the
-    analytical approximations in their presence. They did not affect. The
-    limiting cases for the large-scale contaminants are given by $\beta=-1,\,
-    -3$ in Eq.~\ref{eq:fore}}
-  \label{fig:fore}
+      \includegraphics[width=\textwidth]{./figures/run_chi2_TT_TE_EE_TB_EB_BB.pdf}
+      \caption{\todo{write}}
+      \label{fig:chi2_1bin_flat}
     \end{figure}
+    We have repeated the analysis described in Section \ref{sec:results} on flat-sky realizations, making use of the flat-sky implementation of {\tt NaMaster}.
+We generate Gaussian realizations of the galaxy overdensity and shear maps making use of flat-sky extensions of the methods described in Section \ref{ssec:results.sims}.
+In this case we use a high-resolution mask constructed from the bright-object mask distributed with the first data release of the HSC collaboration \todo{cite} for the VVDS field, which is shown in Fig.
+\ref{fig:mask_flat}.
 
-\subsection{Visual inspection}
-\todo{Rows plot para todas las combinaciones}
-\todo{diff-correlation TTTT BBBB}
-\todo{Eigenvalue plot}
+    We find similar levels of accuracy in the NKA and spin-0 estimators compared to the curved-sky case.
+The higher resolution and smaller area of these simulations allow us to focus on the small-scale galaxy clustering and lensing power spectra, covering the range of multipoles $\ell\in(120,17640)$ in constant bandpowers of width $\Delta \ell=240$.
+On these small scales, the shear power spectrum is more dominated by noise (e.g.
+see Fig.
+\ref{fig:cl-2bins}), and therefore there is roughly the same power in $E$ and $B$ modes.
+This reduces the sensitivity of the method to an inaccurate treatment of $E/B$ leakage, and the agreement between the sample covariance matrix and the NKA estimator improves significantly.
+This can be seen in Fig.
+\ref{fig:chi2_1bin_flat}, which shows the $\chi^2$ distributions for all possible power spectra (including those including $B$ modes) in the case of a single redshift bin.
+In all cases we find a good agreement between the $\chi^2$ distributions derived from all covariance matrix estimates (sample covariance, NKA and spin-0 in blue, orange and green respectively), which also accurately follow the expected $\chi^2$ distribution for the corresponding number of degrees of freedom (red dashed line).
 
-\subsection{Quantitative checks}
-\todo{Plot $\chi^2$ TT-TT, TE-TE, EE-EE}
+  \section{Software implementation}\label{app:namaster}
+    \todo{complete}
 
-\subsection{Contaminant deprojection}
-\todo{}
-
-\subsection{Impact on parameter estimation}
-\todo{$\chi^2$ plot para 2 bines}
-\todo{diff-correlaton plot para 2 bines ver Fig. 8  de 1809.09603}
-\todo{Contours}
-
-\subsection{$\chi^2$ distribution and covariance matrices - old section}
-Apart from checking that we recover the same cosmological predictions from the
-covariance matrices from both the analytical approximation and the
-simulations, we studied the differences between them.
-
-In this sense, one can see in Fig.~\ref{fig:TTTT_EEEE_chi2} that, we fully
-recover the $\chi^2$ distribution (defining $\chi^2 = \cl - \langle \cl
-\rangle C^{-1} \cl - \langle \cl \rangle$) for the TTTT case; whereas for the
-worst case (i.e. EEEE) we are able to recover its shape, although sifted
-toward lower values. This might be consequence of the fact that the inverse of
-the covariance matrix is biased. It is interesting to note that even
-the approximation of the mixing matrix as that of the case with spin 0, gives
-good results. The eigenvalues of the covariance matrix
-(Fig.~\ref{fig:TTTT_EEEE_eigv}) are also recovered within the 1\%, for all
-scales in the TTTT case, and for the smaller scales in the EEEE case. Finally,
-the analytical approximations are shown to describe the scale coupling (see
-Fig.~\ref{fig:TTTT_rows} that comes from the fact we are masking some regions
-of the sky. The larger relative deviation for the better approximations are
-due to the fact that there is a division by $0$, while for the naive
-approximation, which is defined diagonal (see Eq.~\ref{eq:naive}), the
-relative deviation is exactly $-1$ for all values with $l \neq l'$.
-
-\begin{figure*} %[htb]
-  \centering
-  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_TTTT_chi2.png}~
-  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_EEEE_chi2.png}
-  \caption{Comparison of the distributions found using different
-    covariances for the TTTT (left) and EEEE (right) cases.}
-  \label{fig:TTTT_EEEE_chi2}
-\end{figure*}
-
-\begin{figure*} %[htb]
-  \centering
-  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_TTTT_reldev_eigval.png}~
-  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_EEEE_reldev_eigval.png}
-  \caption{Comparison of the covariance matrix eigenvalues for the TTTT (left)
-    and EEEE (right) cases.} 
-  \label{fig:TTTT_EEEE_eigv}
-\end{figure*}
-
-\begin{figure} %[htb]
-  \centering
-  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_TTTT_rows_cov_matrix.png}~
-  \caption{Mask induced scale correlations for the TTTT case.}
-  \label{fig:TTTT_rows}
-\end{figure}
-
-The goodness of the approximation is also seen in the correlation matrices.
-The subtraction of the correlation matrix from the simulations and that from
-the best analytical approximation (which we have just seen gives almost same
-results as the spin-0 one) shows no structure but for the case of the larger
-scales of the EEEE modes (see Fig.~\ref{fig:TTTT_EEEE_corr}).
-
-\begin{figure*} %[htb]
-  \centering
-  \includegraphics[width=\columnwidth]{./figures/run_sph_Efstathiou_TTTT_correlation_difference.png}~
-  \includegraphics[width=\columnwidth]{./figures/run_sph_Efstathiou_EEEE_correlation_difference.png}
-  \caption{Differences between the simulations and analytical approximation of
-    the correlation matrices for the TTTT (left) and EEEE (right) cases.}
-  \label{fig:TTTT_EEEE_corr}
-\end{figure*}
-
-These approximations do not work well for the B-modes, though. We can see in
-Fig.~\ref{fig:BBBB} that we are not able to recover with accuracy the first
-diagonal elements.
-
-\begin{figure} %[htb]
-  \centering
-  \includegraphics[width=\columnwidth]{./figures/run_sph_ALL_BBBB_check_diagonal_terms_4.png}
-  \caption{The analytical approximation are not able to reproduce accurately
-    the terms involving the B-modes. Here we show the most extreme case, the
-    corresponding to the BBBB modes.}
-  \label{fig:BBBB}
-\end{figure}
-
-Finally, the analytical covariance also works for 2-bin correlations. The
-$\chi^2$ distribution for the covariance matrix of the modes TT, TE and EE
-among two bins is shown in Fig.~\ref{fig:TTTEEE_chi2}. The eigenvalues
-are recovered well except for the lower ones that diverge up to $20\%$ (see
-Fig.~\ref{fig:TTTEEE_eigv}, this is not translated to any substantial
-difference in the correlation matrices (see Fig.~\ref{fig:TTTEEE_corr}).
-
-\begin{figure} %[htb]
-  \centering
-  \includegraphics[width=\columnwidth]{./figures/run_sph_2b_same_mask_Efstathiou_TTTEEE_Full_chi2.png}
-  \caption{$\chi^2$ distribution check for the full analytical covariance
-    matrix for the modes TT, TE and EE among two bins.}
-  \label{fig:TTTEEE_chi2}
-\end{figure}
-
-\begin{figure} %[htb]
-  \centering
-  \includegraphics[width=\columnwidth]{./figures/run_sph_2b_same_mask_Efstathiou_TTTEEE_Full_reldev_eigval.png}
-  \caption{Relative difference between the analytical and simulations
-    covariance matrix eigenvalues for the modes TT, TE and EE among two bins.}
-  \label{fig:TTTEEE_eigv}
-\end{figure}
-
-\begin{figure} %[htb]
-  \centering
-  \includegraphics[width=\columnwidth]{./figures/run_sph_2b_same_mask_Efstathiou_TTTEEE_Full_correlation_difference.png}
-  \caption{Difference of the correlation matrices of the simulated $\cl$ and
-    the analytical one for the modes TT, TE and EE among two bins..}
-  \label{fig:TTTEEE_corr}
-\end{figure}
-
-% The resulting power spectra was then used to generate a map of the sky and
-% two random fields (one of spin-0 and other of spin-2), in turn. Their
-% correlations give  are used to obtain the simulated power spectra. In order to
-% find the real power spectra, it is necessary to subtract the contribution of
-% the mask, which mixes the closest modes:
-% \begin{equation}
-%   C^{obs}_l = \sum_{l'} M_{ll'} C_{l'}\,
-% \end{equation}
-% where $M$ depends on the mask. In general, $M$ is not invertible and, in order
-% to be able to recover the true power spectra, $C_l$, one needs to bin the
-% $l$-space. The bin width is given by the characteristics of the window
-% function and must be, approximately, of the size of the range of the mixed
-% modes. That way, each $\tilde C_l$ is almost independent of the others
-% $\tilde C_{l'}$, and $M$ is invertible.
-
-\section{Discussion}\label{sec:discussion}
-
-\section*{Acknowledgements}
-We would like to thank Eva-Maria Mueller for useful discussion. CGG is
-supported the Spanish grant BES-2016-077038, partially funded by the ESF and by
-AYA2015-67854-P from the Ministry of Industry, Science and Innovation of Spain
-and the FEDER funds. He was partially supported by a Balzan Fellowship while
-in Oxford. He would like to thank New College and the Department of Physics at
-Oxford for their hospitality. DA acknowledges support from STFC through an
-Ernest Rutherford Fellowship, grant reference ST/P004474/1.
-
-\appendix
-
-\section{Flat sky}
-\todo{Plot mascara}
-\todo{Plot chi2 TT-TT TE-TE, EE-EE}
-
-\setlength{\bibhang}{2.0em}
-\setlength\labelwidth{0.0em}
-\bibliography{paper.bib}
+  \bibliographystyle{JHEP}
+  \bibliography{paper}
 

### 2019-06-18 draft/paper.tex
added caption of Fig. 13 & 14

@@
       \includegraphics[width=0.6\textwidth]{./figures/mask-lss_flat1.pdf}
-      \caption{\todo{write}} \label{fig:mask_flat}
+      \caption{Flat-sky mask used in the study of the analytical
+        methods presented in this work in the flat-sky regime. The mask was
+        constructed from the bright-object mask distributed with the first
+        data release of the HSC collaboration \todo{cite} for the VVDS
+        field. As in the full-sky case, it is used both for galaxy clustering
+        and shear for simplicity.} \label{fig:mask_flat}
     \end{figure}
@@
       \includegraphics[width=\textwidth]{./figures/run_chi2_TT_TE_EE_TB_EB_BB.pdf}
-      \caption{\todo{write}}
+      \caption{$\chi^2$ distributions for the autocorrelations of the fields
+        of a redshift bin in the flat-sky regime. We compare the distributions
+        obtained with the covariance matrix estimated from simulated power
+        spectra, and the analytical NKA and Spin-0 methods. In addition, we
+        show the theoretical $\chi^2$ distribution, which is almost perfectly
+        followed by the others. The reason why the E and B modes are well
+        behaved in flat sky is because in this regime ($\ell \in (120,
+        17640)$), the shear power spectra is dominated by noise (see
+        Fig.~\ref{fig:cl-2bins}). }
       \label{fig:chi2_1bin_flat}

### 2019-06-21 notebooks/general-plots-paper.py
added method to plot chi2 means & covs

@@ -552,7 +552,136 @@ def chi2_sph_TT_TE_EE():
     axs[0].legend(loc='center right', fontsize='9', frameon=False)
     plt.tight_layout()
     fname = os.path.join(outdir, 'run_sph_2b_1stbin_chi2_TT_TE_EE.pdf')
-    # plt.savefig(fname, dpi=DPI)
+    plt.savefig(fname, dpi=DPI)
+    # plt.show()
+    plt.close()
+
+##############################################################################
+################# TTTT, TETE, EEEE chi2 mean + sigma diff ####################
+##############################################################################
+
+def chi2_sph_1b_chi2_mean_cov_diff():
+    npzfile = os.path.join(outdir, 'run_sph_2b_1stbin_chi2_mean_cov_diff.npz')
+    if os.path.isfile(npzfile):
+        rdmeans = np.load(npzfile)['rdmeans']
+        rdcovs = np.load(npzfile)['rdcovs']
+    else:
+        prefix = 'run_sph_2b'
+        run_path = os.path.join('./simulations_outputs/', prefix, prefix)
+        sims_suffix = '_cl_0001-20000.npz'
+
+        ##############################################################################
+
+        ell = np.loadtxt(run_path + '_ells.txt')
+        lmax = (ell < 2*512).sum()
+        nlbins = lmax
+
+        clsims = np.load(run_path + '_b1' + sims_suffix)['arr_0']
+
+        clTT = clsims[0, 0, :, :lmax]
+        clTE = clsims[0, 1, :, :lmax]
+        clTB = clsims[0, 2, :, :lmax]
+        clEE = clsims[1, 1, :, :lmax]
+        clEB = clsims[1, 2, :, :lmax]
+        clBB = clsims[2, 2, :, :lmax]
+
+        ##############################################################################
+
+        CovSims_path = run_path + '_cov_b1' + sims_suffix
+        Csims = np.load(CovSims_path)['arr_0']
+
+        nlbins_orig = int(Csims.shape[0] / 6)
[…]

### 2019-06-21 notebooks/general-plots-paper.py
added lines to compute rel dev width + mean of chi2 TT-TE-EE 2bins

@@ -816,6 +816,18 @@ def chi2_NKA_TTTEEE_full():
     lbins = ell
     lmax = (ell < 2*512).sum()
 
+    means = np.mean(chi2_list, axis=1)
+    covs = np.array([np.sqrt(np.cov(chi2is)) for chi2is in chi2_list])
+    print("Chi2_NKA_TTTEEE_full:")
+    print("Relative mean diff: {}".format(means[1]/means[0] - 1))
+    print("Relative width diff: {}".format(covs[1]/covs[0] - 1))
+    Ns = chi2_list.shape[1]
+    Nl = lmax * 10
+    H = (Ns - Nl - 2)/(Ns - 1) # arXiv:1601.05786
+    print('Ns =',  Ns, 'Nl =', Nl, 'Hartlap =', H)
+    print("Relative mean diff (Hartlap): {}".format(means[1]/(H * means[0]) - 1))
+    print("Relative width diff (Hartlap): {}".format(covs[1]/(H * covs[0]) - 1))
+
     f, ax = plt.subplots(1, 1, figsize=FSIZE1)
     bins = np.linspace(np.min(chi2_list), np.max(chi2_list), 60)
     _, x, _ = ax.hist(chi2_list[0], bins=bins, histtype='step', density=True,
@@ -846,6 +858,20 @@ def chi2_Spin0_NKA_TTTEEE_full():
     ell = np.loadtxt('../run_sph_2b_same_mask/run_sph_2b_same_mask_ells.txt')
     lbins = ell
     lmax = (ell < 2*512).sum()
+
+    means = np.concatenate([np.mean(chi2_list, axis=1), [np.mean(chi2_spin0)]])
+    covs = np.array([np.sqrt(np.cov(chi2is)) for chi2is in chi2_list] + [np.sqrt(np.cov(chi2_spin0))])
+    print("Chi2_Spin0_NKA_TTTEEE_full:")
+    print("Results for | NKA | Spin0 |")
+    print("Relative mean diff: {}".format(means[1:]/means[0] - 1))
+    print("Relative width diff: {}".format(covs[1:]/covs[0] - 1))
+    Ns = chi2_list.shape[1]
+    Nl = lmax * 10
+    H = (Ns - Nl - 2)/(Ns - 1) # arXiv:1601.05786
+    print('Ns =',  Ns, 'Nl =', Nl, 'Hartlap =', H)
[…]

### 2019-06-21 notebooks/compute_covSims_full.py
accept input folder

@@ -1,10 +1,12 @@
 #!/usr/bin/python
 import numpy as np
 import os
+import sys
 
 # pylint: disable=C0103
 
-prefix = 'run_sph_2b_same_mask'
+# prefix = 'run_sph_2b_same_mask'
+prefix = os.path.basename(sys.argv[1])
 run_path = os.path.join('./simulations_outputs/', prefix, prefix)
 outdir = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
 output_path = os.path.join(outdir, prefix)

### 2019-06-21 notebooks/compute_covSims_full.py
fixed problem when path had / at the end

@@ -6,7 +6,7 @@ import sys
 # pylint: disable=C0103
 
 # prefix = 'run_sph_2b_same_mask'
-prefix = os.path.basename(sys.argv[1])
+prefix = os.path.basename(sys.argv[1].rstrip('/'))
 run_path = os.path.join('./simulations_outputs/', prefix, prefix)
 outdir = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
 output_path = os.path.join(outdir, prefix)

### 2019-06-21 draft/paper.tex
more paper

@@
 The presence of a complicated sky mask causes non-trivial correlations between different Fourier/harmonic modes, which must be accurately characterised in order to obtain reliable cosmological constraints.
+This is particularly relevant for galaxy survey data.
 Unfortunately, an exact calculation of these correlations involve $O(\ell_{\rm max}^6)$ operations that become computationally impractical very quickly.
-We present an implementation of approximate methods to estimate the Gaussian covariance matrix of spin-0 and spin-2 flat- and curved-sky fields, expanding on existing algorithms.
+We present an implementation of approximate methods to estimate the Gaussian covariance matrix of power spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding on existing algorithms.
 These methods achieve an $O(\ell_{\rm max}^3)$ scaling, which makes the computation of the covariance matrix as fast as the computation of the power spectrum itself.
-We quantify the accuracy of these method on large-scale structure and weak lensing data, characterised by comparatively complex survey geometries, making use of a large number of Gaussian but otherwise realistic simulations.
+We quantify the accuracy of these methods on large-scale structure and weak lensing data, making use of a large number of Gaussian but otherwise realistic simulations.
 We show that, using the covariance matrix computing using these approximations we are able to recover the true posterior distribution cosmological parameters to high accuracy.
@@
     In this paper we will present and generalise methods developed in the context of CMB experiments to account for the impact of survey geometry on the Gaussian part of the power spectrum covariance matrix \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}, and will study in detail the performance of these methods for large-scale structure and weak lensing datasets.
-We have also implemented these approximations in the public code {\tt NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}} \cite{2019MNRAS.484.4127A}, making the computation of accurate Gaussian covariance matrices significantly simpler.
+We have also implemented these approximations in the public code {\tt NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}} \cite{2019MNRAS.484.4127A}, making the computation of accurate Gaussian covariance matrices significantly simpler for the community.
 
@@
 We then summarize our results and conclude in Section \ref{sec:discussion}.
-Appendix \ref{app:flat} presents the performance of these methods for flat-sky approximation, and we provide technical details of the software implementation in Appendix \ref{app:namaster}.
+Appendix \ref{app:flat} presents the performance of these methods in the flat-sky approximation, and we provide technical details of the software implementation in Appendix \ref{app:namaster}.
 
@@
 
-      We now make use of Wick's theorem, which states that, for Gaussian fields $\langle a\,b\,f\,g\rangle=\langle a\,b\rangle\langle f\,g\rangle+\langle a\,f\rangle\langle b\,g\rangle+\langle a\,g\rangle\langle b\,f\rangle$.
+      We now make use of Wick's theorem, which states that, for Gaussian fields, $\langle a\,b\,f\,g\rangle=\langle a\,b\rangle\langle f\,g\rangle+\langle a\,f\rangle\langle b\,g\rangle+\langle a\,g\rangle\langle b\,f\rangle$.
 In this case, the expression for $\Sigma^{AF}_{{\bf l}{\bf l'}}$ reads:
@@
 
-      Under the assumption that the coupling coefficients $M_{{\bf l}{\bf k}}$ are sharply peaked around ${\bf l}={\bf k}$, we can simplify the expression above approximating the power spectra as constants within the support of the coupling coefficients.
+      Under the assumption that the coupling coefficients $M_{{\bf l}{\bf k}}$ are sharply peaked around ${\bf l}={\bf k}$, we can simplify the expression above approximating the power spectra as constants within the support of the coupling coefficients \cite{2004MNRAS.349..603E}.
 Explicitly, we approximate
@@
       \begin{equation}
-        \,^{ab}I^{\pm s_a,\pm s_b}_{{\bf l}{\bf l'}}\equiv\summ{k}\summ{x}\summ{y}w_a({\bf x})w_b({\bf y})\,q^{\pm s_a}_{{\bf l}{\bf k}}({\bf x})\,\left(q^{\pm s_b}_{{\bf l}'{\bf k}}({\bf y})\right)^*,
+        \,^{ab}I^{\pm s_a,\pm s_b}_{{\bf l}{\bf l'}}\equiv\summ{k}\summ{x}\summ{y}w_a({\bf x})w_b({\bf y})\,q^{\pm s_a}_{{\bf l}{\bf k}}({\bf x})\,\left[q^{\pm s_b}_{{\bf l}'{\bf k}}({\bf y})\right]^*,
       \end{equation}
@@
       Now, in what follows, we will be concerned with the auto- and cross-correlations of spin-0 and spin-2 fields.
-Thus, to simplify the notation, we will enumerate the different types of coupling coefficients that exist for a spin-0 field with a single component that we will call $\delta$, in analogy to the projected galaxy overensity, and for a spin-2 field $\mathbf{\gamma}$ with $E$ and $B$ components $\gamma_E$ and $\gamma_B$, in analogy to the cosmic shear field.
+Thus, to simplify the notation, we will enumerate the different types of coupling coefficients that exist for a spin-0 field with a single component that we will call $\delta$, in analogy to the projected galaxy overensity, and for a spin-2 field $\boldsymbol{\gamma}$ with $E$ and $B$ components $\gamma_E$ and $\gamma_B$, in analogy to the cosmic shear field.
 With this setup, all the possible non-zero $W^{\alpha\beta,\alpha'\beta'}_{{\bf l}{\bf l}'}$ can be expressed in terms of the $I^{\pm s_a,\pm s_b}_{{\bf l}{\bf l}'}$ as follows:
@@
 
-      Using these identities together with the completeness relation of the basis functions, it is possible to simplify the expressions for the $I^{\pm s_a,\pm s_b}$ \todo{cite}:
+      Using these identities together with the completeness relation of the basis functions, it is possible to simplify the expressions for the $I^{\pm s_a,\pm s_b}$:
       \begin{align}
@@
       \begin{align}\label{eq:covar_coupling_J}
-        &\,^{ab}J^0_{{\bf l}{\bf l}'}=\summ{x}\left(w_a w_b\right)({\bf x})\,q({\bf l},{\bf x})\,\left(q({\bf l}',{\bf x})\right)^*,\\
+        &\,^{ab}J^0_{{\bf l}{\bf l}'}=\summ{x}\left(w_a w_b\right)({\bf x})\,q({\bf l},{\bf x})\,\left[q({\bf l}',{\bf x})\right]^*,\\
         &\,^{ab}J^\pm_{{\bf l}{\bf l}'}=\summ{x}\left(w_a w_b\right)({\bf x})\,q^{\pm2}_{{\bf l}{\bf l}'}({\bf x}).
@@
         \includegraphics[width=0.6\textwidth]{./figures/pz.pdf}
-        \caption{}
+        \caption{Redshift distributions assumed for the simulated maps used in this analysis.}
         \label{fig:pz}
@@
         \includegraphics[width=\textwidth]{./figures/cls-sph-2b.pdf}
-        \caption{Fiducial power spectra. We show only the $T$ and $E$ modes.}
+        \caption{Signal (blue) and noise (orange) power spectra for the different observables used in our simulations.
+Note that the signal for shear $B$-modes ($\gamma_B$) is zero, and the noise is the same as that of $\gamma_E$.
+The different panels show different cross-correlations between $\delta$ and $\gamma_E$ in two redshift bins.}
         \label{fig:cl-2bins}
@@
         \includegraphics[width=0.6\textwidth]{./figures/mask-lss2.pdf}
-        \caption{Sky masks used in our analysis for the two different redshift bins.
-The same mask is assumed for the galaxy overdensity $\delta$ and cosmic shear $\mathbf{\gamma}$ for simplicity.} \label{fig:mask}
+        \caption{Sky masks used in our analysis for bins 1 and 2 (upper and lower panels respectively).
+The same mask is assumed for the galaxy overdensity $\delta$ and cosmic shear $\boldsymbol{\gamma}$ for simplicity.} \label{fig:mask}
       \end{figure}
@@
         \item Spin-2 fields, corresponding to maps of the cosmic shear measured from the projected shapes of galaxies in a given redshift bin.
-We will label these fields $\mathbf{\gamma}^a=(\gamma_E^a,\gamma_B^a)$ where, again, the index $a$ denotes the redshift bin.
+We will label these fields $\boldsymbol{\gamma}^a=(\gamma_E^a,\gamma_B^a)$ where, again, the index $a$ denotes the redshift bin.
       \end{itemize}
@@
       \end{equation}
-      where $\chi$ is the comoving radial distance, $z\equiv z(\chi)$ is the corresponding redshift in the lightcone, $P(k,z)$ is the matter power spectrum and the window functions are given by \cite{2004PhRvD..70d3009H}\footnote{See \cite{2019ApJS..242....2C} and references therein for details about these theory predictions}:
+      where $\chi$ is the comoving radial distance, $z\equiv z(\chi)$ is the corresponding redshift in the lightcone, $P(k,z)$ is the matter power spectrum and the window functions are given by \cite{2004PhRvD..70d3009H}\footnote{See \cite{2019ApJS..242....2C} and references therein for details about these calculations.}:
       \begin{align}
@@
       We have also explored the impact of the presence of sky contaminants on the estimate of the covariance matrix.
-As described in \cite{2019MNRAS.484.4127A}, the presence of a small contamination from observational systematics in the data can be accounted for through a technique known as {\sl mode deprojection} \todo{Cites}.
+As described in \cite{2019MNRAS.484.4127A}, the presence of a small contamination from observational systematics in the data can be accounted for through a technique known as {\sl mode deprojection} \cite{1992ApJ...398..169R,2004PhRvD..69l3003S,2017MNRAS.465.1847E}.
 In this method, the contamination is modeled as a linear contribution at the map level from contaminants with a known template.
@@
       \end{itemize}
-      In both cases we fixed the amplitude of the contaminant power spectrum
-      such that it would yield a $10\%$ contamination in the data power
-      spectrum at $\ell=400$. In both cases, for spin-2 fields, we assumed the
-      same power spectrum for $E$ and $B$ modes, with no cross-correlation
-      between them. The resulting contaminant power spectra are shown in
-      Figure \ref{fig:clcont}.
+      In both cases we fixed the amplitude of the contaminant power spectrum such that it would yield a $10\%$ contamination in the data power spectrum at $\ell=400$.
+For spin-2 fields, we assumed the same power spectrum for $E$ and $B$ modes, with no cross-correlation between them.
+The resulting contaminant power spectra are shown in Figure \ref{fig:clcont}.
 When exploring the effects of mode deprojection, we generate 100 contaminant maps of both types and add them to each simulated realization.
@@
 Note that our fiducial results do not include the effects of mode deprojection.
-These are discussed in detail separately.
+These are discussed separately.
 
@@
         \includegraphics[width=0.7\textwidth]{./figures/sph_1bin_diff_corr.pdf}
-        \caption{Difference between the correlation matrices associated to the sample covariance matrix and its estimate with the NKA method including all possible correlations between $\delta$ and $\mathbf{\gamma}$ in a single redshift bin.} \label{fig:diff_corr_1bin}
+        \caption{Difference between the correlation matrices associated to the sample covariance matrix and its estimate with the NKA method including all possible correlations between $\delta$ and $\boldsymbol{\gamma}$ in a single redshift bin.} \label{fig:diff_corr_1bin}
       \end{figure}
@@
       \end{figure}
-      \begin{figure}
-        \centering
-        \includegraphics[width=0.6\textwidth]{./figures/contaminants_chi2.pdf}
-        \caption{$\chi^2$ distributions for simulations with no systematic contaminants (blue), and for simulations containing contaminants after deprojection of 200 contaminant templates (orange).
+%      \begin{figure}
+%        \centering
+%        \includegraphics[width=0.6\textwidth]{./figures/contaminants_chi2.pdf}
+%        \caption{$\chi^2$ distributions for simulations with no systematic contaminants (blue), and for simulations containing contaminants after deprojection of 200 contaminant templates (orange).
 The covariance matrices associated to both sets of simulations are essentially the same, and the differences between them are much smaller than those associated with the NKA estimator (shown in green).
 The NKA method is therefore sufficient to approximate the covariance even in the presence of contaminant deprojection.}
-        \label{fig:chi2_conts}
-      \end{figure}
+%        \label{fig:chi2_conts}
+%      \end{figure}
       In order to quantify the validity of our analytical approximations, we need to compare the NKA covariance with the sample covariance estimated from the simulations.
@@
       \end{equation}
-      We compute this quantity for a data vector ${\bf d}$ composed of
-      different auto- and cross-correlations for each of the 20,000 Gaussian
-      simulations, with ${\bf m}$ given by the mean over all simulations and
-      different choices of covariance matrix. Figure \ref{fig:chi2_1bin} shows
-      the distribution of $\chi^2$ values for the three non-zero power spectra
-      in the case of a single bin: $\delta$-$\delta$, $\delta$-$\gamma_E$ and
-      $\gamma_E$-$\gamma_E$. The histograms show the distribution for the
-      sample covariance matrix (blue), and the analytical NKA and Spin-0
-      estimators (orange and green respectively). We additionally plot the
-      theoretical $\chi^2$ distribution under the assumption that the
-      underlying data vector is Gaussianly distributed (red dashed lines). In
-      the simplest case of purely spin-0 quantities (leftmost panel), we find
-      an excellent agreement between the different distributions. In the cases
-      involving the spin-2 fields, we see noticeable differences between the
-      distributions found with the sample covariance and the approximate ones.
-      These differences are small, corresponding to less of a $2$ and $4\%$
-      shift in the mean $\chi^2$ for the $\delta$ - $\gamma_E$ and
-      $\gamma_E$ - $\gamma_E$ cases, respectively, and a negligible variation in
-      the width of the distributions. We therefore expect these differents to
-      have a negligible effect on the posterior parameter distributions, as we
-      will show in Section \ref{ssec:results.cosmo}. The fact that these
-      differences appear only for power spectra involving spin-2 quantities
-      indicate that the NKA and spin-0 methods are imperfect at describing the
-      additional mode coupling caused in the presence of a mask for
-      higher-spin fields. This is rather obvious in the case of the spin-0
-      approximation but, interestingly, we find that the NKA and Spin-0
-      predictions yield results that are almost indistinguishable from each
-      other. We therefore conclude that the additional approximation made in
-      the NKA method for spin-2 fields -- neglecting the spatial derivatives
-      of the mask -- is effectively equivalent to ignoring the spin nature of
-      the fields involved. For completeness, Figure \ref{fig:chi2_2bins} shows
-      the distribution of $\chi^2$ values for a data vector composed of all
-      possible auto- and cross-correlations of $\delta$ and $\gamma_E$ for the
-      case of two redshift bins, where similar conclusions hold.
+      We compute this quantity for a data vector ${\bf d}$ composed of different auto- and cross-correlations for each of the 20,000 Gaussian simulations, with ${\bf m}$ given by the mean over all simulations and different choices of covariance matrix.
+Figure \ref{fig:chi2_1bin} shows the distribution of $\chi^2$ values for the three non-zero power spectra in the case of a single bin: $\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$.
+The histograms show the distribution for the sample covariance matrix (blue), and the analytical NKA and Spin-0 estimators (orange and green respectively).
+We additionally plot the theoretical $\chi^2$ distribution under the assumption that the underlying data vector is Gaussianly distributed (red dashed lines).
+In the simplest case of purely spin-0 quantities (leftmost panel), we find an excellent agreement between the different distributions.
+In the cases involving the spin-2 fields, we see noticeable differences between the distributions found with the sample covariance and the approximate ones.
+      These differences are small, corresponding to less of a $2$ and $4\%$ shift in the mean $\chi^2$ for the $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ cases respectively, and a negligible variation in the width of the distributions.
+We therefore expect these differents to have a negligible effect on the posterior parameter distributions, as we show explicitly in Section \ref{ssec:results.cosmo}.
+The fact that these differences appear only for power spectra involving spin-2 quantities indicate that the NKA and spin-0 methods are imperfect at describing the additional mode coupling caused in the presence of a mask for higher-spin fields.
+This is rather obvious in the case of the spin-0 approximation but, interestingly, we find that the NKA and Spin-0 predictions yield results that are almost indistinguishable from each other.
+We therefore conclude that the additional approximation made in the NKA method for spin-2 fields -- neglecting the spatial derivatives of the mask -- is effectively equivalent to ignoring the spin nature of the fields involved.
+For completeness, Figure \ref{fig:chi2_2bins} shows the distribution of $\chi^2$ values for a data vector composed of all possible auto- and cross-correlations of $\delta$ and $\gamma_E$ for the case of two redshift bins, where similar conclusions hold.
 
-      We have also explored the impact of contaminant deprojection on the
-      different covariance matrix estimates. The loss of modes due to
-      deprojection can potentially increase the variance of the power spectrum
-      estimates, affecting the accuracy with which an analytical estimator
-      would be able to recover the covariance. Figure \ref{fig:diag_conts}
-      shows the diagonal of the covariance matrix for the $\delta$-$\delta$,
-      $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ power spectra. Each panel
-      displays the diagonal for the sample covariance matrix estimated from
-      simulations without contaminants or contaminant deprojection (blue
-      line), the sample covariance from simulations with contaminants and
-      contaminant deprojection (dashed orange line) and for the NKA covariance
-      (green line). We see that the power spectrum uncertainties
-      are almost indistinguishable with or without deprojection, and that
-      those relative differences are much smaller than the differences between
-      the sample covariance and the NKA estimator. We therefore conclude that,
-      except in the case where a very large set of contaminant maps are
-      deprojected (comparable with the number of unmasked pixels in the map),
-      the analytical approximation to the covariance matrix should be as
-      accurate as in the absence of contaminants (i.e. accurate enough).
-      This is quantitatively shown in Fig.~\ref{fig:chi2_conts} through
-      comparison of the $\chi^2$ distributions drawn with covariance matrices
-      estimated with simulations without contaminants, simulations with
-      contaminants deprojected and the NKA method.
+      We have also explored the impact of contaminant deprojection on the different covariance matrix estimates.
+The loss of modes due to deprojection can potentially increase the variance of the power spectrum estimates, affecting the accuracy with which an analytical estimator would be able to recover the covariance.
+Figure \ref{fig:diag_conts} shows the diagonal of the covariance matrix for the $\delta$-$\delta$,      $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ power spectra.
+Each panel displays the diagonal for the sample covariance matrix estimated from simulations without contaminants or contaminant deprojection (blue
+      line), the sample covariance from simulations with contaminants and contaminant deprojection (dashed orange line) and for the NKA covariance   (green line).
+We see that the power spectrum uncertainties are almost indistinguishable with or without deprojection, and that those relative differences are much smaller than the differences between the sample covariance and the NKA estimator.
+We therefore conclude that, except in the case where a very large set of contaminant maps are deprojected (comparable with the number of unmasked pixels in the map), the analytical approximation to the covariance matrix should be as accurate as in the absence of contaminants (i.e.
+accurate enough).
+%      This is quantitatively shown in Fig.~\ref{fig:chi2_conts} through comparison of the $\chi^2$ distributions drawn with covariance matrices
+%      estimated with simulations without contaminants, simulations with
+%      contaminants deprojected and the NKA method.
+
+      Figure \ref{fig:diag_conts} serves also to illustrate another important point.
+In all our tests we find that the largest differences between the sample and NKA covariances occur on large scales.
+For $\delta$-$\delta$ correlations, the effect is limited to the first few multipoles ($\ell\lesssim10$), while for $\gamma_E$-$\gamma_E$ we are only able to recover the sample covariance errors within $5\%$ for $\ell\gtrsim40$.
+This is the main source of the small mismatch observed in Fig.
+\ref{fig:chi2_1bin}.
+Note however, that most of the cosmological information is obtained from the higher multipoles, due to their higher statistical weight, and as we will show in the next section, the effect on the final parameter constraints is negligible.
+If accurate covariances bars are needed on these large scales for cosmic shear, they can be estimated alternatively making use of fast, low resolution simulations (e.g.
+{\tt HEALPix} $N_{\rm side}=64$ maps).
 
@@
       \begin{equation}\label{eq:parpos}
-        -2\log p(\vec{\theta})=\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)^T\cdot{\rm Cov}^{-1}\cdot\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)+{\rm const.},
+        -2\log p(\vec{\theta}|\hat{\bf C})=\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)^T\cdot{\rm Cov}^{-1}\cdot\left(\hat{\bf C}-{\bf C}(\vec{\theta})\right)+{\rm const.},
       \end{equation}
@@
 
-      We explore $\log p(\vec{\theta})$ for of two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins.
-In this simple two-dimensional scenario we simply sample the distribution in a regular grid of 200 by 200 points for each parameter.
+      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for of two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins.
+In this simple two-dimensional scenario we simply sample the distribution in a regular grid of 100 by 100 points for each parameter.
 We construct a data vector $\hat{\bf C}$ from the theoretical prediction for the experimental setup described in Section \ref{ssec:results.sims}, and produce theoretical predictions for it at each grid point using {\tt CCL}.
@@
 \ref{fig:contours}.
-We find that both distributions agree with each other remarkably well, and that the $1\sigma$ errors for each parameter agree for both covariances up to \todo{$x\%$}.
+We find that both distributions agree with each other remarkably well, and that the $1\sigma$ errors for each parameter agree for both covariances up to $0.3\%$.
 Note that, since we have not included any statistical noise in the data vector $\hat{\bf C}$, the relative bias between both distributions is zero by construction (since Eq.
@@
 The main impact of survey geometry is the statistical coupling it induces between different Fourier/harmonic modes, which must be accurately characterized in order to obtain reliable estimates of the posterior distribution for cosmological parameters.
-In this paper we have focused on the impact of survey geometry on the Gaussian (i.e.
+In this paper we have focused on the impact of survey geometry on the dominant Gaussian (i.e.
 disconnected) part of the covariance matrix.
@@
 Due to the imperfect treatment of the $E/B$ mixing caused by the sky mask in the NKA estimator, we also find that the predicted covariance matrix for any power spectra involving $B$ modes differs significantly from the true sample covariance, and therefore this approach cannot be used to reliably estimate the uncertainties of $B$-mode power spectra.
+Likewise, we find that the NKA estimator yields inaccurate estimates of the power spectrum uncertainties on the largest scales ($\ell\lesssim50$).
+Although these modes carry a substantially smaller statistical weight, if more accurate covariances are needed on these large scales, they can be easily computed making use of fast low-resolution simulations.
 
-    In spite of these inaccuracies, we find that the approximations described in this paper are able to provide estimates of the power spectrum covariance matrix that are sufficiently accurate for current and future tomographic large-scale structure cosmological datasets.
+    In spite of these shortcomings, we find that the approximations described in this paper are able to provide estimates of the power spectrum covariance matrix that are sufficiently accurate for current and future tomographic large-scale structure cosmological datasets.
 The main advantage of this approach is the computational cost, which is comparable to that of estimating the power spectrum in the first place, and which scales with pixel resolution in a similar way.
@@
 The method, in all its generality, is currently implemented in the public code {\tt NaMaster}.
-Future extensions to this work will focus on improving the estimator for power spectra involving $B$ modes, and incorporating the impact of $E/B$-mode purification \todo{cites}.
+Future extensions to this work will focus on improving the estimator for power spectra involving $B$ modes, and incorporating the impact of $E/B$-mode purification \cite{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S}.
 

### 2019-06-21 get_grid/plot_grid.py

@@ -1,0 +1,74 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+import re, pickle
+
+
+l1s = 2.30
+l2s = 6.18
+
+
+files=np.array(['output/data_ref/chi2_Omega_c_sigma8_ng_100_100_lmin_0_nb_2_cov_sim.pickle',
+                'output/data_ref/chi2_Omega_c_sigma8_ng_100_100_lmin_0_nb_2_cov_th.pickle'])
+folders = np.array([os.path.abspath(os.path.dirname(x)) for x in files])
+sets = np.array([pickle.load(open(fn,'rb')) for fn in files])
+legend = np.array(['{}/{}'.format(folders[x],os.path.basename(sets[x]['grid_fn'])) for x in range(len(sets))])
+legend = np.array([x.replace('/','_') for x in legend])
+legend = np.array([x.replace('.','_') for x in legend])
+legend = np.array([x.split('_') for x in legend])
+legend = legend[:, ~np.all(legend[1:] == legend[:-1], axis=0)]
+legend = np.array(['_'.join(tuple(x)) for x in legend])
+
+
+def prepare_ranges(settings):
+    ranges = settings['range_m_M']
+    ng = settings['n_grid']
+    ranges = np.array([np.linspace(ranges[x,0],ranges[x,1],ng[x]) for x in range(len(ng))])
+    return ranges
+
+def prepare_grids(settings, folder):
+    fn = '{}/{}'.format(folder,os.path.basename(settings['grid_fn']))
+    grid = np.load(fn)['arr_0']
+    grid = grid - grid.min()
+    return grid
+
+
+ranges = np.array([prepare_ranges(s) for s in sets])
+
+grids = np.array([prepare_grids(sets[x],folders[x]) for x in range(len(sets))])
+
+def get_mean_width(grid,rangs):
+    p=np.exp(-0.5*grid)
+    p/=np.sum(p)
+    m1=np.sum(p*rangs[0][None,:])
+    m2=np.sum(p*rangs[1][:,None])
+    m12=np.sum(p*rangs[0][None,:]**2)
+    m22=np.sum(p*rangs[1][:,None]**2)
+    s1=np.sqrt(m12-m1**2)
+    s2=np.sqrt(m22-m2**2)
+    return [s1,s2]
[…]

### 2019-06-21 draft/paper.tex
more paper

@@
 The green line shows the power spectrum used to generate small-scale contaminants.
-The amplitude of the contaminat power spectra was fixed by imposing a $10\%$ contamination level in the power spectrum at $\ell=400$.
+The amplitudes of the contaminant power spectra were fixed by imposing a $10\%$ contamination level in the power spectrum at $\ell=400$.
 The same approach was used for the cosmic shear maps, with independent contaminants defined in $\gamma_E$ and $\gamma_B$.} \label{fig:clcont}
@@
         \includegraphics[width=\textwidth]{./figures/all_rows_sph_1bin.pdf}
-        \caption{Mode coupling of different $C_\ell$'s. The covariances have
-          been normalized by the MC approximation numerator (Eq.~\ref{eq:cov_naive_bpw}).}
+        \caption{Four rows of the covariance matrix for different power spectra.
+The top panels show the cases of spectra with non-zero signal, involving $\delta$ and $\gamma_E$, while the bottom panels show cases involving $B$-modes.
+In each panel we show rows of the covariance matrix for $\ell=60,\,90,\,120$ and $150$, which peak at those central values.
+The different lines show the results for the sample covariance matrix (solid blue), its NKA estimator (dashed orange) and the difference between both (dotted green).
+The black stars show the mode-counting approximation to the covariance matrix (Eq.
+\ref{eq:cov_naive_bpw}).
+To facilitate the visualization of the different rows, we have divided them by $C_\ell^{aa}C_\ell^{bb}+(C^{ab}_\ell)^2$, where $a$ and $b$ are the two fields being correlated in each case.
+The NKA method is able to recover the covariance with high accuracy for all field combinations that do not involve $B$ modes, yielding visibly poorer results otherwise.}
         \label{fig:rows_1bin}
@@
         \includegraphics[width=0.7\textwidth]{./figures/sph_1bin_diff_corr.pdf}
-        \caption{Difference between the correlation matrices associated to the sample covariance matrix and its estimate with the NKA method including all possible correlations between $\delta$ and $\boldsymbol{\gamma}$ in a single redshift bin.} \label{fig:diff_corr_1bin}
+        \caption{Difference between the correlation matrices associated to the sample covariance matrix and its estimate with the NKA method including all possible correlations between $\delta$ and $\boldsymbol{\gamma}$ in a single redshift bin.
+While the NKA estimator is able to recover the covariance matrix to high accuracy in most cases, it is not able to reproduce the off-diagonal correlations between different bandpowers in cases involving $B$ modes.} \label{fig:diff_corr_1bin}
       \end{figure}
@@
         \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_NKA_TTTEEE_reldev_eigval_1stbin.pdf}
-        \caption{Covariance matrix eigenvalues and the relative difference with respect to the eigenvalues from the simulations and NKA covariances.} \label{fig:eigv_1bin}
+        \caption{Eigenvalues of the single-bin covariance matrices for all power spectra involving $\delta$ and $\gamma_E$.
+Results are shown for the sample covariance (solid blue), the NKA estimator (solid orange) and the spin-0 approximation (dashed green).
+The NKA and spin-0 estimators are able to recover the covariance eigenvalues with an accuracy $\lesssim5\%$.} \label{fig:eigv_1bin}
       \end{figure}
@@
         \includegraphics[width=\textwidth]{./figures/run_sph_2b_1stbin_chi2_TT_TE_EE.pdf}
-        \caption{$\chi^2$ distributions for the cases without 
-          $\gamma_B$-modes. We compare the distributions obtained with the
-          covariance matrix computed directly from the simulated power spectra
-          and the two different analytical methods: NKA and Spin-0. In addition,
-          the theoretical $\chi^2$ distribution has been included. We observe
-          the simulations' distribution perfectly follows the former and
-          negligible differences between the approximations. It must be noted
-          their shift with respect to the true distribution when spin-2 fields
-          are considered. It will be shown in Section~\ref{ssec:results.cosmo}
-          it will have no effect on the cosmological predictions.}
+        \caption{$\chi^2$ distributions for the cases without $\gamma_B$-modes.
+We compare the distributions obtained with the covariance matrix computed directly from the simulated power spectra (blue) and the two different analytical methods: NKA (orange) and spin-0 (green).
+In addition, the theoretical $\chi^2$ distribution has been included (dashed red).
+The distribution extracted from the simulation follows the theoretical expectation almost perfectly in all cases.
+In the cases of $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$ we observe small shifts in the peak $\chi^2$, of $\sim2$-$4\%$, while they width is recovered accurately.
+We will show that these differences are due to the inaccuracy of the analytical approximations on the largest scales, and that they have a completely negligible impact on the final cosmological parameter constraints.}
         \label{fig:chi2_1bin} \end{figure}
@@
         \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_same_mask_Spin0_NKA_TTTEEE_Full_chi2.pdf}
-        \caption{Same as Figure \ref{fig:chi2_1bin} for the combination of all correlations between $\delta$ and $\gamma_E$ for two redshift bins.}
+        \caption{Same as Figure \ref{fig:chi2_1bin} for the combination of all correlations between $\delta$ and $\gamma_E$ for two redshift bins, where similar conclusions hold.}
         \label{fig:chi2_2bins}
@@
         \includegraphics[width=\textwidth]{./figures/contaminants_diag_cov.pdf}
-        \caption{Diagonal of the covariance matrix estimated from simulations
-          without contaminants, with 200 contaminants deprojected and the NKA
-          method. The differences between the simulated cases are much
-          smaller than those associated with the NKA estimator (shown in
-          green). The NKA method is therefore sufficient to approximate the
-          covariance even in the presence of contaminant deprojection.}
+        \caption{Diagonal of the covariance matrix estimated from simulations without contaminants (solid blue), with 200 contaminants deprojected (dashed orange) and the NKA estimator (solid green).
+The differences between the simulated cases are significantly smaller than those associated with the NKA estimator.
+The NKA method is therefore equally valid to approximate the covariance even in the presence of contaminant deprojection.}
         \label{fig:diag_conts}
       \end{figure}
-%      \begin{figure}
-%        \centering
-%        \includegraphics[width=0.6\textwidth]{./figures/contaminants_chi2.pdf}
-%        \caption{$\chi^2$ distributions for simulations with no systematic contaminants (blue), and for simulations containing contaminants after deprojection of 200 contaminant templates (orange).
-The covariance matrices associated to both sets of simulations are essentially the same, and the differences between them are much smaller than those associated with the NKA estimator (shown in green).
-The NKA method is therefore sufficient to approximate the covariance even in the presence of contaminant deprojection.}
-%        \label{fig:chi2_conts}
-%      \end{figure}
       In order to quantify the validity of our analytical approximations, we need to compare the NKA covariance with the sample covariance estimated from the simulations.
@@
 accurate enough).
-%      This is quantitatively shown in Fig.~\ref{fig:chi2_conts} through comparison of the $\chi^2$ distributions drawn with covariance matrices
-%      estimated with simulations without contaminants, simulations with
-%      contaminants deprojected and the NKA method.
 
@@
         \caption{1$\sigma$ and 2$\sigma$ contours for $\Omega_M$ and $\sigma_8$ found from two different estimates of the power spectrum likelihood for a dataset containing all possible cross-power spectra between $\delta$ and $\gamma_E$ for two redshift bins.
-The solid black and dashed red contours correspond to Gaussian likelihoods estimated using the sample covariance matrix and the covariance estimated using the NKA method respectively.
-We find a remarkably good agreement between both likelihoods, highlighting the negligible impact of the approximations involved in the NKA method on the final parameter estimates.
-\todo{finish this plot.}} \label{fig:contours}
+The filled contours corresponds to the parameter likelihood evaluated using the NKA estimator for the power spectrum covariance matrix, while the dashed black contours correspond to the same calculation done with the sample covariance matrix corrected by the factor in Eq.
+\ref{eq:hartlap}.
+In both cases we assume a Gaussian likelihood.
+We find a remarkably good agreement between both likelihoods, highlighting the negligible impact of the approximations involved in the NKA method on the final parameter estimates.} \label{fig:contours}
       \end{figure}
@@
 In most situations this can be done simply by rescaling the inverse covariance matrix by a factor  given by \cite{2007A&A...464..399H}
-      \begin{equation}
+      \begin{equation}\label{eq:hartlap}
         {\rm Cov}^{-1}\,\longrightarrow \frac{N_s-2-N_{\rm data}}{N_s-1}\,{\rm Cov}^{-1},
@@
       \includegraphics[width=\textwidth]{./figures/run_chi2_TT_TE_EE_TB_EB_BB.pdf}
-      \caption{$\chi^2$ distributions for the autocorrelations of the fields
-        of a redshift bin in the flat-sky regime. We compare the distributions
-        obtained with the covariance matrix estimated from simulated power
-        spectra, and the analytical NKA and Spin-0 methods. In addition, we
-        show the theoretical $\chi^2$ distribution, which is almost perfectly
-        followed by the others. The reason why the E and B modes are well
-        behaved in flat sky is because in this regime ($\ell \in (120,
-        17640)$), the shear power spectra is dominated by noise (see
-        Fig.~\ref{fig:cl-2bins}). }
+      \caption{Same as Figure \ref{fig:chi2_1bin} for flat-sky fields.
+This figure additionally shows the $\chi^2$ distributions for power spectra involving $B$-modes.
+For the smaller scales covered by these flat-sky simulations, the NKA and spin-0 estimators work remarkably well, even for $B$-mode spectra.}
       \label{fig:chi2_1bin_flat}

### 2019-06-21 notebooks/compute_covSpin0_sph_full.py
added script to compute full spin0 cov

@@ -1,0 +1,247 @@
+#!/usr/bin/python
+
+import pymaster as nmt
+import numpy as np
+import healpy as hp
+import os
+import sys
+
+#pylint: disable=C0103
+
+##############################################################################
+##############################################################################
+##############################################################################
+
+# prefix = 'run_sph_2b_same_mask'
+prefix = os.path.basename(sys.argv[1].strip('/'))
+run_path = os.path.join('./simulations_outputs/', prefix, prefix)
+output_dir_NKA = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
+output_dir = os.path.join('./simulations_outputs/', prefix, 'full_covariance_spin0')
+
+if not os.path.isdir(output_dir):
+    os.makedirs(output_dir)
+out_run_path_NKA = os.path.join(output_dir_NKA, prefix)
+out_run_path = os.path.join(output_dir, prefix)
+data_folder = './data/'
+
+nside = 512
+
+##############################################################################
+# Load theoretical cls and masks
+##############################################################################
+
+f = np.load(os.path.join(data_folder, "cls_lss_2bins.npz"))
+lTh = f['ls']
+clTh = (f['cls'] + f['nls'])
+clTh_all = (f['cls'] + f['nls'])[np.triu_indices(f['cls'].shape[0])]
+
+fmasks = ["data/mask_lss_sph1.fits", "data/mask_lss_sph1.fits"]
+
+mask_lss_ar = []
+
+for fmask in fmasks:
+    mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=nside)
+    mask_lss_ar.append(mask_lss)
+
+#Set up binning scheme
+fsky = np.mean(np.product(mask_lss_ar, axis=0))
+d_ell = int(1./fsky)
+b = nmt.NmtBin(nside, nlb=d_ell)
+
+if not os.path.isfile(run_path + '_ells.txt'):
+    np.savetxt(run_path + '_ells.txt', b.get_effective_ells())
+
[…]

### 2019-06-21 notebooks/join_covsTh_2bins.py
implemented arbitrary input

@@ -1,6 +1,7 @@
 #!/usr/bin/python
 import numpy as np
 import os
+import sys
 
 # pylint: disable=C0103
 
@@ -8,7 +9,8 @@ import os
 ##############################################################################
 ##############################################################################
 
-prefix = 'run_sph_2b_same_mask'
+# prefix = 'run_sph_2b_same_mask'
+prefix = os.path.basename(sys.argv[1].strip('/'))
 run_path = os.path.join('./simulations_outputs/', prefix, prefix)
 output_dir = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
 out_run_path = os.path.join(output_dir, prefix)

### 2019-06-21 notebooks/join_covsTh_2bins_spin0.py

@@ -1,6 +1,7 @@
 #!/usr/bin/python
 import numpy as np
 import os
+import sys
 
 # pylint: disable=C0103
 
@@ -8,7 +9,8 @@ import os
 ##############################################################################
 ##############################################################################
 
-prefix = 'run_sph_2b_same_mask'
+# prefix = 'run_sph_2b_same_mask'
+prefix = os.path.basename(sys.argv[1].strip('/'))
 run_path = os.path.join('./simulations_outputs/', prefix, prefix)
 output_dir = os.path.join('./simulations_outputs/', prefix, 'full_covariance_spin0')
 out_run_path = os.path.join(output_dir, prefix)

### 2019-06-21 notebooks/general-plots-paper.py
generalized to use 2b diff masks

@@ -809,10 +809,15 @@ def chi2_flat_TT_TE_EE_TB_EB_BB():
 ############################ Chi2 TTTEEE Full ################################
 ##############################################################################
 
-def chi2_NKA_TTTEEE_full():
-    chi2_list = np.load('../run_sph_2b_same_mask/full_covariance/run_sph_2b_same_mask_chi2_sims_th_TTTEEE_short_Full.npz')['arr_0']
+def chi2_NKA_TTTEEE_full(same_mask=False):
+    suffix = ''
+    if same_mask:
+        suffix = '_same_mask'
 
-    ell = np.loadtxt('../run_sph_2b_same_mask/run_sph_2b_same_mask_ells.txt')
+    chi2_list = np.load('../run_sph_2b{0}/full_covariance/run_sph_2b{0}_chi2_sims_th_TTTEEE_short_Full.npz'.format(suffix))['arr_0']
+
+    ell = np.loadtxt('../run_sph_2b{0}/run_sph_2b{0}_ells.txt'.format(suffix))
+
     lbins = ell
     lmax = (ell < 2*512).sum()
 
@@ -846,16 +851,20 @@ def chi2_NKA_TTTEEE_full():
 
     ax.legend(loc='upper right', fontsize='9', frameon=False)
     plt.tight_layout()
-    fname = os.path.join(outdir, 'run_sph_2b_same_mask_NKA_TTTEEE_Full_chi2.pdf')
+    fname = os.path.join(outdir, 'run_sph_2b{}_NKA_TTTEEE_Full_chi2.pdf'.format(suffix))
     plt.savefig(fname, dpi=DPI)
     # plt.show()
     plt.close()
 
-def chi2_Spin0_NKA_TTTEEE_full():
-    chi2_list = np.load('../run_sph_2b_same_mask/full_covariance/run_sph_2b_same_mask_chi2_sims_th_TTTEEE_short_Full.npz')['arr_0']
-    chi2_spin0 = np.load('../run_sph_2b_same_mask/full_covariance_spin0/run_sph_2b_same_mask_chi2_sims_th_TTTEEE_short_Full_spin0.npz')['arr_0'][1]
+def chi2_Spin0_NKA_TTTEEE_full(same_mask=False):
+    suffix = ''
+    if same_mask:
+        suffix = '_same_mask'
 
[…]

### 2019-06-21 notebooks/run_sph_2b_do_all_checks_full_cov_general.py

@@ -2,10 +2,12 @@
 import common as co
 import numpy as np
 import os
+import sys
 
 # pylint: disable=C0103
 
-prefix = 'run_sph_2b_same_mask'
+# prefix = 'run_sph_2b_same_mask'
+prefix = os.path.basename(sys.argv[1].strip('/'))
 run_path = os.path.join('./simulations_outputs/', prefix, prefix)
 outdir = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
 output_path = os.path.join(outdir, prefix)

### 2019-06-21 notebooks/run_sph_2b_do_all_checks_full_cov_general_spin0.py

@@ -2,10 +2,12 @@
 import common as co
 import numpy as np
 import os
+import sys
 
 # pylint: disable=C0103
 
-prefix = 'run_sph_2b_same_mask'
+# prefix = 'run_sph_2b_same_mask'
+prefix = os.path.basename(sys.argv[1].strip('/'))
 run_path = os.path.join('./simulations_outputs/', prefix, prefix)
 
 outdir = os.path.join('./simulations_outputs/', prefix, 'full_covariance_spin0')

### 2019-06-23 notebooks/compute_covTh_sph_full.py
added script to compute the NKA cov matrix for 2bin case

@@ -1,0 +1,247 @@
+#!/usr/bin/python
+
+import pymaster as nmt
+import numpy as np
+import healpy as hp
+import os
+import sys
+
+#pylint: disable=C0103
+
+##############################################################################
+##############################################################################
+##############################################################################
+
+# prefix = 'run_sph_2b_same_mask'
+prefix = os.path.basename(sys.argv[1].strip('/'))
+run_path = os.path.join('./simulations_outputs/', prefix, prefix)
+output_dir = os.path.join('./simulations_outputs/', prefix, 'full_covariance')
+
+if not os.path.isdir(output_dir):
+    os.makedirs(output_dir)
+out_run_path = os.path.join(output_dir, prefix)
+data_folder = './data/'
+
+nside = 512
+
+##############################################################################
+# Load theoretical cls and masks
+##############################################################################
+
+f = np.load(os.path.join(data_folder, "cls_lss_2bins.npz"))
+lTh = f['ls']
+clTh = (f['cls'] + f['nls'])
+clTh_all = (f['cls'] + f['nls'])[np.triu_indices(f['cls'].shape[0])]
+
+fmasks = ["data/mask_lss_sph1.fits", "data/mask_lss_sph1.fits"]
+
+mask_lss_ar = []
+
+for fmask in fmasks:
+    mask_lss = hp.ud_grade(hp.read_map(fmask, verbose=False), nside_out=nside)
+    mask_lss_ar.append(mask_lss)
+
+#Set up binning scheme
+fsky = np.mean(np.product(mask_lss_ar, axis=0))
+d_ell = int(1./fsky)
+b = nmt.NmtBin(nside, nlb=d_ell)
+
+if not os.path.isfile(run_path + '_ells.txt'):
+    np.savetxt(run_path + '_ells.txt', b.get_effective_ells())
+
+##############################################################################
+# Generate field to compute the covariance workspaces
[…]

### 2019-06-23 draft/paper.tex

@@
 \documentclass[a4paper,11pt]{article}
-
 \pdfoutput=1
@@
 \newcommand{\wtj}[6]{\left(\begin{array}{ccc} #1 & #2 & #3\\#4 & #5 & #6\end{array} \right)}
-
-\newcommand{\todo}[1]{{\bf TODO: #1}}
-\newcommand{\red}[1]{{\color{red} #1}}
 
@@
         Making use of the completeness relation for the Wigner 3-j symbols, it is possible to show, that these can be written as:
-        \begin{align}
+        \begin{align}\label{eq:coeff_cmcm}
           &\,^{af}_{bg}\mathcal{J}^{00}_{\ell\ell'}=\Xi^{00}_{\ell\ell'}(w_aw_f,w_bw_g),\hspace{12pt}
@@
         and, as before
-        \begin{align}
+        \begin{align}\label{eq:coeff_cmcm_flat}
           &\,^{af}_{bg}\bar{\mathcal{J}}^{00}_{{\bf l}{\bf l}'}=\bar{\Xi}^{00}_{{\bf l}{\bf l}'}(w_aw_f,w_bw_g),\hspace{12pt}
@@
   \section{Discussion}\label{sec:discussion}
-    Estimating accurate covariance matrices for projected two-point correlators is an ubiquitous problem in modern cosmology \todo{cites}, but it is particularly relevant for large-scale structure datasets.
+    Estimating accurate covariance matrices for projected two-point correlators is an ubiquitous problem in modern cosmology \cite{2013PhRvD..88f3537D,2013MNRAS.432.1928T,2017arXiv170609359K}, but it is particularly relevant for large-scale structure datasets.
 This problem is further complicated in this case, in comparison with e.g.
@@
       \includegraphics[width=0.6\textwidth]{./figures/mask-lss_flat1.pdf}
-      \caption{Flat-sky mask used in the study of the analytical
-        methods presented in this work in the flat-sky regime. The mask was
-        constructed from the bright-object mask distributed with the first
-        data release of the HSC collaboration \todo{cite} for the VVDS
-        field. As in the full-sky case, it is used both for galaxy clustering
-        and shear for simplicity.} \label{fig:mask_flat}
+      \caption{Flat-sky mask used in the study of the analytical methods presented in this work in the flat-sky regime.
+The mask was constructed from the bright-object mask distributed with the first data release of the HSC collaboration \cite{2018PASJ...70S...4A} for the VVDS field.
+As in the full-sky case, it is used both for galaxy clustering and shear for simplicity.} \label{fig:mask_flat}
     \end{figure}
@@
 We generate Gaussian realizations of the galaxy overdensity and shear maps making use of flat-sky extensions of the methods described in Section \ref{ssec:results.sims}.
-In this case we use a high-resolution mask constructed from the bright-object mask distributed with the first data release of the HSC collaboration \todo{cite} for the VVDS field, which is shown in Fig.
+In this case we use a high-resolution mask constructed from the bright-object mask distributed with the first data release of the HSC collaboration \cite{2018PASJ...70S...4A} for the VVDS field, which is shown in Fig.
 \ref{fig:mask_flat}.
@@
   \section{Software implementation}\label{app:namaster}
-    \todo{complete}
+    In addition to the code functionality described in Section 3 of \cite{2019MNRAS.484.4127A}, we have now included the capability to estimate Gaussian covariance matrices using the NKA method.
+This functionality is structured around a {\tt python} class called {\tt NmtCovarianceWorkspace}.
+These objects are used to compute and store the covariance mode-coupling coefficients in Eqs.
+\ref{eq:coeff_cmcm} and \ref{eq:coeff_cmcm_flat}.
+They are initialized from two pairs of fields corresponding to the two power spectra for which the covariance is required.
+Once initialized, these coefficients can be reused for any other set of fields with the same combination of sky masks.
+{\tt NaMaster} then provides routines to estimate covariance matrix elements making use of the coupling coefficients stored in a {\tt NmtCovarianceWorkspace} object and best-guess power spectra for the fields involved using Eqs.
+\ref{eq:cov_sph} and \ref{eq:cov_flat}.
+Further details about the implementation and practical examples can be found in \url{https://namaster.readthedocs.io/en/latest/sample_covariance.html}.
 

### 2019-06-23 notebooks/compute_covSpin0_sph_full.py
fixed not changed code for same_mask

@@ -35,7 +35,10 @@ lTh = f['ls']
 clTh = (f['cls'] + f['nls'])
 clTh_all = (f['cls'] + f['nls'])[np.triu_indices(f['cls'].shape[0])]
 
-fmasks = ["data/mask_lss_sph1.fits", "data/mask_lss_sph1.fits"]
+if 'same_mask' in prefix:
+    fmasks = ["data/mask_lss_sph1.fits", "data/mask_lss_sph1.fits"]
+else:
+    fmasks = ["data/mask_lss_sph1.fits", "data/mask_lss_sph2.fits"]
 
 mask_lss_ar = []
 

### 2019-06-23 notebooks/compute_covTh_sph_full.py

@@ -33,7 +33,11 @@ lTh = f['ls']
 clTh = (f['cls'] + f['nls'])
 clTh_all = (f['cls'] + f['nls'])[np.triu_indices(f['cls'].shape[0])]
 
-fmasks = ["data/mask_lss_sph1.fits", "data/mask_lss_sph1.fits"]
+if 'same_mask' in prefix:
+    fmasks = ["data/mask_lss_sph1.fits", "data/mask_lss_sph1.fits"]
+else:
+    fmasks = ["data/mask_lss_sph1.fits", "data/mask_lss_sph2.fits"]
+
 
 mask_lss_ar = []
 

### 2019-06-23 notebooks/general-plots-paper.py

@@ -950,18 +950,18 @@ def eigv_sph_1bin():
     plt.close()
 
 if __name__ == '__main__':
-    pz()
-    masks()
-    foregrounds()
-    chi2_foregrounds()
-    # corr_diff_2bins_same_mask()
-    # compare_cls_1bin()
-    cls_2b()
-    chi2_sph_TT_TE_EE()
-    chi2_flat_TT_TE_EE_TB_EB_BB()
-    chi2_NKA_TTTEEE_full()
-    eigv_sph_1bin()
+    # pz()
+    # masks()
+    # foregrounds()
+    # chi2_foregrounds()
+    # # corr_diff_2bins_same_mask()
+    # # compare_cls_1bin()
+    # cls_2b()
+    # chi2_sph_TT_TE_EE()
+    # chi2_flat_TT_TE_EE_TB_EB_BB()
+    # chi2_NKA_TTTEEE_full()
+    # eigv_sph_1bin()
     chi2_Spin0_NKA_TTTEEE_full()
-    foregrounds_cov_diag()
-    corr_diff_2bins_diff_mask()
-    chi2_sph_1b_chi2_mean_cov_diff()
+    # foregrounds_cov_diag()
+    # corr_diff_2bins_diff_mask()
+    # chi2_sph_1b_chi2_mean_cov_diff()

### 2019-06-23 notebooks/join_covsTh_2bins.py

@@ -91,7 +91,7 @@ cov_mat[i, :, j, :] = cov_arr
 cov_mat[j, :, i, :] = cov_arr.swapaxes(1, 2)
 cov_mat = cov_mat.reshape((len(cl_bins) * l_bpw, len(cl_bins) * l_bpw))
 
-fname = out_run_path + '_covTh_TTTEEE_2bins_same_mask.npz'
+fname = out_run_path + '_covTh_TTTEEE_2bins.npz'
 np.savez_compressed(fname, cov_mat)
 
 ####################### Short version #############################
@@ -102,7 +102,7 @@ cov_mat[i, :, j, :] = cov_arr[:, :lmax, :lmax]
 cov_mat[j, :, i, :] = cov_arr[:, :lmax, :lmax].swapaxes(1, 2)
 cov_mat = cov_mat.reshape((len(cl_bins) * lmax, len(cl_bins) * lmax))
 
-fname = out_run_path + '_covTh_TTTEEE_short_2bins_same_mask.npz'
+fname = out_run_path + '_covTh_TTTEEE_short_2bins.npz'
 np.savez_compressed(fname, cov_mat)
 
 # cov_mat = np.empty((len(cl_bins), len(cl_bins), lmax, lmax))

### 2019-06-23 notebooks/run_sph_2b_do_all_checks_full_cov_general.py
removed same_mask from filename

@@ -21,7 +21,7 @@ ells = np.loadtxt(run_path + '_ells.txt')
 lmax = (ells < 2*512).sum()
 
 CovSims_path = output_path + '_covSims_TTTEEE_short_0001-20000.npz'
-CovTh_path = output_path + '_covTh_TTTEEE_short_2bins_same_mask.npz'
+CovTh_path = output_path + '_covTh_TTTEEE_short_2bins.npz'
 
 CovSims_Full = np.load(CovSims_path)['arr_0']
 CovTh_Full = np.load(CovTh_path)['arr_0']

### 2019-06-23 notebooks/run_sph_2b_do_all_checks_full_cov_general_spin0.py

@@ -25,7 +25,7 @@ ells = np.loadtxt(run_path + '_ells.txt')
 lmax = (ells < 2*512).sum()
 
 CovSims_path = output_path_NKA + '_covSims_TTTEEE_short_0001-20000.npz'
-CovTh_path = output_path + '_covTh_TTTEEE_short_2bins_same_mask_spin0.npz'
+CovTh_path = output_path + '_covTh_TTTEEE_short_2bins_spin0.npz'
 
 CovSims_Full = np.load(CovSims_path)['arr_0']
 CovTh_Full = np.load(CovTh_path)['arr_0']

### 2019-06-23 notebooks/join_covsTh_2bins_spin0.py
removed same_mask from filename

@@ -71,7 +71,7 @@ cov_mat[i, :, j, :] = cov_arr
 cov_mat[j, :, i, :] = cov_arr.swapaxes(1, 2)
 cov_mat = cov_mat.reshape((len(cl_bins) * l_bpw, len(cl_bins) * l_bpw))
 
-fname = out_run_path + '_covTh_TTTEEE_2bins_same_mask_spin0.npz'
+fname = out_run_path + '_covTh_TTTEEE_2bins_spin0.npz'
 np.savez_compressed(fname, cov_mat)
 
 ####################### Short version #############################
@@ -82,7 +82,7 @@ cov_mat[i, :, j, :] = cov_arr[:, :lmax, :lmax]
 cov_mat[j, :, i, :] = cov_arr[:, :lmax, :lmax].swapaxes(1, 2)
 cov_mat = cov_mat.reshape((len(cl_bins) * lmax, len(cl_bins) * lmax))
 
-fname = out_run_path + '_covTh_TTTEEE_short_2bins_same_mask_spin0.npz'
+fname = out_run_path + '_covTh_TTTEEE_short_2bins_spin0.npz'
 np.savez_compressed(fname, cov_mat)
 
 # cov_mat = np.empty((len(cl_bins), len(cl_bins), lmax, lmax))

### 2019-06-23 notebooks/general-plots-paper.py
added chi2 for 2bins diff mask

@@ -419,6 +419,73 @@ def corr_diff_2bins_same_mask():
     plt.close()
 
 ##############################################################################
+###################### 2bins corr diff (diff mask) ###########################
+##############################################################################
+
+def corr_diff_2bins_diff_mask2():
+    Cth = np.load('../run_sph_2b/full_covariance/run_sph_2b_covTh_TTTEEE_short_2bins.npz')['arr_0']
+    Csims = np.load('../run_sph_2b/full_covariance/run_sph_2b_covSims_TTTEEE_short_0001-20000.npz')['arr_0']
+
+    ell = np.loadtxt('../run_sph_2b/run_sph_2b_ells.txt')
+
+    CorrSims = Csims/np.sqrt(np.diag(Csims)[:, None] * np.diag(Csims)[None, :])
+    CorrTh = Cth/np.sqrt(np.diag(Cth)[:, None] * np.diag(Cth)[None, :])
+
+
+    labels = [r'$\delta_1$', r'$\gamma_{E,1}$', r'$\delta_2$', r'$\gamma_{E,2}$']
+    labels_cov = []
+    for i in range(4):
+        for j in range(i, 4):
+            labels_cov.append(labels[i] + labels[j])
+
+
+    f, ax = plt.subplots(1, 1, figsize=(8, 8))
+    nlbins = int(CorrSims.shape[0] / 10)
+    cb = ax.imshow(CorrTh - CorrSims, vmin=-0.02, vmax=0.02)
+    divider = make_axes_locatable(ax)
+    cax = divider.append_axes("right", size="5%", pad=0.05)
+    f.colorbar(cb, cax=cax)
+    for i in range(1, 10):
+        ax.plot([0, 10*nlbins-1], [i*nlbins, i*nlbins], 'k-', lw=0.5)
+        ax.plot([i*nlbins, i*nlbins], [0, 10*nlbins-1], 'k-', lw=0.5)
+
+    ticks = [(i + 0.5) * nlbins for i in range(10)]
+    ax.set_xticks(ticks)
+    ax.set_xticklabels(labels_cov)
+    ax.set_yticks(ticks)
+    ax.set_yticklabels(labels_cov)
+
+    fname = os.path.join(outdir, 'run_sph_2b_NKA_diff_corr.pdf')
+    plt.savefig(fname, dpi=DPI)
+    # plt.show()
[…]

### 2019-06-24 notebooks/general-plots-paper.py
removed same_mask = True as default

@@ -1028,7 +1028,7 @@ if __name__ == '__main__':
     chi2_flat_TT_TE_EE_TB_EB_BB()
     chi2_NKA_TTTEEE_full()
     eigv_sph_1bin()
-    chi2_Spin0_NKA_TTTEEE_full(True)
+    chi2_Spin0_NKA_TTTEEE_full()
     foregrounds_cov_diag()
     corr_diff_2bins_diff_mask()
     chi2_sph_1b_chi2_mean_cov_diff()

### 2019-06-24 notebooks/general-plots-paper.py
output more info

@@ -938,17 +938,23 @@ def chi2_Spin0_NKA_TTTEEE_full(same_mask=False):
     means = np.concatenate([np.mean(chi2_list, axis=1), [np.mean(chi2_spin0)]])
     covs = np.array([np.sqrt(np.cov(chi2is)) for chi2is in chi2_list] + [np.sqrt(np.cov(chi2_spin0))])
     print("Chi2_Spin0_NKA_TTTEEE_full:")
-    print("Results for | NKA | Spin0 |")
-    print("Relative mean diff: {}".format(means[1:]/means[0] - 1))
-    print("Relative width diff: {}".format(covs[1:]/covs[0] - 1))
+    print("Results for Sims | NKA | Spin0 |")
+    print("Mean : {}".format(means))
+    print("Width : {}".format(covs))
+    print("Relative mean diff: {}".format(means/means[0] - 1))
+    print("Relative mean diff: {}".format(means/means[0] - 1))
+    print("Relative width diff: {}".format(covs/covs[0] - 1))
     Ns = chi2_list.shape[1]
     Nl = lmax * 10
     H = (Ns - Nl - 2)/(Ns - 1) # arXiv:1601.05786
     print('Ns =',  Ns, 'Nl =', Nl, 'Hartlap =', H)
     # covs = np.array([np.sqrt(np.cov(H * chi2_list[0])), np.sqrt(np.cov(chi2_list[1])), np.sqrt(np.cov(chi2_spin0))])
-    print("Relative mean diff (Hartlap): {}".format(means[1:]/(H * means[0]) - 1))
-    print("Relative width diff (Hartlap): {}".format(covs[1:]/(H * covs[0]) - 1))
+    print("Mean (Hartlap): {}".format([H, 1, 1] * means))
+    print("Width (Hartlap): {}".format([H, 1, 1] * covs))
+    print("Relative mean diff (Hartlap): 0, {}".format(means[1:]/(H * means[0]) - 1))
+    print("Relative width diff (Hartlap): 0, {}".format(covs[1:]/(H * covs[0]) - 1))
 
+    return
     f, ax = plt.subplots(1, 1, figsize=FSIZE1)
     bins = np.linspace(np.min(chi2_list), np.max(chi2_list), 60)
     _, x, _ = ax.hist(chi2_list[0], bins=bins, histtype='step', density=True,

### 2019-06-24 notebooks/general-plots-paper.py
plot rel dev mean & width chi2 flat

@@ -872,6 +872,55 @@ def chi2_flat_TT_TE_EE_TB_EB_BB():
     # plt.show()
     plt.close()
 
+    ############## Plot mean & cov rel dev #################
+
+    rdmeans = np.empty((6, chi2_TT_ar.shape[0]))
+    rdcovs = np.empty((6, chi2_TT_ar.shape[0]))
+    for i, chi2_list in enumerate([chi2_TT_ar, chi2_TE_ar, chi2_EE_ar, chi2_TB_ar, chi2_EB_ar, chi2_BB_ar]):
+        means = np.mean(chi2_list, axis=1)
+        covs = np.array([np.sqrt(np.cov(chi2is)) for chi2is in chi2_list])
+
+        rdmeans[i] = (means / means[0] - 1)
+        rdcovs[i] = (covs / covs[0] - 1)
+
+
+
+    label = [r"$(\delta \delta)$", r"$(\delta \gamma_E)$", r"$(\gamma_E \gamma_E)$",
+             r"$(\delta \gamma_B)$",
+             r"$(\gamma_E \gamma_B)$",
+             r"$(\gamma_B \gamma_B)$"]
+
+
+    f, axs = plt.subplots(2, 1, figsize=(4, 4), sharex=True, gridspec_kw={'hspace': 0, 'wspace': 0})
+
+    X = np.arange(rdmeans.shape[0])
+    # axs[0].plot(X, np.abs(rdmeans[:, 0]), label='Sims.')
+    axs[0].plot(X, 100 * np.abs(rdmeans[:, 1]), c=DEFAULT_COLOR_CYCLE[1], label='NKA')
+    axs[0].plot(X, 100 * np.abs(rdmeans[:, 2]), c=DEFAULT_COLOR_CYCLE[2], label='Spin-0')
+
+    # axs[1].plot(X, np.abs(rdcovs[:, 0]), label='Sims.')
+    axs[1].plot(X, 100 * np.abs(rdcovs[:, 1]), c=DEFAULT_COLOR_CYCLE[1], label='NKA')
+    axs[1].plot(X, 100 * np.abs(rdcovs[:, 2]), c=DEFAULT_COLOR_CYCLE[2], label='Spin-0')
+
+    # axs[0].set_yscale('log')
+    # axs[1].set_yscale('log')
+
[…]

### 2019-06-24 draft/paper.tex
added plots in paper + new Fig for flat sky chi2 reldev mean & width

@@
         \centering
-        \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_same_mask_Spin0_NKA_TTTEEE_Full_chi2.pdf}
+        \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_1stbin_chi2_mean_cov_diff.pdf}
+        \caption{$\chi^2$ distributions' mean and width relative difference
+          between that obtained with a covariance matrix estimated from
+          analytical methods (NKA and Spin-0) and simulations. This is for the
+          case with 1 bin.}
+        \label{fig:chi2_1bin_diffs}
+      \end{figure}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_Spin0_NKA_TTTEEE_Full_chi2.pdf}
         \caption{Same as Figure \ref{fig:chi2_1bin} for the combination of all correlations between $\delta$ and $\gamma_E$ for two redshift bins, where similar conclusions hold.}
@@
       \end{figure}
+      \begin{table}
+        \centering
+        \begin{tabular}{|l|c|c|}
+          \hline
+              & NKA & Spin-0 \\
+          \hline
+          Rel. Dev. Mean [\%]  & -2.8 & -2.8 \\ 
+          Rel. Dev. Width [\%] & 6.5  & 6.9  \\
+          \hline
+        \end{tabular}
+        \caption{$\chi^2$ mean and width relative difference between that
+          obtained with a covariance matrix estimated from simulations and
+          the analytical methods NKA and Spin-0.} 
+        \label{tab:chi2_2bins_diffs}
+      \end{table}
       \begin{figure}
@@
 In the cases involving the spin-2 fields, we see noticeable differences between the distributions found with the sample covariance and the approximate ones.
-      These differences are small, corresponding to less of a $2$ and $4\%$ shift in the mean $\chi^2$ for the $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ cases respectively, and a negligible variation in the width of the distributions.
-We therefore expect these differents to have a negligible effect on the posterior parameter distributions, as we show explicitly in Section \ref{ssec:results.cosmo}.
-The fact that these differences appear only for power spectra involving spin-2 quantities indicate that the NKA and spin-0 methods are imperfect at describing the additional mode coupling caused in the presence of a mask for higher-spin fields.
-This is rather obvious in the case of the spin-0 approximation but, interestingly, we find that the NKA and Spin-0 predictions yield results that are almost indistinguishable from each other.
-We therefore conclude that the additional approximation made in the NKA method for spin-2 fields -- neglecting the spatial derivatives of the mask -- is effectively equivalent to ignoring the spin nature of the fields involved.
-For completeness, Figure \ref{fig:chi2_2bins} shows the distribution of $\chi^2$ values for a data vector composed of all possible auto- and cross-correlations of $\delta$ and $\gamma_E$ for the case of two redshift bins, where similar conclusions hold.
+      These differences are small, corresponding to less of a $2$ and $4\%$
+      shift in the mean $\chi^2$ for the $\delta$-$\gamma_E$ and
+      $\gamma_E$-$\gamma_E$ cases respectively, and $\lesssim 2\%$ variation
+      in the width of the distributions (see
+      Figure~\ref{fig:chi2_1bin_diffs}).  We therefore expect these
+      differences to have a negligible effect on the posterior parameter
+      distributions, as we show explicitly in Section
+      \ref{ssec:results.cosmo}. The fact that these differences appear only
+      for power spectra involving spin-2 quantities indicate that the NKA and
+      spin-0 methods are imperfect at describing the additional mode coupling
+      caused in the presence of a mask for higher-spin fields. This is rather
+      obvious in the case of the spin-0 approximation but, interestingly, we
+      find that the NKA and Spin-0 predictions yield results that are almost
+      indistinguishable from each other. We therefore conclude that the
+      additional approximation made in the NKA method for spin-2 fields --
+      neglecting the spatial derivatives of the mask -- is effectively
+      equivalent to ignoring the spin nature of the fields involved. For
+      completeness, Figure \ref{fig:chi2_2bins} shows the distribution of
+      $\chi^2$ values for a data vector composed of all possible auto- and
+      cross-correlations of $\delta$ and $\gamma_E$ for the case of two
+      redshift bins, where similar conclusions hold.
 
@@
     \end{figure}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.6\textwidth]{./figures/run_chi2_mean_cov_diff.pdf}
+        \caption{$\chi^2$ distributions' mean and width relative difference
+          between that obtained with a covariance matrix estimated from
+          analytical methods (NKA and Spin-0) and simulations. This is for the
+          case with 1 bin in the flat-sky approximation.}
+        \label{fig:chi2_1bin_flat_diffs}
+      \end{figure}
     We have repeated the analysis described in Section \ref{sec:results} on flat-sky realizations, making use of the flat-sky implementation of {\tt NaMaster}.
@@
 
-    We find similar levels of accuracy in the NKA and spin-0 estimators compared to the curved-sky case.
-The higher resolution and smaller area of these simulations allow us to focus on the small-scale galaxy clustering and lensing power spectra, covering the range of multipoles $\ell\in(120,17640)$ in constant bandpowers of width $\Delta \ell=240$.
-On these small scales, the shear power spectrum is more dominated by noise (e.g.
-see Fig.
-\ref{fig:cl-2bins}), and therefore there is roughly the same power in $E$ and $B$ modes.
-This reduces the sensitivity of the method to an inaccurate treatment of $E/B$ leakage, and the agreement between the sample covariance matrix and the NKA estimator improves significantly.
-This can be seen in Fig.
-\ref{fig:chi2_1bin_flat}, which shows the $\chi^2$ distributions for all possible power spectra (including those including $B$ modes) in the case of a single redshift bin.
-In all cases we find a good agreement between the $\chi^2$ distributions derived from all covariance matrix estimates (sample covariance, NKA and spin-0 in blue, orange and green respectively), which also accurately follow the expected $\chi^2$ distribution for the corresponding number of degrees of freedom (red dashed line).
-
+    We find similar levels of accuracy in the NKA and spin-0 estimators
+    compared to the curved-sky case. The higher resolution and smaller area of
+    these simulations allow us to focus on the small-scale galaxy clustering
+    and lensing power spectra, covering the range of multipoles
+    $\ell\in(120,17640)$ in constant bandpowers of width $\Delta \ell=240$. On
+    these small scales, the shear power spectrum is more dominated by noise
+    (e.g. see Fig. \ref{fig:cl-2bins}), and therefore there is roughly the
+    same power in $E$ and $B$ modes. This reduces the sensitivity of the
+    method to an inaccurate treatment of $E/B$ leakage, and the agreement
+    between the sample covariance matrix and the NKA estimator improves
+    significantly. This can be seen in Figs.~\ref{fig:chi2_1bin_flat} and
+    \ref{fig:chi2_1bin_flat_diffs}, which
+    shows the $\chi^2$ distributions for all possible power spectra (including
+    those including $B$ modes) in the case of a single redshift bin. In all
+    cases we find a good agreement (difference $\lesssim 2 \%$) between the
+    $\chi^2$ distributions derived from all covariance matrix estimates
+    (sample covariance, NKA and spin-0 in blue, orange and green
+    respectively), which also accurately follow the expected $\chi^2$
+    distribution for the corresponding number of degrees of freedom (red
+    dashed line).     
   \section{Software implementation}\label{app:namaster}

### 2019-06-24 notebooks/general-plots-paper.py
removed duplicated line

@@ -991,7 +991,6 @@ def chi2_Spin0_NKA_TTTEEE_full(same_mask=False):
     print("Mean : {}".format(means))
     print("Width : {}".format(covs))
     print("Relative mean diff: {}".format(means/means[0] - 1))
-    print("Relative mean diff: {}".format(means/means[0] - 1))
     print("Relative width diff: {}".format(covs/covs[0] - 1))
     Ns = chi2_list.shape[1]
     Nl = lmax * 10

### 2019-06-24 draft/paper.tex
fixed decimal
reverted addition plots to David version + use 2bin diff mask chi2 full plot + added relative differences of width and mean.

--- 2019-06-23 draft/paper.tex
@@
         \includegraphics[width=\textwidth]{./figures/run_sph_2b_1stbin_chi2_TT_TE_EE.pdf}
-        \caption{$\chi^2$ distributions for the cases without $\gamma_B$-modes.
-We compare the distributions obtained with the covariance matrix computed directly from the simulated power spectra (blue) and the two different analytical methods: NKA (orange) and spin-0 (green).
-In addition, the theoretical $\chi^2$ distribution has been included (dashed red).
-The distribution extracted from the simulation follows the theoretical expectation almost perfectly in all cases.
-In the cases of $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$ we observe small shifts in the peak $\chi^2$, of $\sim2$-$4\%$, while they width is recovered accurately.
-We will show that these differences are due to the inaccuracy of the analytical approximations on the largest scales, and that they have a completely negligible impact on the final cosmological parameter constraints.}
-        \label{fig:chi2_1bin} \end{figure}
+        \caption{$\chi^2$ distributions for the cases without
+          $\gamma_B$-modes. We compare the distributions obtained with the
+          covariance matrix computed directly from the simulated power spectra
+          (blue) and the two different analytical methods: NKA (orange) and
+          spin-0 (green). In addition, the theoretical $\chi^2$ distribution
+          has been included (dashed red). The distribution extracted from the
+          simulation follows the theoretical expectation almost perfectly in
+          all cases. In the cases of $\delta\times\gamma_E$ and
+          $\gamma_E\times\gamma_E$ we observe small shifts in the peak
+          $\chi^2$, of $\sim2$-$4\%$, while their width is recovered
+          accurately (a difference $\lesssim 2\%$). We will show that these
+          differences are due to the inaccuracy of the analytical
+          approximations on the largest scales, and that they have a
+          completely negligible impact on the final cosmological parameter
+          constraints.} 
+        \label{fig:chi2_1bin}
+      \end{figure}
       \begin{figure}
         \centering
-        \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_same_mask_Spin0_NKA_TTTEEE_Full_chi2.pdf}
-        \caption{Same as Figure \ref{fig:chi2_1bin} for the combination of all correlations between $\delta$ and $\gamma_E$ for two redshift bins, where similar conclusions hold.}
+        \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_Spin0_NKA_TTTEEE_Full_chi2.pdf}
+        \caption{Same as Figure \ref{fig:chi2_1bin} for the combination of all
+          correlations between $\delta$ and $\gamma_E$ for two redshift bins,
+          where similar conclusions hold. In this case, the distribution
+          shift and width difference is less than a $3\%$ and $7\%$,
+          respectively. 
         \label{fig:chi2_2bins}
@@
       \includegraphics[width=\textwidth]{./figures/run_chi2_TT_TE_EE_TB_EB_BB.pdf}
-      \caption{Same as Figure \ref{fig:chi2_1bin} for flat-sky fields.
-This figure additionally shows the $\chi^2$ distributions for power spectra involving $B$-modes.
-For the smaller scales covered by these flat-sky simulations, the NKA and spin-0 estimators work remarkably well, even for $B$-mode spectra.}
+      \caption{Same as Figure \ref{fig:chi2_1bin} for flat-sky fields. This
+        figure additionally shows the $\chi^2$ distributions for power spectra
+        involving $B$-modes. For the smaller scales covered by these flat-sky
+        simulations, the NKA and spin-0 estimators work remarkably well, even
+        for $B$-mode spectra. The distribution mean shift and width difference
+        is of about $\lesssim 2\%$ for all cases.}
       \label{fig:chi2_1bin_flat}

### 2019-06-24 draft/paper.tex
fixed missing closing } + added ICCUB affiliation
fixed affiliation

@@
 
-\author[a]{Carlos Garc\'{i}a-Garc\'{i}a}
-\author[b]{David Alonso}
-\author[b]{Emilio Bellini}
+\author[a,b]{Carlos Garc\'{i}a-Garc\'{i}a}
+\author[c]{David Alonso}
+\author[c]{Emilio Bellini}
 
 \affiliation[a]{Instituto de Física Fundamental, Consejo Superior de Investigaciones Científicas, c/. Serrano 123, E–28006, Madrid, Spain}
-\affiliation[b]{Oxford Astrophysics, Department of Physics, Keble Road, Oxford, OX1 3RH, UK}
+\affiliation[b]{Institut de Ci\`{e}ncies del Cosmos (UB–IEEC), c/. Martí i Franqués 1, E–08028, Barcelona, Spain}
+\affiliation[c]{Oxford Astrophysics, Department of Physics, Keble Road, Oxford, OX1 3RH, UK}
 
@@
           shift and width difference is less than a $3\%$ and $7\%$,
-          respectively. 
+          respectively.}
         \label{fig:chi2_2bins}

### 2019-06-24 draft/paper.tex
abstract: change in sentence

@@
 
-\abstract{The disconnected part of the power spectrum covariance matrix (also known as the ``Gaussian'' covariance) is the dominant contribution on large scales for galaxy clustering and weak lensing datasets.
-The presence of a complicated sky mask causes non-trivial correlations between different Fourier/harmonic modes, which must be accurately characterised in order to obtain reliable cosmological constraints.
-This is particularly relevant for galaxy survey data.
-Unfortunately, an exact calculation of these correlations involve $O(\ell_{\rm max}^6)$ operations that become computationally impractical very quickly.
-We present an implementation of approximate methods to estimate the Gaussian covariance matrix of power spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding on existing algorithms.
-These methods achieve an $O(\ell_{\rm max}^3)$ scaling, which makes the computation of the covariance matrix as fast as the computation of the power spectrum itself.
-We quantify the accuracy of these methods on large-scale structure and weak lensing data, making use of a large number of Gaussian but otherwise realistic simulations.
-We show that, using the covariance matrix computing using these approximations we are able to recover the true posterior distribution cosmological parameters to high accuracy.
+\abstract{The disconnected part of the power spectrum covariance matrix (also
+  known as the ``Gaussian'' covariance) is the dominant contribution on large
+  scales for galaxy clustering and weak lensing datasets. The presence of a
+  complicated sky mask causes non-trivial correlations between different
+  Fourier/harmonic modes, which must be accurately characterised in order to
+  obtain reliable cosmological constraints. This is particularly relevant for
+  galaxy survey data. Unfortunately, an exact calculation of these
+  correlations involve $O(\ell_{\rm max}^6)$ operations that become
+  computationally impractical very quickly. We present an implementation of
+  approximate methods to estimate the Gaussian covariance matrix of power
+  spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding
+  on existing algorithms. These methods achieve an $O(\ell_{\rm max}^3)$
+  scaling, which makes the computation of the covariance matrix as fast as the
+  computation of the power spectrum itself. We quantify the accuracy of these
+  methods on large-scale structure and weak lensing data, making use of a
+  large number of Gaussian but otherwise realistic simulations. We show that,
+  using the approximated covariance matrix, we are able to recover
+  the true posterior distribution of cosmological parameters to high accuracy.
 We also quantify the shortcomings of these methods, which become unreliable on the very largest scales, as well as for covariance matrix elements involving lensing $B$ modes.

### 2019-06-24 draft/paper.tex
introduction: some minor changes

@@
   \section{Introduction}\label{sec:intro}
-    The two-point correlation of different fields projected on the celestial sphere is one of the most common observables used in the analysis of large datasets in astrophysics, from studies of the Cosmic Microwave Background \citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,1998PhRvD..57.2117B,2001PhRvD..64f3001T,2002ApJ...567....2H,2003PhRvD..67b3001W} to large-scale structure and weak lensing surveys \cite{2000MNRAS.317L..23H,2001ApJ...555..547H,2001MNRAS.325.1603E,2011arXiv1112.5723H,2016MNRAS.456.1508K,2018MNRAS.476.1050B,2018arXiv181208182X,2019MNRAS.tmp.1446C,2019PASJ...71...43H}.
-Using these two-point functions achieves a high level of data compression (with respect to the size of the raw datasets -- time-ordered data, images or catalogs), and they can be directly used to constrain cosmological and astrophysical parameters assuming that one can model their likelihood.
-This is usually done by assuming that the two-point functions are Gaussianly distributed, which is often good approximation due to the central limit theorem \cite{2008PhRvD..77j3013H,2018MNRAS.477.4879S}.
-In this case, the only obstacle that remains is being able to estimate the covariance matrix of a set of two-point correlators.
-Since the form of this covariance directly affects the posterior parameter uncertainties, a precise determination of it is of paramount importance.
-In large-scale structure experiments this has often been resolved by making use of one's own data through resampling techniques \cite{tukey1958,1984MNRAS.210P..19B,1993ApJ...406L..47H,2002ApJ...571..172Z,2016arXiv160600233E}, or by generating a large number of mock realizations \cite{2013MNRAS.428.1036M,2018MNRAS.479...94A,2019MNRAS.485.2806B}.
+    The two-point correlation of different fields projected on the celestial
+    sphere is one of the most common observables used in the analysis of large
+    datasets in astrophysics, from studies of the Cosmic Microwave Background
+    \citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,1998PhRvD..57.2117B,2001PhRvD..64f3001T,2002ApJ...567....2H,2003PhRvD..67b3001W}
+    to large-scale structure and weak lensing surveys
+    \cite{2000MNRAS.317L..23H,2001ApJ...555..547H,2001MNRAS.325.1603E,2011arXiv1112.5723H,2016MNRAS.456.1508K,2018MNRAS.476.1050B,2018arXiv181208182X,2019MNRAS.tmp.1446C,2019PASJ...71...43H}.
+    Using these two-point functions, one achieves a high level of data
+    compression (with respect to the size of the raw datasets -- time-ordered
+    data, images or catalogs). They can also be directly used to constrain
+    cosmological and astrophysical parameters assuming that one can model
+    their likelihood. This is usually done by assuming that the two-point
+    functions are Gaussianly distributed, which is often a good approximation
+    due to the central limit theorem
+    \cite{2008PhRvD..77j3013H,2018MNRAS.477.4879S}. In this case, the only
+    obstacle that remains is being able to estimate the covariance matrix of a
+    set of two-point correlators. Since the form of this covariance directly
+    affects the posterior parameter uncertainties, a precise determination of
+    it is of paramount importance.
+In large-scale structure experiments, this has often been resolved by making use of one's own data through resampling techniques \cite{tukey1958,1984MNRAS.210P..19B,1993ApJ...406L..47H,2002ApJ...571..172Z,2016arXiv160600233E}, or by generating a large number of mock realizations \cite{2013MNRAS.428.1036M,2018MNRAS.479...94A,2019MNRAS.485.2806B}.
 With the advent of the larger current and future surveys, the increasing size of the data vector and of the volume to be simulated has made this solution impractical, and fully analytical and hybrid estimators are now being used.
@@
 
-    In this paper we will present and generalise methods developed in the context of CMB experiments to account for the impact of survey geometry on the Gaussian part of the power spectrum covariance matrix \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}, and will study in detail the performance of these methods for large-scale structure and weak lensing datasets.
+    In this paper, we will present and generalise methods developed in the context of CMB experiments to account for the impact of survey geometry on the Gaussian part of the power spectrum covariance matrix \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}, and will study in detail the performance of these methods for large-scale structure and weak lensing datasets.
 We have also implemented these approximations in the public code {\tt NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}} \cite{2019MNRAS.484.4127A}, making the computation of accurate Gaussian covariance matrices significantly simpler for the community.

### 2019-06-24 draft/paper.tex
2.1 Preliminaries: explained what is spin-0/2

@@
     \subsection{Preliminaries}\label{ssec:theory.prelim}
-      We will deal with spin-0 and spin-2 fields defined on a 2-dimensional space.
-We will use a generic notation that can then be applied to both curved and flat skies.
-Given a field ${\bf a}({\bf x})$, with 1 (spin-0) or 2 components (spin-2), and defined on the coordinates ${\bf x}$, we define its generalized Fourier coefficients as
+      We will deal with spin-0 and spin-2 fields defined on a 2-dimensional
+      space. We will use a generic notation that can be applied to both
+      curved and flat skies. Given a field ${\bf a}({\bf x})$, with 1
+      (spin-0; i.e. an scalar field) or 2 components (spin-2; i.e. a field
+      that remains invariant under a transformation $e^{i2\phi}$), defined on the coordinates ${\bf x}$, we define its generalized Fourier coefficients as
       \begin{equation}

### 2019-06-24 draft/paper.tex
2.2 The pseudo-Cl method: fixed typo

@@
 
-      Before we move on to covariances, it is worth considering the case of unmasked field (i.e $w^a({\bf x})=1$ everywhere).
-In this case $\,^a{\sf M}_{{\bf l}{\bf l}'}=\mathds{1}\Delta^x({\bf k},{\bf l})$, and therefore different modes are uncorrelated (as should have been obvious).
+      Before we move on to covariances, it is worth considering the case of
+      unmasked field (i.e $w^a({\bf x})=1$ everywhere). In this case $\,^a{\sf
+        M}_{{\bf l}{\bf l}'}=\mathds{1}\Delta^x({\bf l},{\bf l}')$, and therefore different modes are uncorrelated (as should have been obvious).
 In a non-ideal case where the mask is still sufficiently well behaved (i.e.

### 2019-06-24 draft/paper.tex
2.3. Cov mat: commas, etc

@@
 
-      Let $A$ and $F$ be the vector indices corresponding to the pairs of field indices $(\alpha,\beta)$ and $(\phi,\gamma)$ respectively, and let us start by considering the covariance
+      Let $A$ and $F$ be the vector indices corresponding to the pairs of
+      field indices $(\alpha,\beta)$ and $(\phi,\gamma)$, respectively, and let us start by considering the covariance
       \begin{equation}
@@
 
-      Now, in what follows, we will be concerned with the auto- and cross-correlations of spin-0 and spin-2 fields.
-Thus, to simplify the notation, we will enumerate the different types of coupling coefficients that exist for a spin-0 field with a single component that we will call $\delta$, in analogy to the projected galaxy overensity, and for a spin-2 field $\boldsymbol{\gamma}$ with $E$ and $B$ components $\gamma_E$ and $\gamma_B$, in analogy to the cosmic shear field.
+      Now, in what follows, we will be concerned with the auto- and
+      cross-correlations of spin-0 and spin-2 fields. Thus, to simplify the
+      notation, we will enumerate the different types of coupling coefficients
+      that exist for a spin-0 field with a single component that we will call
+      $\delta$, in analogy to the projected galaxy overensity, and for a
+      spin-2 field, $\boldsymbol{\gamma}$, with $E$ and $B$ components, $\gamma_E$ and $\gamma_B$, in analogy to the cosmic shear field.
 With this setup, all the possible non-zero $W^{\alpha\beta,\alpha'\beta'}_{{\bf l}{\bf l}'}$ can be expressed in terms of the $I^{\pm s_a,\pm s_b}_{{\bf l}{\bf l}'}$ as follows:
@@
       \end{align}
-      Thus, in principle we only need to compute 7 different types of terms ($I^{0,0}$, $I^{0,\pm2}$, $I^{\pm2,\pm2}$ and $I^{\pm2,\mp2})$).
-In order to simplify these expressions further we follow \cite{2017A&A...602A..41C} and neglect all gradients of the masks, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
+      Thus, in principle, we only need to compute 7 different types of terms
+      ($I^{0,0}$, $I^{0,\pm2}$, $I^{\pm2,\pm2}$ and $I^{\pm2,\mp2})$). In
+      order to simplify these expressions further, we follow \cite{2017A&A...602A..41C} and neglect all gradients of the masks, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
       \begin{align}\nonumber

### 2019-06-24 draft/paper.tex
2 pseudo Cl: Delta(l, l) -> Delta(0)
2.1 Preliminaries: use more precise word

@@
 
-      Finally, we will assume that all fields are Gaussian stochastic fields that are additionally statistically isotropic.
-As a consequence of this, different generalised Fourier modes are uncorrelated:
+      Finally, we will assume that all fields are Gaussian stochastic fields
+      that are additionally statistically isotropic. As a consequence of the
+      latter, different generalised Fourier modes are uncorrelated:
       \begin{equation}\label{eq:iso}
@@
         \end{equation}
-        where the bandpower weights are normalized such that $\sum_{{\bf l}\in q}B_q^{\bf l}=(K\Delta^x({\bf l},{\bf l}))^{-1}$.
+        where the bandpower weights are normalized such that $\sum_{{\bf l}\in
+          q}B_q^{\bf l}=(K\Delta^x({\bf 0}))^{-1}$.
         \item Then, the correlation between bandpowers induced by the mode-coupling coefficients is partially reversed by multiplying $\tilde{\sf C}^{ab}_q$ by the so-called binned ``mode-coupling matrix'' $\mathcal{M}$, giving the final estimator

### 2019-06-24 draft/paper.tex
2.3.1 Cov sph: comma + removed extra =

@@
       \subsubsection{Covariances for curved skies}
-        As described in Section \ref{ssec:theory.pcl}, in the curved-sky estimator it is common to first average over the harmonic number $m$ as part of the bandpower binning operation.
-Let us therefore define
+        As described in Section \ref{ssec:theory.pcl}, in the curved-sky
+        estimator it is common to first average over the harmonic number $m$
+        as part of the bandpower binning operation. Let us, therefore, define
         \begin{equation}
@@
         \end{equation}
-        The covariance of these objects under the approximation in Eq. \ref{eq:covar_efstathiou}, can be computed as:
+        The covariance of these objects, under the approximation in Eq. \ref{eq:covar_efstathiou}, can be computed as:
         \begin{equation}\label{eq:cov_sph}
@@
         \begin{equation}
-          ^{af}_{bg}\mathcal{W}^{\alpha\phi,\alpha'\phi'}_{\beta\gamma,\beta'\gamma'}(\ell,\ell')=\equiv\sum_{m=-\ell}^\ell \sum_{m'=-\ell'}^{\ell'}\frac{\,^{af}W^{\alpha\phi,\alpha'\phi'}_{{\bf l}{\bf l}'}\left(\,^{bg}W^{\beta\gamma,\beta'\gamma'}_{{\bf l}{\bf l}'}\right)^*}{(2\ell+1)(2\ell'+1)}.
+          ^{af}_{bg}\mathcal{W}^{\alpha\phi,\alpha'\phi'}_{\beta\gamma,\beta'\gamma'}(\ell,\ell')\equiv\sum_{m=-\ell}^\ell \sum_{m'=-\ell'}^{\ell'}\frac{\,^{af}W^{\alpha\phi,\alpha'\phi'}_{{\bf l}{\bf l}'}\left(\,^{bg}W^{\beta\gamma,\beta'\gamma'}_{{\bf l}{\bf l}'}\right)^*}{(2\ell+1)(2\ell'+1)}.
         \end{equation}

### 2019-06-24 draft/paper.tex
2.3.2 Cov flat: commas + missing Eq.

@@
       \subsubsection{Covariances for flat skies}
-        Similar results hold in the case of flat skies.
-As mentioned in Section \ref{ssec:theory.pcl}, in this case averaging over the Fourier-space azimuth happens while binning into bandpowers.
-Under the assumption that the underlying power spectra are roughly constant within each bandpower, in this case the covariance matrix of the bandpowers defined in Eq.
-\ref{eq:bandpowers} takes the form:
+        Similar results hold in the case of flat skies. As mentioned in
+        Section \ref{ssec:theory.pcl}, in this case averaging over the
+        Fourier-space azimuth happens while binning into bandpowers. Under the
+        assumption that the underlying power spectra are roughly constant
+        within each bandpower, in this case, the covariance matrix of the bandpowers defined in Eq. \ref{eq:bandpowers} takes the form:
         \begin{equation}\label{eq:cov_flat}
@@
         \end{equation}
-        where the coefficients $\bar{\mathcal{W}}$ are related to the mode-coupling coefficients $\bar{\Xi}$ defined in \ref{eq:coeff_mcm_flat} in the same way that the curved-sky coefficients $\mathcal{W}$ were related to the $\Xi$:
+        where the coefficients $\bar{\mathcal{W}}$ are related to the
+        mode-coupling coefficients $\bar{\Xi}$ defined in
+        Eq.~\ref{eq:coeff_mcm_flat}, in the same way that the curved-sky coefficients $\mathcal{W}$ were related to the $\Xi$:
         \begin{align}
@@
         \end{align}
-        and, as before
+        and, as before,
         \begin{align}\label{eq:coeff_cmcm_flat}
@@
            \,^{af}_{bg}\bar{\mathcal{J}}^{-+}_{{\bf l}{\bf l}'}=
-           \,^{af}_{bg}\bar{\mathcal{J}}^{0-}_{{\bf l}{\bf l}'}=0,
+           \,^{af}_{bg}\bar{\mathcal{J}}^{0-}_{{\bf l}{\bf l}'}=0.
         \end{align}

### 2019-06-24 draft/paper.tex
24 Approx covs: added comma
3.1 Simulations

@@
         \item The {\bf spin-0} approximation corresponds to a simplified version of the NKA in which the spin nature of all fields involved is completely ignored, and all fields, including the $E$- and $B$-mode components of a spin-2 field, are treated as spin-0 quantities.
-        \item The mode-counting approximation (labelled {\bf MC} here), commonly known as the {\sl Knox formula} \cite{1995PhRvD..52.4307K}, which applies the result found for full-sky observations (Eq.
-\ref{eq:cov_naive}) to masked fields corrected by an overall factor that accounts to the loss of modes due to the sky mask.
-In this case, the covariance is simply given by:
+        \item The mode-counting approximation (labelled {\bf MC} here),
+          commonly known as the {\sl Knox formula} \cite{1995PhRvD..52.4307K},
+          which applies the result found for full-sky observations (Eq.
+          \ref{eq:cov_naive}) to masked fields, corrected by an overall factor that accounts to the loss of modes due to the sky mask. In this case, the covariance is simply given by:
         \begin{equation}\label{eq:cov_naive_bpw}
@@
 
-      We consider the case of auto- and cross-correlations between all fields in two redshift bins, with redshift distributions modelled as Gaussians with width $\sigma_z=0.13$ centered around redshifts $0.75$ and $0.95$, with number densities $n_\Omega=7.5\,{\rm arcmin}^{-2}$.
-The corresponding redshift distributions $p_z(z)$ are shown in Figure \ref{fig:pz}.
-For simplicity we use a constant bias $b(z)=1$.
-For these specifications, we generate signal power spectra using the Core Cosmology Library ({\tt CCL} \cite{2019ApJS..242....2C}) for cosmological parameters $(\Omega_c,\Omega_b,h,A_s,n_s)=(0.27,0.045,0.67,2.1\times10^{-9},0.96)$.
-The resulting non-zero signal and noise power spectra are shown in Fig.
-\ref{fig:cl-2bins}.
+      We consider the case of auto- and cross-correlations between all fields
+      in two redshift bins, with redshift distributions modelled as Gaussians
+      with width $\sigma_z=0.13$ centered around redshifts $0.75$ and $0.95$,
+      with number densities $n_\Omega=7.5\,{\rm arcmin}^{-2}$. The
+      corresponding redshift distributions $p_z(z)$ are shown in Figure
+      \ref{fig:pz}. For simplicity we use a constant bias $b(z)=1$. For these
+      specifications, we generate signal power spectra using the Core
+      Cosmology Library ({\tt CCL} \cite{2019ApJS..242....2C}) for
+      cosmological parameters
+      $(\Omega_c,\Omega_b,h,A_s,n_s)=(0.27,0.045,0.67,2.1\times10^{-9},0.96)$.
+      The resulting non-zero signal and noise power spectra are shown in
+      Figure~\ref{fig:cl-2bins}.
 
@@
       \end{itemize}
-      In both cases we fixed the amplitude of the contaminant power spectrum such that it would yield a $10\%$ contamination in the data power spectrum at $\ell=400$.
-For spin-2 fields, we assumed the same power spectrum for $E$ and $B$ modes, with no cross-correlation between them.
-The resulting contaminant power spectra are shown in Figure \ref{fig:clcont}.
-When exploring the effects of mode deprojection, we generate 100 contaminant maps of both types and add them to each simulated realization.
-We then deproject the full set of 100 contaminant templates from the simulated maps and compute the corresponding unbiased power spectra.
-Note that our fiducial results do not include the effects of mode deprojection.
-These are discussed separately.
+      In both cases, we fixed the amplitude of the contaminant power spectrum
+      such that it would yield a $10\%$ contamination in the data power
+      spectrum at $\ell=400$. For spin-2 fields, we assumed the same power
+      spectrum for $E$ and $B$ modes, with no cross-correlation between them.
+      The resulting contaminant power spectra are shown in Figure
+      \ref{fig:clcont}. When exploring the effects of mode deprojection, we
+      generated 100 contaminant maps of both types and added them to each
+      simulated realization. We then deprojected the full set of 100
+      contaminant templates from the simulated maps and computed the
+      corresponding unbiased power spectra. Note that, in what follows, our
+      fiducial results do not include the effects of mode deprojection. These
+      are discussed separately.
 

### 2019-06-24 draft/paper.tex
3.2 Qualitative comp: commas + caption Fig 5 + Fig -> Figure

@@
         \includegraphics[width=\textwidth]{./figures/all_rows_sph_1bin.pdf}
-        \caption{Four rows of the covariance matrix for different power spectra.
-The top panels show the cases of spectra with non-zero signal, involving $\delta$ and $\gamma_E$, while the bottom panels show cases involving $B$-modes.
-In each panel we show rows of the covariance matrix for $\ell=60,\,90,\,120$ and $150$, which peak at those central values.
-The different lines show the results for the sample covariance matrix (solid blue), its NKA estimator (dashed orange) and the difference between both (dotted green).
-The black stars show the mode-counting approximation to the covariance matrix (Eq.
-\ref{eq:cov_naive_bpw}).
-To facilitate the visualization of the different rows, we have divided them by $C_\ell^{aa}C_\ell^{bb}+(C^{ab}_\ell)^2$, where $a$ and $b$ are the two fields being correlated in each case.
+        \caption{Four rows of the covariance matrix for different power
+          spectra. The top panels show the cases of spectra with non-zero
+          signal, involving $\delta$ and $\gamma_E$, while the bottom panels
+          show cases involving $B$-modes. In each panel we show rows of the
+          covariance matrix for $\ell=60,\,90,\,120$ and $150$, which peak at
+          those central values. The different lines show the results for the
+          sample covariance matrix (solid blue), its NKA estimator (dashed
+          orange) and the difference between both (dotted green). The black
+          stars show the mode-counting approximation to the covariance matrix
+          (Eq. \ref{eq:cov_naive_bpw}). To facilitate the visualization of the
+          different rows, we have divided them by
+          $C_\ell^{aa}C_\ell^{bb}+(C^{ab}_\ell)^2$, the numerator of
+          Eq.~\ref{eq:cov_naive_bpw}, for the two fields, $a$ and $b$, that are being correlated in each case.
 The NKA method is able to recover the covariance with high accuracy for all field combinations that do not involve $B$ modes, yielding visibly poorer results otherwise.}
@@
       \end{figure}
-      As a first step, we visually compare the main properties of the sample covariance matrix estimated from the simulations and the NKA, Spin-0 and MC approximations described in the previous section.
-Figure \ref{fig:rows_1bin} shows four rows of the covariance matrix of different auto- and cross-correlations.
-The rows correspond to bandpowers centered on multipoles $\ell_q\simeq60,\,90,\,120$ and $150$.
-The upper panels show results for the non-zero power spectra ($\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$), with the solid blue, dashed orange and dotted green lines showing results for the sample covariance matrix, the analytical covariance using the NKA approximation and their difference respectively.
+      As a first step, we visually compare the main properties of the sample
+      covariance matrix estimated from the simulations and the NKA, Spin-0 and
+      MC approximations described in the previous section. Figure
+      \ref{fig:rows_1bin} shows four rows of the covariance matrix of
+      different auto- and cross-correlations. The rows correspond to
+      bandpowers centered on multipoles $\ell_q\simeq60,\,90,\,120$ and $150$.
+      The upper panels show results for the non-zero power spectra
+      ($\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$), with
+      the solid blue, dashed orange and dotted green lines showing results for
+      the sample covariance matrix, the analytical covariance using the NKA
+      approximation and their difference, respectively.
 For comparison, the black stars show the diagonal covariance matrix elements predicted by the MC approximation (Eq.
@@
 
-      The bottom panel in Fig.
-\ref{fig:rows_1bin} shows the same rows of the covariance matrix for power spectra involving $B$-modes (and therefore with zero signal expectation value).
+      The bottom panel in Figure~\ref{fig:rows_1bin} shows the same rows of the covariance matrix for power spectra involving $B$-modes (and therefore with zero signal expectation value).
 In this case we find significant differences, at the level of $30-50\%$, on the covariance matrix elements, with the analyticial prediction underestimating the error bars overall.
@@
 
-      This is further illustrated by Fig.
-\ref{fig:diff_corr_1bin}.
+      This is further illustrated by Figure~\ref{fig:diff_corr_1bin}.
 The figure shows, for the case of a single redshift bin, the difference between the correlation matrices associated with the sample covariance matrix and the NKA estimate\footnote{The correlation matrix is defined as $r_{ij} = {\rm Cov}_{ij}/\sqrt{{\rm Cov}_{ii}{\rm Cov}_{jj}}$}.
@@
 
-      We thus conclude that while the NKA estimator is able to recover the covariance matrix for the non-zero power spectrum elements (i.e.
-those involving $\delta$ and $\gamma_E$) with high accuracy, a more sophisticated approach would be needed in order to obtain a precise estimate of the uncertainties for components involving $B$-modes.
-This is not a major concern, since $B$-mode power spectra are predominantly used as null tests, while cosmological parameter constraints are driven by the analysis of $\delta$ and $\gamma_E$.
-For completeness, Fig.
-\ref{fig:corr_diff_2bins} shows the difference between the correlation matrices for the sample covariance and the NKA estimator for all non-zero cross-correlations between different bins in the case of two redshift bins with different small-scale masks, where we find a similarly good agreement.
+      We thus conclude that while the NKA estimator is able to recover the
+      covariance matrix for the non-zero power spectrum elements (i.e. those
+      involving $\delta$ and $\gamma_E$) with high accuracy, a more
+      sophisticated approach would be needed in order to obtain a precise
+      estimate of the uncertainties for components involving $B$-modes. This
+      is not a major concern, since $B$-mode power spectra are predominantly
+      used as null tests, while cosmological parameter constraints are driven
+      by the analysis of $\delta$ and $\gamma_E$.
+For completeness, Figure~\ref{fig:corr_diff_2bins} shows the difference between the correlation matrices for the sample covariance and the NKA estimator for all non-zero cross-correlations between different bins in the case of two redshift bins with different small-scale masks, where we find a similarly good agreement.
 

### 2019-06-24 draft/paper.tex
3.3 Quantitaive: added short sentence about difference between spin0 and NKA + commas

@@
           all cases. In the cases of $\delta\times\gamma_E$ and
-          $\gamma_E\times\gamma_E$ we observe small shifts in the peak
+          $\gamma_E\times\gamma_E$, we observe small shifts in the peak
           $\chi^2$, of $\sim2$-$4\%$, while their width is recovered
@@
       In order to quantify the validity of our analytical approximations, we need to compare the NKA covariance with the sample covariance estimated from the simulations.
-However, comparing two matrices is not as straightforward as comparing their matrix elements one by one.
+However, comparing two matrices is not as straightforward as comparing their elements one by one.
 The covariance between far-away bandpowers is expected to be very close to zero, and therefore a direct comparison of those elements would easily yield large relative differences simply due to the statistical noise in the sample covariance matrix.
@@
 
-      As a first test to quantify the differences between covariance matrices, we compute the relative difference between their eigenvalues.
-This is shown in Fig.
-\ref{fig:eigv_1bin} for a data vector combining all auto- and cross-correlations between $\delta$ and $\gamma_E$ for a single redshift bin.
+      As a first test to quantify the differences between covariance matrices,
+      we compute the relative difference between their eigenvalues. This is
+      shown in Figure~\ref{fig:eigv_1bin} for a data vector combining all auto- and cross-correlations between $\delta$ and $\gamma_E$ for a single redshift bin.
 The eigenvalues of both matrices are roughly similar, with relative differences of about $5\%$.
@@
       \end{equation}
-      We compute this quantity for a data vector ${\bf d}$ composed of different auto- and cross-correlations for each of the 20,000 Gaussian simulations, with ${\bf m}$ given by the mean over all simulations and different choices of covariance matrix.
-Figure \ref{fig:chi2_1bin} shows the distribution of $\chi^2$ values for the three non-zero power spectra in the case of a single bin: $\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$.
-The histograms show the distribution for the sample covariance matrix (blue), and the analytical NKA and Spin-0 estimators (orange and green respectively).
+      We compute this quantity for a data vector ${\bf d}$ composed of
+      different auto- and cross-correlations for each of the 20,000 Gaussian
+      simulations, with ${\bf m}$ given by the mean over all simulations and
+      different choices of covariance matrix. Figure \ref{fig:chi2_1bin} shows
+      the distribution of $\chi^2$ values for the three non-zero power spectra
+      in the case of a single bin: $\delta$-$\delta$, $\delta$-$\gamma_E$ and
+      $\gamma_E$-$\gamma_E$. The histograms show the distribution for the
+      sample covariance matrix (blue), and the analytical NKA and Spin-0
+      estimators (orange and green, respectively).
 We additionally plot the theoretical $\chi^2$ distribution under the assumption that the underlying data vector is Gaussianly distributed (red dashed lines).
@@
 In the cases involving the spin-2 fields, we see noticeable differences between the distributions found with the sample covariance and the approximate ones.
-      These differences are small, corresponding to less of a $2$ and $4\%$ shift in the mean $\chi^2$ for the $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ cases respectively, and a negligible variation in the width of the distributions.
-We therefore expect these differents to have a negligible effect on the posterior parameter distributions, as we show explicitly in Section \ref{ssec:results.cosmo}.
-The fact that these differences appear only for power spectra involving spin-2 quantities indicate that the NKA and spin-0 methods are imperfect at describing the additional mode coupling caused in the presence of a mask for higher-spin fields.
-This is rather obvious in the case of the spin-0 approximation but, interestingly, we find that the NKA and Spin-0 predictions yield results that are almost indistinguishable from each other.
-We therefore conclude that the additional approximation made in the NKA method for spin-2 fields -- neglecting the spatial derivatives of the mask -- is effectively equivalent to ignoring the spin nature of the fields involved.
-For completeness, Figure \ref{fig:chi2_2bins} shows the distribution of $\chi^2$ values for a data vector composed of all possible auto- and cross-correlations of $\delta$ and $\gamma_E$ for the case of two redshift bins, where similar conclusions hold.
+      These differences are small, corresponding to less of a $2$ and $4\%$
+      shift in the mean $\chi^2$ for the $\delta$-$\gamma_E$ and
+      $\gamma_E$-$\gamma_E$ cases, respectively, and a negligible variation in
+      the width of the distributions. We therefore expect these differents to
+      have a negligible effect on the posterior parameter distributions, as we
+      show explicitly in Section \ref{ssec:results.cosmo}. The fact that these
+      differences appear only for power spectra involving spin-2 quantities
+      indicate that the NKA and spin-0 methods are imperfect at describing the
+      additional mode coupling caused in the presence of a mask for
+      higher-spin fields. This is rather obvious in the case of the spin-0
+      approximation but, interestingly, we find that the NKA and Spin-0
+      predictions yield results that are almost indistinguishable from each
+      other. We therefore conclude that the additional approximation made in
+      the NKA method for spin-2 fields -- neglecting the spatial derivatives
+      of the mask -- is effectively equivalent to ignoring the spin nature of
+      the fields involved \textcolor{red}{for the $\delta$ and $\gamma_E$ cases. The B
+        modes break this similarity and worsen the spin-0 approximation,
+        yielding, in the most extreme case ($\gamma_b$- $\gamma_b$), around an
+        order of magnitude worse results than the NKA method}. For
+        completeness, Figure \ref{fig:chi2_2bins} shows the distribution of
+        $\chi^2$ values for a data vector composed of all possible auto- and
+        cross-correlations of $\delta$ and $\gamma_E$ for the case of two
+        redshift bins, where similar conclusions hold.
 
@@
 
-      Figure \ref{fig:diag_conts} serves also to illustrate another important point.
-In all our tests we find that the largest differences between the sample and NKA covariances occur on large scales.
-For $\delta$-$\delta$ correlations, the effect is limited to the first few multipoles ($\ell\lesssim10$), while for $\gamma_E$-$\gamma_E$ we are only able to recover the sample covariance errors within $5\%$ for $\ell\gtrsim40$.
-This is the main source of the small mismatch observed in Fig.
-\ref{fig:chi2_1bin}.
-Note however, that most of the cosmological information is obtained from the higher multipoles, due to their higher statistical weight, and as we will show in the next section, the effect on the final parameter constraints is negligible.
+      Figure \ref{fig:diag_conts} serves also to illustrate another important
+      point. In all our tests we find that the largest differences between the
+      sample and NKA covariances occur on large scales. For $\delta$-$\delta$
+      correlations, the effect is limited to the first few multipoles
+      ($\ell\lesssim10$), while for $\gamma_E$-$\gamma_E$ we are only able to
+      recover the sample covariance errors within $5\%$ for $\ell\gtrsim40$.
+      This is the main source of the small mismatch observed in
+      Figure~\ref{fig:chi2_1bin}. Note, however, that most of the cosmological
+      information is obtained from the higher multipoles, due to their higher
+      statistical weight, and, as we will show in the next section, the effect on the final parameter constraints is negligible.
 If accurate covariances bars are needed on these large scales for cosmic shear, they can be estimated alternatively making use of fast, low resolution simulations (e.g.

### 2019-06-24 draft/paper.tex
added comma + Fig -> Figure

@@
 
-      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for of two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins.
-In this simple two-dimensional scenario we simply sample the distribution in a regular grid of 100 by 100 points for each parameter.
+      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for of two parameters
+      $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of
+      all possible auto- and cross-correlations between $\delta$ and
+      $\gamma_E$ in the case of two redshift bins. In this simple
+      two-dimensional scenario, we simply sample the distribution in a regular grid of 100 by 100 points for each parameter.
 We construct a data vector $\hat{\bf C}$ from the theoretical prediction for the experimental setup described in Section \ref{ssec:results.sims}, and produce theoretical predictions for it at each grid point using {\tt CCL}.
@@
 
-      The 68\% and 95\% confidence level contours associated with the posterior distributions for the sample covariance and the NKA estimator are shown in Fig.
-\ref{fig:contours}.
+      The 68\% and 95\% confidence level contours associated with the
+      posterior distributions for the sample covariance and the NKA estimator
+      are shown in Figure~\ref{fig:contours}.
 We find that both distributions agree with each other remarkably well, and that the $1\sigma$ errors for each parameter agree for both covariances up to $0.3\%$.

### 2019-06-24 draft/paper.tex
Discussion: added commas

@@
 
-    More in detail, we have also found that the impact of contaminant deprojection on the covariance matrix, through the corresponding loss of modes, is negligible unless a very large number of contaminant templates are removed.
-This simplifies the procedure to estimate covariance matrices for galaxy clustering data, which are particularly sensitive to a large number of astrophysical and observational systematics.
-Additionally, we have found that, although the NKA estimator is accurate enough, it is not able to perfectly capture the additional effects of mode coupling that are present for spin-2 fields, and that a simpler approach treating the shear $E$ modes as a spin-0 object is able to reach similar levels of accuracy.
-Due to the imperfect treatment of the $E/B$ mixing caused by the sky mask in the NKA estimator, we also find that the predicted covariance matrix for any power spectra involving $B$ modes differs significantly from the true sample covariance, and therefore this approach cannot be used to reliably estimate the uncertainties of $B$-mode power spectra.
+    More in detail, we have also found that the impact of contaminant
+    deprojection on the covariance matrix, through the corresponding loss of
+    modes, is negligible unless a very large number of contaminant templates
+    are removed. This simplifies the procedure to estimate covariance matrices
+    for galaxy clustering data, which are particularly sensitive to a large
+    number of astrophysical and observational systematics. Additionally, we
+    have found that, although the NKA estimator is accurate enough, it is not
+    able to perfectly capture the additional effects of mode coupling that are
+    present for spin-2 fields, and that a simpler approach treating the shear
+    $E$ modes as a spin-0 object is able to reach similar levels of accuracy.
+    Due to the imperfect treatment of the $E/B$ mixing caused by the sky mask
+    in the NKA estimator, we also find that the predicted covariance matrix
+    for any power spectra involving $B$ modes differs significantly from the
+    true sample covariance, and, therefore, this approach cannot be used to reliably estimate the uncertainties of $B$-mode power spectra.
 Likewise, we find that the NKA estimator yields inaccurate estimates of the power spectrum uncertainties on the largest scales ($\ell\lesssim50$).
@@
 
-    In spite of these shortcomings, we find that the approximations described in this paper are able to provide estimates of the power spectrum covariance matrix that are sufficiently accurate for current and future tomographic large-scale structure cosmological datasets.
-The main advantage of this approach is the computational cost, which is comparable to that of estimating the power spectrum in the first place, and which scales with pixel resolution in a similar way.
-This is therefore significantly less time-consuming than generating large numbers of mock datasets (even simple Gaussian or log-normal realizations), and more reliable than the traditional jackknife resampling techniques.
+    In spite of these shortcomings, we find that the approximations described
+    in this paper are able to provide estimates of the power spectrum
+    covariance matrix that are sufficiently accurate for current and future
+    tomographic large-scale structure cosmological datasets. The main
+    advantage of this approach is the computational cost, which is comparable
+    to that of estimating the power spectrum in the first place, and which
+    scales with pixel resolution in a similar way.
+This is, therefore, significantly less time-consuming than generating large numbers of mock datasets (even simple Gaussian or log-normal realizations), and more reliable than the traditional jackknife resampling techniques.
 The method, in all its generality, is currently implemented in the public code {\tt NaMaster}.

### 2019-06-24 draft/paper.tex
Appendix A: comma + Fig -> Figure

@@
     \end{figure}
-    We have repeated the analysis described in Section \ref{sec:results} on flat-sky realizations, making use of the flat-sky implementation of {\tt NaMaster}.
-We generate Gaussian realizations of the galaxy overdensity and shear maps making use of flat-sky extensions of the methods described in Section \ref{ssec:results.sims}.
-In this case we use a high-resolution mask constructed from the bright-object mask distributed with the first data release of the HSC collaboration \cite{2018PASJ...70S...4A} for the VVDS field, which is shown in Fig.
-\ref{fig:mask_flat}.
-
-    We find similar levels of accuracy in the NKA and spin-0 estimators compared to the curved-sky case.
-The higher resolution and smaller area of these simulations allow us to focus on the small-scale galaxy clustering and lensing power spectra, covering the range of multipoles $\ell\in(120,17640)$ in constant bandpowers of width $\Delta \ell=240$.
-On these small scales, the shear power spectrum is more dominated by noise (e.g.
-see Fig.
-\ref{fig:cl-2bins}), and therefore there is roughly the same power in $E$ and $B$ modes.
-This reduces the sensitivity of the method to an inaccurate treatment of $E/B$ leakage, and the agreement between the sample covariance matrix and the NKA estimator improves significantly.
-This can be seen in Fig.
-\ref{fig:chi2_1bin_flat}, which shows the $\chi^2$ distributions for all possible power spectra (including those including $B$ modes) in the case of a single redshift bin.
+    We have repeated the analysis described in Section \ref{sec:results} on
+    flat-sky realizations, making use of the flat-sky implementation of {\tt
+      NaMaster}. We generate Gaussian realizations of the galaxy overdensity
+    and shear maps making use of flat-sky extensions of the methods described
+    in Section \ref{ssec:results.sims}. In this case we use a high-resolution
+    mask constructed from the bright-object mask distributed with the first
+    data release of the HSC collaboration \cite{2018PASJ...70S...4A} for the
+    VVDS field, which is shown in Figure~\ref{fig:mask_flat}.
+
+    We find similar levels of accuracy in the NKA and spin-0 estimators
+    compared to the curved-sky case. The higher resolution and smaller area of
+    these simulations allow us to focus on the small-scale galaxy clustering
+    and lensing power spectra, covering the range of multipoles
+    $\ell\in(120,17640)$ in constant bandpowers of width $\Delta \ell=240$. On
+    these small scales, the shear power spectrum is more dominated by noise
+    (e.g. see Fig. \ref{fig:cl-2bins}), and therefore there is roughly the
+    same power in $E$ and $B$ modes. This reduces the sensitivity of the
+    method to an inaccurate treatment of $E/B$ leakage, and the agreement
+    between the sample covariance matrix and the NKA estimator improves
+    significantly.
+This can be seen in Figure~\ref{fig:chi2_1bin_flat}, which shows the $\chi^2$ distributions for all possible power spectra (including those including $B$ modes) in the case of a single redshift bin.
 In all cases we find a good agreement between the $\chi^2$ distributions derived from all covariance matrix estimates (sample covariance, NKA and spin-0 in blue, orange and green respectively), which also accurately follow the expected $\chi^2$ distribution for the corresponding number of degrees of freedom (red dashed line).

### 2019-06-25 draft/paper.tex
small mods

@@
 
-\abstract{The disconnected part of the power spectrum covariance matrix (also
-  known as the ``Gaussian'' covariance) is the dominant contribution on large
-  scales for galaxy clustering and weak lensing datasets. The presence of a
-  complicated sky mask causes non-trivial correlations between different
-  Fourier/harmonic modes, which must be accurately characterised in order to
-  obtain reliable cosmological constraints. This is particularly relevant for
-  galaxy survey data. Unfortunately, an exact calculation of these
-  correlations involve $O(\ell_{\rm max}^6)$ operations that become
-  computationally impractical very quickly. We present an implementation of
-  approximate methods to estimate the Gaussian covariance matrix of power
-  spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding
-  on existing algorithms. These methods achieve an $O(\ell_{\rm max}^3)$
-  scaling, which makes the computation of the covariance matrix as fast as the
-  computation of the power spectrum itself. We quantify the accuracy of these
-  methods on large-scale structure and weak lensing data, making use of a
-  large number of Gaussian but otherwise realistic simulations. We show that,
-  using the approximated covariance matrix, we are able to recover
-  the true posterior distribution of cosmological parameters to high accuracy.
-We also quantify the shortcomings of these methods, which become unreliable on the very largest scales, as well as for covariance matrix elements involving lensing $B$ modes.
+\abstract{The disconnected part of the power spectrum covariance matrix (also known as the ``Gaussian'' covariance) is the dominant contribution on large scales for galaxy clustering and weak lensing datasets.
+The presence of a complicated sky mask causes non-trivial correlations between different Fourier/harmonic modes, which must be accurately characterised in order to obtain reliable cosmological constraints.
+This is particularly relevant for galaxy survey data.
+Unfortunately, an exact calculation of these correlations involves $O(\ell_{\rm max}^6)$ operations that become computationally impractical very quickly.
+We present an implementation of approximate methods to estimate the Gaussian covariance matrix of power spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding on existing algorithms.
+These methods achieve an $O(\ell_{\rm max}^3)$ scaling, which makes the computation of the covariance matrix as fast as the computation of the power spectrum itself.
+We quantify the accuracy of these methods on large-scale structure and weak lensing data, making use of a large number of Gaussian but otherwise realistic simulations.
+We show that, using the approximate covariance matrix, we are able to recover the true posterior distribution of cosmological parameters to high accuracy.
+We also quantify the shortcomings of these methods, which become unreliable on the very largest scales, as well as for covariance matrix elements involving cosmic shear $B$ modes.
 The algorithms presented here are implemented in the public code {\tt NaMaster} \url{https://github.com/LSSTDESC/NaMaster}.}
@@
   \section{Introduction}\label{sec:intro}
-    The two-point correlation of different fields projected on the celestial
-    sphere is one of the most common observables used in the analysis of large
-    datasets in astrophysics, from studies of the Cosmic Microwave Background
-    \citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,1998PhRvD..57.2117B,2001PhRvD..64f3001T,2002ApJ...567....2H,2003PhRvD..67b3001W}
-    to large-scale structure and weak lensing surveys
-    \cite{2000MNRAS.317L..23H,2001ApJ...555..547H,2001MNRAS.325.1603E,2011arXiv1112.5723H,2016MNRAS.456.1508K,2018MNRAS.476.1050B,2018arXiv181208182X,2019MNRAS.tmp.1446C,2019PASJ...71...43H}.
-    Using these two-point functions, one achieves a high level of data
-    compression (with respect to the size of the raw datasets -- time-ordered
-    data, images or catalogs). They can also be directly used to constrain
-    cosmological and astrophysical parameters assuming that one can model
-    their likelihood. This is usually done by assuming that the two-point
-    functions are Gaussianly distributed, which is often a good approximation
-    due to the central limit theorem
-    \cite{2008PhRvD..77j3013H,2018MNRAS.477.4879S}. In this case, the only
-    obstacle that remains is being able to estimate the covariance matrix of a
-    set of two-point correlators. Since the form of this covariance directly
-    affects the posterior parameter uncertainties, a precise determination of
-    it is of paramount importance.
+    The two-point correlation of different fields projected on the celestial sphere is one of the most common observables used in the analysis of large datasets in astrophysics, from studies of the Cosmic Microwave Background \citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,1998PhRvD..57.2117B,2001PhRvD..64f3001T,2002ApJ...567....2H,2003PhRvD..67b3001W} to large-scale structure and weak lensing surveys \cite{2000MNRAS.317L..23H,2001ApJ...555..547H,2001MNRAS.325.1603E,2011arXiv1112.5723H,2016MNRAS.456.1508K,2018MNRAS.476.1050B,2018arXiv181208182X,2019MNRAS.tmp.1446C,2019PASJ...71...43H}.
+Using these two-point functions, one achieves a high level of data compression (with respect to the size of the raw datasets -- time-ordered data, images or catalogs).
+They can also be directly used to constrain cosmological and astrophysical parameters assuming that one can model their likelihood.
+This is usually done by assuming that the two-point functions are Gaussianly distributed, which is often a good approximation due to the central limit theorem \cite{2008PhRvD..77j3013H,2018MNRAS.477.4879S}.
+In this case, the only obstacle that remains is being able to estimate the covariance matrix of a set of two-point correlators.
+Since the form of this covariance directly affects the posterior parameter uncertainties, a precise determination of it is of paramount importance.
 In large-scale structure experiments, this has often been resolved by making use of one's own data through resampling techniques \cite{tukey1958,1984MNRAS.210P..19B,1993ApJ...406L..47H,2002ApJ...571..172Z,2016arXiv160600233E}, or by generating a large number of mock realizations \cite{2013MNRAS.428.1036M,2018MNRAS.479...94A,2019MNRAS.485.2806B}.
@@
 In simpler terms, this is the covariance matrix one would obtain if all fields involved were Gaussianly distributed.
-      \item {\bf Connected non-Gaussian covariance:} this is the contribution
-        from the connected trispectrum (which would vanish if all fields were
-        Gaussianly distributed).
+      \item {\bf Connected non-Gaussian covariance:} this is the contribution from the connected trispectrum (which would vanish if all fields were Gaussianly distributed).
       \item {\bf Super-sample covariance:} this is the additional coupling between different scales induced by density fluctuations on scales larger than the volume mapped.
+This term also vanishes for Gaussian fields.
     \end{itemize}
@@
     \subsection{Preliminaries}\label{ssec:theory.prelim}
-      We will deal with spin-0 and spin-2 fields defined on a 2-dimensional
-      space. We will use a generic notation that can be applied to both
-      curved and flat skies. Given a field ${\bf a}({\bf x})$, with 1
-      (spin-0; i.e. an scalar field) or 2 components (spin-2; i.e. a field
-      that remains invariant under a transformation $e^{i2\phi}$), defined on the coordinates ${\bf x}$, we define its generalized Fourier coefficients as
+      We will deal with spin-0 and spin-2 fields defined on a 2-dimensional space.
+In two dimensions, spin-$s$ fields in general have two components ${\bf a}({\bf x})=(a_1({\bf x}),a_2({\bf x}))$\footnote{E.g.
+for CMB polarisation, a spin-2 field, these components are the Stokes parameters $(Q,U)$, while for cosmic shear the two components are usually labelled $(\gamma_1,\gamma_2)$.}.
+Forming a complex number from these components, $a_1+ia_2$, spin-$s$ fields transform, under a coordinate rotation with angle $\psi$, as $a_1+ia_2\rightarrow (a_1+ia_2)\exp(i\,s\psi)$.
+Thus, spin-0 fields are invariant under rotations, and are usually expressed as real-valued fields with a single component.
+
+      Given a field ${\bf a}({\bf x})$, with 1 (spin-0) or 2 components (spin-2), defined on the coordinates ${\bf x}$, we define its generalized Fourier coefficients as
       \begin{equation}
@@
       \end{align}
-      where $\summ{k}$ denotes an integral over all possible generalised Fourier coefficients ${\bf k}$, and $\Delta^x$ and $\Delta^k$ are generalized delta functions, defined through their action on functions of ${\bf x}$ or ${\bf k}$:
+      where $\summ{k}$ denotes an integral over all possible generalised Fourier coefficients ${\bf k}$, and $\Delta^x$ and $\Delta^k$ are generalized delta functions, defined through their actionl on functions of ${\bf x}$ or ${\bf k}$:
       \begin{align}
@@
 
-      Finally, we will assume that all fields are Gaussian stochastic fields
-      that are additionally statistically isotropic. As a consequence of the
-      latter, different generalised Fourier modes are uncorrelated:
+      Finally, we will assume that all fields are Gaussian stochastic fields that are additionally statistically isotropic.
+As a consequence of the latter, different generalised Fourier modes are uncorrelated:
       \begin{equation}\label{eq:iso}
@@
           \(\displaystyle \bar{\eth}\,_sf\) & \(\displaystyle -(\sin\theta)^{-s}\left(\partial_\theta-i\frac{\partial_\varphi}{\sin\theta}\right)(\sin\theta)^s\,_sf(\nv)\) & \(\displaystyle (\partial_x+i\partial_y)\,_sf\) \\
-          \(\displaystyle \beta_{\ell,s}\) & \(\displaystyle \sqrt{\frac{(\ell-2)!}{(\ell+2)!}}\) & \(\displaystyle \ell^{-s}\) \\
+          \(\displaystyle \beta_{\ell,s}\) & \(\displaystyle \sqrt{\frac{(\ell-s)!}{(\ell+s)!}}\) & \(\displaystyle \ell^{-s}\) \\
           \(\displaystyle K \) & \(\displaystyle 1 \) & \(\displaystyle (2\pi)^{-1}\) \\
@@
     \subsection{Simulations}\label{ssec:results.sims}
+      In order to quantitatively study the performance of the analytical approximations for the power spectrum covariance matrix introduced in Section \ref{sec:theory}, we generate a large number of Gaussian simulations including a number of realistic observational effects.
+
+      Each simulation is a set of maps corresponding to a number of spin-0 and spin-2 fields that are drawn as Gaussian random fields following a set of input power spectra that include all relevant cross-correlations between different fields.
+We generate simulations for two types of fields modelled after the two main large-scale structure observables of photometric redshift surveys:
+      \begin{itemize}
+        \item Spin-0 fields, corresponding to maps of the overdensity of galaxies within a given redshift bin projected on the sphere.
+In keeping with the notation introduced in Section \ref{ssec:theory.pclcov}, we will label these fields as $\delta^a$, where the index $a$ denotes the redshift bin.
+        \item Spin-2 fields, corresponding to maps of the cosmic shear measured from the projected shapes of galaxies in a given redshift bin.
+We will label these fields $\boldsymbol{\gamma}^a=(\gamma_E^a,\gamma_B^a)$ where, again, the index $a$ denotes the redshift bin.
+      \end{itemize}
+      The cross-correlation between two of these fields can be written as:
+      \begin{equation}
+        C^{ab}_\ell = S^{ab}_\ell+N^{ab}_\ell,
+      \end{equation}
+      where $S$ and $N$ are the power spectra of the cosmological signal and noise respectively. We model the signal part for $\delta$ and $\gamma_E$ as:
+      \begin{equation}
+        S^{ab}_\ell = \int d\chi \frac{W^a_\ell(\chi)\,W^b_\ell(\chi)}{\chi^2}\,P\left(k=\frac{\ell+1/2}{\chi},z\right),
+      \end{equation}
+      where $\chi$ is the comoving radial distance, $z\equiv z(\chi)$ is the corresponding redshift in the lightcone, $P(k,z)$ is the matter power spectrum.
+The window functions are given by \cite{2004PhRvD..70d3009H}\footnote{See \cite{2019ApJS..242....2C} and references therein for details about these calculations.}:
+      \begin{align}
+        \nonumber
+        &W^\delta_\ell(\chi)=b(z)\,H(z)\,p_z(z),\\
+        &W^\gamma_\ell(\chi)=f_\ell\,\frac{3H_0^2\Omega_M}{2}(1+z)\chi\,\int dz'\,p_z(z')\,\frac{\chi(z')-\chi}{\chi(z')},
+      \end{align}
+      where $H(z)$ is the expansion rate in units where the speed of light is $c=1$, $H_0\equiv H(z=0)$, $b(z)$ is the linear galaxy bias, $p_z$ is the normalized redshift distribution of galaxies within the redshift bin, and
+      \begin{equation}
+        f_\ell\equiv\frac{\sqrt{(\ell+2)(\ell+1)\ell(\ell-1)}}{(\ell+1/2)^2}.
+      \end{equation}
+      We assume zero signal for the shear $B$-modes.
       \begin{figure}
@@
         \includegraphics[width=0.6\textwidth]{./figures/pz.pdf}
-        \caption{Redshift distributions assumed for the simulated maps used in this analysis.}
+        \caption{Redshift distributions assumed for the Gaussian simulations used in this analysis.}
         \label{fig:pz}
       \end{figure}
-
       \begin{figure*}
@@
       \end{figure*}      
-
       \begin{figure}
@@
       \end{figure}
-      In order to quantitatively study the performance of the analytical approximations for the power spectrum covariance matrix introduced in Section \ref{sec:theory}, we generate a large number of Gaussian simulations including a number of realistic observational effects.
-
-      Each simulation is a set of maps corresponding to a number of spin-0 and spin-2 fields that are drawn as Gaussian random fields following a set of input power spectra that include all relevant cross-correlations between different fields.
-We generate simulations for two types of fields modelled after the two main large-scale structure observables of photometric redshift surveys:
-      \begin{itemize}
-        \item Spin-0 fields, corresponding to maps of the overdensity of galaxies within a given redshift bin projected on the sphere.
-In keeping with the notation introduced in Section \ref{ssec:theory.pclcov}, we will label these fields as $\delta^a$, where the index $a$ denotes the redshift bin.
-        \item Spin-2 fields, corresponding to maps of the cosmic shear measured from the projected shapes of galaxies in a given redshift bin.
-We will label these fields $\boldsymbol{\gamma}^a=(\gamma_E^a,\gamma_B^a)$ where, again, the index $a$ denotes the redshift bin.
-      \end{itemize}
-      The cross-correlation between two of these fields can be written as:
-      \begin{equation}
-        C^{ab}_\ell = S^{ab}_\ell+N^{ab}_\ell,
-      \end{equation}
-      where $S$ and $N$ are the power spectra of the cosmological signal and noise. We model the signal part for $\delta$ and $\gamma_E$ as:
-      \begin{equation}
-        S^{ab}_\ell = \int d\chi \frac{W^a_\ell(\chi)\,W^b_\ell(\chi)}{\chi^2}\,P\left(k=\frac{\ell+1/2}{\chi},z\right),
-      \end{equation}
-      where $\chi$ is the comoving radial distance, $z\equiv z(\chi)$ is the corresponding redshift in the lightcone, $P(k,z)$ is the matter power spectrum and the window functions are given by \cite{2004PhRvD..70d3009H}\footnote{See \cite{2019ApJS..242....2C} and references therein for details about these calculations.}:
-      \begin{align}
-        \nonumber
-        &W^\delta_\ell(\chi)=b(z)\,H(z)\,p_z(z),\\
-        &W^\gamma_\ell(\chi)=f_\ell\,\frac{3H_0^2\Omega_M}{2}(1+z)\chi\,\int dz'\,p_z(z')\,\frac{\chi(z')-\chi}{\chi(z')},
-      \end{align}
-      where $H(z)$ is the expansion rate in units where the speed of light is $c=1$, $H_0\equiv H(z=0)$, $b(z)$ is the linear galaxy bias, $p_z$ is the normalized redshift distribution of galaxies within the redshift bin, and
-      \begin{equation}
-        f_\ell\equiv\frac{\sqrt{(\ell+2)(\ell+1)\ell(\ell-1)}}{(\ell+1/2)^2}.
-      \end{equation}
-      We assume zero signal for the shear $B$-modes.
 

### 2019-06-25 draft/paper.tex
done with text

@@
 
-    In this paper, we will present and generalise methods developed in the context of CMB experiments to account for the impact of survey geometry on the Gaussian part of the power spectrum covariance matrix \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}, and will study in detail the performance of these methods for large-scale structure and weak lensing datasets.
+    In this paper, we will present and generalise methods developed in the context of CMB experiments to account for the impact of survey geometry on the Gaussian part of the power spectrum covariance matrix \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}\footnote{See also \cite{2019JCAP...01..016L} for a similar application to the problem of 3D power spectrum covariances.}, and will study in detail the performance of these methods for large-scale structure and weak lensing datasets.
 We have also implemented these approximations in the public code {\tt NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}} \cite{2019MNRAS.484.4127A}, making the computation of accurate Gaussian covariance matrices significantly simpler for the community.
@@
 
-      We consider the case of auto- and cross-correlations between all fields
-      in two redshift bins, with redshift distributions modelled as Gaussians
-      with width $\sigma_z=0.13$ centered around redshifts $0.75$ and $0.95$,
-      with number densities $n_\Omega=7.5\,{\rm arcmin}^{-2}$. The
-      corresponding redshift distributions $p_z(z)$ are shown in Figure
-      \ref{fig:pz}. For simplicity we use a constant bias $b(z)=1$. For these
-      specifications, we generate signal power spectra using the Core
-      Cosmology Library ({\tt CCL} \cite{2019ApJS..242....2C}) for
-      cosmological parameters
-      $(\Omega_c,\Omega_b,h,A_s,n_s)=(0.27,0.045,0.67,2.1\times10^{-9},0.96)$.
-      The resulting non-zero signal and noise power spectra are shown in
-      Figure~\ref{fig:cl-2bins}.
+      We consider the case of auto- and cross-correlations between all fields in two redshift bins, with redshift distributions modelled as Gaussians with width $\sigma_z=0.13$ centered around redshifts $0.75$ and $0.95$, with number densities $n_\Omega=7.5\,{\rm arcmin}^{-2}$.
+The corresponding redshift distributions $p_z(z)$ are shown in Figure \ref{fig:pz}.
+For simplicity we use a constant bias $b(z)=1$.
+For these specifications, we generate signal power spectra using the Core Cosmology Library ({\tt CCL} \cite{2019ApJS..242....2C}) for cosmological parameters $(\Omega_c,\Omega_b,h,A_s,n_s)=(0.27,0.045,0.67,2.1\times10^{-9},0.96)$.
+The resulting non-zero signal and noise power spectra are shown in Figure~\ref{fig:cl-2bins}.
 
@@
       \end{itemize}
-      In both cases, we fixed the amplitude of the contaminant power spectrum
-      such that it would yield a $10\%$ contamination in the data power
-      spectrum at $\ell=400$. For spin-2 fields, we assumed the same power
-      spectrum for $E$ and $B$ modes, with no cross-correlation between them.
-      The resulting contaminant power spectra are shown in Figure
-      \ref{fig:clcont}. When exploring the effects of mode deprojection, we
-      generated 100 contaminant maps of both types and added them to each
-      simulated realization. We then deprojected the full set of 100
-      contaminant templates from the simulated maps and computed the
-      corresponding unbiased power spectra. Note that, in what follows, our
-      fiducial results do not include the effects of mode deprojection. These
-      are discussed separately.
+      In both cases, we fixed the amplitude of the contaminant power spectrum such that it would yield a $10\%$ contamination in the data power spectrum at $\ell=400$.
+For spin-2 fields, we assumed the same power spectrum for $E$ and $B$ modes, with no cross-correlation between them.
+The resulting contaminant power spectra are shown in Figure \ref{fig:clcont}.
+When exploring the effects of mode deprojection, we generated 100 contaminant maps of both types and added them to each simulated realization.
+We then deprojected the full set of 100 contaminant templates from the simulated maps and computed the corresponding unbiased power spectra.
+Note that, in what follows, our fiducial results do not include the effects of mode deprojection.
+These are discussed separately.
 
@@
         \includegraphics[width=\textwidth]{./figures/all_rows_sph_1bin.pdf}
-        \caption{Four rows of the covariance matrix for different power
-          spectra. The top panels show the cases of spectra with non-zero
-          signal, involving $\delta$ and $\gamma_E$, while the bottom panels
-          show cases involving $B$-modes. In each panel we show rows of the
-          covariance matrix for $\ell=60,\,90,\,120$ and $150$, which peak at
-          those central values. The different lines show the results for the
-          sample covariance matrix (solid blue), its NKA estimator (dashed
-          orange) and the difference between both (dotted green). The black
-          stars show the mode-counting approximation to the covariance matrix
-          (Eq. \ref{eq:cov_naive_bpw}). To facilitate the visualization of the
-          different rows, we have divided them by
-          $C_\ell^{aa}C_\ell^{bb}+(C^{ab}_\ell)^2$, the numerator of
-          Eq.~\ref{eq:cov_naive_bpw}, for the two fields, $a$ and $b$, that are being correlated in each case.
+        \caption{Four rows of the covariance matrix for different power spectra.
+The top panels show the cases of spectra with non-zero signal, involving $\delta$ and $\gamma_E$, while the bottom panels show cases involving $B$-modes.
+In each panel we show rows of the covariance matrix for $\ell=60,\,90,\,120$ and $150$, which peak at those central values.
+The different lines show the results for the sample covariance matrix (solid blue), its NKA estimator (dashed orange) and the difference between both (dotted green).
+The black stars show the mode-counting approximation to the covariance matrix (Eq.
+\ref{eq:cov_naive_bpw}).
+To facilitate the visualization of the different rows, we have divided them by $C_\ell^{aa}C_\ell^{bb}+(C^{ab}_\ell)^2$, the numerator of Eq.~\ref{eq:cov_naive_bpw}, for the two fields, $a$ and $b$, that are being correlated in each case.
 The NKA method is able to recover the covariance with high accuracy for all field combinations that do not involve $B$ modes, yielding visibly poorer results otherwise.}
@@
       \end{figure}
-      As a first step, we visually compare the main properties of the sample
-      covariance matrix estimated from the simulations and the NKA, Spin-0 and
-      MC approximations described in the previous section. Figure
-      \ref{fig:rows_1bin} shows four rows of the covariance matrix of
-      different auto- and cross-correlations. The rows correspond to
-      bandpowers centered on multipoles $\ell_q\simeq60,\,90,\,120$ and $150$.
-      The upper panels show results for the non-zero power spectra
-      ($\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$), with
-      the solid blue, dashed orange and dotted green lines showing results for
-      the sample covariance matrix, the analytical covariance using the NKA
-      approximation and their difference, respectively.
+      As a first step, we visually compare the main properties of the sample covariance matrix estimated from the simulations and the NKA, Spin-0 and MC approximations described in the previous section.
+Figure \ref{fig:rows_1bin} shows four rows of the covariance matrix of different auto- and cross-correlations.
+The rows correspond to bandpowers centered on multipoles $\ell_q\simeq60,\,90,\,120$ and $150$.
+The upper panels show results for the non-zero power spectra ($\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$), with the solid blue, dashed orange and dotted green lines showing results for the sample covariance matrix, the analytical covariance using the NKA approximation and their difference, respectively.
 For comparison, the black stars show the diagonal covariance matrix elements predicted by the MC approximation (Eq.
@@
 
-      We thus conclude that while the NKA estimator is able to recover the
-      covariance matrix for the non-zero power spectrum elements (i.e. those
-      involving $\delta$ and $\gamma_E$) with high accuracy, a more
-      sophisticated approach would be needed in order to obtain a precise
-      estimate of the uncertainties for components involving $B$-modes. This
-      is not a major concern, since $B$-mode power spectra are predominantly
-      used as null tests, while cosmological parameter constraints are driven
-      by the analysis of $\delta$ and $\gamma_E$.
+      We thus conclude that while the NKA estimator is able to recover the covariance matrix for the non-zero power spectrum elements (i.e.
+those involving $\delta$ and $\gamma_E$) with high accuracy, a more sophisticated approach would be needed in order to obtain a precise estimate of the uncertainties for components involving $B$-modes.
+This is not a major concern, since $B$-mode power spectra are predominantly used as null tests, while cosmological parameter constraints are driven by the analysis of $\delta$ and $\gamma_E$.
 For completeness, Figure~\ref{fig:corr_diff_2bins} shows the difference between the correlation matrices for the sample covariance and the NKA estimator for all non-zero cross-correlations between different bins in the case of two redshift bins with different small-scale masks, where we find a similarly good agreement.
@@
       \end{figure}
-      \begin{figure}
-        \centering
-        \includegraphics[width=\textwidth]{./figures/run_sph_2b_1stbin_chi2_TT_TE_EE.pdf}
-        \caption{$\chi^2$ distributions for the cases without
-          $\gamma_B$-modes. We compare the distributions obtained with the
-          covariance matrix computed directly from the simulated power spectra
-          (blue) and the two different analytical methods: NKA (orange) and
-          spin-0 (green). In addition, the theoretical $\chi^2$ distribution
-          has been included (dashed red). The distribution extracted from the
-          simulation follows the theoretical expectation almost perfectly in
-          all cases. In the cases of $\delta\times\gamma_E$ and
-          $\gamma_E\times\gamma_E$, we observe small shifts in the peak
-          $\chi^2$, of $\sim2$-$4\%$, while their width is recovered
-          accurately (a difference $\lesssim 2\%$). We will show that these
-          differences are due to the inaccuracy of the analytical
-          approximations on the largest scales, and that they have a
-          completely negligible impact on the final cosmological parameter
-          constraints.} 
-        \label{fig:chi2_1bin}
-      \end{figure}
-      \begin{figure}
-        \centering
-        \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_Spin0_NKA_TTTEEE_Full_chi2.pdf}
-        \caption{Same as Figure \ref{fig:chi2_1bin} for the combination of all
-          correlations between $\delta$ and $\gamma_E$ for two redshift bins,
-          where similar conclusions hold. In this case, the distribution
-          shift and width difference is less than a $3\%$ and $7\%$,
-          respectively.}
-        \label{fig:chi2_2bins}
-      \end{figure}
-      \begin{figure}
-        \centering
-        \includegraphics[width=\textwidth]{./figures/contaminants_diag_cov.pdf}
-        \caption{Diagonal of the covariance matrix estimated from simulations without contaminants (solid blue), with 200 contaminants deprojected (dashed orange) and the NKA estimator (solid green).
-The differences between the simulated cases are significantly smaller than those associated with the NKA estimator.
-The NKA method is therefore equally valid to approximate the covariance even in the presence of contaminant deprojection.}
-        \label{fig:diag_conts}
-      \end{figure}
       In order to quantify the validity of our analytical approximations, we need to compare the NKA covariance with the sample covariance estimated from the simulations.
@@
 
-      As a first test to quantify the differences between covariance matrices,
-      we compute the relative difference between their eigenvalues. This is
-      shown in Figure~\ref{fig:eigv_1bin} for a data vector combining all auto- and cross-correlations between $\delta$ and $\gamma_E$ for a single redshift bin.
+      As a first test to quantify the differences between covariance matrices, we compute the relative difference between their eigenvalues.
+This is shown in Figure~\ref{fig:eigv_1bin} for a data vector combining all auto- and cross-correlations between $\delta$ and $\gamma_E$ for a single redshift bin.
 The eigenvalues of both matrices are roughly similar, with relative differences of about $5\%$.
@@
       \end{equation}
-      We compute this quantity for a data vector ${\bf d}$ composed of
-      different auto- and cross-correlations for each of the 20,000 Gaussian
-      simulations, with ${\bf m}$ given by the mean over all simulations and
-      different choices of covariance matrix. Figure \ref{fig:chi2_1bin} shows
-      the distribution of $\chi^2$ values for the three non-zero power spectra
-      in the case of a single bin: $\delta$-$\delta$, $\delta$-$\gamma_E$ and
-      $\gamma_E$-$\gamma_E$. The histograms show the distribution for the
-      sample covariance matrix (blue), and the analytical NKA and Spin-0
-      estimators (orange and green, respectively).
+      \begin{figure}
+        \centering
+        \includegraphics[width=\textwidth]{./figures/run_sph_2b_1stbin_chi2_TT_TE_EE.pdf}
+        \caption{$\chi^2$ distributions for the cases without $\gamma_B$-modes.
+We compare the distributions obtained with the covariance matrix computed directly from the simulated power spectra (blue) and the two different analytical methods: NKA (orange) and spin-0 (green).
+In addition, the theoretical $\chi^2$ distribution has been included (dashed red).
+The distribution extracted from the simulation follows the theoretical expectation almost perfectly in all cases.
+In the cases of $\delta\times\gamma_E$ and $\gamma_E\times\gamma_E$, we observe small shifts in the peak $\chi^2$, of $\sim2$-$4\%$, while their width is recovered accurately (a difference $\lesssim 2\%$).
+We will show that these differences are due to the inaccuracy of the analytical approximations on the largest scales, and that they have a completely negligible impact on the final cosmological parameter constraints.}
+        \label{fig:chi2_1bin}
+      \end{figure}
+      \begin{figure}
+        \centering
+        \includegraphics[width=0.6\textwidth]{./figures/run_sph_2b_Spin0_NKA_TTTEEE_Full_chi2.pdf}
+        \caption{Same as Figure \ref{fig:chi2_1bin} for the combination of all correlations between $\delta$ and $\gamma_E$ for two redshift bins, where similar conclusions hold.
+In this case, the differences in the distribution means and widths are less than $3\%$ and $7\%$, respectively.}
+        \label{fig:chi2_2bins}
+      \end{figure}
+      We compute this quantity for a data vector ${\bf d}$ composed of different auto- and cross-correlations for each of the 20,000 Gaussian simulations, with ${\bf m}$ given by the mean over all simulations and different choices of covariance matrix.
+Figure \ref{fig:chi2_1bin} shows the distribution of $\chi^2$ values for the three non-zero power spectra in the case of a single bin: $\delta$-$\delta$, $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$.
+The histograms show the distribution for the sample covariance matrix (blue), and the analytical NKA and Spin-0 estimators (orange and green, respectively).
 We additionally plot the theoretical $\chi^2$ distribution under the assumption that the underlying data vector is Gaussianly distributed (red dashed lines).
@@
 In the cases involving the spin-2 fields, we see noticeable differences between the distributions found with the sample covariance and the approximate ones.
-      These differences are small, corresponding to less of a $2$ and $4\%$
-      shift in the mean $\chi^2$ for the $\delta$-$\gamma_E$ and
-      $\gamma_E$-$\gamma_E$ cases, respectively, and a negligible variation in
-      the width of the distributions. We therefore expect these differents to
-      have a negligible effect on the posterior parameter distributions, as we
-      show explicitly in Section \ref{ssec:results.cosmo}. The fact that these
-      differences appear only for power spectra involving spin-2 quantities
-      indicate that the NKA and spin-0 methods are imperfect at describing the
-      additional mode coupling caused in the presence of a mask for
-      higher-spin fields. This is rather obvious in the case of the spin-0
-      approximation but, interestingly, we find that the NKA and Spin-0
-      predictions yield results that are almost indistinguishable from each
-      other. We therefore conclude that the additional approximation made in
-      the NKA method for spin-2 fields -- neglecting the spatial derivatives
-      of the mask -- is effectively equivalent to ignoring the spin nature of
-      the fields involved \textcolor{red}{for the $\delta$ and $\gamma_E$ cases. The B
-        modes break this similarity and worsen the spin-0 approximation,
-        yielding, in the most extreme case ($\gamma_b$- $\gamma_b$), around an
-        order of magnitude worse results than the NKA method}. For
-        completeness, Figure \ref{fig:chi2_2bins} shows the distribution of
-        $\chi^2$ values for a data vector composed of all possible auto- and
-        cross-correlations of $\delta$ and $\gamma_E$ for the case of two
-        redshift bins, where similar conclusions hold.
+These differences are small, corresponding to less of a $2$ and $4\%$ shift in the mean $\chi^2$ for the $\delta$-$\gamma_E$ and $\gamma_E$-$\gamma_E$ cases, respectively, and a negligible variation in the width of the distributions.
+We therefore expect these differents to have a negligible effect on the posterior parameter distributions, as we show explicitly in Section \ref{ssec:results.cosmo}.
+The fact that these differences appear only for power spectra involving spin-2 quantities indicate that the NKA and spin-0 methods are imperfect at describing the additional mode coupling caused in the presence of a mask for higher-spin fields.
+This is rather obvious in the case of the spin-0 approximation but, interestingly, we find that the NKA and Spin-0 predictions yield results that are almost indistinguishable from each other.
+We therefore conclude that the additional approximation made in the NKA method for spin-2 fields -- neglecting the spatial derivatives of the mask -- is effectively equivalent to ignoring the spin nature of the fields involved.
+Note, however, that this is not the case for $B$-modes, where the NKA estimator outperforms the spin-0 approximation by up to one order of magnitude, even though its accuracy is very poor (as we described in the previous section).
+For completeness, Figure \ref{fig:chi2_2bins} shows the distribution of $\chi^2$ values for a data vector composed of all possible auto- and cross-correlations of $\delta$ and $\gamma_E$ for the case of two redshift bins, where similar conclusions hold.
 
@@
 
-      Figure \ref{fig:diag_conts} serves also to illustrate another important
-      point. In all our tests we find that the largest differences between the
-      sample and NKA covariances occur on large scales. For $\delta$-$\delta$
-      correlations, the effect is limited to the first few multipoles
-      ($\ell\lesssim10$), while for $\gamma_E$-$\gamma_E$ we are only able to
-      recover the sample covariance errors within $5\%$ for $\ell\gtrsim40$.
-      This is the main source of the small mismatch observed in
-      Figure~\ref{fig:chi2_1bin}. Note, however, that most of the cosmological
-      information is obtained from the higher multipoles, due to their higher
-      statistical weight, and, as we will show in the next section, the effect on the final parameter constraints is negligible.
+      \begin{figure}
+        \centering
+        \includegraphics[width=\textwidth]{./figures/contaminants_diag_cov.pdf}
+        \caption{Diagonal of the covariance matrix estimated from simulations without contaminants (solid blue), with 200 contaminants deprojected (dashed orange) and the NKA estimator (solid green).
+The differences between the simulated cases are significantly smaller than those associated with the NKA estimator.
+The NKA method is therefore equally valid to approximate the covariance even in the presence of contaminant deprojection.}
+        \label{fig:diag_conts}
+      \end{figure}
+      Figure \ref{fig:diag_conts} serves also to illustrate another important point.
+In all our tests we find that the largest differences between the sample and NKA covariances occur on large scales.
+For $\delta$-$\delta$ correlations, the effect is limited to the first few multipoles ($\ell\lesssim10$), while for $\gamma_E$-$\gamma_E$ we are only able to recover the sample covariance errors within $5\%$ for $\ell\gtrsim40$.
+This is the main source of the small mismatch observed in Figure~\ref{fig:chi2_1bin}.
+Note, however, that most of the cosmological information is obtained from the higher multipoles, due to their higher statistical weight, and, as we will show in the next section, the effect on the final parameter constraints is negligible.
 If accurate covariances bars are needed on these large scales for cosmic shear, they can be estimated alternatively making use of fast, low resolution simulations (e.g.
@@
         \caption{1$\sigma$ and 2$\sigma$ contours for $\Omega_M$ and $\sigma_8$ found from two different estimates of the power spectrum likelihood for a dataset containing all possible cross-power spectra between $\delta$ and $\gamma_E$ for two redshift bins.
-The filled contours corresponds to the parameter likelihood evaluated using the NKA estimator for the power spectrum covariance matrix, while the dashed black contours correspond to the same calculation done with the sample covariance matrix corrected by the factor in Eq.
+The filled contours correspond to the parameter likelihood evaluated using the NKA estimator for the power spectrum covariance matrix, while the dashed black contours correspond to the same calculation done with the sample covariance matrix corrected by the factor in Eq.
 \ref{eq:hartlap}.
@@
 
-      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for of two parameters
-      $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of
-      all possible auto- and cross-correlations between $\delta$ and
-      $\gamma_E$ in the case of two redshift bins. In this simple
-      two-dimensional scenario, we simply sample the distribution in a regular grid of 100 by 100 points for each parameter.
+      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for of two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins described in Section \ref{ssec:results.sims}.
+In this simple two-dimensional scenario, we simply sample the distribution in a regular grid of 100 by 100 points for each parameter.
 We construct a data vector $\hat{\bf C}$ from the theoretical prediction for the experimental setup described in Section \ref{ssec:results.sims}, and produce theoretical predictions for it at each grid point using {\tt CCL}.
@@
 
-      The 68\% and 95\% confidence level contours associated with the
-      posterior distributions for the sample covariance and the NKA estimator
-      are shown in Figure~\ref{fig:contours}.
+      The 68\% and 95\% confidence level contours associated with the posterior distributions for the sample covariance and the NKA estimator are shown in Figure~\ref{fig:contours}.
 We find that both distributions agree with each other remarkably well, and that the $1\sigma$ errors for each parameter agree for both covariances up to $0.3\%$.
-Note that, since we have not included any statistical noise in the data vector $\hat{\bf C}$, the relative bias between both distributions is zero by construction (since Eq.
+Note that, since we have not included any statistical noise in the data vector $\hat{\bf C}$, the relative difference in the means of both distributions is zero by construction (since Eq.
 \ref{eq:parpos} is bounded from below by zero).
@@
 
-    More in detail, we have also found that the impact of contaminant
-    deprojection on the covariance matrix, through the corresponding loss of
-    modes, is negligible unless a very large number of contaminant templates
-    are removed. This simplifies the procedure to estimate covariance matrices
-    for galaxy clustering data, which are particularly sensitive to a large
-    number of astrophysical and observational systematics. Additionally, we
-    have found that, although the NKA estimator is accurate enough, it is not
-    able to perfectly capture the additional effects of mode coupling that are
-    present for spin-2 fields, and that a simpler approach treating the shear
-    $E$ modes as a spin-0 object is able to reach similar levels of accuracy.
-    Due to the imperfect treatment of the $E/B$ mixing caused by the sky mask
-    in the NKA estimator, we also find that the predicted covariance matrix
-    for any power spectra involving $B$ modes differs significantly from the
-    true sample covariance, and, therefore, this approach cannot be used to reliably estimate the uncertainties of $B$-mode power spectra.
+    More in detail, we have also found that the impact of contaminant deprojection on the covariance matrix, through the corresponding loss of modes, is negligible unless a very large number of contaminant templates are removed.
+This simplifies the procedure to estimate covariance matrices for galaxy clustering data, which are particularly sensitive to a large number of astrophysical and observational systematics.
+Additionally, we have found that, although the NKA estimator is accurate enough, it is not able to perfectly capture the additional effects of mode coupling that are present for spin-2 fields, and that a simpler approach treating the shear $E$ modes as a spin-0 object is able to reach similar levels of accuracy.
+Due to the imperfect treatment of the $E/B$ mixing caused by the sky mask in the NKA estimator, we also find that the predicted covariance matrix for any power spectra involving $B$ modes differs significantly from the true sample covariance, and, therefore, this approach cannot be used to reliably estimate the uncertainties of $B$-mode power spectra.
 Likewise, we find that the NKA estimator yields inaccurate estimates of the power spectrum uncertainties on the largest scales ($\ell\lesssim50$).
@@
 
-    In spite of these shortcomings, we find that the approximations described
-    in this paper are able to provide estimates of the power spectrum
-    covariance matrix that are sufficiently accurate for current and future
-    tomographic large-scale structure cosmological datasets. The main
-    advantage of this approach is the computational cost, which is comparable
-    to that of estimating the power spectrum in the first place, and which
-    scales with pixel resolution in a similar way.
+    In spite of these shortcomings, we find that the approximations described in this paper are able to provide estimates of the power spectrum covariance matrix that are sufficiently accurate for current and future tomographic large-scale structure cosmological datasets.
+The main advantage of this approach is the computational cost, which is comparable to that of estimating the power spectrum in the first place, and which  scales with pixel resolution in a similar way.
 This is, therefore, significantly less time-consuming than generating large numbers of mock datasets (even simple Gaussian or log-normal realizations), and more reliable than the traditional jackknife resampling techniques.
@@
 The mask was constructed from the bright-object mask distributed with the first data release of the HSC collaboration \cite{2018PASJ...70S...4A} for the VVDS field.
-As in the full-sky case, it is used both for galaxy clustering and shear for simplicity.} \label{fig:mask_flat}
+As in the curved-sky case, it is used both for galaxy clustering and shear for simplicity.} \label{fig:mask_flat}
     \end{figure}
@@
       \includegraphics[width=\textwidth]{./figures/run_chi2_TT_TE_EE_TB_EB_BB.pdf}
-      \caption{Same as Figure \ref{fig:chi2_1bin} for flat-sky fields. This
-        figure additionally shows the $\chi^2$ distributions for power spectra
-        involving $B$-modes. For the smaller scales covered by these flat-sky
-        simulations, the NKA and spin-0 estimators work remarkably well, even
-        for $B$-mode spectra. The distribution mean shift and width difference
-        is of about $\lesssim 2\%$ for all cases.}
+      \caption{Same as Figure \ref{fig:chi2_1bin} for flat-sky fields.
+This figure additionally shows the $\chi^2$ distributions for power spectra involving $B$-modes.
+For the smaller scales covered by these flat-sky simulations, the NKA and spin-0 estimators work remarkably well, even for $B$-mode spectra.
+The distribution mean shift and width difference is of about $\lesssim 2\%$ for all cases.}
       \label{fig:chi2_1bin_flat}
     \end{figure}
-    We have repeated the analysis described in Section \ref{sec:results} on
-    flat-sky realizations, making use of the flat-sky implementation of {\tt
-      NaMaster}. We generate Gaussian realizations of the galaxy overdensity
-    and shear maps making use of flat-sky extensions of the methods described
-    in Section \ref{ssec:results.sims}. In this case we use a high-resolution
-    mask constructed from the bright-object mask distributed with the first
-    data release of the HSC collaboration \cite{2018PASJ...70S...4A} for the
-    VVDS field, which is shown in Figure~\ref{fig:mask_flat}.
+    We have repeated the analysis described in Section \ref{sec:results} on flat-sky realizations, making use of the flat-sky implementation of {\tt NaMaster}.
+We generate Gaussian realizations of the galaxy overdensity and shear maps making use of flat-sky extensions of the methods described in Section \ref{ssec:results.sims}.
+In this case we use a high-resolution mask constructed from the bright-object mask distributed with the first data release of the HSC collaboration \cite{2018PASJ...70S...4A} for the VVDS field, which is shown in Figure~\ref{fig:mask_flat}.
 
-    We find similar levels of accuracy in the NKA and spin-0 estimators
-    compared to the curved-sky case. The higher resolution and smaller area of
-    these simulations allow us to focus on the small-scale galaxy clustering
-    and lensing power spectra, covering the range of multipoles
-    $\ell\in(120,17640)$ in constant bandpowers of width $\Delta \ell=240$. On
-    these small scales, the shear power spectrum is more dominated by noise
-    (e.g. see Fig. \ref{fig:cl-2bins}), and therefore there is roughly the
-    same power in $E$ and $B$ modes. This reduces the sensitivity of the
-    method to an inaccurate treatment of $E/B$ leakage, and the agreement
-    between the sample covariance matrix and the NKA estimator improves
-    significantly.
+    We find similar levels of accuracy in the NKA and spin-0 estimators compared to the curved-sky case.
+The higher resolution and smaller area of these simulations allow us to focus on the small-scale galaxy clustering and lensing power spectra, covering the range of multipoles $\ell\in(120,17640)$ in constant bandpowers of width $\Delta \ell=240$.
+On these small scales, the shear power spectrum is more dominated by noise (e.g.
+see Fig.
+\ref{fig:cl-2bins}), and therefore there is roughly the same power in $E$ and $B$ modes.
+This reduces the sensitivity of the method to an inaccurate treatment of $E/B$ leakage, and the agreement between the sample covariance matrix and the NKA estimator improves significantly.
 This can be seen in Figure~\ref{fig:chi2_1bin_flat}, which shows the $\chi^2$ distributions for all possible power spectra (including those including $B$ modes) in the case of a single redshift bin.

### 2019-06-28 draft/preprint_JCAP_v0.1/paper.tex
added folder with first version submitted to arXiv and JCAP

--- 2019-06-25 draft/paper.tex
@@
 \abstract{The disconnected part of the power spectrum covariance matrix (also known as the ``Gaussian'' covariance) is the dominant contribution on large scales for galaxy clustering and weak lensing datasets.
-The presence of a complicated sky mask causes non-trivial correlations between different Fourier/harmonic modes, which must be accurately characterised in order to obtain reliable cosmological constraints.
+The presence of a complicated sky mask causes non-trivial correlations between different Fourier/harmonic modes, which must be accurately characterized in order to obtain reliable cosmological constraints.
 This is particularly relevant for galaxy survey data.
@@
 
-    In this paper, we will present and generalise methods developed in the context of CMB experiments to account for the impact of survey geometry on the Gaussian part of the power spectrum covariance matrix \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}\footnote{See also \cite{2019JCAP...01..016L} for a similar application to the problem of 3D power spectrum covariances.}, and will study in detail the performance of these methods for large-scale structure and weak lensing datasets.
+    In this paper, we will present and generalize methods developed in the context of CMB experiments to account for the impact of survey geometry on the Gaussian part of the power spectrum covariance matrix \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}\footnote{See also \cite{2019JCAP...01..016L} for a similar application to the problem of 3D power spectrum covariances.}, and will study in detail the performance of these methods for large-scale structure and weak lensing datasets.
 We have also implemented these approximations in the public code {\tt NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}} \cite{2019MNRAS.484.4127A}, making the computation of accurate Gaussian covariance matrices significantly simpler for the community.
@@
 In two dimensions, spin-$s$ fields in general have two components ${\bf a}({\bf x})=(a_1({\bf x}),a_2({\bf x}))$\footnote{E.g.
-for CMB polarisation, a spin-2 field, these components are the Stokes parameters $(Q,U)$, while for cosmic shear the two components are usually labelled $(\gamma_1,\gamma_2)$.}.
+for CMB polarization, a spin-2 field, these components are the Stokes parameters $(Q,U)$, while for cosmic shear the two components are usually labeled $(\gamma_1,\gamma_2)$.}.
 Forming a complex number from these components, $a_1+ia_2$, spin-$s$ fields transform, under a coordinate rotation with angle $\psi$, as $a_1+ia_2\rightarrow (a_1+ia_2)\exp(i\,s\psi)$.
@@
       \end{align}
-      where $\summ{k}$ denotes an integral over all possible generalised Fourier coefficients ${\bf k}$, and $\Delta^x$ and $\Delta^k$ are generalized delta functions, defined through their actionl on functions of ${\bf x}$ or ${\bf k}$:
+      where $\summ{k}$ denotes an integral over all possible generalized Fourier coefficients ${\bf k}$, and $\Delta^x$ and $\Delta^k$ are generalized delta functions, defined through their actionl on functions of ${\bf x}$ or ${\bf k}$:
       \begin{align}
@@
       Finally, we will assume that all fields are Gaussian stochastic fields that are additionally statistically isotropic.
-As a consequence of the latter, different generalised Fourier modes are uncorrelated:
+As a consequence of the latter, different generalized Fourier modes are uncorrelated:
       \begin{equation}\label{eq:iso}
@@
 
-      All the functions and operators above can be specialised to fields defined on the sphere or the 2D plane (flat sky approximation) as described in Table \ref{tab:notation}.
+      All the functions and operators above can be specialized to fields defined on the sphere or the 2D plane (flat sky approximation) as described in Table \ref{tab:notation}.
 
@@
       \begin{enumerate}
-        \item The narrow-kernel approximation (labelled {\bf NKA} here), described in the previous sections.
+        \item The narrow-kernel approximation (labeled {\bf NKA} here), described in the previous sections.
 This approximation assumes that the support of the harmonic-space masks (represented by the mode-coupling coefficients in e.g.
@@
         \item The {\bf spin-0} approximation corresponds to a simplified version of the NKA in which the spin nature of all fields involved is completely ignored, and all fields, including the $E$- and $B$-mode components of a spin-2 field, are treated as spin-0 quantities.
-        \item The mode-counting approximation (labelled {\bf MC} here),
+        \item The mode-counting approximation (labeled {\bf MC} here),
           commonly known as the {\sl Knox formula} \cite{1995PhRvD..52.4307K},
@@
       Each simulation is a set of maps corresponding to a number of spin-0 and spin-2 fields that are drawn as Gaussian random fields following a set of input power spectra that include all relevant cross-correlations between different fields.
-We generate simulations for two types of fields modelled after the two main large-scale structure observables of photometric redshift surveys:
+We generate simulations for two types of fields modeled after the two main large-scale structure observables of photometric redshift surveys:
       \begin{itemize}
@@
 
-      We consider the case of auto- and cross-correlations between all fields in two redshift bins, with redshift distributions modelled as Gaussians with width $\sigma_z=0.13$ centered around redshifts $0.75$ and $0.95$, with number densities $n_\Omega=7.5\,{\rm arcmin}^{-2}$.
+      We consider the case of auto- and cross-correlations between all fields in two redshift bins, with redshift distributions modeled as Gaussians with width $\sigma_z=0.13$ centered around redshifts $0.75$ and $0.95$, with number densities $n_\Omega=7.5\,{\rm arcmin}^{-2}$.
 The corresponding redshift distributions $p_z(z)$ are shown in Figure \ref{fig:pz}.

### 2019-09-26 draft/preprint_JCAP_v0.2/paper.tex
first inclusion of revisor comments to v0.1

--- 2019-06-28 draft/preprint_JCAP_v0.1/paper.tex
@@
 
+\newcommand{\cgg}[1]{{\color{olive} #1}}
+
 \newcommand{\nv}{\hat{\boldsymbol{\theta}}}
-\newcommand{\summ}[1]{\sum_{\bf #1}\Delta #1^2}
+\newcommand{\summ}[1]{\sum_{\bf #1}\Delta^2 #1\,}
 \newcommand{\Ylm}[3]{{\sf E}^{#1}_{\bf #2}({\bf #3})}
@@
 
-\abstract{The disconnected part of the power spectrum covariance matrix (also known as the ``Gaussian'' covariance) is the dominant contribution on large scales for galaxy clustering and weak lensing datasets.
-The presence of a complicated sky mask causes non-trivial correlations between different Fourier/harmonic modes, which must be accurately characterized in order to obtain reliable cosmological constraints.
-This is particularly relevant for galaxy survey data.
-Unfortunately, an exact calculation of these correlations involves $O(\ell_{\rm max}^6)$ operations that become computationally impractical very quickly.
-We present an implementation of approximate methods to estimate the Gaussian covariance matrix of power spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding on existing algorithms.
-These methods achieve an $O(\ell_{\rm max}^3)$ scaling, which makes the computation of the covariance matrix as fast as the computation of the power spectrum itself.
-We quantify the accuracy of these methods on large-scale structure and weak lensing data, making use of a large number of Gaussian but otherwise realistic simulations.
-We show that, using the approximate covariance matrix, we are able to recover the true posterior distribution of cosmological parameters to high accuracy.
-We also quantify the shortcomings of these methods, which become unreliable on the very largest scales, as well as for covariance matrix elements involving cosmic shear $B$ modes.
-The algorithms presented here are implemented in the public code {\tt NaMaster} \url{https://github.com/LSSTDESC/NaMaster}.}
+\abstract{The disconnected part of the power spectrum covariance matrix (also
+  known as the ``Gaussian'' covariance) is the dominant contribution on large
+  scales for galaxy clustering and weak lensing datasets. The presence of a
+  complicated sky mask causes non-trivial correlations between different
+  Fourier/harmonic modes, which must be accurately characterized in order to
+  obtain reliable cosmological constraints. This is particularly relevant for
+  galaxy survey data. Unfortunately, an exact calculation of these
+  correlations involves $O(\ell_{\rm max}^6)$ operations that become
+  computationally impractical very quickly. We present an implementation of
+  approximate methods to estimate the Gaussian covariance matrix of power
+  spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding
+  on existing algorithms \cgg{for CMB}. These methods achieve an $O(\ell_{\rm max}^3)$
+  scaling, which makes the computation of the covariance matrix as fast as the
+  computation of the power spectrum itself. We quantify the accuracy of these
+  methods on large-scale structure and weak lensing data, making use of a
+  large number of Gaussian but otherwise realistic simulations. We show that,
+  using the approximate covariance matrix, we are able to recover the true
+  posterior distribution of cosmological parameters to high accuracy. We also
+  quantify the shortcomings of these methods, which become unreliable on the
+  very largest scales, as well as for covariance matrix elements involving
+  cosmic shear $B$ modes. The algorithms presented here are implemented in the
+  public code {\tt NaMaster} \url{https://github.com/LSSTDESC/NaMaster}.}
 
@@
   \section{Introduction}\label{sec:intro}
-    The two-point correlation of different fields projected on the celestial sphere is one of the most common observables used in the analysis of large datasets in astrophysics, from studies of the Cosmic Microwave Background \citep{1995PhRvL..74.4369B,1997PhRvD..55.5895T,1998PhRvD..57.2117B,2001PhRvD..64f3001T,2002ApJ...567....2H,2003PhRvD..67b3001W} to large-scale structure and weak lensing surveys \cite{2000MNRAS.317L..23H,2001ApJ...555..547H,2001MNRAS.325.1603E,2011arXiv1112.5723H,2016MNRAS.456.1508K,2018MNRAS.476.1050B,2018arXiv181208182X,2019MNRAS.tmp.1446C,2019PASJ...71...43H}.
-Using these two-point functions, one achieves a high level of data compression (with respect to the size of the raw datasets -- time-ordered data, images or catalogs).
-They can also be directly used to constrain cosmological and astrophysical parameters assuming that one can model their likelihood.
-This is usually done by assuming that the two-point functions are Gaussianly distributed, which is often a good approximation due to the central limit theorem \cite{2008PhRvD..77j3013H,2018MNRAS.477.4879S}.
-In this case, the only obstacle that remains is being able to estimate the covariance matrix of a set of two-point correlators.
-Since the form of this covariance directly affects the posterior parameter uncertainties, a precise determination of it is of paramount importance.
-In large-scale structure experiments, this has often been resolved by making use of one's own data through resampling techniques \cite{tukey1958,1984MNRAS.210P..19B,1993ApJ...406L..47H,2002ApJ...571..172Z,2016arXiv160600233E}, or by generating a large number of mock realizations \cite{2013MNRAS.428.1036M,2018MNRAS.479...94A,2019MNRAS.485.2806B}.
-With the advent of the larger current and future surveys, the increasing size of the data vector and of the volume to be simulated has made this solution impractical, and fully analytical and hybrid estimators are now being used.
+    The two-point correlation of different fields projected on the celestial
+    sphere is one of the most common observables used in the analysis of large
+    datasets in astrophysics, from studies of the Cosmic Microwave Background
+    \cite{Gorski:1994ye,Gorski:1994uu,1995PhRvL..74.4369B,Gorski:1996cf,Gorski:1996ti,1997PhRvD..55.5895T,1998PhRvD..57.2117B,Wandelt:2000av,2001PhRvD..64f3001T,2002ApJ...567....2H,2003PhRvD..67b3001W}
+    to large-scale structure and weak lensing surveys
+    \cite{2000MNRAS.317L..23H,2001ApJ...555..547H,2001MNRAS.325.1603E,2011arXiv1112.5723H,2016MNRAS.456.1508K,2018MNRAS.476.1050B,2018arXiv181208182X,2019MNRAS.tmp.1446C,2019PASJ...71...43H}.
+    Using these two-point functions, one achieves a high level of data
+    compression (with respect to the size of the raw datasets -- time-ordered
+    data, images or catalogs). They can also be directly used to constrain
+    cosmological and astrophysical parameters assuming that one can model
+    their likelihood. This is usually done by assuming that the two-point
+    functions are Gaussianly distributed, which is often a good approximation
+    due to the central limit theorem
+    \cite{2008PhRvD..77j3013H,2018MNRAS.477.4879S}. In this case, the only
+    obstacle that remains is being able to estimate the covariance matrix of a
+    set of two-point correlators. Since the form of this covariance directly
+    affects the posterior parameter uncertainties, a precise determination of
+    it is of paramount importance. In large-scale structure experiments, this
+    has often been resolved by making use of one's own data through resampling
+    techniques
+    \cite{tukey1958,1984MNRAS.210P..19B,1993ApJ...406L..47H,2002ApJ...571..172Z,2016arXiv160600233E},
+    or by generating a large number of mock realizations
+    \cite{2013MNRAS.428.1036M,2018MNRAS.479...94A,2019MNRAS.485.2806B}. With
+    the advent of the larger current and future surveys, the increasing size
+    of the data vector and of the volume to be simulated has made this
+    solution impractical, and fully analytical and hybrid estimators are now
+    being used.
 
@@
       ($I^{0,0}$, $I^{0,\pm2}$, $I^{\pm2,\pm2}$ and $I^{\pm2,\mp2})$). In
-      order to simplify these expressions further, we follow \cite{2017A&A...602A..41C} and neglect all gradients of the masks, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
+      order to simplify these expressions further, we follow \cite{Efstathiou:2006eb,2017A&A...602A..41C} and neglect all gradients of the masks, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
       \begin{align}\nonumber
@@
 
-      For our fiducial results, we generated a set of 20,000 random simulations.
-This number was chosen in order to recover the covariance matrix for all possible auto- and cross-correlations with sufficient accuracy.
-All simulations were generated as {\tt HEALPix} maps with resolution $N_{\rm side}=512$.
+      For our fiducial results, we generated a set of 20,000 random
+      simulations. This number was chosen in order to recover the covariance
+      matrix for all possible auto- and cross-correlations with sufficient
+      accuracy. All simulations were generated as {\tt
+        HEALPix}\footnote{\url{http://healpix.sourceforge.net}} maps with resolution $N_{\rm side}=512$.
 We then used {\tt NaMaster} to compute all possible power spectra for each simulation using narrow bandpowers of width $\Delta\ell=3$ from $\ell=2$ to $\ell=1023$.
@@
       \end{figure}
-      Figure \ref{fig:diag_conts} serves also to illustrate another important point.
-In all our tests we find that the largest differences between the sample and NKA covariances occur on large scales.
-For $\delta$-$\delta$ correlations, the effect is limited to the first few multipoles ($\ell\lesssim10$), while for $\gamma_E$-$\gamma_E$ we are only able to recover the sample covariance errors within $5\%$ for $\ell\gtrsim40$.
-This is the main source of the small mismatch observed in Figure~\ref{fig:chi2_1bin}.
-Note, however, that most of the cosmological information is obtained from the higher multipoles, due to their higher statistical weight, and, as we will show in the next section, the effect on the final parameter constraints is negligible.
-If accurate covariances bars are needed on these large scales for cosmic shear, they can be estimated alternatively making use of fast, low resolution simulations (e.g.
-{\tt HEALPix} $N_{\rm side}=64$ maps).
+      Figure \ref{fig:diag_conts} serves also to illustrate another important
+      point. In all our tests we find that the largest differences between the
+      sample and NKA covariances occur on large scales. For $\delta$-$\delta$
+      correlations, the effect is limited to the first few multipoles
+      ($\ell\lesssim10$), while for $\gamma_E$-$\gamma_E$ we are only able to
+      recover the sample covariance errors within $5\%$ for $\ell\gtrsim40$.
+      This is the main source of the small mismatch observed in
+      Figure~\ref{fig:chi2_1bin}. Note, however, that most of the cosmological
+      information is obtained from the higher multipoles, due to their higher
+      statistical weight, and, as we will show in the next section, the effect
+      on the final parameter constraints is negligible. If accurate
+      \cgg{covariance matrix elements} are needed on these large scales for cosmic shear, they
+      can be estimated alternatively making use of fast, low resolution
+      simulations (e.g. {\tt HEALPix} $N_{\rm side}=64$ maps) \cgg{or computed
+      exactly as in Ref.~\cite{Efstathiou:2006eb,2017A&A...602A..41C}}.
 
@@
       \end{figure}
-      Ultimately, the most important test to judge the accuracy of the analytical covariance matrix estimators implemented here is to study their impact on the posterior distribution of cosmological parameters derived from power spectrum measurements.
-Assuming flat priors, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
+      Ultimately, the most important test to judge the accuracy of the
+      analytical covariance matrix estimators implemented here is to study
+      their impact on the posterior distribution of cosmological parameters
+      derived from power spectrum measurements. Assuming flat priors \cgg{and
+      a fiducial Gaussian likelihood approximation~\cite{2008PhRvD..77j3013H}}, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
       \begin{equation}\label{eq:parpos}
@@
 
-      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for of two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins described in Section \ref{ssec:results.sims}.
+      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for \cgg{the} two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins described in Section \ref{ssec:results.sims}.
 In this simple two-dimensional scenario, we simply sample the distribution in a regular grid of 100 by 100 points for each parameter.
@@
 
-    More in detail, we have also found that the impact of contaminant deprojection on the covariance matrix, through the corresponding loss of modes, is negligible unless a very large number of contaminant templates are removed.
-This simplifies the procedure to estimate covariance matrices for galaxy clustering data, which are particularly sensitive to a large number of astrophysical and observational systematics.
-Additionally, we have found that, although the NKA estimator is accurate enough, it is not able to perfectly capture the additional effects of mode coupling that are present for spin-2 fields, and that a simpler approach treating the shear $E$ modes as a spin-0 object is able to reach similar levels of accuracy.
-Due to the imperfect treatment of the $E/B$ mixing caused by the sky mask in the NKA estimator, we also find that the predicted covariance matrix for any power spectra involving $B$ modes differs significantly from the true sample covariance, and, therefore, this approach cannot be used to reliably estimate the uncertainties of $B$-mode power spectra.
-Likewise, we find that the NKA estimator yields inaccurate estimates of the power spectrum uncertainties on the largest scales ($\ell\lesssim50$).
-Although these modes carry a substantially smaller statistical weight, if more accurate covariances are needed on these large scales, they can be easily computed making use of fast low-resolution simulations.
+    More in detail, we have also found that the impact of contaminant
+    deprojection on the covariance matrix, through the corresponding loss of
+    modes, is negligible unless a very large number of contaminant templates
+    are removed. This simplifies the procedure to estimate covariance matrices
+    for galaxy clustering data, which are particularly sensitive to a large
+    number of astrophysical and observational systematics. Additionally, we
+    have found that, although the NKA estimator is accurate enough, it is not
+    able to perfectly capture the additional effects of mode coupling that are
+    present for spin-2 fields, and that a simpler approach treating the shear
+    $E$ modes as a spin-0 object is able to reach similar levels of accuracy.
+    Due to the imperfect treatment of the $E/B$ mixing caused by the sky mask
+    in the NKA estimator, we also find that the predicted covariance matrix
+    for any power spectra involving $B$ modes differs significantly from the
+    true sample covariance, and, therefore, this approach cannot be used to
+    reliably estimate the uncertainties of $B$-mode power spectra. Likewise,
+    we find that the NKA estimator yields inaccurate estimates of the power
+    spectrum uncertainties on the largest scales ($\ell\lesssim50$). Although
+    these modes carry a substantially smaller statistical weight, if more
+    accurate covariances are needed on these large scales, they can be easily
+    computed making use of fast low-resolution simulations \cgg{or exactly
+    following Refs.~\cite{Efstathiou:2006eb,2017A&A...602A..41C}}.
 
@@
   \acknowledgments
-  We thank Thibaut Louis, Eva-Maria Mueller, Francisco Javier S\'anchez and An\v ze Slosar for useful comments and discussion.
-CGG is supported the Spanish grant BES-2016-077038, partially funded by the ESF and by AYA2015-67854-P from the Ministry of Industry, Science and Innovation of Spain and the FEDER funds.
-He was partially supported by a Balzan Fellowship while in Oxford.
-He would like to thank New College and the Department of Physics at Oxford for their hospitality.
-DA acknowledges support from STFC through an Ernest Rutherford Fellowship, grant reference ST/P004474/1.
-EB acknowledges support from the Beecroft Trust.
+  We thank Thibaut Louis, Eva-Maria Mueller, Francisco Javier S\'anchez and
+  An\v ze Slosar for useful comments and discussion. CGG is supported the
+  Spanish grant BES-2016-077038, partially funded by the ESF and by
+  AYA2015-67854-P from the Ministry of Industry, Science and Innovation of
+  Spain and the FEDER funds. He was partially supported by a Balzan Fellowship
+  while in Oxford. He would like to thank New College and the Department of
+  Physics at Oxford for their hospitality. DA acknowledges support from STFC
+  through an Ernest Rutherford Fellowship, grant reference ST/P004474/1. EB
+  acknowledges support from the Beecroft Trust.  \cgg{Some of the results in this
+  paper have been derived using the HEALPix~\cite{Gorski:2004by} package}
 

### 2019-09-26 draft/preprint_JCAP_v0.2/reply-to-comments-to-v0.1/reply-to-comments-to-v0.1.tex
first inclusion of revisor comments to v0.1

\documentclass[a4paper]{article}
\usepackage[left=2cm, right=2cm, top=3cm, bottom=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}

\definecolor{midgreen}{rgb}{0.0,0.675,0.0}
\newcommand{\cgg}[1]{\textcolor{midgreen}{#1}}

\title{Addressing the reviser concerns}
\author{Carlos Garc\'ia-Garc\'ia, Pilar Ru\'iz-Lapuente, David Alonso and\\ Miguel
  Zumalac\'arregui}

\begin{document}
\maketitle
We have carefully addressed the referee's questions and concerns. The
modifications we have performed, for each of the numbered requests, are as
follow:
\begin{enumerate}
  \item  Abstract. Line 9. Added explicit reference to CMB. We have changed
    \textit{expanding on existing algorithms} to \textit{expanding on existing algorithms \cgg{for CMB}}.
  \item  Page 8. After Eq. 2.44. Added GPE2 reference.
  \item  \dots
  \item  Last sentence of 3.3. Added \textit{or computed exactly as in [41,
      44]}. In the second to last paragraph of Section 4, we added at the end
    \textit{or exactly following Refs.~[41, 44]}.
  \item  In Section 3.3, before previous sentence, we changed \textit{If accurate
    covariances bars} to \textit{If accurate covariance matrix \cgg{elements}}.
\item  In Section 3.4, before Eq. 3.8, I have mentioned the Gaussian
  approximation of the likelihood: \textit{Assuming flat priors \cgg{and a
      fiducial Gaussian likelihood approximation [21]}}.
  \item  Second paragraph of Section 3.4. Changed \textit{for of two
      parameters} to \textit{for the \cgg{two} parameters}.
\end{enumerate}

Changes related to the CMB background:
\begin{enumerate}
  \item  
\end{enumerate}

Extra changes:
\begin{enumerate}
  \item  Changed $\Delta x^2$ to $\Delta^2 x$.
  \item  Introduction. Line 3. Updated the set of references for CMB.
  \item  Added HEALPix project url and credits in the Acknowledgments.
\end{enumerate}

\end{document}

### 2019-09-27 notebooks/general-plots-paper.py
modified limits of Fig 6 and 7 in paper

@@ -503,7 +503,8 @@ def corr_diff_2bins_diff_mask():
 
     f, ax = plt.subplots(1, 1, figsize=(4, 4))
     nlbins = int(CorrSims.shape[0] / 4)
-    cb = ax.imshow(CorrTh - CorrSims, vmin=-0.02, vmax=0.02)
+    # cb = ax.imshow(CorrTh - CorrSims, vmin=-0.02, vmax=0.02)
+    cb = ax.imshow(CorrTh - CorrSims, vmin=-0.01, vmax=0.01)
     divider = make_axes_locatable(ax)
     cax = divider.append_axes("right", size="5%", pad=0.05)
     cbar = f.colorbar(cb, cax=cax)
@@ -519,7 +520,8 @@ def corr_diff_2bins_diff_mask():
     ax.set_yticklabels(labels, fontsize=8)
 
     plt.tight_layout()
-    fname = os.path.join(outdir, 'run_sph_2b_NKA_diff_corr_short.pdf')
+    # fname = os.path.join(outdir, 'run_sph_2b_NKA_diff_corr_short.pdf')
+    fname = os.path.join(outdir, 'run_sph_2b_NKA_diff_corr_short-lims-v0.2.pdf')
     plt.savefig(fname, dpi=DPI)
     # plt.show()
     plt.close()

### 2019-09-27 notebooks/row-sph-1bin-cov.py

@@ -182,7 +182,8 @@ CorrTh = Cth/np.sqrt(np.diag(Cth)[:, None] * np.diag(Cth)[None, :])
 
 f, ax = plt.subplots(1, 1, figsize=(4, 4))
 
-cb = ax.imshow(CorrTh - CorrSims, vmin=-0.02, vmax=0.02)
+# cb = ax.imshow(CorrTh - CorrSims, vmin=-0.02, vmax=0.02)
+cb = ax.imshow(CorrTh - CorrSims, vmin=-0.01, vmax=0.01)
 divider = make_axes_locatable(ax)
 cax = divider.append_axes("right", size="5%", pad=0.05)
 cbar = f.colorbar(cb, cax=cax)
@@ -198,7 +199,8 @@ ax.set_yticks(ticks)
 ax.set_yticklabels(labels, fontsize=8)
 
 plt.tight_layout()
-fname = os.path.join(outdir, 'sph_1bin_diff_corr.pdf')
+# fname = os.path.join(outdir, 'sph_1bin_diff_corr.pdf')
+fname = os.path.join(outdir, 'sph_1bin_diff_corr-lims-v0.2.pdf')
 plt.savefig(fname, dpi=DPI)
 # plt.show()
 plt.close()

### 2019-09-27 draft/preprint_JCAP_v0.2/paper.tex
implemented more changes from revision to v0.1

@@
       \begin{enumerate}
-        \item We first bin different ${\bf l}$ modes into sets of them called bandpowers (typically bands of similar $\ell$ or annuli of flat-sky Fourier modes spanning a range of radii).
-Let us denote a given bandpower by its index $q$.
-The binned pseudo-power spectrum is:
+        \item We first bin different ${\bf l}$ modes into sets of them called
+          bandpowers (typically bands of similar $\ell$ or annuli of flat-sky
+          Fourier modes spanning a range of radii). Let us denote a given
+          bandpower by its index $q$. \cgg{The number of bandpowers and their
+            width are determined by the true underlying power spectra, which
+            must not vary much within the bandpower}. The binned pseudo-power spectrum is:
         \begin{equation}\label{eq:bandpowers}
@@
       accuracy. All simulations were generated as {\tt
-        HEALPix}\footnote{\url{http://healpix.sourceforge.net}} maps with resolution $N_{\rm side}=512$.
+        HEALPix}\footnote{\url{http://healpix.sourceforge.net}}~\cite{Gorski:2004by} maps with resolution $N_{\rm side}=512$.
 We then used {\tt NaMaster} to compute all possible power spectra for each simulation using narrow bandpowers of width $\Delta\ell=3$ from $\ell=2$ to $\ell=1023$.
@@
         \centering
-        \includegraphics[width=0.7\textwidth]{./figures/sph_1bin_diff_corr.pdf}
+        \includegraphics[width=0.7\textwidth]{./figures/sph_1bin_diff_corr-lims-v02.pdf}
         \caption{Difference between the correlation matrices associated to the sample covariance matrix and its estimate with the NKA method including all possible correlations between $\delta$ and $\boldsymbol{\gamma}$ in a single redshift bin.
@@
         \centering
-        \includegraphics[width=0.9\textwidth]{./figures/run_sph_2b_NKA_diff_corr_short.pdf}
+        \includegraphics[width=0.9\textwidth]{./figures/run_sph_2b_NKA_diff_corr_short-lims-v02.pdf}
         \caption{Same as Fig. \ref{fig:diff_corr_1bin} for all cross-correlations between $\delta$ and $\gamma_E$ measured in two different redshift bins.}
@@
       where $\hat{\bf C}$ is a vector of power spectrum measurements, ${\rm Cov}$ is their covariance matrix, and ${\bf C}(\vec{\theta})$ is their theoretical prediction for parameters $\vec{\theta}$.
+      \cgg{Note that the low-$\ell$ non-gaussianities of real power spectra
+        are not being taking into account with this approximation. They,
+        however, have a lower statistical weight and will contribute less to
+        the posterior distribution.}
 
@@
       \end{equation}
-      where $N_s=20,000$ is the number of samples used to estimate the covariance, and $N_{\rm data}=3510$ is the number of data points.
+      where $N_s=20,000$ is the number of samples used to estimate the
+      covariance, and $N_{\rm data}=3510$ is the number of data points.
 
@@
   \section{Discussion}\label{sec:discussion}
-    Estimating accurate covariance matrices for projected two-point correlators is an ubiquitous problem in modern cosmology \cite{2013PhRvD..88f3537D,2013MNRAS.432.1928T,2017arXiv170609359K}, but it is particularly relevant for large-scale structure datasets.
-This problem is further complicated in this case, in comparison with e.g.
-CMB experiments, by two factors: the fact that the fields involved are non-Gaussian at some level and the higher complexity of the sky masks used in optical datasets.
-The main impact of survey geometry is the statistical coupling it induces between different Fourier/harmonic modes, which must be accurately characterized in order to obtain reliable estimates of the posterior distribution for cosmological parameters.
-In this paper we have focused on the impact of survey geometry on the dominant Gaussian (i.e.
+    Estimating accurate covariance matrices for projected two-point
+    correlators is an ubiquitous problem in modern cosmology
+    \cite{2013PhRvD..88f3537D,2013MNRAS.432.1928T,2017arXiv170609359K}, but it
+    is particularly relevant for large-scale structure datasets. This problem
+    is further complicated in this case, in comparison with e.g. CMB
+    experiments, by 
+    %two factors: 
+    the fact that the fields involved are
+    non-Gaussian at some level
+    %and the higher complexity of the sky masks used in optical datasets.
+    The main impact of survey geometry is the statistical
+    coupling it induces between different Fourier/harmonic modes, which must
+    be accurately characterized in order to obtain reliable estimates of the
+    posterior distribution for cosmological parameters. In this paper we have
+    focused on the impact of survey geometry on the dominant Gaussian (i.e.
     disconnected) part of the covariance matrix.
 
-    We have described and generalized existing analytical approaches to estimate the covariance matrix for pseudo-$C_\ell$ power spectrum estimators \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}, and implemented them in the public code {\tt NaMaster} \cite{2019MNRAS.484.4127A}, making it straightforward to fully account for the effects of survey geometry on the data uncertainties.
-With these approximation, computationally speaking, the problem of estimating a covariance matrix is as complex as that of computing the power spectrum itself, and scales with the number of pixels in the map as $N_{\rm pix}^{3/2}$.
+    We have described and generalized existing analytical approaches to
+    estimate the covariance matrix for pseudo-$C_\ell$ power spectrum
+    estimators
+    \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}, and
+    implemented them in the public code {\tt NaMaster}
+    \cite{2019MNRAS.484.4127A}, making it straightforward to fully account for
+    the effects of survey geometry on the data uncertainties. With these
+    approximation\cgg{s}, computationally speaking, the problem of estimating a
+    covariance matrix is as complex as that of computing the power spectrum
+    itself, and scales with the number of pixels in the map as $N_{\rm
+      pix}^{3/2}$. \cgg{We leave for future work the study of the impact of
+      extra complications, as position-dependent noise, on their performance.}
 
-    The main finding of this paper is the excellent performance of the analytical methods described in Section \ref{ssec:theory.pclcov}.
-We have shown that, in spite of the higher complexity of large-scale structure masks, the NKA estimator (see Section \ref{ssec:theory.approx}) is able to recover the covariance matrix for all power spectra with non-zero signal expectation value (i.e.
-those involving the galaxy overdensity $\delta$ or the $E$-mode shear $\gamma_E$), as well as the posterior distribution for cosmological parameters, to a high degree of accuracy.
+    The main finding of this paper is the excellent performance of the
+    analytical methods described in Section \ref{ssec:theory.pclcov}. We have
+    shown that
+    %, in spite of the higher complexity of large-scale structure masks,
+    the NKA estimator (see Section \ref{ssec:theory.approx}) is able to
+    recover the covariance matrix for all power spectra with non-zero signal
+    expectation value (i.e. those involving the galaxy overdensity $\delta$ or
+    the $E$-mode shear $\gamma_E$), as well as the posterior distribution for
+    cosmological parameters, to a high degree of accuracy.
 

### 2019-09-27 draft/preprint_JCAP_v0.2/reply-to-comments-to-v0.1/reply-to-comments-to-v0.1.tex
implemented more changes from revision to v0.1

@@
 \title{Addressing the reviser concerns}
-\author{Carlos Garc\'ia-Garc\'ia, Pilar Ru\'iz-Lapuente, David Alonso and\\ Miguel
-  Zumalac\'arregui}
+\author{Carlos Garc\'ia-Garc\'ia, David Alonso and Emilio Bellini}
 
@@
   \item  Page 8. After Eq. 2.44. Added GPE2 reference.
-  \item  \dots
+  \item  Changed limits of Figure 6 and 7.
   \item  Last sentence of 3.3. Added \textit{or computed exactly as in [41,
@@
       parameters} to \textit{for the \cgg{two} parameters}.
+  \item  At the end of the first paragraph of Section 3.4,  after Eq. 3.8, I
+    have added \textit{\cgg{Note that the low-$\ell$ non-gaussianities of real power spectra
+        are not being taking into account with this approximation. They,
+        however, have a lower statistical weight and will contribute less to
+        the posterior distribution.}
+}
+
 \end{enumerate}
@@
   \item  
+  \item In Section 4, removed the statements about the higher complexity of LSS
+    masks.
+  \item  In the second paragraph of Section 4, we have added a comment on the
+    effect of "real-wold" complications: \textit{With these
+    approximation\cgg{s}, computationally speaking, the problem of estimating a
+    covariance matrix is as complex as that of computing the power spectrum
+    itself, and scales with the number of pixels in the map as $N_{\rm
+      pix}^{3/2}$. \cgg{We leave for future work the study of the impact of
+      extra complications, as position-dependent noise, on their performance.}
+}
 \end{enumerate}
@@
   \item  Changed $\Delta x^2$ to $\Delta^2 x$.
+  \item  Before Eq. 2.12, added a comment on how the true power spectra must
+    not change much over a bandpower. It says now \textit{Let us denote a given
+          bandpower by its index $q$. \cgg{The number of bandpowers and their
+            width are determined by the true underlying power spectra, which
+            must not vary much within the bandpower}.}
   \item  Introduction. Line 3. Updated the set of references for CMB.
-  \item  Added HEALPix project url and credits in the Acknowledgments.
+  \item  Added HEALPix project url, reference and credits in the
+    Acknowledgments.
 \end{enumerate}

### 2019-10-17 draft/preprint_JCAP_v0.2/paper.tex
finished comments

@@
 
-\newcommand{\cgg}[1]{{\color{olive} #1}}
+\newcommand{\cgg}[1]{{\color{red} #1}}
 
@@
   spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding
-  on existing algorithms \cgg{for CMB}. These methods achieve an $O(\ell_{\rm max}^3)$
+  on existing algorithms \cgg{developed in the context of CMB analyses}.
+  These methods achieve an $O(\ell_{\rm max}^3)$
   scaling, which makes the computation of the covariance matrix as fast as the
@@
 
-    In this paper, we will present and generalize methods developed in the context of CMB experiments to account for the impact of survey geometry on the Gaussian part of the power spectrum covariance matrix \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2017A&A...602A..41C}\footnote{See also \cite{2019JCAP...01..016L} for a similar application to the problem of 3D power spectrum covariances.}, and will study in detail the performance of these methods for large-scale structure and weak lensing datasets.
+    In this paper, we will present and generalize methods developed in the context of CMB experiments to account for the impact of survey geometry on the Gaussian part of the power spectrum covariance matrix \cite{2004MNRAS.349..603E,2005MNRAS.360.1262B,2005MNRAS.360..509C,Efstathiou:2006eb,2014A&A...571A..15P,2016A&A...594A..11P,2017A&A...602A..41C,2019arXiv190712875P}\footnote{See also \cite{2019JCAP...01..016L} for a similar application to the problem of 3D power spectrum covariances.}, and will study in detail the performance of these methods for large-scale structure and weak lensing datasets.
 We have also implemented these approximations in the public code {\tt NaMaster}\footnote{\url{https://github.com/LSSTDESC/NaMaster}} \cite{2019MNRAS.484.4127A}, making the computation of accurate Gaussian covariance matrices significantly simpler for the community.
@@
       \begin{enumerate}
-        \item We first bin different ${\bf l}$ modes into sets of them called
-          bandpowers (typically bands of similar $\ell$ or annuli of flat-sky
-          Fourier modes spanning a range of radii). Let us denote a given
-          bandpower by its index $q$. \cgg{The number of bandpowers and their
-            width are determined by the true underlying power spectra, which
-            must not vary much within the bandpower}. The binned pseudo-power spectrum is:
+        \item We first bin different ${\bf l}$ modes into sets of them called bandpowers (typically bands of similar $\ell$ or annuli of flat-sky Fourier modes spanning a range of radii).
+Let us denote a given bandpower by its index $q$.
+\cgg{We must note that it is more appropriate to use bandpower-averaged spectra when the underlying spectrum does not vary much within each $\ell$ bin.
+When this is not the case, it is often useful to apply $\ell$-weights (such as $D_\ell\equiv\ell(\ell+1)C_\ell/(2\pi)$).
+The large-scale structure spectra discussed here are sufficiently flat that the binning used is appropriate.} The binned pseudo-power spectrum is:
         \begin{equation}\label{eq:bandpowers}
@@
       ($I^{0,0}$, $I^{0,\pm2}$, $I^{\pm2,\pm2}$ and $I^{\pm2,\mp2})$). In
-      order to simplify these expressions further, we follow \cite{Efstathiou:2006eb,2017A&A...602A..41C} and neglect all gradients of the masks, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
+      order to simplify these expressions further, we follow \cite{Efstathiou:2006eb,2017A&A...602A..41C} and neglect all gradients of the masks\footnote{\cgg{It is worth noting that it is possible in principle to avoid this approximation, as demonstrated in \cite{2005MNRAS.360..509C}}.}, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
       \begin{align}\nonumber
@@
       can be estimated alternatively making use of fast, low resolution
-      simulations (e.g. {\tt HEALPix} $N_{\rm side}=64$ maps) \cgg{or computed
+      simulations (e.g. {\tt HEALPix} $N_{\rm side}=64$ maps)\cgg{, or computed
       exactly as in Ref.~\cite{Efstathiou:2006eb,2017A&A...602A..41C}}.
@@
       derived from power spectrum measurements. Assuming flat priors \cgg{and
-      a fiducial Gaussian likelihood approximation~\cite{2008PhRvD..77j3013H}}, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
+      a fiducial Gaussian likelihood approximation~\cite{2008PhRvD..77j3013H}, in which the covariance matrix is computed only once for the fiducial model}, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
       \begin{equation}\label{eq:parpos}
@@
       where $\hat{\bf C}$ is a vector of power spectrum measurements, ${\rm Cov}$ is their covariance matrix, and ${\bf C}(\vec{\theta})$ is their theoretical prediction for parameters $\vec{\theta}$.
-      \cgg{Note that the low-$\ell$ non-gaussianities of real power spectra
-        are not being taking into account with this approximation. They,
-        however, have a lower statistical weight and will contribute less to
-        the posterior distribution.}
+\cgg{It is worth noting that this Gaussian likelihood is not accurate on large scales, where the small number of modes invalidates the application of the central limit theorem.
+Even in this regime, the likelihood can be approximated through the method described in \cite{2008PhRvD..77j3013H}, which still requires an accurate estimate of the power spectrum covariance.}
 
@@
     is further complicated in this case, in comparison with e.g. CMB
-    experiments, by 
-    %two factors: 
+    experiments, by two factors: 
     the fact that the fields involved are
-    non-Gaussian at some level
-    %and the higher complexity of the sky masks used in optical datasets.
+    non-Gaussian at some level and the \cgg{arguably} higher complexity of the sky masks used in optical datasets.
     The main impact of survey geometry is the statistical

### 2019-11-25 draft/preprint_JCAP_v0.2.1-for-arXiv/paper.tex
accepted version of paper in JCAP for arXiv without coloured changes

--- 2019-10-17 draft/preprint_JCAP_v0.2/paper.tex
@@
   spectra involving spin-0 and spin-2 flat- and curved-sky fields, expanding
-  on existing algorithms \cgg{developed in the context of CMB analyses}.
+  on existing algorithms developed in the context of CMB analyses.
   These methods achieve an $O(\ell_{\rm max}^3)$
@@
       \end{align}
-      where $\summ{k}$ denotes an integral over all possible generalized Fourier coefficients ${\bf k}$, and $\Delta^x$ and $\Delta^k$ are generalized delta functions, defined through their actionl on functions of ${\bf x}$ or ${\bf k}$:
+      where $\summ{k}$ denotes an integral over all possible generalized Fourier coefficients ${\bf k}$, and $\Delta^x$ and $\Delta^k$ are generalized delta functions, defined through their action on functions of ${\bf x}$ or ${\bf k}$:
       \begin{align}
@@
 Let us denote a given bandpower by its index $q$.
-\cgg{We must note that it is more appropriate to use bandpower-averaged spectra when the underlying spectrum does not vary much within each $\ell$ bin.
+We must note that it is more appropriate to use bandpower-averaged spectra when the underlying spectrum does not vary much within each $\ell$ bin.
 When this is not the case, it is often useful to apply $\ell$-weights (such as $D_\ell\equiv\ell(\ell+1)C_\ell/(2\pi)$).
-The large-scale structure spectra discussed here are sufficiently flat that the binning used is appropriate.} The binned pseudo-power spectrum is:
+The large-scale structure spectra discussed here are sufficiently flat that the binning used is appropriate.
+The binned pseudo-power spectrum is:
         \begin{equation}\label{eq:bandpowers}
@@
       ($I^{0,0}$, $I^{0,\pm2}$, $I^{\pm2,\pm2}$ and $I^{\pm2,\mp2})$). In
-      order to simplify these expressions further, we follow \cite{Efstathiou:2006eb,2017A&A...602A..41C} and neglect all gradients of the masks\footnote{\cgg{It is worth noting that it is possible in principle to avoid this approximation, as demonstrated in \cite{2005MNRAS.360..509C}}.}, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
+      order to simplify these expressions further, we follow \cite{Efstathiou:2006eb,2017A&A...602A..41C} and neglect all gradients of the masks\footnote{It is worth noting that it is possible in principle to avoid this approximation, as demonstrated in \cite{2005MNRAS.360..509C}.}, which allows us to relate the different $I^{\pm s_a,\pm s_b}$ through the following set of identities:
       \begin{align}\nonumber
@@
       on the final parameter constraints is negligible. If accurate
-      \cgg{covariance matrix elements} are needed on these large scales for cosmic shear, they
+      covariance matrix elements are needed on these large scales for cosmic shear, they
       can be estimated alternatively making use of fast, low resolution
-      simulations (e.g. {\tt HEALPix} $N_{\rm side}=64$ maps)\cgg{, or computed
-      exactly as in Ref.~\cite{Efstathiou:2006eb,2017A&A...602A..41C}}.
+      simulations (e.g. {\tt HEALPix} $N_{\rm side}=64$ maps), or computed
+      exactly as in Ref.~\cite{Efstathiou:2006eb,2017A&A...602A..41C}.
 
@@
       their impact on the posterior distribution of cosmological parameters
-      derived from power spectrum measurements. Assuming flat priors \cgg{and
-      a fiducial Gaussian likelihood approximation~\cite{2008PhRvD..77j3013H}, in which the covariance matrix is computed only once for the fiducial model}, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
+      derived from power spectrum measurements. Assuming flat priors and
+      a fiducial Gaussian likelihood approximation~\cite{2008PhRvD..77j3013H}, in which the covariance matrix is computed only once for the fiducial model, the posterior distribution for parameters $\vec{\theta}$ is simply given by:
       \begin{equation}\label{eq:parpos}
@@
       where $\hat{\bf C}$ is a vector of power spectrum measurements, ${\rm Cov}$ is their covariance matrix, and ${\bf C}(\vec{\theta})$ is their theoretical prediction for parameters $\vec{\theta}$.
-\cgg{It is worth noting that this Gaussian likelihood is not accurate on large scales, where the small number of modes invalidates the application of the central limit theorem.
-Even in this regime, the likelihood can be approximated through the method described in \cite{2008PhRvD..77j3013H}, which still requires an accurate estimate of the power spectrum covariance.}
+It is worth noting that this Gaussian likelihood is not accurate on large scales, where the small number of modes invalidates the application of the central limit theorem.
+Even in this regime, the likelihood can be approximated through the method described in \cite{2008PhRvD..77j3013H}, which still requires an accurate estimate of the power spectrum covariance.
 
-      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for \cgg{the} two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins described in Section \ref{ssec:results.sims}.
+      We explore $\log p(\vec{\theta}|\hat{\bf C})$ for the two parameters $\vec{\theta}\equiv(\Omega_M,\sigma_8)$, for a data vector composed of all possible auto- and cross-correlations between $\delta$ and $\gamma_E$ in the case of two redshift bins described in Section \ref{ssec:results.sims}.
 In this simple two-dimensional scenario, we simply sample the distribution in a regular grid of 100 by 100 points for each parameter.
@@
     the fact that the fields involved are
-    non-Gaussian at some level and the \cgg{arguably} higher complexity of the sky masks used in optical datasets.
+    non-Gaussian at some level and the arguably higher complexity of the sky masks used in optical datasets.
     The main impact of survey geometry is the statistical
@@
     the effects of survey geometry on the data uncertainties. With these
-    approximation\cgg{s}, computationally speaking, the problem of estimating a
+    approximations, computationally speaking, the problem of estimating a
     covariance matrix is as complex as that of computing the power spectrum
     itself, and scales with the number of pixels in the map as $N_{\rm
-      pix}^{3/2}$. \cgg{We leave for future work the study of the impact of
-      extra complications, as position-dependent noise, on their performance.}
+      pix}^{3/2}$. We leave for future work the study of the impact of
+      extra complications, as position-dependent noise, on their performance.
 
@@
     accurate covariances are needed on these large scales, they can be easily
-    computed making use of fast low-resolution simulations \cgg{or exactly
-    following Refs.~\cite{Efstathiou:2006eb,2017A&A...602A..41C}}.
+    computed making use of fast low-resolution simulations or exactly
+    following Refs.~\cite{Efstathiou:2006eb,2017A&A...602A..41C}.
 
@@
   through an Ernest Rutherford Fellowship, grant reference ST/P004474/1. EB
-  acknowledges support from the Beecroft Trust.  \cgg{Some of the results in this
-  paper have been derived using the HEALPix~\cite{Gorski:2004by} package}
+  acknowledges support from the Beecroft Trust.  Some of the results in this
+  paper have been derived using the HEALPix~\cite{Gorski:2004by} package.
 

### 2019-11-25 draft/preprint_JCAP_v0.2.1-for-arXiv/reply-to-comments-to-v0.1/reply-to-comments-to-v0.1.tex
accepted version of paper in JCAP for arXiv without coloured changes

\documentclass[a4paper]{article}
\usepackage[left=2cm, right=2cm, top=3cm, bottom=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}

\definecolor{midgreen}{rgb}{0.0,0.675,0.0}
\newcommand{\cgg}[1]{\textcolor{midgreen}{#1}}

\title{Addressing the reviser concerns}
\author{Carlos Garc\'ia-Garc\'ia, David Alonso and Emilio Bellini}

\begin{document}
\maketitle
We would like to thank the referee for their careful reading of our manuscript.
We have addressed the points raised as described below. All modifications are
highlighted in red in the resubmitted version of the manuscript.

With regards to the specific points raised by the referee:
\begin{enumerate}
  \item We have added an explicit reference to the CMB expertise in the abstract.
  \item We have added a citation to GPE2 after Eq. 2.44.
  \item We have reduced the range of Figures 6 and 7 to highlight the differences in the correlation matrices.
  \item We have included a reference to GPE2 in the context of low-ell covariance estimation.
  \item Thanks for noticing this. This is a typo (left over from a previous version of this sentence that read ``error bars'' instead of ``covariance''). We have changed this sentence as suggested.
  \item We have added the {\sl fiducial Gaussian likelihood} caveat in Section 3.4. We are aware of the biases associated with varying the covariance at each point in parameter space.
  \item Thanks for noticing this typo. This has been fixed.
  \item We have added a caveat about the appropriateness of the Gaussian approximation. We have also noted that the approximate likelihood of Hamimeche \& Lewis also require an accurate evaluation of the power spectrum covariance, and therefore these results are still valuable in that case.
\end{enumerate}

Regarding the additional points raised by the referee:
\begin{enumerate}
 \item We have changed $\Delta x^2$ to $\Delta^2 x$ ($\Delta x^2$ made sense in the case of flat skies, where this is just $\Delta x\,\Delta y$).
 \item We have added a comment about bandpower averaging before Eq. 2.12.
\end{enumerate}

Changes related to the CMB background:
\begin{enumerate}
  \item We have expanded the list of references to the CMB legacy for covariance matrix estimation in the introduction, by including a reference to GPE2, to the Planck 2013, 2015 and 2018 likelihood papers, and to Challinog \& Chon 2005 (we also cite the latter in section 2 in the context of mask gradients).
  \item We have toned down the statement about the higher complexity of LSS masks.
  \item We have added a statement about the role of anisotropic noise and other complications at the end of Section 4.
\end{enumerate}

Other modifications:
\begin{enumerate}
  \item We have updated the references to CMB in line 3 of the introduction.
  \item We have added a citation to the HEALPix URL and the corresponding references and acknowledgements.
\end{enumerate}

\end{document}
